
# General agriculture

## Nepal agriculture and geography

**Ecological regions**

- Himalayan region: 35%
- Hilly region: 41.67%
- Terai region: 23.11%

**Agricultural land**

- Cultivated land: 3091000 ha (21%)
- Cultivated land uncultivated: 1030 (6.99%)
- Jungle and shrubs: 5828 (39.597%); Shrub alone: 1560
- Pasture land: 1766000 (11.99%)
- Total Jungle = Shrub + Pasture

**GDP of Agriculture and releated sector/subsectors**

```{r}
dplyr::tribble(~"sector", ~"fy_2074-75", ~"fy_2073-74", ~"fy_2072-73", 
               "Agriculture and forestry", 27.10, 28.25, 31.08,
               "Fishery", 0.49, 0.51, 0.53,
               "Non-agriculture", 72.41, 71.24, 68.39) %>% 
  knitr::kable(booktabs = TRUE, caption = "Percentage contribution of different sector and subsectors to total national GDP", col.names = c("Sector", "FY 2074/75", "FY 2073/74", "FY 2072/73")) %>% 
  kableExtra::kable_styling(latex_options = "striped")
```

**Agricultural growth rates**

```{r}
dplyr::tribble(~"Year", ~"Growth rate (%)", 
               2072/73, 0.01, 
               2073/74, 5.14,
               2074/75, 2.72) %>% 
  knitr::kable(caption = "Agriculture growth rates of recent years in Nepal.", booktabs = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped")
```

## Chronology of Agriculture development in Nepal

```{r agriculture-chronology}
readxl::read_xlsx("./data/agriculture_development_chronology.xlsx", skip = 1) %>% 
  select(-1) %>% 
  knitr::kable(caption = "Chronology of Agriculture development in Nepal", booktabs = TRUE) %>% 
  kableExtra::kable_styling(latex_options = c("striped", "HOLD_position"))
```


## Development of cooperatives in Nepal

- Traditionally, custom of _Parma_, _Mankakhal_, _Dharmabhakari_, _Dhikuri_/ _Dhukuti_ were on place very early.
- In 2010, Cooperative development department was established.
- In 2013, government implemented executive guidelines.
- In 2016, Cooperative organization act came into force.
- In 2018, Cooperative organization regulations came into force.
- In 2019, Cooperative training center was established.
- In 2020, Cooperative bank was established.
- In 2024, "Village return campaign" was initiated under cooperative program.
- In 2027, Agriculture development bank started managing cooperative organizations.
- In 2033, "Sajha" program was implemented in 30 districts establishing multipurpose cooperatives in VDCs.
- In 2035, Agriculture development bank handed over the managerial responsibility of cooperatives to management board.
- In 2041, Cooperative ("Sajha") organization act came into force.
- In 2048, National cooperative development board was set up.
- In 2048, Cooperative act came into force (new)
- In 2049, National Cooperative Association was established.
- In 2069, Ministry of cooperative and poverty alleviation was formed.

- Cooperative campaign was first initiated when implementing first five year plan (In 2013), wherein Rapti Dunn Project was launched and 378 cooperative organizations were established in the course.

Importance of cooperatives

- Poverty alleviation
- Social and economic support to low income households
- Combines efforts of small producers and consumers and leads them to a larger scale commercial operation
- Reduces transaction cost, when compared to individual efforts
- Reduces chances of suppression and exploitation by large scale acting merchants
- Effective utilization of available resources and supplies
- Empowerment of local community
- Qualitative development of labor
- Improves bargaining power
- Has a bigger role in operationizing in market demand and supply forces
- Members have ownership in rural development
- Promotion of community welfare
- Community development based on justice and equality
- Produce can have more accessible market
- Quality service could be delivered
- Effective mobilization of consumable goods
- Support in national economy

- 10th five year plan focused on emphasized agriculture based commerce and industries, cooperative farming, cooperative based agriculture input supply, and cooperative based small farmer irrigation project implementation, dairy collection and supply and other cooperative based approach to community development.
- The plan promoted active participation of cooperative groups and private sectors for the efforts.
- National agriculture policy, 2061 has prioritized cooperative based agriculture industries and commerce promotion, farmer group cooperative formation and agriculture wholesale market and haat bazzar established and management via cooperatives.
- To legalize a cooperative organization establishment, it should be registered:
  - Under Organization act in District Administration Office, or
  - According to Cooperative act and regulations under Cooperative Division Office.

## Budget speech, 2076-77 BS

Total budget announced was Rs 1.53 trillion for fiscal year 2019/20 (76/77). Economic growth rate was set at 7% as target. The budget focuses on social justice, increment of export to reduce trade deficit and increase in general productivity.

NRs 130 billion is to be distributed from revenue between provincial and local levels. Education will be freed upto secondary level. In terms of literacy, 70 districts will be designated the status of fully literate districts. To that end, NRs 2 billion will be appropriated for colorful textbooks for primary level. Day meals for 2.2 million school children will be provisioned and sanitary pads will be free for female students attending public schools.

Over 10 billion was allocated for Madan Bhandari science and technology university. It is also stated that science and technology laboratory will be established in each province.

NRs 6 billion will be allocated for free insurance in all districts.

Health grants will be increased for treatment of 8 types of severe illnesses. Likewise, NRs 2.2 billion is appropriated for new-mothers travel expenses. 52000 female community health volunteers will be provided Rs 3000 annual allowance. NRs 5 billion is dedicated to establish health service providing facilities at local levels. Similarly, Rs 400 milion is allocated for Ramraja Prasad Singh Hospital in Rajbiraj. Rs 400 million appropriated for betterment of Bir Hospital.

Addressing issue of public health, smoking and drinking will be banned in public places. 92% of the population will be provided full access to drinking water in coming fiscal year. NRs 7 billion is allocated for completion of Melamchi drinking water project. Over 43 billion NRs is allocated for drinking water and hygiene.

Social security allowance to pregnant women and elderly senior citizens allowance sees an increment of Rs 1000 (increased from Rs 2000 to 3000). There have been betterment of employment schemes for the peoples with disabilities.

Coming fiscal year to marked as youth mobilization year. Youth scientists conference to be held in the coming year.

In agriculture, grants will be provided for purchase of agricultural products and technology. Schemes for achieving self sufficiency in dairy, poultry and fresh vegetables will be prepared and implemented. Grants will be provided for fertilizer input purchase. Organic farming wil be encouraged. With the dubling of fruit cultivation in next 5 years, food quality labs will also be set in every province. Rs 500 million is appropriated for community farming. Rs 34 billion is allocated for agriculture.

A budget of 23.6 billion is allocated for irrigation programmes. 960 million allocated for irrigation programmes in Terai. Sunkoshi Marine to be developed as a national pride project, with 2.05 billion NRs allocated for program initiation. NRs 5.6 billion is allocated for construction of dams.

Next fiscal year will be marked as tree plantation year. Security to be beefed up in forest areas. Newer programs/practices to be launced in livestock management. Rs 1 billion is dedicated for 'Rastrapati Chure Programme'.

Under land management, a revised land management act will be introduced for sustainable utilization of land. Encroached public lands will be brought under government management within next fiscal year. Online issuance of land ownership certificate will have been started by next two years.

Tourism sector will be prioritized, with stress on infrastructure development of main tourist areas. Trail connecting Darchula to taplejung to be developed. Operation of cable cars will be encouraged in mountainous regions.

Government officials will be mandated to only gift homemade (domestic) products as and when needed.

Private and cooperative sector to be encouraged for production of necessary commodities. Local cement and wire frames to be encouraged in construction. To be self sufficient in production of at least 2 dozen products.
  - Hetauda and Udayapur cement factories to be made more efficient 
  - Local products will be promoted in construction inputs
  - Import of luxury goods and health unfriendly products to be discouraged
  - Economic zone to be established in Kavre and Nuwakot. 50% concession to Nepali textile industry on electricity.

Infrastructure development at trade transit points in the north. Business with third countries to begin via Chinese port. Completion of pending works on pipeline by the next year to facilitate import of petroleum products.

In energy, at least two big hydropower projects will be embarked on in all seven provinces. NRs 13 billion is allocated for Budigandaki, 2.02 for Budhiganga-Tamor project to be a national pride project. Province 2, Karnali province and Sudurpaschim province to have full access to electricity. Rs. 4.5 billion allocated for rural electrification. Waste-to-energy programmes will be encouraged.

Rs 163 billion is appropriated for railway and waterways. Digital payment system will be installed in public transport. 1.5 billion for transport management. Electric buses to be introduced in "major" cities. NEA to install charging stations. Additional budget is allocated for development of Narayani and Koshi waterways.

Under urban development, feasibility study will be conducted for development of mega-city and smart city. 530 million allocated to replace thatched roofs of 20000 houses. 4.3 billion for construction of 30000 houses under housing project.

Convention center with a capacity to accomodate 3000 people to be constructed in the valley, this year. City halls to be constructed in local levels. Compulsory footpath, underground cable management. Rs 47.7 billion allocated for housing and urban development. MP's constituency budget (MP fund) increased to Rs 60 million. 1 billion for Dharahara reconstruction (to be completed within the next two years). 141 billion appropriated for reconstruction. Rs 9 billion isolated for infrastructural development in each electoral constituency.

Online visa services will be furnished for foreign nationals. National Defence University to be established. 18-20% (non-gazetted and gazetted) increment in salaries of public service personnel. National knowledge centre to be established at Khumaltar, Tribhuvan International Airport to be upgraded to a boutique airport. Gautam Buddha International airport to come into operation next year. Contractors to be held responsible for the upkeep of their projects for years after completion.

VAT and other taxes to be made more effective through improved taxation system VAT rates to stay intact changes in customs rate. Import reliance will be significantly reduced. 

The budget size is 1.53 trillion, recurrent budget is 957.1 billion, capital expenditure is 408.59 billion and financing 167.5 billion.

## Food security

For a more fundamental discussion of food security topic, refer to [lecture handout](https://github.com/DeependraD/life_science_presentations/raw/master/food_security.pdf), in link to life science presentations.

FAO defined food security in 1996 A.D. In 1996, convention held in Rome approved seven points of food soverignty. Nepal has been a member of FAO since 2007 BS (1950 AD).

McMichael's projection:
  - 2 billion people suffer hidden hunger
  - 1.5 billion people suffer over nutrition related problems

Since 1981 (2038 BS), FAO started celebrating World Food Day on October 16th. In Nepal, 34th World Food Day was celebrated in 2014 with the slogan: "Feeding the world, caring for the earth".

According to FAO's projection, 52% of the population in South Asia are dependent on Agriculture while agriculture contributes 22% to the total GDP in the region.

Pillars of food security:

1. Food availability
2. Food access
  - Nepal's poverty rate in 2001/2002 AD: 30%
  - Nepal's poverty rate in 2010 AD: 26%
  - Nepal's poverty rate at the end of 12th plan: 23.8%
3. Food utilization
  - Per person per day recommendation by WHO: 250 ml milk (91 ltr per year)
  - Nepal's aim for per day milk consumption: 156 ml (57 ltr per year)
  - Nepal's aim for per year meat consumption: 14 kg
  - Nepal's aim for per year egg consumption: 48
  - 30% of total daily protein requirement should be met by animal sources.
4. Food stability
  
- In 2011 AD, Nepal ranked 157th among 187 nations in HDI
- In 2014 AD, Nepal ranked 145th in HDI
- Poverty rate at village/rural areas: 27% (Survey, 2012)
- Poverty rate at cities: 15% (Survey, 2012)
- Monthly poverty:
  - Maximum at Chaitra-Baisakh
  - Minimum at Mangsir-Poush
  
State of food security in Nepal:

- Untill 2042 BS (1985), production was twice that satisfied the population's need.
- Upto 2047 BS, food insufficieny was absent.
- Currently 40 districts are declared food minimum and more than 27 districts of high hill and mountain districts are food insecure.
- Calorie deficiency is prevalent in 41% of population.
- In 19 districts of Midwestern Development Region and Eastern Development Region, food security project is launched.
- 46% of the total cultivated area is under rice.
- Of the total food production, contribution of respective crops are:
  - Rice: 56%
  - Maize: 24%
  - Wheat: 17%
  - Millet: 3.6%
  - Barley: 0.29%

## Nepal agriculture research council

- Has following organizational structure:
  - Executive director
  - Director, Planning and cooperation
  - Director, Crop and horticulture research
  - Director, Livestock and fisheries research
  - Director, Employee administration
  - Director, Economic adminsistration
  - Head, Planning commission
  - Communication, publication and inventory commission
  - Socio-economic and agricultural research policy commission
  - External research directorate
  - Agriculture environment directorate
  
- Programs under Agronomic and horticultural crop research programs
  - Rice crop research program, Baniniya, Dhanusha
  - Maize research program, Rampur, Chitwan
  - Wheat crop research program, Bhairahawa, Rupandehi
  - Legume crop research program, Nepalgunj
  - Oil crop research program, Nawalpur, Sarlahi
  - Hilly crop research program, Kavre, Dolakha
  - Sugarcane research program, Jitpur, Bara
  - Potato research program, Khumaltar
  - Ginger research program, Salyan
  - Orange variety research program, Dhankuta
  - Jute crop research program, Itahari, Sunsari
  - National commercial agriculture research program, Pakhribas, Dhankuta

- Divisions situated in Khumaltar
  - Crop science division
  - Crop disease science division
  - Entomology science division
  - Soil science division
  - Agri-engineering division
  - Agri-botany division
  - Horticulture research division
  - Food research division
  - Biotechnology division
  - Commercial crop division
  - Seed science and technology research division

Regional agriculture research centres and programs

1. Regional agriculture research centre, Nepalgunj
  - ARC, Surkhet
  - ARC, Doti
  - ARC, Jumla, Bijayanagar
  - ARC, Jumla, Rajikot
  - ARC, Dolakha

2. Regional agriculture research centre, Lumle
  - ARC, Baidam, Pokhara $\rightarrow$ Aquaculture
  - ARC, Begnas, Pokhara $\rightarrow$ Aquaculture
  - ARC, Malepatan, Pokhara $\rightarrow$ Horticulture
  - ARC, Bandipur, Tanahun $\rightarrow$ Goat

3. Regional agriculture research centre, Parwanipur
  - ARC, Rasuwa $\rightarrow$ Pasture
  - ARC, Ranighat, Birgung, Parsa $\rightarrow$ Agri-machinery
  - ARC, Trishuli $\rightarrow$ Acquaculture
  - ARC, Belachhapi, Dhanusha $\rightarrow$ Tobaccoo

4. Regional agriculture research centre, Tarahara
  - Pakhribas, Dhankuta
  - Tarahara $\rightarrow$ Aquaculture

## ADS

- MoAD, ADB, Rastriya Kishan Sanjal etc. including 12 development cooperatives (technical team) were involved in preparation of the strategy. 
- Vision (7 sub heads):
  - Self reliance
  - Sustainablity
  - Competition
  - Inclusive
  - Growth
  - Livelihood
  - Food and nutrition security

- There are 16 indicators ??

## National Farmers Commission

- Chariman: Chitra Bahadur Shrestha
- Members: 7 total
- 2073/10/06: National Farmers' Commission organization directives
- Has central office in Hariharbhawan, Lalitpur

## Prime minister agriculture modernization project

Initiation, project details ...

### Zones and superzones


## Crop and livestock insurance

## One village one commodity (ODOP)

## One district one commodity

## History of Agricultural Insitutions in Nepal

Institute of Agriculture and Animal Science (IAAS)

- Began as a school of Agriculture under ministry of Agriculture in 1957 to train JTA's in agriculture.
- School of agriculture was upgraded to college level in 1968.
- It was entitled IAAS in 1972 as a constituent institute of Tribhuwan University and located the specific place at Rampur in the same year.
- The institute has 230 ha of land for its academic and research programs at Rampur.
- Lamjung Campus located at Sundarbazar, Lamjung was established in 1975 and has about 18 hac of land.
- Paklihawa Campus at Rupandehi established in 1978 with area of about 22 hac of land.
- In 1998 the institute started M.Sc.Ag program.
- In 2001 the semester system of study in bachelor level was introduced.
- Ph.D degree course started in 2002.
- IAAS day: 4th of Bhadra
- Logo of IAAS was designed by Thakur Prasad Mainali

Agriculture and Forestry University (AFU)

- The first agriculture university in Nepal.
- Established in June 2010 at Rampur Chitwan.
- University formed merging the then IAAS Rampur and the then IOF Hetauda.
- Rampur academic complex extends to an areas of 280 hac
- Hetauda forestry campus has an area of 95 hac.
- 17th June (3rd Ashad) is celebrated as AFU day.
- Chancellor: Prime Minister of Nepal.
- 1st VC of AFU: Prof.Dr Kailash Nath pyakural
- 1st Registrar: Dr.Surya Kant Ghimire
- Vice-Chancellor: Ishwori Prasad Dhakal
- Registrar: Man Raj Kulakshyapati

## National Commodity Research Programs under NARC and their locations

- National Rice Research Program; Hardinath, Dhanusa
- National Wheat Research Program; Bhairahawa, Rupandehi
- National Maize Research Program; Rampur, Chitwan
- National Grain Legume Research Program; Khajura,Banke (recently transferred from Rampur, Chitwan)
- National Potato Research Program; Khumaltar, Lalitpur
- National Swine and Avian Research Program; Khumaltar, Lalitpur
- National Sheep and Goat Research Program; Jumla
- National Oil Seeds Research Program; Nawalpur, Sarlahi
- National Sugarcane Research Program; Jitpur, Bara
- National Citrus Research Program; Paripatle, Dhankuta
- National Jute Research Program; Itahari, Sunsari
- National Ginger Research Program; Kapurkot, Salyan
- Hill Crop Research Program; Kavre, Dolakha
- National Bovine Research Program; Khumaltar, Lalitpur
- National Tobacco Research Program; Belachapi, Janakpur
- Apple Processing Centre; Jumla
- Cotton Research Centre; Khajura, Banke

## Miscellaneous

- Nepal ranks 18th position among rice cultivating nation in the world.
- Father of Agronomy in Nepal: Netra Bahadur Basnyat
- Golden rice is genetically modified rice with yellow colored seeds rich in vitamin A
- Date palm is the oldest fruit.
- Rye is known as grain of poverty.
- Tomato is known as poor man's orange.
- DDT was the first discovered insecticide by Paul Muller.
- Maximum amount of sugar in honey is 73%.
- Mandarin is alled fancy fruits.
- Cardamom production highest in Nepal in the world.
- Dance of bee is called waggle dance.
- Apple production is highest in China.
- King of cereals is Wheat and Queen of cereals is Maize.
- Orchidaceae is the largest family among flowering plants.
- National Agriculture Policy in 2061 BS
- Gaurab hybrid and Rampur hybrid are the only hybrid varieties in Nepal in maize.
- Halo kranti started in 2006 BS from Duradanda, Lamung.
- OVOP (One Village One Product) practice in Nepal base on PPP (Public Private Partnership) approach started from Japan.
- One Village Two Product practice is prevailing in Bhutan.
- Year of Family farming: 2014
- Yellow color of Papaya is due to Cryptoxanthin.
- Edible part of apple is enlarged thalamus and edible part of cabbage is bud while that of cauliflower is inflorescence.
- Earthworm is farmer's friend.
- 39.6% of area of Nepal's land is covered by forest.
- Japan is Nepal's largest bilateral donor.
- Lentil is a major pulse crop with efficient export.
- Crop and livestock insurance started in Nepal in 2069 Magh 1
- Triticale is artificial cereal crop.
- China is the world's highest rice producer and Vietnam is the world's highest rice exporter.
- Azolla is widely used bio - fertilizer in Rice Crop.
- Amendment of acid soil is done by lime and of alkaline soil is done by CaSO4
- Tomato is known as poor man's orange.
- Air layering in litchi is to epicotyl grafting in Mango
- Anadi, Jumli Marshi, Pokhreli Jetho Budo are indigenous variety of Rice in Nepal.
- Readily available form of water to plants is capillary water.
- Limiting amino acid in Cereals are Lysine and threonine whereas limiting amino acid in legumes are methionine and tryptophan.
- Stubble is the portion of stem that is left on the field during harvest.
- Leguminous crop is considered as biological plough.
- First certified organic commodity exported from Nepal: Coffee
- Genetic dwarfism can be overcome by the use of Gibberelic Acid.
- Peripneustic type (having spiracles in a row on each side of the body) respiratory system is found in Lepidoptera larvae.
- Gilson's glands is found in order Trichopter.
- Growth curve of kiwi is Triple sigmoid type.
- Which produces cocoon in silkworm ? Pupa
- Chroma refers to: Relative purity of color
- Pointed gourd is commonly propagated through stem cutting.
- Allophane is an amorphous to poorly crystalline hydrous aluminium silicate clay mineraloid. It is common in vocanic soil and typically forms under mildly acidic to neutral pH (5–7). Its structure has been debated, but it is similar to clay minerals and is composed of curved alumina octahedral and silica tetrahedral layers.
- Horizon of maximum accumulation of Iron (Fe) and Aluminium oxides is B-horizon.
- Spinach can hybridize with Swiss chard and Garden beet.
- Soil with highest CEC is Vermiculite.
- Dioecious species: Spinach, pointed gourd and yam.
- Mo is available at slightly basic pH.
- GA is flowering hormone of longday and chilling requiring plants.
- Swelling of Smectite is more because of Mg binding.
- Chinese cabbage can hybridize with Turnip and Broad leaf mustard.
- Paclobutrazole is also known as anti GA hormone.
- Color of semi-aerated iron containing soil is Yellow.
- Sodicity is measured by SAR and ESP.
- Chayote amongst all vegetables has the highest productivity.
- Salt resistant crops: Sugarbeet, cotton, sweet potato.


Contributors in Agriculture

- Sociology: Auguste Comte
- Statistics: R.A. Fisher
- Economics: Adam Smith
- Soil Science: V.V. Dokuchaev
- Tillage: Jethro Tull
- Green Revolution: Norman E Borlaug
- Modern Agronomy: Peter De Crescent
- Organic Farming: Rudolf Steiner

- Dr. Norman E. Borlaug is the pioneer of Green Revolution and was awarded Nobel Peace Prize in 1970.
- Rice, wheat and maize are cereal crops while buckwheat and amaranthus are pseudocereals.
- Wheat and maize show hypogeal germination but soyabean and chickpea show epigeal germination.

- Tree of Paradise/wisdom: Banana
- King of Fruits/ Bathroom Fruit: Mango
- Bitter taste of Bitter Gourd due to: Cucurbitacin/ Bitter Glucoside
- Rice is semi - aquatic in nature.
- Critical stage for irrigation in wheat is CRI
- Cotton fiber is also known as White Gold.
- Apple of tropics/Poor man's apple: Guava
- Poor man's meat: Soyabean
- Poor man's friend/ King of Vegetable: Potato
- World's highest cereal production: Wheat
- Nepal's highest cereal production: Rice
- Queen of spice crops: Cardamom
- King of spice: Black pepper
- Queen of Flower: Gladiolus
- King of flower: Dahlia
- Butter fruit: Avocado
- Queen of Beverage: Tea
- King of Temperate fruit: Apple
- Drosophila of crop plants: Maize
- Vegetable meat: Cowpea

**NPK Content of fertilizer**

- Urea: 46%  Nitrogen
- DAP(Diammonium Phosphate): 18% N, 46% Phosphorus
- MoP(Muriate of Potash): 60% potash
- SSP(Single Superphosphate): 16% P
- DSP(Double Sugar Phosphate): 32% P
- TSP(Triple Sugar Phosphate): 48% P
- ZnSO4: 22-35% Zinc

**Kishan call center**

- Phone number: 1660-5652-999
- Active hours: 10 AM - 4:30 PM

- Goat is called poor man's cow.
- Male adult of horse is stallion and female adult of horse is mare.
- Heifer is young female cattle.
- Young male poultry is cockrel and young female poultry is pullet.
- Black gold: Buffalo
- Yellow gold: colostrums milk
- King of fodder: Elephant grass/Napier
- Queen of fodder: Leucerne
- Freshly drawn milk has the PH value in range of 6.5-6.7
- Evening milk is higher in fat than morning milk
- Amount of milk left in udder after normal milking is called residual milk.
- Water content of milk is 87-88%
- Gizzard in poultry is used grinding of food whereas proventiculus in poultry is the true stomach.
- Abomasum is true stomach in ruminants.
- Omasum is absent in camel and rumen is water storing sac.
- Sweetness in milk is due to Lactose.
- Milk holiday is the day when the dairy does not buy milk from the farmers.
- Milk storage is done at $5^\circ C$.
- 1 lit of milk can be synthesized by 500 lit. of blood circulation in body.
- Milk is complete nutrient. It lacks only Fe.
- Meat of calf is k/a veal.
- Ear notching is the best identification method in pig.
- Gall bladder is absent in horse.
- Elephant has longest gestation period i.e. 18-23 months(624days)
- Kavrepalanchowk is the highest milk producing district in Nepal.
- Semen is stored in liquid Nitrogen in $-196^\circ C$; the process of storage is k/a Cryopreservation.

Common animal diseases

- Bacterial diseases: Anthrax, Brucellosis (Bang's disease), Black Quarter(Black leg), Haemorrhagic Septicaemia, Mastitis
- Fungal diseases: Aspergillosis, Candidiasis, Ringworm
- Protozoal diseases: Coccidiosis, Babasiosis
- Viral diseases: Foot and Mouth disease, Swine fever(hog cholera), New castle disease, Marek's

Common Poultry Diseases

- New Castle Disease/Ranikhet
- Mareck's Disease (MD)
- Gumboro
- Coccidiosis
- Chronic Respiratory Disease (CRD)
- Avian Influenza/Bird Flu- H5N1, H9N2

**Agriculture and child labor**

According to the latest ILO estimates, at least 250 million children of between 5-14 years of age work in developing countries. Almost half of these children work on a full-time basis. The participation rates of children in economic activities in much higher in rural areas than in urban centres. Rural children, in particular girls, tend to start working at an early age. In Latin America and the Carribbean, out of 15 million children involved in the labor market, 56 percent work in agriculture sector from the age of 5 to 7 onwards. In some countries, children account for as much as 30 percent of the agricultural workforce. Most children work seven days a week and are paid less than the prevailing rates in their localities. They work long hours, and a very high proportion of these children are injured at work. Exposure to poor working conditions has serious repercussions on children's growth, development and health. The most common injuries include: cuts and wounds, eye infection, skin problems, fever, and headaches caused by excessive heat or by exposure to pesticides while working in agricultural fields.

**Hazards in agriculture occupation**

Despite the fact that certain developing countries have reached higher levels of economic development, nutrition and health are still the problem areas. This situation provokes a vicious circle of low productivity, low wages, malnutrition, ill-health and low working capacity. The interaction between poor living and working conditions determines a distinctive morbidity-mortality pattern among agricultural workers, which is due to the combination of malnutrition, general and occupational diseases, and complications arising from undiagnosed or untreated diseases. Low working capacity is closely related to workers' malnutrition and poor health. 

Agriculture related hazards mostly include:

1. Machinery such as tractors, trucks and harvestors, and cutting and piercing tools;
2. Hazardous chemicals: Pesticides, fertilizers, antibiotics and other veterinarian products;
3. Toxic or allergenic agents: Plants, flowers, dusts, animal waste, gloves, chrome, oils;
4. Carcinogenic substances or agents: Certain pesticides such as arsenicals and phenoxy-acetic herbicides, UV radiations, parasitic diseases such as bilharziasis and facioliasis;
5. Transmissible animal diseases: brucellosis, bovine tuberculosis, hydatid disease, tularaemia, rabies;
6. Other infectious and parasitic diseases: leishmaniasis, bilharziasis, facioliasis, malaria, tetanus, mycosis;
7. Confined spaces such as silos, pits, cellars and tanks;
8. Noise and vibration;
9. Ergonomic hazards: use of inadequate equipment and tools, unnatural body position or prolonged static postures, carrying of heavy loads, repetitive work, excessive long hours;
10. Extreme temperatures and adverse weather conditions;
11. Contact with wild and poisonous animals: insect, spider, scorpion, snake and certain wild mammals.

## Top exports by Nepal

1. Man made staple fiber (11.9%) -- US \$ 87.9 million
2. Textile floor covering (11.8%) -- US \$ 86 million
3. Coffee, tea, spices (6.4%) -- US \$ 47.6 million

Agriculture export

1. Cardamom
2. Fruit juice
3. Black tea
4. Forest and vegetable products
5. Betal nuts
6. Oilcakes and bran/feeds
7. Noodles
8. Medicinal plant
9. Lentil
10. Bovine meat (Buffalo)
