
# General agriculture

## Nepal agriculture and geography

**Ecological regions**

- Himalayan region: 35%
- Hilly region: 41.67%
- Terai region: 23.11%

**Agricultural land**

- Cultivated land: 3091000 ha (21%)
- Cultivated land uncultivated: 1030 (6.99%)
- Jungle and shrubs: 5828 (39.597%); Shrub alone: 1560
- Pasture land: 1766000 (11.99%)
- Total Jungle = Shrub + Pasture

**GDP of Agriculture and releated sector/subsectors**

```{r}
dplyr::tribble(~"sector", ~"fy_2074-75", ~"fy_2073-74", ~"fy_2072-73", 
               "Agriculture and forestry", 27.10, 28.25, 31.08,
               "Fishery", 0.49, 0.51, 0.53,
               "Non-agriculture", 72.41, 71.24, 68.39) %>% 
  knitr::kable(booktabs = TRUE, caption = "Percentage contribution of different sector and subsectors to total national GDP", col.names = c("Sector", "FY 2074/75", "FY 2073/74", "FY 2072/73")) %>% 
  kableExtra::kable_styling(latex_options = "striped")
```

**Agricultural growth rates**

```{r}
dplyr::tribble(~"Year", ~"Growth rate (%)", 
               2072/73, 0.01, 
               2073/74, 5.14,
               2074/75, 2.72) %>% 
  knitr::kable(caption = "Agriculture growth rates of recent years in Nepal.", booktabs = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped")
```

## Chronology of Agriculture development in Nepal

```{r agriculture-chronology}
readxl::read_xlsx("./data/agriculture_development_chronology.xlsx", skip = 1) %>% 
  select(-1) %>% 
  rename(`Year (BS)` = `Year_BS`) %>% 
  knitr::kable(caption = "Chronology of Agriculture development in Nepal", booktabs = TRUE, longtable = TRUE) %>% 
  kableExtra::kable_styling(latex_options = c("striped", "HOLD_position"), font_size = 10) %>% 
  kableExtra::column_spec(column = 5, width = "16em")
```


## Development of cooperatives in Nepal

- Traditionally, custom of _Parma_, _Mankakhal_, _Dharmabhakari_, _Dhikuri_/ _Dhukuti_ were on place very early.
- In 2010, Cooperative development department was established.
- In 2013, government implemented executive guidelines.
- In 2016, Cooperative organization act came into force.
- In 2018, Cooperative organization regulations came into force.
- In 2019, Cooperative training center was established.
- In 2020, Cooperative bank was established.
- In 2024, "Village return campaign" was initiated under cooperative program.
- In 2027, Agriculture development bank started managing cooperative organizations.
- In 2033, "Sajha" program was implemented in 30 districts establishing multipurpose cooperatives in VDCs.
- In 2035, Agriculture development bank handed over the managerial responsibility of cooperatives to management board.
- In 2041, Cooperative ("Sajha") organization act came into force.
- In 2048, National cooperative development board was set up.
- In 2048, Cooperative act came into force (new)
- In 2049, National Cooperative Association was established.
- In 2069, Ministry of cooperative and poverty alleviation was formed.

- Cooperative campaign was first initiated when implementing first five year plan (In 2013), wherein Rapti Dunn Project was launched and 378 cooperative organizations were established in the course.

Importance of cooperatives

- Poverty alleviation
- Social and economic support to low income households
- Combines efforts of small producers and consumers and leads them to a larger scale commercial operation
- Reduces transaction cost, when compared to individual efforts
- Reduces chances of suppression and exploitation by large scale acting merchants
- Effective utilization of available resources and supplies
- Empowerment of local community
- Qualitative development of labor
- Improves bargaining power
- Has a bigger role in operationizing in market demand and supply forces
- Members have ownership in rural development
- Promotion of community welfare
- Community development based on justice and equality
- Produce can have more accessible market
- Quality service could be delivered
- Effective mobilization of consumable goods
- Support in national economy

- 10th five year plan focused on emphasized agriculture based commerce and industries, cooperative farming, cooperative based agriculture input supply, and cooperative based small farmer irrigation project implementation, dairy collection and supply and other cooperative based approach to community development.
- The plan promoted active participation of cooperative groups and private sectors for the efforts.
- National agriculture policy, 2061 has prioritized cooperative based agriculture industries and commerce promotion, farmer group cooperative formation and agriculture wholesale market and haat bazzar established and management via cooperatives.
- To legalize a cooperative organization establishment, it should be registered:
  - Under Organization act in District Administration Office, or
  - According to Cooperative act and regulations under Cooperative Division Office.

## Budget speech, 2076-77 BS

Total budget announced was Rs 1.53 trillion for fiscal year 2019/20 (76/77). Economic growth rate was set at 7% as target. The budget focuses on social justice, increment of export to reduce trade deficit and increase in general productivity.

NRs 130 billion is to be distributed from revenue between provincial and local levels. Education will be freed upto secondary level. In terms of literacy, 70 districts will be designated the status of fully literate districts. To that end, NRs 2 billion will be appropriated for colorful textbooks for primary level. Day meals for 2.2 million school children will be provisioned and sanitary pads will be free for female students attending public schools.

Over 10 billion was allocated for Madan Bhandari science and technology university. It is also stated that science and technology laboratory will be established in each province.

NRs 6 billion will be allocated for free insurance in all districts.

Health grants will be increased for treatment of 8 types of severe illnesses. Likewise, NRs 2.2 billion is appropriated for new-mothers travel expenses. 52000 female community health volunteers will be provided Rs 3000 annual allowance. NRs 5 billion is dedicated to establish health service providing facilities at local levels. Similarly, Rs 400 milion is allocated for Ramraja Prasad Singh Hospital in Rajbiraj. Rs 400 million appropriated for betterment of Bir Hospital.

Addressing issue of public health, smoking and drinking will be banned in public places. 92% of the population will be provided full access to drinking water in coming fiscal year. NRs 7 billion is allocated for completion of Melamchi drinking water project. Over 43 billion NRs is allocated for drinking water and hygiene.

Social security allowance to pregnant women and elderly senior citizens allowance sees an increment of Rs 1000 (increased from Rs 2000 to 3000). There have been betterment of employment schemes for the peoples with disabilities.

Coming fiscal year to marked as youth mobilization year. Youth scientists conference to be held in the coming year.

In agriculture, grants will be provided for purchase of agricultural products and technology. Schemes for achieving self sufficiency in dairy, poultry and fresh vegetables will be prepared and implemented. Grants will be provided for fertilizer input purchase. Organic farming wil be encouraged. With the doubling of fruit cultivation in next 5 years, food quality labs will also be set in every province. Rs 500 million is appropriated for community farming. Rs 34 billion is allocated for agriculture.

A budget of 23.6 billion is allocated for irrigation programmes. 960 million allocated for irrigation programmes in Terai. Sunkoshi Marine to be developed as a national pride project, with 2.05 billion NRs allocated for program initiation. NRs 5.6 billion is allocated for construction of dams.

Next fiscal year will be marked as tree plantation year. Security to be beefed up in forest areas. Newer programs/practices to be launced in livestock management. Rs 1 billion is dedicated for 'Rastrapati Chure Programme'.

Under land management, a revised land management act will be introduced for sustainable utilization of land. Encroached public lands will be brought under government management within next fiscal year. Online issuance of land ownership certificate will have been started by next two years.

Tourism sector will be prioritized, with stress on infrastructure development of main tourist areas. Trail connecting Darchula to taplejung to be developed. Operation of cable cars will be encouraged in mountainous regions.

Government officials will be mandated to only gift homemade (domestic) products as and when needed.

Private and cooperative sector to be encouraged for production of necessary commodities. Local cement and wire frames to be encouraged in construction. To be self sufficient in production of at least 2 dozen products.

- Hetauda and Udayapur cement factories to be made more efficient 
- Local products will be promoted in construction inputs
- Import of luxury goods and health unfriendly products to be discouraged
- Economic zone to be established in Kavre and Nuwakot. 50% concession to Nepali textile industry on electricity.

Infrastructure development at trade transit points in the north. Business with third countries to begin via Chinese port. Completion of pending works on pipeline by the next year to facilitate import of petroleum products.

In energy, at least two big hydropower projects will be embarked on in all seven provinces. NRs 13 billion is allocated for Budigandaki, 2.02 for Budhiganga-Tamor project to be a national pride project. Province 2, Karnali province and Sudurpaschim province to have full access to electricity. Rs. 4.5 billion allocated for rural electrification. Waste-to-energy programmes will be encouraged.

Rs 163 billion is appropriated for railway and waterways. Digital payment system will be installed in public transport. 1.5 billion for transport management. Electric buses to be introduced in "major" cities. NEA to install charging stations. Additional budget is allocated for development of Narayani and Koshi waterways.

Under urban development, feasibility study will be conducted for development of mega-city and smart city. 530 million allocated to replace thatched roofs of 20000 houses. 4.3 billion for construction of 30000 houses under housing project.

Convention center with a capacity to accomodate 3000 people to be constructed in the valley, this year. City halls to be constructed in local levels. Compulsory footpath, underground cable management. Rs 47.7 billion allocated for housing and urban development. MP's constituency budget (MP fund) increased to Rs 60 million. 1 billion for Dharahara reconstruction (to be completed within the next two years). 141 billion appropriated for reconstruction. Rs 9 billion isolated for infrastructural development in each electoral constituency.

Online visa services will be furnished for foreign nationals. National Defence University to be established. 18-20% (non-gazetted and gazetted) increment in salaries of public service personnel. National knowledge centre to be established at Khumaltar, Tribhuvan International Airport to be upgraded to a boutique airport. Gautam Buddha International airport to come into operation next year. Contractors to be held responsible for the upkeep of their projects for years after completion.

VAT and other taxes to be made more effective through improved taxation system VAT rates to stay intact changes in customs rate. Import reliance will be significantly reduced. 

The budget size is 1.53 trillion, recurrent budget is 957.1 billion, capital expenditure is 408.59 billion and financing 167.5 billion.

## Food security

For a more fundamental discussion of food security topic, refer to [lecture handout](https://github.com/DeependraD/life_science_presentations/raw/master/food_security.pdf), in link to life science presentations.

FAO defined food security in 1996 A.D. In 1996, convention held in Rome approved seven points of food soverignty. Nepal has been a member of FAO since 2007 BS (1950 AD).

McMichael's projection:

- 2 billion people suffer hidden hunger
- 1.5 billion people suffer over nutrition related problems

Since 1981 (2038 BS), FAO started celebrating World Food Day on October 16th. In Nepal, 34th World Food Day was celebrated in 2014 with the slogan: "Feeding the world, caring for the earth".

According to FAO's projection, 52% of the population in South Asia are dependent on Agriculture while agriculture contributes 22% to the total GDP in the region.

Pillars of food security:

1. Food availability
2. Food access
  - Nepal's poverty rate in 2001/2002 AD: 30%
  - Nepal's poverty rate in 2010 AD: 26%
  - Nepal's poverty rate at the end of 12th plan: 23.8%
3. Food utilization
  - Per person per day recommendation by WHO: 250 ml milk (91 ltr per year)
  - Nepal's aim for per day milk consumption: 156 ml (57 ltr per year)
  - Nepal's aim for per year meat consumption: 14 kg
  - Nepal's aim for per year egg consumption: 48
  - 30% of total daily protein requirement should be met by animal sources.
4. Food stability
  
- In 2011 AD, Nepal ranked 157th among 187 nations in HDI
- In 2014 AD, Nepal ranked 145th in HDI
- Poverty rate at village/rural areas: 27% (Survey, 2012)
- Poverty rate at cities: 15% (Survey, 2012)
- Monthly poverty:
  - Maximum at Chaitra-Baisakh
  - Minimum at Mangsir-Poush
  
State of food security in Nepal:

- Untill 2042 BS (1985), production was twice that satisfied the population's need.
- Upto 2047 BS, food insufficieny was absent.
- Currently 40 districts are declared food minimum and more than 27 districts of high hill and mountain districts are food insecure (when ??).
- Calorie deficiency is prevalent in 41% of population.
- In 19 districts of Midwestern Development Region and Eastern Development Region, food security project is launched.
- 46% of the total cultivated area is under rice.
- Of the total food production, contribution of respective crops are:
  - Rice: 56%
  - Maize: 24%
  - Wheat: 17%
  - Millet: 3.6%
  - Barley: 0.29%

One worrying aspect related to poverty is malnutrition. Indicators of malnutrition, particularly of children are still high not only in traditionally food deficit areas but increasingly also in food surplus areas. About 42% of children less than 5 years old suffer from stunting (NLSS 2010/11).

Three and half million people in Nepal, 13% of the population, are considered to be moderately to severely food insecure, and 42 out of 75 districts are classified as food insecure with respect to food grains (Draft Food and Nutrition Security Plan, MOAD, 2012).

The national cereal balance witnessed surplus of about 886,000 tons (equivalent to over 17% of total requirement) in 2011/12 ?? but in the past decade Nepal's average food grain import was about 5% of domestic production annually. Despite surplus, about 1.8 million people received staple food supplements from government.

## Nepal agriculture research council

- Has following organizational structure:
  - Executive director
  - Director, Planning and cooperation
  - Director, Crop and horticulture research
  - Director, Livestock and fisheries research
  - Director, Employee administration
  - Director, Economic adminsistration
  - Head, Planning commission
  - Communication, publication and inventory commission
  - Socio-economic and agricultural research policy commission
  - External research directorate
  - Agriculture environment directorate
  
- Programs under Agronomic and horticultural crop research programs
  - Rice crop research program, Baniniya, Dhanusha
  - Maize research program, Rampur, Chitwan
  - Wheat crop research program, Bhairahawa, Rupandehi
  - Legume crop research program, Nepalgunj
  - Oil crop research program, Nawalpur, Sarlahi
  - Hilly crop research program, Kavre, Dolakha
  - Sugarcane research program, Jitpur, Bara
  - Potato research program, Khumaltar
  - Ginger research program, Salyan
  - Orange variety research program, Dhankuta
  - Jute crop research program, Itahari, Sunsari
  - National commercial agriculture research program, Pakhribas, Dhankuta

- Divisions situated in Khumaltar
  - Crop science division
  - Crop disease science division
  - Entomology science division
  - Soil science division
  - Agri-engineering division
  - Agri-botany division
  - Horticulture research division
  - Food research division
  - Biotechnology division
  - Commercial crop division
  - Seed science and technology research division

Regional agriculture research centres and programs

1. Regional agriculture research centre, Nepalgunj
  - ARC, Surkhet
  - ARC, Doti
  - ARC, Jumla, Bijayanagar
  - ARC, Jumla, Rajikot
  - ARC, Dolakha

2. Regional agriculture research centre, Lumle
  - ARC, Baidam, Pokhara $\rightarrow$ Aquaculture
  - ARC, Begnas, Pokhara $\rightarrow$ Aquaculture
  - ARC, Malepatan, Pokhara $\rightarrow$ Horticulture
  - ARC, Bandipur, Tanahun $\rightarrow$ Goat

3. Regional agriculture research centre, Parwanipur
  - ARC, Rasuwa $\rightarrow$ Pasture
  - ARC, Ranighat, Birgung, Parsa $\rightarrow$ Agri-machinery
  - ARC, Trishuli $\rightarrow$ Acquaculture
  - ARC, Belachhapi, Dhanusha $\rightarrow$ Tobaccoo

4. Regional agriculture research centre, Tarahara
  - Pakhribas, Dhankuta
  - Tarahara $\rightarrow$ Aquaculture

## Agricultural Perspective Plan

- Finalized: 1995 with technical assistance of ADB
- Implemented for the period: 2052/53 -- 2071/72 BS

The Agricultural Perspective Plan (APP) was designed to increase agricultural growth whereby per capita AGDP will grow from its 1995 level of 0.5% to 4% per year. This growth was expected to stimulate nonagricultural growth in employment-intensive goods and services in both urban and rural areas. This would open up job opportunities for the poor, particularly poor women, and thereby help reduce the number of rural poor. With implementation of the APP the incidence of poverty was expected to come down from 42% in 1991/92 to 14% in 2014/15, whereas the latter figure without the APP would have been 29%. Poverty in 2011 is estimated at 25%, which is still far from the APP target. The increase in agricultural productivity was also expected to help protect the environment by removing the most fragile land resources from agriculture and putting them under suitable forest cover and other sustainable uses.

The overall objectives of APP were as follows:

1. Accelerate the growth rate in agriculture through increased factor productivity;

2. Alleviate poverty and achieve significant improvement in the standard of living through accelerated growth and expanded employment opportunities;

3. transform agriculture from subsistence to commercial orientation through diversification and realization of comparative advantage; 

4. Expand opportunities for overall economic transformation by fulfilling the preconditions of agricultural development; 

5. Identify immediate, short-term and long-term strategies for implementation, and provide clear guidelines for preparing future periodic plans and programs.

The APP strategy is to accelerate the agricultural growth rate sufficiently to obtain strong multiplier effects on growth and employment, in both the agricultural and non-agricultural sectors. This growth would occur through technological change to be achieved through investment in research and extension. The APP aims for a broad-based participatory growth across regions and income classes and emphasizes sub-sectors particularly important to women.

The following six strategic thrusts are identified as essential to achieve APP objectives: 

1. A technology-based green revolution in agriculture which becomes the initial engine of accelerated growth;
2. Accelerated agricultural growth which creates a demand-pull for the production of high-value commodities in agriculture, as well as for non-agricultural commodities, with consequent large multiplier effects on other sectors of the economy;
3. Broadly based high employment growth, which then becomes the mechanism for achieving societal objectives;
4. Public policy and investment focus on a small number of priorities, building on past investment in human capital and physical and institutional infrastructure;
5. A package approach to development, which would be different for the Terai, Hills and Mountains and would recognize the powerful complementation between public and private investment and priorities, and would ensure their co-ordination;
6. A regionally balanced and gender-balanced approach that explicitly ensures the participation of women and therefore achieves broad participation.

Targets

1. Increase country's agricultural growth by 2% (3% -- 5%)
2. Reduce poverty povery from 49% to 14%
3. Increase per person agriculture income from 0.5% to 3% (6 folds)
4. Increase contribution of livestock farming from 31% to 45%
5. Increase in livestock products growth rate from 2.3 to 6.2%
6. Contribution of agriculture in GDP was expected to decrease from 41.7% to 25%.
7. HVC production rate increase from 4.8% to 5.8%
8. Full unemployment decrease from 4.7% to 3% and half unemployment decrease to 10% from 47%.
9. Crop growth rate increase from 3.17% to 3.71%.
10. Horticulture production increase from 4.84% to 5.8%.
11. Fruit production increment per year from 5.13% to 6.29%.
12. Vegetable production increment from 4.63% to 5.4%.

- Priority animal production: 1. Milk, 2. Meat, 3. Fish
- Priority crop production: 1. Fruits, 2. Vegetables, 3. Ornamentals
- Priority production according to geography:
  - High hills: Apple, livestock raising, bee farming, NTFPs and potato seed production
  - Mid hills: Citrus and livestock raising
  - Terai: Offseason vegetables, cereals, cash, industrial crops and livestock

### Priority inputs of APP

**Irrigation**

- Irrigate from 459,000 ha at the base year to 1126,000 ha by the completion of plan.
- Annual increase in 34,000 ha of year round irrigated area.
- Share to terai 70-79% in well controlled year round irrigation.
- Increase in groundwater 141,000 --> 612,000 ha (24,000 ha per year of which 22,000 ha will be irrigated by shallow tube well and 2000 by deep tube well).
- On average, 8800 shallow tube wells and 40 deep wells per year will be established.
- Coverage:
  - Shallow tube well: 2-5 ha
  - Deep tube well: 50 ha
- Decrease unirrigated area from 57% to 14% (in terai)
- Decrease unirrigated area from 76% to 67% (in hills)
- Total cropped area will increase from 4103,000 ha to 4815,000 ha.
- Annual addition of 35,000 ha area to cropped area due to cropping intensification.

**Fertilizer**

- Increase use from 101,000 mt to 628,000 mt (26,000 mt/year increase).
- Increase use:
  - In terai: from 70,000 mt to 436,000 mt.
  - In hills and mountains: 31,000 mt to 192,000 mt

- At the end of plan fertilizer used per crop area:
  - Terai: 152 kg/ha
  - Hills: 101 kg/ha
  - Mountain: 38 kg/ha

**Road and power**

- Terai: 20 km per 100 sq km of mapped agricultural land (50% higher than base year)
  - length of agriculture road in terai: 34,000 km.
- Hill: 11 km per 100 sq km of mapped agricultural land
  - length of agricultural road in hill: 1950 km
- Mountain: 4 km per 100 sq km of mapped agricultural land
  - length of agricultural road in mountain: 850 km
  
- By the end of 2014/15 total length of Agricultural road in Nepal will be 6200 km
- 22 districts without road will need 1837 km of roads to connect their headquarters
- Electrification: Total length of rural electrification will be 9496 of distribution length.

**Technology**

- Green revolution (developing, improving and disseminating technology suitable for Nepal)
- Improved technology brings increased specialization and requires low transaction costs
- Production oriented applied research program on increse in fertilizer efficiency through proper timing and placement, complementary use of organic fertilizer, proper balance of nutrients, attenuation of trace elements' deficiencies, keyed to the different land utilization categories.

### Priority outputs

1. Livestock

- Share of livestock to AGDP: 31 % --> 33 % by the end of plan.
- Per capita livestock GDP of hills and mountains two times as much as in terai (Livestock growth rate: 2.9 % --> 6%)

2. High value crops (HVCs)

- The income from HVCs will be three times by the end of the plan.
- Annual growth rate: 4.8% - 5.8%
- Share of AGDP: 13% - 15%
- Commodity priorities:
  - Mid hills: Citrus
  - High hills/inner himalayan zone: Apple
  - Hill and mountain: Vegetable and flower seed
  - Hill and mountain: Bee keeping
  - Hills: Raw silk
  - Hill and terai: Off season vegetable

3. Agribusiness

- Provides opportunity for women to achieve some degree of independence.
- Dominant subsectors: Sericulture, dry ginger processing, cardamom drying, fruit processing, cut flower and saffron
- Target: Institutional development and private investment.

4. Forestry

- Top four forestry priorities:
  i. Community forestry in hills and mountains
  ii. Commercial management in terai
  iii. Private and leasehold forestry
  iv. Training, research and development
- Share of forestry will be 2.3% of AGDP through establishment of Community Forest User Groups (CFUGs).

- Cropping intensity to be increased from 150 to 250
- Fertilizer used per cropped area will be increased from 24 nutrients/ha to 94 nutrients per ha.
- Number of people below poverty line will be reduced from 9.8 million to 3.8 million.

### Progress summary

APP design gave priority to certain key inputs (i.e., irrigation, fertilizer, technology, roads and power, and financial credit for agriculture), and key outputs (i.e. livestock, high value crops, agribusiness and forestry). Agricultural sector growth has been less than the APP target of 4 % annual AGDP growth, achieving an average of 3.0 %, slightly below the national GDP average growth of 3.5% in the past decade and rising only after 2010 (National Accounts Estimates, Nepal, 2012). 

The agricultural sector in Nepal has made progress in several indicators of well-being and development. For example, income per capita and productivity of agricultural labor has increased, poverty has reduced, and malnutrition has declined. The road network has considerably expanded and irrigation cover has increased as well. In almost all agriculture subsectors (crop, livestock, fishery, and agroforestry) there has been progress in terms of production or/and productivity. However, the sector is in a low development stage as highlighted by a number of indicators including labor productivity, productivity gaps, trade and competitiveness, poverty and malnutrition, and infrastructure. There are however positive signals that show not only the potential for growth but also opportunities that the ADS should build upon.

```{r}
tribble(
  ~"Indicator", ~"1995/96", ~"2010/11", 
  "Agricultural GDP", "\\$3.4 billion", "\\$5.2 billion", 
  "Productivity of agricultural labor (\\$ per person)", "\\$466", "\\$700", 
  "Agricultural land per household (ha/hh)", "1.1", "0.7", 
  "Percentage of hodings operating less than 0.5 ha", "40.1", "51.6", 
  "Agricultural land use (Cereal as percentage of cultivated land)", "80", "80", 
  "Seed turnover", "8", "8", 
  "Employment in agriculure", "66", "60", 
  "Agricultural exports", "\\$32 million", "\\$248 million", 
  "Agricultural imports", "\\$157 million", "\\$621 million", 
  "Poverty (2010/11)", "42", "25", 
  "Percentage of households reporting inadequacy of food consumption", "50.9", "15.7", 
  "Stunting of children (less than 5 years age)", "60", "42", 
  "Irrigation coverage (\\% of cultivated area)", "39.6", "54", 
  "Infrastructure (Rural road network (km) and strategic road network (km))", "SRN = 10000 km", "RRN = 40000 km, SRN = 20000 km", 
  "ICT reach", "Less than 10\\% connected", "46\\% connected"
) %>% 
  knitr::kable(booktabs = TRUE, escape = FALSE) %>% 
  kableExtra::kable_styling(font_size = 10) %>% 
  kableExtra::column_spec(column = 1, width = "12em")
```


Growth of agricultural GDP since the beginning of the APP (1995/96) has been slow (about 3%), highly variable from year to year, and with a slight upward trend due to stronger growth in 2011 and 2012. With a growth of population of around 2 percent over the same period, the increase in agricultural GDP per capita has been too slow to create strong dynamics leading to sustained poverty reduction and structural transformation from subsistence to commercialization.

Over the period of APP, agriculture sector (and overal economy, in general) has seen slow growth when compared to neighboring countries, most of which had considerably faster GDP growth then Nepal. 

Agriculture absorbs the majority of the labor force (61% self-employed in agriculture, 3% earning wages from agriculture) (NLSSIII, 2010/11). But this labor force is characterized by low productivity relatively to the rest of the economy. The estimate of labor productivity in agriculture in Nepal ($794/unit of agricultural labor) is about one fourth of the productivity in the rest of the economy. The weak performance of agriculture has created strong incentives for a large part of the most productive labor force (the ones in 20-40 age group) to seek employment abroad. The departure of migrants reached the level of almost 300000 in 2010.

Agricultural land per capital also decreased as the combined effect of several factors including inheritances, loss of agricultural land to urbanization, and degradation of land. Smaller size and more fragmented farms make it more difficult to realize economics of scale and also to provide sufficient livelihood for smallholder farm families. Even though GDP per hectare of cultivated land is about \$1800, the average farm size is only about 0.7 ha per household on average and more than 50% of households have farm size less than 0.5 ha. Although, it should be noted that, aggregate land productivity does not compare too badly with several neighbors of Nepal. During 2011 Nepal had 0.082 ha arable land per capita which yielded 2481 kg of cereal per hectare and fetched ADGP equivalent of \$2651 per hectare (India had \$2139 per hectare).

```{r}
tribble(
  ~"Selected variables", ~"1995/96", ~"2003/04", ~"2010/11", 
  "Average size of agricultural land (in hectares)", 1.1, 0.8, 0.7, 
  "Holdings operating less than 0.5 hectare (% of total holdings)", 40.1, 44.8, 51.6
  ) %>% 
  knitr::kable(booktabs = TRUE)
```

Among several factors reported to contribute towards low productivity, low adoption of improved agricultural technology remains a key one. The productivity gap between current and potential production is significant and this related to high level of farming for subsistence, access and adoption of suitable technology (both on farm and post-harvest), availability of inputs (planting material, breeds, feed, plant and animal health protection, irrigation, electricity, finance), and limited investment in the sector. Nepal has an estimated 44.7% of agricultural entities commercialized and 55.3% under subsistence farming entities. Staple commodities such as rice, wheat, potato and vegetables have higher commercialization rates (30-50%) than maize and fruits (15-25%).

Huge productivity gaps are noticed in several subsectors of agriculture including fish (Current production: 3.6 t/ha/year, Potential production: 10 t/ha/year), timber (CP: 0.337 cum/year, PP: 13.4 cum/year), paddy (CP: 2.72 t/ha/year, PP: 10-12 t/ha/year).

One set of constraints to the realization of such potential is the availability of inputs. For example feed ingredient supply is the major input to poultry and 60% is imported at relatively high cost.

The use of chemical NPK fertilizer (nitrogen, phosphorous, and potassium) in Nepal 2011/12 was 422,547 MT of which only about 25% is imported and officially declared to customs. This amount is sold on a subsidized rate by the state only through agricultural cooperatives. However, price does not substantially affect farmers' demand, and the other 75% is imported informally.

SRR has remained very low. Against the desirable SRR for crops at 25% to 30%, average SRR is only 4% to 8% for wheat followed by 4.4% in rice, 3.8% in maize and 1.6% for pulses.

Nepal's agricultural trade is in deficit. The growth of imports has outpaced exports and the agricultural trade deficit has increased over the years from \$124 million to \$373 million (TEPC, 2009/10). Nepal imports primary and industrial raw materials (due to declining domestic raw material production), and processed agriculture products (due to limited investment and competitiveness in high-quality, high-value agro-processing). The range of exports is concentrated in a narrow set of manufactured and agricultural products, such as carpets, ready-made garments, pashmina, handicrafts, lentils, tea, cardamom, fruit, ginger, vegetable ghee and medicinal and aromatic plant products (MAPs).

Nepal's agricultural import and export trade comprises about 15.6% of total trade, which includes items such as petroleum, construction materials, vehicles and equipment, consumer goods and others. Export value of the top three high value crops exceeds the value of cereal and diary imports.

Nepal has comparative advantages in export markets in resource- and labor-intensive low technology agriculture products such as dried vegetables, coffee, tea, vegetable and roots, ginger, and cardamom.

The Industrial Policy, 2010 has prioritized agriculture and agro-forestry industries for investment, and provides additional incentives and facilities to these industries. Foreign investment in agriculture sector (including processing and retailing) is less than 1% of total foreign investment.

The overall performance of APP has been mixed. The implementation of the APP was not helped by the conflict that plagued the country during approximately the first 10 years of the APP period. In addition to the reduced government and donors' support there are other factors that contributed to the weak performance of the APP including:

- Lack of Coordination
- Withdrawal of Subsidies on Fertilizer and Tube Wells (shallow and deep tube wells)
- Faults in the Design and Economic Assumptions of the APP
- Weak institutional capacity on project/program implementation
- Lack of Attention to Legal Issues
- Lack of Attention to Social and Geographic Inclusiveness
- Low Attention to Land Management Issues
- Inadequate Consideration of Regional Trade

Broadly, the explanation of the slow progress is based on eight sets of interrelated factors: Implementation, legal aspects, design aspects (rigid targets, simplistic view of technology, narrow green revolution perspective, emphasis on guiding inputs and outputs, lack of systems for re-planning), conflict, politics and policies, investment, capacity and human resources, and plans.

## Agriculture Development Strategy (ADS)

- Was developed by MoAD with support from technical assistance of ADB, Rastriya Kishan Sanjal (National peasants' coalition) comprising a technical team of 12 other development co-partners in preparation (IFAD, EU, FAO, SDC, JICA, AUSAID, DANIDA, WFP, DFID, USAID, WB, UN women).
- The ADS report includes a 10-year Action Plan and Roadmap and a rationale based on the assessment of the current and past performance of the agriculture sector.
- Key issues for the ADS: Productivity; competitiveness; trade; commercialization; infrastructure; credit, insurance and taxes; subsidies; land; food and nutrition security; institutions and human resources; climate change and natural resource management; social and geographic inclusion; and legal.

- Vision (7 sub heads):
  - Self reliance
  - Sustainability
  - Competition
  - Inclusive
  - Economic growth
  - Improved livelihood
  - Food and nutrition security

- There are 16 indicators and corresponding targets for ADS vision (last 4 indicator labels in table below \ref{tab:ads-indicator-targets} are grouped under single indicator).

```{r ads-indicator-targets}
readxl::read_xlsx("./data/ads_indicators_targets.xlsx") %>% 
  select(-7, -8) %>% 
  knitr::kable(booktabs = TRUE, caption = "Preliminary indicators and targets for ADS vision (ADS final report, 2014)", longtable = TRUE, format = "latex") %>% 
  kableExtra::kable_styling(font_size = 6, latex_options = "striped") %>% 
  kableExtra::column_spec(column = 1:6, width = c("6em", rep("12em", 5)))
```


## Prime minister agriculture modernization project

Initiation, project details ...

### Zones and superzones

## One village one commodity (ODOP)

## One district one commodity

## History of academic instiutions for Agriculture in Nepal

Institute of Agriculture and Animal Science (IAAS)

- Began as a school of Agriculture under ministry of Agriculture in 1957 to train JTA's in agriculture.
- School of agriculture was upgraded to college level in 1968.
- It was entitled IAAS in 1972 as a constituent institute of Tribhuwan University and located the specific place at Rampur in the same year.
- The institute has 230 ha of land for its academic and research programs at Rampur.
- Lamjung Campus located at Sundarbazar, Lamjung was established in 1975 and has about 18 hac of land.
- Paklihawa Campus at Rupandehi established in 1978 with area of about 22 hac of land.
- In 1998 the institute started M.Sc.Ag program.
- In 2001 the semester system of study in bachelor level was introduced.
- Ph.D degree course started in 2002.
- IAAS day: 4th of Bhadra
- Logo of IAAS was designed by Thakur Prasad Mainali

Agriculture and Forestry University (AFU)

- The first agriculture university in Nepal.
- Established in June 2010 at Rampur Chitwan.
- University formed merging the then IAAS Rampur and the then IOF Hetauda.
- Rampur academic complex extends to an areas of 280 hac
- Hetauda forestry campus has an area of 95 hac.
- 17th June (3rd Ashad) is celebrated as AFU day.
- Chancellor: Prime Minister of Nepal.
- 1st VC of AFU: Prof.Dr Kailash Nath pyakural
- 1st Registrar: Dr.Surya Kant Ghimire
- Vice-Chancellor: Ishwori Prasad Dhakal
- Registrar: Man Raj Kulakshyapati

## Nepal Agricultural Research Council (NARC)

### National Commodity Research Programs under NARC and their locations

- National Rice Research Program; Hardinath, Dhanusa
- National Wheat Research Program; Bhairahawa, Rupandehi
- National Maize Research Program; Rampur, Chitwan
- National Grain Legume Research Program; Khajura,Banke (recently transferred from Rampur, Chitwan)
- National Potato Research Program; Khumaltar, Lalitpur
- National Swine and Avian Research Program; Khumaltar, Lalitpur
- National Sheep and Goat Research Program; Jumla
- National Oil Seeds Research Program; Nawalpur, Sarlahi
- National Sugarcane Research Program; Jitpur, Bara
- National Citrus Research Program; Paripatle, Dhankuta
- National Jute Research Program; Itahari, Sunsari
- National Ginger Research Program; Kapurkot, Salyan
- Hill Crop Research Program; Kavre, Dolakha
- National Bovine Research Program; Khumaltar, Lalitpur
- National Tobacco Research Program; Belachapi, Janakpur
- Apple Processing Centre; Jumla
- Cotton Research Centre; Khajura, Banke

## Department of Food Technology and Quality Control

Department of Food Technology and Quality Control (DFTQC) is one of the three departments under the Ministry of Agriculture and Livestocks Development of Government of Nepal. It was established in 1961 AD as Department of Food and placed in Singha Durbar, which later in 1965 was shifted to present location of Babarmahal, Kathmandu. In 1966, the Department of Food then was renamed as Food Research Laboratory. The laboratory later in 1980 was again converted to Central Food Research Laboratory (CFRL) and known by this name until it became the present department in 2000 under MoAD.

DFTQC played a pioneer role to lay down foundation stone for food quality control system, research and development in the field of food science and technology and nutrition support program. As mandated by Government of Nepal, DFTQC is the apex organization responsible for the enforcement and implementation of of Food Act and Regulations. The main aim is to ensure and enhance the quality and safety of food and feed products in the country. Further, the department has a paramount role in augmenting appropriate food processing and post harvest techniques to promote agribusinesses. Similarly, the department has been implementing various food and nutrition activitie for the reduction of various forms and types of malnutition in the country. DFTQC, as has been entrusted as CODEX contact point for Nepal for more than three decades, has also been given the role of National SPS enquiry point in 2004.

Organization structure

Currently DFTQC has the following Divisions/sections in the central level:

1. Food and feed safety and quality control division
2. Food technology development and nutrition division
3. National food and feed reference laboratory
4. Planning, monitoring and evaluation section
5. SPS enquiry point
6. Codex and Infosan secretariat
7. Administration section

The structure of DFTQC also accompanies:

- 6 offices at Biratnagar, Janakpur, Hetauda, Bhairahawa, Nepalgunj and Dhangadhi; 
- 24 Division offices
- 12 food Import-Export Quality Certification Offices at various districts and custom points all over Nepal.

**Vision**

Create environment for availability of wholesome, safe and nutritious food to all Nepali citizens and facilitate food trade.

**Mission**

- Safeguard and protect the health of consumers by assuring the availability of safe and nutritious food with the creation of awareness among producers, traders as well as consumers and facilitate food trade in a coordinated and collaborative approach based on scientific evidence with the updated regulatory framework.
- Promote food businesses by developing appropriate food processing technology through R and D and disseminating and adopting the outcome of research for the development of food processing enterprises.
- Promote nutritional status of people through the development of food based nutritional approaches.

**Goal**

- Ensure availability of wholesome, safe and nutritious food to the consumers by assuring the quality in food production and trade.
- Develop and disseminate appropriate food processing technology to promote entrepreneurship.

**Objectives**

DFTQC has the following three major objectives:

- Maintain safety and quality of food and feed products in the country by implementing updated food (2023) and feed act (2033) and regulations.
- Promote entrepreneurship by developing and disseminating appropriate technologies.
- Improve nutritional status of the people through food-based approaches.

**Activities of DFTQC**

- Food and feed quality regulation
  - Recommendation for establishment of food and feed production industry, or withdrawl of such recommendation and monitoring.
  - Food import-export and quality certification
  - Food nutritional status regulation and dietary recommendation
  - Categorization of food and feed materials and making changes to that.
  - Food safety and quality related training and promotion
  - Market inspection of food and feeds, sample collection and testing and legal action to offenders.
  - Monitoring and inspection of hotel, restaurant and confectionery stores.
  - Awareness programs on food safety and quality for consumers.
  - Quality certification of food and feed materials.
  - Organization of special programs related to food safety in auspices of festivals and celebrations.
  
- Development of food technology and nutrition
  - Conduct research on food technology and nutrition
  - Provide food technology and nutrition laboratory facilities
  - Training, promotion and communication programs on food processing and nutrition
  - Food technology development program
  - Determination of nutritional attributes of food
  - Skill training programs for self employment
  - Development of entreprise technologies and advisory services to entrepreneurs.
  - Development of food items (mainly for infants/children)
  - Food and nutrition radio program
  - Community nutrition improvement program, identification and promotion of locally available nutritious food items, recommendations for healthy eating habits, and conduction of exhibition and fairs.
  
- Central food and feed reference laboratory
  - Testing and analytic service for domestic food and feed materials
  - Testing and analytic serive for imported and exported food and feed materials (Physical and chemical testing laboratory, microbiological testing laboratory, dietary food and feed laboratory, food residue and chemical adultration testing laboratory, reference laboratory related activities)
  - Laboratory training and skill development programs
  - Internal and external laboratory testing and analysis
  - Development of analytical procedure and protocols and method validation program
  - Laboratory accreditation program
  
## Miscellaneous

### National Farmers Commission

- Chariman: Chitra Bahadur Shrestha
- Members: 7 total
- 2073/10/06: National Farmers' Commission organization directives
- Has central office in Hariharbhawan, Lalitpur

### Weekly radio and television broadcast on Agriculture Program

```{r}
dplyr::tribble(~"SN", ~"Day of week", ~"Radio Nepal (6:40 to 6:55)", ~"NTV (60:40 to 6:57)",
               1, "Sunday", "Saptahik krishi gatibidhi", "Krishi sambad", 
               2, "Monday", "Safalta ko katha", "Krishi sambad", 
               3, "Tuesday", "Food and nutrition", "Samaya sandarva", 
               4, "Wednesday", "Krishi sambad", "Ajako krishi", 
               5, "Thursday", "Krishak ko sarokar", "Krishak ko sarokar", 
               6, "Friday", "JTA ra budhi ama", "Saptahik krishi gatibidhi", 
               7, "Saturday", "Radio patrika", "Krishi teleserial"
               ) %>% 
  knitr::kable(caption = "Schedule of weekly telecasts of Agriculture programs in Radio and Television", booktabs = TRUE) %>% 
  kableExtra::kable_styling(latex_options = c("striped", "HOLD_position"))
```

### Kishan call center

The public agricultural extension service is able to reach only so far to 15% of the total farm households. While the need for professionalizing agriculture sector and improving livelihood standards of farming community is imminent. The issue of food security, besides being a hot topic of discussion in national and internatioal communitity, is closely associated with the citizen's rights. In this regard, it becomes all the more essential for traditional agricultural systems to be transformed to a information-technology based commercial system wherein farmers can connect with greater ease to subject specialists for consultation. For serving the purpose of addressing farmers' inquiries about crop and livestock, mainly concerning problems such as disease and pests, Kishan Call Center was established in 2072/73.

**Objectives**

- Addressing problems relating to crop, livestock and business enterprise faced by farmers and entrepreneurs through telephone contact with subject matter specialists.
- Minimize the geophysical barriers in communication and information exchange so that service delivery could be streamlined.

- Phone number: 16600195000
- Subjects dealt in respective days and schedule:

```{r kishan-call-center-schedule}
tribble(
  ~"Day", ~"Subject", ~"Time of day", 
  "Sunday", "Cereal crop, apiary, silkworm farming, mushroom cultivation, crop protection", "11 AM to 4 PM", 
  "Tuesday", "Vegetable, fruits, soil and market", "11 AM to 4 PM", 
  "Thursday", "Fisheries, livestock farming, poultry farming, production technology, fodder and forage management", "11 AM to 4 PM"
) %>% 
  knitr::kable(booktabs = TRUE) %>% 
  kableExtra::kable_styling(font_size = 10) %>% 
  kableExtra::column_spec(column = 2, width = "16em")
```


### Crop and livestock insurance

### NPK Content of fertilizer

- Urea: 46%  Nitrogen
- DAP(Diammonium Phosphate): 18% N, 46% Phosphorus
- MoP(Muriate of Potash): 60% potash
- SSP(Single Superphosphate): 16% P
- DSP(Double Sugar Phosphate): 32% P
- TSP(Triple Sugar Phosphate): 48% P
- ZnSO4: 22-35% Zinc

### Contributors in Agriculture

- Sociology: Auguste Comte
- Statistics: R.A. Fisher
- Economics: Adam Smith
- Soil Science: V.V. Dokuchaev
- Tillage: Jethro Tull
- Green Revolution: Norman E Borlaug
- Modern Agronomy: Peter De Crescent
- Organic Farming: Rudolf Steiner
- Dr. Norman E. Borlaug is the pioneer of Green Revolution and was awarded Nobel Peace Prize in 1970.

### Common animal diseases

- Bacterial diseases: Anthrax, Brucellosis (Bang's disease), Black Quarter(Black leg), Haemorrhagic Septicaemia, Mastitis
- Fungal diseases: Aspergillosis, Candidiasis, Ringworm
- Protozoal diseases: Coccidiosis, Babasiosis
- Viral diseases: Foot and Mouth disease, Swine fever(hog cholera), New castle disease, Marek's

### Common Poultry Diseases

- New Castle Disease/Ranikhet
- Mareck's Disease (MD)
- Gumboro
- Coccidiosis
- Chronic Respiratory Disease (CRD)
- Avian Influenza/Bird Flu- H5N1, H9N2

### Agriculture and child labor

According to the latest ILO estimates, at least 250 million children of between 5-14 years of age work in developing countries. Almost half of these children work on a full-time basis. The participation rates of children in economic activities in much higher in rural areas than in urban centres. Rural children, in particular girls, tend to start working at an early age. In Latin America and the Carribbean, out of 15 million children involved in the labor market, 56 percent work in agriculture sector from the age of 5 to 7 onwards. In some countries, children account for as much as 30 percent of the agricultural workforce. Most children work seven days a week and are paid less than the prevailing rates in their localities. They work long hours, and a very high proportion of these children are injured at work. Exposure to poor working conditions has serious repercussions on children's growth, development and health. The most common injuries include: cuts and wounds, eye infection, skin problems, fever, and headaches caused by excessive heat or by exposure to pesticides while working in agricultural fields.

### Hazards in agriculture occupation

Despite the fact that certain developing countries have reached higher levels of economic development, nutrition and health are still the problem areas. This situation provokes a vicious circle of low productivity, low wages, malnutrition, ill-health and low working capacity. The interaction between poor living and working conditions determines a distinctive morbidity-mortality pattern among agricultural workers, which is due to the combination of malnutrition, general and occupational diseases, and complications arising from undiagnosed or untreated diseases. Low working capacity is closely related to workers' malnutrition and poor health. 

Agriculture related hazards mostly include:

1. Machinery such as tractors, trucks and harvestors, and cutting and piercing tools;
2. Hazardous chemicals: Pesticides, fertilizers, antibiotics and other veterinarian products;
3. Toxic or allergenic agents: Plants, flowers, dusts, animal waste, gloves, chrome, oils;
4. Carcinogenic substances or agents: Certain pesticides such as arsenicals and phenoxy-acetic herbicides, UV radiations, parasitic diseases such as bilharziasis and facioliasis;
5. Transmissible animal diseases: brucellosis, bovine tuberculosis, hydatid disease, tularaemia, rabies;
6. Other infectious and parasitic diseases: leishmaniasis, bilharziasis, facioliasis, malaria, tetanus, mycosis;
7. Confined spaces such as silos, pits, cellars and tanks;
8. Noise and vibration;
9. Ergonomic hazards: use of inadequate equipment and tools, unnatural body position or prolonged static postures, carrying of heavy loads, repetitive work, excessive long hours;
10. Extreme temperatures and adverse weather conditions;
11. Contact with wild and poisonous animals: insect, spider, scorpion, snake and certain wild mammals.

### Gender mainstreaming/Women in agriculture development

- In decisions regarding agriculture, contribution of,
  - Women: 25%
  - Men: 32%
  - Both: 43%
- In labor contribution to agricultural activities,
  - Men: 5 hrs
  - Women: 9 hrs
- Proportion of women engaged in agriculture: 72.8%
- Farm household labor accounts for 50-60% of total cost of production
- Proportion of women engaged in livestock farming: 70%
- Labor participation of female in agriculture: 48.9%
- Proportion of female having ownership of land: 10.8%
- Proportion of women participants in,
  - Horticulture: 42.3%
  - Agronomy: 50.6%
  - Vegetable farming: 67%
- Worldwide, female contribution of cereal production is more than 50%
- 10th five year (2059/60 - 2063/64) plan oversaw participation of 40% female in agricultural programs
- Major income generating activities of female include: Fresh vegetable production, goat farming, sericulture and apiary.
- Females have largest labor contribution to silkworm rearing: 69%
- Proportion of female engaged in,
  - Goat farming: 55%
  - Pig farming: 40%
  - Poultry farming: 46%
- Share of female in gross national income: 30%
- Proportion of female among total population (census, 2068): 51.5%
- The literacy rate was 54.3 % for men and 21.3 for women (NLSS, 1996) and 65.8 % and 35.4 %, respectively in 2000.

**Gender mainstreaming**

- 6th five year plan (2036-2042) laid emphasis to women participation
- 7th five yera plan (2042/43 - 2046/47) had policy of equal participation of men and women in development programmes in order to improve their skills and increase productivity.
- 7th five year plan allocated 10% quota for women trainings.
- 8th five year plan (2049/50-53/54) declared policy for mainstreaming of women in development to empower socio-economically.
  - Women Farmer Development Division was established for the same purpose in 2049
  - The division prepared strategy paper for women farmer development (1994-1999) and developed programs for engaging women farmers in cereal production, horticulture, poultry farming and rabbit farming.
- Gender mainstreaming has been continued to be placed in agriculture and livestock development policy papers since 8th five year plan.
- 9th five year plan had 3 major objectives -- empowerment, equality and mainstreaming.
  - The plan envisaged participation of 33% female at minimum in every agricultural development activities.
- 10th five year plan (2059/60 - 2063/64) envisaged involvement of women in the agricultural programmes at average of 40% while expecting more than 60% in programmes such as vegetable farming, horticulture and silk farming.
- In 2061 BS, Gender equality and environment division was established.

**National Agricultural Policy and Gender mainstreaming**

- While conducting agriculture related programs, where possible in every activities, women participation will be increased to 50%.
- In order to increase women farmer participation, revolving Women Farmer Training Program will be encouraged to ensure physical access to the training.
- MoAD prepared Gender Mainstreaming Stragegy, 2064 with assistance of ADB. Strategy highlights of the include,
  - Ensure opportunity for 33% women agriculture technicians for higher education and abroad visits.
  - Maintain 50-50 participation of men-women in farmer visit and training.
  - Ensure 33% women participation in trainings conducted for JT, JTA.
  - Promote women's group and mixed gender group, while latter ensuring participation of 60% women.

**Participation of women farmers in decision making**

```{r women-farmer-participation}
tribble(
  ~"SN", ~"activity", ~"decision_making_role_percentage", 
  1, "Crop selection", 60, 
  2, "Seed selection", 52, 
  3, "Organic manure type and amount", 96, 
  4, "Chemical fertilizer type and amount", 29, 
  5, "Irrigation (timeing and frequency)", 50, 
  6, "Interculture operation timing", 62, 
  7, "Crop harvest", 63, 
  8, "Storage technique", 73, 
  9, "Allocation for family consumption", 73, 
  10, "Processing", 73, 
  11, "Allocation for sales", 51, 
  12, "Vegetable sales", 58
) %>% 
  rename_at(2:3, function(x)str_replace_all(x, "_", " ") %>% str_to_sentence()) %>% 
  knitr::kable(booktabs = TRUE, caption = "Women farmer participation in decision making") %>% 
  kableExtra::kable_styling(latex_options = "hold_position")
```


**Agriculture training types based on levels of delivery**

```{r agriculture-training-types}
tribble(
  ~"sn", ~"level/strata", ~"period_days", ~"participants", 
  1, "Service center level on-site training", "1", "20-25", 
  2, "Revolving on-site farmer's training", "1-2", "15-25", 
  3, "Farmers' group mobilization training", "1", "All members of a group", 
  4, "District level farmer's training", "1-3", "20-25", 
  5, "District level pesticide retail trader training", "3", "20-25 (25 max)", 
  6, "Regional level farmer's training", "5-14 days", "10-20", 
  7, "Central level farmer's training", NA, "10-20"
) %>% 
  rename_at(2:4, function(x)str_replace_all(x, "_", " ") %>% str_to_sentence()) %>% 
  knitr::kable(booktabs = TRUE) %>% 
  kableExtra::kable_styling(font_size = 8, latex_options = c("striped", "hold_position"))
```


- In a "Ghumti stalgat" training, 50% of the participants should be women.


### Simple facts

- "Agriculture" is derived from Latin.
- "Agronomy" is derived from the Greek.
- A maximum of 10 participants can parttake in a crop production competition.
- There is a provision of 25% subsidy in production of foundation seed to the producer.
- With the use of improved seed 15-25% grain in production can be achieved.
- Small irrigation projects in Nepal are conducted by farmers' groups and farmers' cooperatives both.
- A small irrigation project for Terai constitutes 200 hectares of land.
- A small irrigation project in Hills constitutes 25 hectares of land.
- Nepal ranks 18th position among rice cultivating nation in the world.
- Father of Agronomy in Nepal: Netra Bahadur Basnyat
- Golden rice is genetically modified rice with yellow colored seeds rich in vitamin A
- Date palm is the oldest fruit.
- Rye is known as grain of poverty.
- Tomato is known as poor man's orange.
- DDT was the first discovered insecticide by Paul Muller.
- Maximum amount of sugar in honey is 73%.
- Mandarin is alled fancy fruits.
- Cardamom production highest in Nepal in the world.
- Dance of bee is called waggle dance.
- Apple production is highest in China.
- King of cereals is Wheat and Queen of cereals is Maize.
- Orchidaceae is the largest family among flowering plants.
- National Agriculture Policy in 2061 BS
- Gaurab hybrid and Rampur hybrid are the only hybrid varieties in Nepal in maize.
- Halo kranti started in 2006 BS from Duradanda, Lamung.
- OVOP (One Village One Product) practice in Nepal base on PPP (Public Private Partnership) approach started from Japan.
- One Village Two Product practice is prevailing in Bhutan.
- Year of Family farming: 2014
- Yellow color of Papaya is due to Cryptoxanthin.
- Edible part of apple is enlarged thalamus and edible part of cabbage is bud while that of cauliflower is inflorescence.
- Earthworm is farmer's friend.
- 39.6% of area of Nepal's land is covered by forest.
- Japan is Nepal's largest bilateral donor.
- Lentil is a major pulse crop with efficient export.
- Crop and livestock insurance started in Nepal in 2069 Magh 1
- Triticale is artificial cereal crop.
- China is the world's highest rice producer and Vietnam is the world's highest rice exporter.
- Azolla is widely used bio - fertilizer in Rice Crop.
- Amendment of acid soil is done by lime and of alkaline soil is done by CaSO4
- Tomato is known as poor man's orange.
- Air layering in litchi is to epicotyl grafting in Mango
- Anadi, Jumli Marshi, Pokhreli Jetho Budo are indigenous variety of Rice in Nepal.
- Readily available form of water to plants is capillary water.
- Limiting amino acid in Cereals are Lysine and threonine whereas limiting amino acid in legumes are methionine and tryptophan.
- Stubble is the portion of stem that is left on the field during harvest.
- Leguminous crop is considered as biological plough.
- First certified organic commodity exported from Nepal: Coffee
- Genetic dwarfism can be overcome by the use of Gibberelic Acid.
- Peripneustic type (having spiracles in a row on each side of the body) respiratory system is found in Lepidoptera larvae.
- Gilson's glands is found in order Trichopter.
- Growth curve of kiwi is Triple sigmoid type.
- Which produces cocoon in silkworm ? Pupa
- Chroma refers to: Relative purity of color
- Pointed gourd is commonly propagated through stem cutting.
- Allophane is an amorphous to poorly crystalline hydrous aluminium silicate clay mineraloid. It is common in vocanic soil and typically forms under mildly acidic to neutral pH (57). Its structure has been debated, but it is similar to clay minerals and is composed of curved alumina octahedral and silica tetrahedral layers.
- Horizon of maximum accumulation of Iron (Fe) and Aluminium oxides is B-horizon.
- Spinach can hybridize with Swiss chard and Garden beet.
- Soil with highest CEC is Vermiculite.
- Dioecious species: Spinach, pointed gourd and yam.
- Mo is available at slightly basic pH.
- GA is flowering hormone of longday and chilling requiring plants.
- Swelling of Smectite is more because of Mg binding.
- Chinese cabbage can hybridize with Turnip and Broad leaf mustard.
- Paclobutrazole is also known as anti GA hormone.
- Color of semi-aerated iron containing soil is Yellow.
- Sodicity is measured by SAR and ESP.
- Chayote amongst all vegetables has the highest productivity.
- Salt resistant crops: Sugarbeet, cotton, sweet potato.
- Rice, wheat and maize are cereal crops while buckwheat and amaranthus are pseudocereals.
- Wheat and maize show hypogeal germination but soyabean and chickpea show epigeal germination.
- Tree of Paradise/wisdom: Banana
- King of Fruits/ Bathroom Fruit: Mango
- Bitter taste of Bitter Gourd due to: Cucurbitacin/ Bitter Glucoside
- Rice is semi - aquatic in nature.
- Critical stage for irrigation in wheat is CRI
- Cotton fiber is also known as White Gold.
- Apple of tropics/Poor man's apple: Guava
- Poor man's meat: Soyabean
- Poor man's friend/ King of Vegetable: Potato
- World's highest cereal production: Wheat
- Nepal's highest cereal production: Rice
- Queen of spice crops: Cardamom
- King of spice: Black pepper
- Queen of Flower: Gladiolus
- King of flower: Dahlia
- Butter fruit: Avocado
- Queen of Beverage: Tea
- King of Temperate fruit: Apple
- Drosophila of crop plants: Maize
- Vegetable meat: Cowpea
- Goat is called poor man's cow.
- Male adult of horse is stallion and female adult of horse is mare.
- Heifer is young female cattle.
- Young male poultry is cockrel and young female poultry is pullet.
- Black gold: Buffalo
- Yellow gold: colostrums milk
- King of fodder: Elephant grass/Napier
- Queen of fodder: Leucerne
- Freshly drawn milk has the PH value in range of 6.5-6.7
- Evening milk is higher in fat than morning milk
- Amount of milk left in udder after normal milking is called residual milk.
- Water content of milk is 87-88%
- Gizzard in poultry is used grinding of food whereas proventiculus in poultry is the true stomach.
- Abomasum is true stomach in ruminants.
- Omasum is absent in camel and rumen is water storing sac.
- Sweetness in milk is due to Lactose.
- Milk holiday is the day when the dairy does not buy milk from the farmers.
- Milk storage is done at $5^\circ C$.
- 1 lit of milk can be synthesized by 500 lit. of blood circulation in body.
- Milk is complete nutrient. It lacks only Fe.
- Meat of calf is k/a veal.
- Ear notching is the best identification method in pig.
- Gall bladder is absent in horse.
- Elephant has longest gestation period i.e. 18-23 months(624days)
- Kavrepalanchowk is the highest milk producing district in Nepal.
- Semen is stored in liquid Nitrogen in $-196^\circ C$; the process of storage is k/a Cryopreservation.
- Poor man's food -- Bajra
- King of coarse cereals -- Sorghum
- Banana is the major cultivated crop after Wheat, rice and maize worldwide. India is the largest producer of Banana.
- Potato variety cultivated in largest acerage in Nepal -- Cardinal
- Nepal imported 40% of total consumed orange from India.
- Districts famous for industrial crops cultivation:
  - Cotton: Dang, Banke, Bardiya
  - Jute: Jhapa, Morang, Sunsari, Saptari, Udayapur, Siraha
  - Mustard: Sunsari, Dhanusha, Sarlahi, Chitwan, Dang, Banke, Bardiya
- China first started the fish rearing.
- Wheat is mostly stored in metal bins.
- Training session for artificial insemination is conducted for 1 week.
- Cowpea is considered vegetable meat.


## Top exports by Nepal

(when ??)
1. Man made staple fiber (11.9%) -- US \$ 87.9 million
2. Textile floor covering (11.8%) -- US \$ 86 million
3. Coffee, tea, spices (6.4%) -- US \$ 47.6 million

Agriculture export

1. Cardamom
2. Fruit juice
3. Black tea
4. Forest and vegetable products
5. Betal nuts
6. Oilcakes and bran/feeds
7. Noodles
8. Medicinal plant
9. Lentil
10. Bovine meat (Buffalo)

## Accredited food commodities and parameters of National Food and Feed Reference Laboratory, DFTQC, Nepal

National Accreditation Board of Testing and Callibration Laboratories (NABL), India on ISO/IEC 17025:2017.

```{r}
tribble(
  ~"SN", ~"Testing group", ~"Commodity group", ~"Commodities", ~"Parameters", 
  1, "Biological testing", "Food and agricultural products", "Fruits and vegetable products: Jam, juice, jelly, pickles and candies", "1. Total bacteria count, 2. Yeast and mould count, 3. Coliform count, 4. Escherichia coli, 5. Enterobacteriaceae",
  2, "Biological testing", "Food and agricultural products", "Bakery and confectionary products (Pulses and cereal): Instant noodles, Biscuits and Infant foods", "1. Total bacteria count, 2. Yeast and mould count, 3. Coliform count, 4. Escherichia coli, 5. Enterobacteriaceae",
  3, "Biological testing", "Food and agricultural products", "Milk and dairy products: Fluid milk, skimmed while milk powder, condensed milk, icecream, and yogurt sweets", "1. Total bacteria count, 2. Yeast and mould count, 3. Coliform count, 4. Escherichia coli, 5. Enterobacteriaceae",
  4, "Biological testing", "Food and agricultural products", "Processed meat and meat products", "1. Total bacteria count, 2. Yeast and mould count, 3. Coliform count, 4. Escherichia coli, 5. Enterobacteriaceae",
  5, "Biological testing", "Water", "Processed drinking water", "1. Total bacteria count, 2. Yeast and mould count, 3. Coliform count",
  1, "Chemical testing", "Food and agricultural products", "Fats and oils", "1. Free fatty acid, 2. Refractive index, 3. Acid value, 4. Peroxide value", 
  2, "Chemical testing", "Food and agricultural products", "Fruits and vegetables (processed products and sweets)", "1. Total soluble solids (TSS), 2. Acidity, 3. Sulphur dioxide (SO2), 4. Benzoic acid, 5. Tartazine, 6. Sunset yellow", 
  3, "Chemical testing", "Food and agricultural products", "Spices and condiments: Ginger cardamom and turmeric", "1. Volatile oils, 2. Crude fiber, 3. Total ash", 
  4, "Chemical testing", "Food and agricultural products", "Tea and coffee", "1. Total ash, 2. Water extract, 3. Crude fiber, 4. Lead, 5. Caffine", 
  5, "Chemical testing", "Food and agricultural products", "Cereal and cereal products", "1. Moisture, 2. Protein", 
  6, "Chemical testing", "Food and agricultural products", "Honey", "1. Moisture, 2. Acidity as formic acid, 3. Hydroxymethyl furfural (HMF)", 
  7, "Chemical testing", "Food and agricultural products", "Milk and milk products: Processed milk, skimmed milk, whole milk powder, infant food, condensed milk", "1. Milk fat, 2. Moisture, 3. Protein, 4. Ash content, 5. Fat", 
  8, "Chemical testing", "Food and agricultural products", "Meat products", "Sodium nitrite",
  9, "Chemical testing", "Residues in food products", "Fruits and vegetables", "Organochlorine pesticides 1. Aldrin (SS), 2. Alpha-BHC (SS), 3. Alpha-Chlordane (SS), 4. Beta, BHC (SS), 5. Delta BHC (SS), 6. Dieldrin (SS), 7. Endosulfan I (Alpha), 8. Endosulphan II (Beta, SS), 9. Endosulfan sulphate (SS), 10. Endrin aldehyde (SS), 11. Endrin ketone (SS), 12. Endrin (SS), 13. Gamma-BHC (Lindane, SS), 14. Gamma chlordane (SS), 15. Heptachlor (99\\%, SS), 16. Heptachlor-epoxide, 17. Isomer (B, SS), 18. $4, 4^{\\prime}$-DDD (SS), 19. 4, $4^{\\prime}$-DDE (SS), 20. 4, $4^{\\prime}$-DDT, 21. Methoxychlor", 
  10, "Chemical testing", "Residues in food products", "Fruits and vegetables", "Organophosphate pesticides 1. O, O, O-Trimethylphosphorothioate, 2. Thionazin, 3. Sulfotep, 4. Phorate, 5. Dimethoate, 6. Disulfoton, 7. Parathion, 8. Methyl parathion, 9. Famphur", 
  11, "Chemical testing", "Residues in food products", "Cereals and cereal products: Instant noodles, biscuits, snacks", "Mycotoxins 1. Total aflatoxin, 2. Aflatoxin B1, 3. Aflatoxin G2; Trace elements 1. Zinc, 2. Calcium, 3. Magnesium, 4. Iron", 
  12, "Chemical testing", "Water", "Processed dinking water", "1. pH, 2. Hardness, 3. Alkalinity, 4. Chloride content", 
  13, "Chemical testing", "Residues in water", "Processed drinking water", "Trace metal elements 1. Lead, 2. Cadmium, 3. Arsenic, 4. Calcium, 5. Copper, 6. Iron, 7. Magnesium, 8. Zinc"
) %>% 
  knitr::kable(escape = FALSE, format = "latex", booktabs = TRUE, longtable = TRUE) %>% 
  kableExtra::kable_styling(font_size = 8, latex_options = c("striped")) %>%
  kableExtra::column_spec(column = 2:5, width = c("8em", "8em", "10em", "22em"))

```

