
# General agriculture

## Nepal agriculture and geography

**Ecological regions**
- Himalayan region: 35%
- Hilly region: 41.67%
- Terai region: 23.11%

**Agricultural land**

- Cultivated land: 3091000 ha (21%)
- Cultivated land uncultivated: 1030 (6.99%)
- Jungle and shrubs: 5828 (39.597%); Shrub alone: 1560
- Pasture land: 1766000 (11.99%)
- Total Jungle = Shrub + Pasture

**GDP of Agriculture and releated sector/subsectors**

```{r}
dplyr::tribble(~"sector", ~"fy_2074-75", ~"fy_2073-74", ~"fy_2072-73", 
               "Agriculture and forestry", 27.10, 28.25, 31.08,
               "Fishery", 0.49, 0.51, 0.53,
               "Non-agriculture", 72.41, 71.24, 68.39) %>% 
  knitr::kable(booktabs = TRUE, caption = "Percentage contribution of different sector and subsectors to total national GDP")
```

**Agricultural growth rates**

```{r}
dplyr::tribble(~"Year", ~"Growth rate (%)", 
               2072/73, 0.01, 
               2073/74, 5.14,
               2074/75, 2.72)
```

## Population census, 2068

```{r}
dplyr::tribble(~"Indicator", "Unit", ~"Value", 
               "Male", "count", 12849041,
               "Female", "count", 13645463,
               "Annual population growth rate", "percentage", 1.35,
               "Population density", "individuals per msq", 180,
               "Total engagement in agriculture", "percentage", 65.6,
               "Male engagement in agriculture", "percentage", 60.2,
               "Female engagement in agriculture", "percentage", 72.8, 
               "Crude birth rate", "per thousand", 21.8,
               "Crude death rate", "per thousand", 7.3,
               "Child death rate", "per thousand", 40.5,
               "Net fertility rate", "percentage", 2.5, 
               "Average life expectancy", "years", 66.6)

```


## Kishan call center

- Phone number: 1660-5652-999
- Active hours: 10 AM - 4:30 PM

## Development of cooperatives in Nepal

- Traditionally, custom of _Parma_, _Mankakhal_, _Dharmabhakari_, _Dhikuri_/ _Dhukuti_ were on place very early.
- In 2010, Cooperative development department was established.
- In 2013, government implemented executive guidelines.
- In 2016, Cooperative organization act came into force.
- In 2018, Cooperative organization regulations came into force.
- In 2019, Cooperative training center was established.
- In 2020, Cooperative bank was established.
- In 2024, "Village return campaign" was initiated under cooperative program.
- In 2027, Agriculture development bank started managing cooperative organizations.
- In 2033, "Sajha" program was implemented in 30 districts establishing multipurpose cooperatives in VDCs.
- In 2035, Agriculture development bank handed over the managerial responsibility of cooperatives to management board.
- In 2041, Cooperative ("Sajha") organization act came into force.
- In 2048, National cooperative development board was set up.
- In 2048, Cooperative act came into force (new)
- In 2049, National Cooperative Association was established.
- In 2069, Ministry of cooperative and poverty alleviation was formed.

- Cooperative campaign was first initiated when implementing first five year plan (In 2013), wherein Rapti Dunn Project was launched and 378 cooperative organizations were established in the course.

Importance of cooperatives

- Poverty alleviation
- Social and economic support to low income households
- Combines efforts of small producers and consumers and leads them to a larger scale commercial operation
- Reduces transaction cost, when compared to individual efforts
- Reduces chances of suppression and exploitation by large scale acting merchants
- Effective utilization of available resources and supplies
- Empowerment of local community
- Qualitative development of labor
- Improves bargaining power
- Has a bigger role in operationizing in market demand and supply forces
- Members have ownership in rural development
- Promotion of community welfare
- Community development based on justice and equality
- Produce can have more accessible market
- Quality service could be delivered
- Effective mobilization of consumable goods
- Support in national economy

- 10th five year plan focused on emphasized agriculture based commerce and industries, cooperative farming, cooperative based agriculture input supply, and cooperative based small farmer irrigation project implementation, dairy collection and supply and other cooperative based approach to community development.
- The plan promoted active participation of cooperative groups and private sectors for the efforts.
- National agriculture policy, 2061 has prioritized cooperative based agriculture industries and commerce promotion, farmer group cooperative formation and agriculture wholesale market and haat bazzar established and management via cooperatives.
- To legalize a cooperative organization establishment, it should be registered:
  - Under Organization act in District Administration Office, or
  - According to Cooperative act and regulations under Cooperative Division Office.

## Budget speech, 2076-77 BS

Total budget announced was Rs 1.53 trillion for fiscal year 2019/20 (76/77). Economic growth rate was set at 7% as target. The budget focuses on social justice, increment of export to reduce trade deficit and increase in general productivity.

NRs 130 billion is to be distributed from revenue between provincial and local levels. Education will be freed upto secondary level. In terms of literacy, 70 districts will be designated the status of fully literate districts. To that end, NRs 2 billion will be appropriated for colorful textbooks for primary level. Day meals for 2.2 million school children will be provisioned and sanitary pads will be free for female students attending public schools.

Over 10 billion was allocated for Madan Bhandari science and technology university. It is also stated that science and technology laboratory will be established in each province.

NRs 6 billion will be allocated for free insurance in all districts.

Health grants will be increased for treatment of 8 types of severe illnesses. Likewise, NRs 2.2 billion is appropriated for new-mothers travel expenses. 52000 female community health volunteers will be provided Rs 3000 annual allowance. NRs 5 billion is dedicated to establish health service providing facilities at local levels. Similarly, Rs 400 milion is allocated for Ramraja Prasad Singh Hospital in Rajbiraj. Rs 400 million appropriated for betterment of Bir Hospital.

Addressing issue of public health, smoking and drinking will be banned in public places. 92% of the population will be provided full access to drinking water in coming fiscal year. NRs 7 billion is allocated for completion of Melamchi drinking water project. Over 43 billion NRs is allocated for drinking water and hygiene.

Social security allowance to pregnant women and elderly senior citizens allowance sees an increment of Rs 1000 (increased from Rs 2000 to 3000). There have been betterment of employment schemes for the peoples with disabilities.

Coming fiscal year to marked as youth mobilization year. Youth scientists conference to be held in the coming year.

In agriculture, grants will be provided for purchase of agricultural products and technology. Schemes for achieving self sufficiency in dairy, poultry and fresh vegetables will be prepared and implemented. Grants will be provided for fertilizer input purchase. Organic farming wil be encouraged. With the dubling of fruit cultivation in next 5 years, food quality labs will also be set in every province. Rs 500 million is appropriated for community farming. Rs 34 billion is allocated for agriculture.

A budget of 23.6 billion is allocated for irrigation programmes. 960 million allocated for irrigation programmes in Terai. Sunkoshi Marine to be developed as a national pride project, with 2.05 billion NRs allocated for program initiation. NRs 5.6 billion is allocated for construction of dams.

Next fiscal year will be marked as tree plantation year. Security to be beefed up in forest areas. Newer programs/practices to be launced in livestock management. Rs 1 billion is dedicated for 'Rastrapati Chure Programme'.

Under land management, a revised land management act will be introduced for sustainable utilization of land. Encroached public lands will be brought under government management within next fiscal year. Online issuance of land ownership certificate will have been started by next two years.

Tourism sector will be prioritized, with stress on infrastructure development of main tourist areas. Trail connecting Darchula to taplejung to be developed. Operation of cable cars will be encouraged in mountainous regions.

Government officials will be mandated to only gift homemade (domestic) products as and when needed.

Private and cooperative sector to be encouraged for production of necessary commodities. Local cement and wire frames to be encouraged in construction. To be self sufficient in production of at least 2 dozen products.
  - Hetauda and Udayapur cement factories to be made more efficient 
  - Local products will be promoted in construction inputs
  - Import of luxury goods and health unfriendly products to be discouraged
  - Economic zone to be established in Kavre and Nuwakot. 50% concession to Nepali textile industry on electricity.

Infrastructure development at trade transit points in the north. Business with third countries to begin via Chinese port. Completion of pending works on pipeline by the next year to facilitate import of petroleum products.

In energy, at least two big hydropower projects will be embarked on in all seven provinces. NRs 13 billion is allocated for Budigandaki, 2.02 for Budhiganga-Tamor project to be a national pride project. Province 2, Karnali province and Sudurpaschim province to have full access to electricity. Rs. 4.5 billion allocated for rural electrification. Waste-to-energy programmes will be encouraged.

Rs 163 billion is appropriated for railway and waterways. Digital payment system will be installed in public transport. 1.5 billion for transport management. Electric buses to be introduced in "major" cities. NEA to install charging stations. Additional budget is allocated for development of Narayani and Koshi waterways.

Under urban development, feasibility study will be conducted for development of mega-city and smart city. 530 million allocated to replace thatched roofs of 20000 houses. 4.3 billion for construction of 30000 houses under housing project.

Convention center with a capacity to accomodate 3000 people to be constructed in the valley, this year. City halls to be constructed in local levels. Compulsory footpath, underground cable management. Rs 47.7 billion allocated for housing and urban development. MP's constituency budget (MP fund) increased to Rs 60 million. 1 billion for Dharahara reconstruction (to be completed within the next two years). 141 billion appropriated for reconstruction. Rs 9 billion isolated for infrastructural development in each electoral constituency.

Online visa services will be furnished for foreign nationals. National Defence University to be established. 18-20% (non-gazetted and gazetted) increment in salaries of public service personnel. National knowledge centre to be established at Khumaltar, Tribhuvan International Airport to be upgraded to a boutique airport. Gautam Buddha International airport to come into operation next year. Contractors to be held responsible for the upkeep of their projects for years after completion.

VAT and other taxes to be made more effective through improved taxation system VAT rates to stay intact changes in customs rate. Import reliance will be significantly reduced. 

The budget size is 1.53 trillion, recurrent budget is 957.1 billion, capital expenditure is 408.59 billion and financing 167.5 billion.

## Food security

For a more fundamental discussion of food security topic, refer to [lecture handout](https://github.com/DeependraD/life_science_presentations/raw/master/food_security.pdf), in link to life science presentations.

FAO defined food security in 1996 A.D. In 1996, convention held in Rome approved seven points of food soverignty. Nepal has been a member of FAO since 2007 BS (1950 AD).

McMichael's projection:
  - 2 billion people suffer hidden hunger
  - 1.5 billion people suffer over nutrition related problems

Since 1981 (2038 BS), FAO started celebrating World Food Day on October 16th. In Nepal, 34th World Food Day was celebrated in 2014 with the slogan: "Feeding the world, caring for the earth".

According to FAO's projection, 52% of the population in South Asia are dependent on Agriculture while agriculture contributes 22% to the total GDP in the region.

Pillars of food security:

1. Food availability
2. Food access
  - Nepal's poverty rate in 2001/2002 AD: 30%
  - Nepal's poverty rate in 2010 AD: 26%
  - Nepal's poverty rate at the end of 12th plan: 23.8%
3. Food utilization
  - Per person per day recommendation by WHO: 250 ml milk (91 ltr per year)
  - Nepal's aim for per day milk consumption: 156 ml (57 ltr per year)
  - Nepal's aim for per year meat consumption: 14 kg
  - Nepal's aim for per year egg consumption: 48
  - 30% of total daily protein requirement should be met by animal sources.
4. Food stability
  
- In 2011 AD, Nepal ranked 157th among 187 nations in HDI
- In 2014 AD, Nepal ranked 145th in HDI
- Poverty rate at village/rural areas: 27% (Survey, 2012)
- Poverty rate at cities: 15% (Survey, 2012)
- Monthly poverty:
  - Maximum at Chaitra-Baisakh
  - Minimum at Mangsir-Poush
  
State of food security in Nepal:

- Untill 2042 BS (1985), production was twice that satisfied the population's need.
- Upto 2047 BS, food insufficieny was absent.
- Currently 40 districts are declared food minimum and more than 27 districts of high hill and mountain districts are food insecure.
- Calorie deficiency is prevalent in 41% of population.
- In 19 districts of Midwestern Development Region and Eastern Development Region, food security project is launched.
- 46% of the total cultivated area is under rice.
- Of the total food production, contribution of respective crops are:
  - Rice: 56%
  - Maize: 24%
  - Wheat: 17%
  - Millet: 3.6%
  - Barley: 0.29%

## Nepal agriculture research council

- Has following organizational structure:
  - Executive director
  - Director, Planning and cooperation
  - Director, Crop and horticulture research
  - Director, Livestock and fisheries research
  - Director, Employee administration
  - Director, Economic adminsistration
  - Head, Planning commission
  - Communication, publication and inventory commission
  - Socio-economic and agricultural research policy commission
  - External research directorate
  - Agriculture environment directorate
  
- Programs under Agronomic and horticultural crop research programs
  - Rice crop research program, Baniniya, Dhanusha
  - Maize research program, Rampur, Chitwan
  - Wheat crop research program, Bhairahawa, Rupandehi
  - Legume crop research program, Nepalgunj
  - Oil crop research program, Nawalpur, Sarlahi
  - Hilly crop research program, Kavre, Dolakha
  - Sugarcane research program, Jitpur, Bara
  - Potato research program, Khumaltar
  - Ginger research program, Salyan
  - Orange variety research program, Dhankuta
  - Jute crop research program, Itahari, Sunsari
  - National commercial agriculture research program, Pakhribas, Dhankuta

- Divisions situated in Khumaltar
  - Crop science division
  - Crop disease science division
  - Entomology science division
  - Soil science division
  - Agri-engineering division
  - Agri-botany division
  - Horticulture research division
  - Food research division
  - Biotechnology division
  - Commercial crop division
  - Seed science and technology research division

Regional agriculture research centres and programs

1. Regional agriculture research centre, Nepalgunj
  - ARC, Surkhet
  - ARC, Doti
  - ARC, Jumla, Bijayanagar
  - ARC, Jumla, Rajikot
  - ARC, Dolakha

2. Regional agriculture research centre, Lumle
  - ARC, Baidam, Pokhara $\rightarrow$ Aquaculture
  - ARC, Begnas, Pokhara $\rightarrow$ Aquaculture
  - ARC, Malepatan, Pokhara $\rightarrow$ Horticulture
  - ARC, Bandipur, Tanahun $\rightarrow$ Goat

3. Regional agriculture research centre, Parwanipur
  - ARC, Rasuwa $\rightarrow$ Pasture
  - ARC, Ranighat, Birgung, Parsa $\rightarrow$ Agri-machinery
  - ARC, Trishuli $\rightarrow$ Acquaculture
  - ARC, Belachhapi, Dhanusha $\rightarrow$ Tobaccoo

4. Regional agriculture research centre, Tarahara
  - Pakhribas, Dhankuta
  - Tarahara $\rightarrow$ Aquaculture

## ADS

- MoAD, ADB, Rastriya Kishan Sanjal etc. including 12 development cooperatives (technical team) were involved in preparation of the strategy. 
- Vision (7 sub heads):
  - Self reliance
  - Sustainablity
  - Competition
  - Inclusive
  - Growth
  - Livelihood
  - Food and nutrition security

- There are 16 indicators ??

## National Farmers Commission

- Chariman: Chitra Bahadur Shrestha
- Members: 7 total
- 2073/10/06: National Farmers' Commission organization directives
- Has central office in Hariharbhawan, Lalitpur

## Prime minister agriculture modernization project

Initiation, project details ...

### Zones and superzones


## Crop and livestock insurance

## One village one commodity (ODOP)

## One district one commodity

## Impacts of climate change

What are the impacts of climate change ? Discuss on detail about the initiative taken by Nepal Government in tackling the impacts of climate change.

Climate change is now affecting every country in every continent. It is disrupting national economies and affecting lives, of individuals, communities and countries today and is likely to do so in the ensuing future. Weather pattern are changing, snowline are melting due to global warming and sea level are rising. 

**Negative impacts**

1. Numerous mountains are seeing accelerated the melting of snow due to global warming.
2. Changes in monsoon patterns, such as erratic rainfall with alternate drought periods are the serious impacts of climate change in Nepal which hampers the agricultural production. The changes in monsoon pattern in Nepal greatly exacerbate the situation of poverty and food insecurity in the country.
3. Flood, landslides, forest fires, snow melt are the disasters that occur as a result of climate change without any prewarning and causes heavy loss.
4. High hills and mountains are more vulnerable to climate change compared to terai, thus climate change make large scale economic and ecolgical disruption in upper latitudes.
5. Loss of plant and animal biodiversity.
6. Increase in number of insects, pest and diseases and sporadic outbreaks becoming much frequent.

The positive impact of global warming, a characteristic feature of climate change, is that the crops that could in the past be only growth in certain low altitudes, requiring prolonged exposure to heat can now be growth in higher regions such as Manang and Mustang districts of Nepal. The is evidently due to increase average temperature of the region, thus allowing the crop to perform well. However, as noted earlier, there are more prominent negative impacts than the benifits we can derive from the climate change.

Nepal became signatory of UN Framework Convention on Climate Change (UNFCC) in 1994. Ever since then the issue of climate change has been routinely incorporated into policy and strategic frameworks by Government of Nepal. Following policies and strategies are currently in place to help mitigate the impacts of climate change in Nepal:

1. Constitutional approach

Article 30(1) of the constitution of Nepal guarantees a "clean environment" as fundamental right and elaborates that "every citizen shall have the right to live in a clean and healthy environment." Article 30(2) of the constitution also encourages the state to formulate necessary legal framework to balance environment and development.

2. Approaches in periodic plans
3. Climate change policy, 2067 (2011)
4. National climate change policy, 2076 (2019)
6. National adaptation plan of action (NAPA, 2010)
7. Framework for local adaptation plan for action (LAPA, 2011)

### Highlights of NCCP, 2076

Reasons for Nepal's vulnerability to climate change are Topographical fragility, fragile geological structure, sensitive ecosystems and diversity of climate and micro-climate zones prevalence in Nepal. Additionally, poverty, illiteracy, social disparity and high dependence of communities on natural resources for livelihood.

- Nepal National Red Strategy implemented to check greenhouse gases emmission from deforestation and forest destruction.
- Provision of climate change coordination committee in Chairmanship of Prime minister.
- Establishment of Climate change management section under Ministry of Forest and Environment as focal point for implementation of international treaty.
- Provision of Climate budget code by National planning commission.
- Alternative energy promotion center has established climate and carbon unit for executing Clean Development Mechanism (CDM) related activities.

**Goal**

To contribute to socio-economic prosperity of the nation by building a climate resilient society.

**Objectives**

a. To enhance climate change adaptation capacity of persons, families, groups and communities vulnerable to, and at risk of, climate change;
b. To build resilience of ecosystems that are at risk of adverse impacts of climate change;
c. To promote green economy by adopting the concept of low carbon emmission development;
d. To mobilize national and international financial resources for climate change mitigation and adaptation in just manner;
e. To conduct research, make effective technology development and information service delivery related to climate change;
f. To mainstream or integrate climate change issues into policies, strategies, plans and programs at all levels of State and sectoral areas;
g. To mainstream gender equality and social inclusion (GESI) into climate change mitigation and adaptation programs.

**Policies** (Sectoral policy, strategies and working policies)

1. Agriculture and food security
2. Forest, biodiversity and watershed conservation
3. Water resources and energy
4. Rural and urban habitats
5. Industry, transport and physical infrastructure
6. Tourism and natural and cultural heritage
7. Health and drinking water and sanitation
8. Disaster risk reduction and management
9. Gender equality and social inclusion, livelihoods and good governance
10. Awareness raising and capacity development
11. Research, technology development and expansion
12. Climate finance management

Policy number 9-12 are inter-thematic areas.
