
## Mango cultivation

### Varieties of mango and classification

```{r mango-varieties}
tribble(
  ~"classification", ~"group", ~"cultivar",
  "Based on maturity", "Very early varieties", "Bangalora, Banganpalli",
  "Based on maturity", "Early varieties", "Bombay green, Bombay yellow, Himsagar, Kesar, Swarnarekha",
  "Based on maturity", "Mid season varieties", "Dashehari, Langra, Zardalu, Krishnabhog, Alphonso",
  "Based on maturity", "Late varieties", "Samarbahisth, Chausa, Fazri, Neelum, Mulgoa, Amrapali, Kaitki",
  "Based on firmness of pulp", "Soft pulp type", "Langra, Dashehari",
  "Based on firmness of pulp", "Firm pulp type", "Bangalora, Mallika",
  "Based on firmness of pulp", "Loose pulp type", "Sukul, Baramasia",
  NA, "Sucking type", "Luknow safeda, Mithua, Ghazipur, Gilus",
  "Based on embryony", "Polyembryonic varieties", "Bellary, Bappakai, Chandrakaran, Goa, Kurukkan, Olour, Nileshwar dwarf",
  "Based on embryony", "Monoembryonic varieties", "Almost all commercial varieties",
  "Based on fibre content of fruit", "Fibrous", "Sukul, Baramasia",
  "Based on fibre content of fruit", "Fibreless", "Langra, Dashehari, Zardalu, Ranipasand", 
  "Based on fibre content of fruit", "Moderately fibrous", "Fazil, Sipia", 
  NA, "Regular bearing varieties", "Amrapali, Neelum, Bangalora, Mallika",
  NA, "Colored varieties", "Sindhuria, Gulabhas, Zafrani",
  NA, "Hybrid varieties", "Mallika (Neelam x Dashehari)",
  NA, "Hybrid varieties", "Amrapalli (Dashehari x Neelam)",
  NA, "Hybrid varieties", "Ratna (Neelum x Alphonso)",
  NA, "Hybrid varieties", "Neeluddin (Neelum x Himayuddin)",
  NA, "Hybrid varieties", "Arka aruna (Baganpalli x Alphonso)",
  NA, "Hybrid varieties", "Nelphonso (Neelum x Alphonso)",
  NA, "Hybrid varieties", "Neelshan (Neelum x Baneshan)",
  NA, "Hybrid varieties", "Neeleshwari (Neelum x Dashehari)"
) %>% 
  rename_all(function(x)str_to_sentence(str_replace_all(x, "_", " "))) %>% 
  knitr::kable(booktabs = TRUE, caption = "Mango varieties and its classification", longtable = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped") %>% 
  kableExtra::column_spec(column = 1:3, width = c("8em", "8em", "16em"))
```

## Citrus cultivation

### Botany and classification

Citrus belongs to the family Rutaceae. Almost all the cultivated species, varieties and forms of citrus, fortunella and poincitrus are diploid, the monoploid number of chromosome being 9 i.e. 2n = 18.

#### Related genera

- Trifoliate orange ( _Poncitrus trifoliata_): Very thorny parts, fruits inedible, used as root stock and for inter-generic crosses, polyembryonic. Susceptible to blight but tolerant to most viral diseases, Phytophthora root rot and citrus nematode.
- Fortunella (Kumquat) ( _F. japonica_, _F. hindsii_, _F. margarita_, _F. crassifolia_) used for ornamental purpose. Trees and shrubby. Fruits are made into candy.

#### Orange group

- Sweet orange ( _Citrus sinensis_): Some of the varieties are Junar, Shamauti and Pineapple, Valencia late, Hamlin, Jaffa, Blood red malta, Mosambi, Washington navel orange.
- Sour orange ( _C. aurantium_): World's principal root stock, cold hardy and resistant to several diseases.

#### Mandarin group

```{r mandarin-citrus}
tribble(
  ~"mandarin_group", ~"scientific_name", ~"cultivars", 
  "Mandarin orange", "\\textit{Citrus reticulata}", "Includes all loose skinned orange", 
  "King mandarin", "\\textit{Citrus nobilis}", "\\textit{Citrus reticulata} x \\textit{Citrus sinensis}", 
  "Willow leaf mandarin", "\\textit{Citrus deliciosa}", NA,
  "Kinnow mandarin", NA, "Hybrid of king and willow leaf mandarin", 
  "Nagpur santra", NA, "Ponkon",
  "Nepali suntala", NA, NA,
  "Satsuma mandarin", "\\textit{Citrus unshiu}", NA
) %>% 
  rename_all(function(x)str_to_sentence(str_replace_all(x, "_", " "))) %>% 
  knitr::kable(booktabs = TRUE, caption = "Mandarin group and some examples", escape = FALSE, longtable = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped")
```

The introduced loose skinned orange are Yoshida Ponkan, Clementine, Kinnow and Israeli.

#### Acid group

```{r acid-citrus}
tribble(
  ~"mandarin_group", ~"scientific_name", ~"type",
  "Citron", "\\textit{Citrus medica}", "Monoembryonic seed",
  "Lemon", "\\textit{Citrus limon}", "Eureka lemon, Lisbon lemon, Lucknow seedless, Nepali oblong, Hill lemon",
  "Lime", "\\textit{Citrus aurantifolia}", "Acid lime/kagzi lime (Paper thin rind)", 
  "Rangapur lime", "\\textit{Citrus limonia}", "Mainly used for root stock. It is considered to be hybrid between mandarin and lemon like fruit", 
  "Sweet lime", "\\textit{Citrus limettoides}", NA, 
  "Rough lemon", "\\textit{Citrus jambhiri}", "It is indigenouly found in Nepal. Mainly used for making concentrated Nibu juice, an important root stock",
  "Pomelo", "\\textit{Citrus maxima}, \\textit{Citrus grandis}", "Strictly monoembryonic", 
  "Grapefruit", "\\textit{Citrus paradisi}", "Closely related to pomelo. As the fruits are borne in cluster (upto 12) like grapes"
) %>% 
  rename_all(function(x)str_to_sentence(str_replace_all(x, "_", " "))) %>% 
  knitr::kable(booktabs = TRUE, caption = "Acid group of citrus fruits", escape = FALSE, longtable = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped") %>% 
  kableExtra::column_spec(column = 2:3, width = c("8em", "18em"))
```

#### Citrus hybrids

```{r hybrids-citrus}
tribble(
  ~"hybrid_name", ~"parentage",
  "Citrange", "\\textit{Poncitrus trifoliata} x \\textit{Citrus sinensis}",
  "Citrange quat", "Citrange x Kumquat",
  "Citrangors", "Backcross between Citrange and \\textit{Citrus sinensis}",
  "Cicitrange", "Citrange x \\textit{Poncitrus trifoliata}",
  "Citrumelo", "\\textit{Poncitrus trifoliata} x \\textit{Citrus paradisi}",
  "Citandarin", "\\textit{Poncitrus trifoliata} x \\textit{Citrus reticulata}",
  "Limequat", "\\textit{Citrus aurantifolia} x \\textit{Fortunella japonica}",
  "Lemonimes", "\\textit{Citrus limon} x \\textit{Citrus convantifolia}",
  "Tangors (Temple orange)", "\\textit{Citrus sinensis} x \\textit{Citrus reticulata}",
  "Tangelos", "\\textit{Citrus reticulata} x \\textit{Citrus paradisi}"
) %>% 
  rename_all(function(x)str_to_sentence(str_replace_all(x, "_", " "))) %>% 
  knitr::kable(booktabs = TRUE, caption = "Citrus hybrids and their parentage", escape = FALSE, longtable = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped") %>% 
  kableExtra::column_spec(column = 2, width = "16em")
```


### Principal rootstocks

Today, five types of rootstock predominate in relatively not cool climates where cold or freezing weather is not probable, especially Florida and southern Europe.

Sour orange rootstock: it is the only rootstock that truly is an orange (the Citrus aurantium or bitter orange). It is vigorous and highly drought-resistant.

Poncirus trifoliata: it is a close relative of the Citrus genus, sometimes classified as Citrus trifoliata. It is especially resistant to cold, the tristeza virus, and the fungus Phytophthora parasitica (root rot) and grows well in loam soil. Among its disadvantages are its slow growth-it is the slowest growing rootstock-and its poor resistance to heat and drought. It is primarily used in China, Japan, and areas of California with heavy soils.

Swingle citrumelo: it is tolerant of tristeza virus and Phytophthora parasitica and moderately resistant to salt and freezing. This rootstock selection was hybridized from the Duncan grapefruit (Citrus paradisi Macfadyen) and the Poncirus trifoliata (L.) Raf. by Walter Tennyson Swingle in Eustis, Florida, in 1907. It was released by the US Department of Agriculture to nurserymen in 1974.

Troyer citrange and Carrizo citrange: these reasonably vigorous rootstocks are resistant to Phytophthora parasitica, nematodes, and tristeza virus and show good cold tolerance. They also are highly polyembryonic, so growers can obtain multiple plants from a single seed. Citrange, however, does not do well in clay, calcareous or high-pH soils, and is sensitive to salinity. It is not feasible as rootstock for mandarin scions, as it overgrows them by producing branches of its own in competition with the grafted budwood.[3] Citranges are hybrids of the Washington navel orange and the Poncirus trifoliata. The original crosses, made in the early 1900s by the U.S. Department of Agriculture with the intention of producing cold tolerant scion varieties, were later identified as suitable for use as rootstocks. The commercial use of these rootstocks began in Australia in the 1960s. The Troyer variety generally is found in California, while the Carrizo variety is used in Florida.

Cleopatra mandarin: it is tolerant of salinity and soil alkalinity and also suitable for shallow soils. It is used primarily in Spain, Australia, and Florida. Dade County, for example, has 85% calcareous soil, a typical trait of land that has been under water.[4] The Cleopatra mandarin, originated in India and introduced into Florida from Jamaica in the mid-nineteenth century, has been distributed and tested as a rootstock throughout the world. Nowadays, however, it is considered an inferior rootstock because it is sensitive to many diseases, grows slowly, and is difficult to propagate.

## Grape cultivation

### History, uses and production

- Cultivated all over the world, viticulture is a major horticultural industry with most grapes grown for cultivation of wine.
- Today fruit is used as fresh fruit, preserved fruit, wine and raisins.
- The vine grows vigorously; in some places, it is considered invasive weed.
- It is non-climacteric fruit that grows on perennial and deciduous woody vines.
- Fruit grow in cluster of 6-300, can be crimson, black, dark blue, yellow green and pink.
- Most cultivars of grape are generated from variants of _Vitis vinifera_.
- Wild form of the crop and germplasm resource may arise from _V. labrusca_, _V. riparia_, _V. rotundifolia_ and _V. amurensis_.

### Botany and morphology

- Flowers and fruit clusters are borne only on new shoots arising from dormant buds. These buds are formed in the axil of leaves the previous season and are called compound buds. These contain a group of three separate buds.
- When growth starts in the spring, the primary or central bud breaks dormancy and produces the fruiting shoot. On young vines, this shoot may remain entirely vegetative and produce no fruits. Grape shoots do not form terminal buds, and the tip typically dies back in the winter to the lignified or ripened portion of the cane.
- Flowering or bloom occurs after the shoot reach 18-24 inches or about 4 weeks after bud break.
- Number of flower clusters that may develop from a single dormant bud is determined to a great extent by vine vigor and especially by exposure to sunlight during the previous season.
- Grapes flower later than most tree fruit and are mostly pollinated by wind. Fruit set is influenced by the weather.
- Frequently, during the growing season, a short, weak shoot is produced in the axil of the leaf adjacent to the bud and is known as a summer lateral. These can form new leaves, but often don't mature well enough to survive the winter and are usually removed in dormant pruning.
- After the leaves drop in fall, the term cane is used to refer to the mature, dormant shoot. It is from these canes that next year's fruiting wood is selected at pruning time and from which propagation wood is taken.
- The root system is concentrated in the well-aerated upper 18-24 inches of soil but can penetrate deep into the soil if not impeded by a hard pan or high water table.
- Grapevine require some kind of structural support for commercial and home production. The trellis or arbor provides support so the vine can be managed efficiently.
- Vines must be pruned annually to maintain productivity and fruit quality.

### Climate

- Require frost free period of 150-180 days. Impacts fruit quality (lowered sugar content) if growing season is short.
- Vineyard diseases may be related to climate, especially humidity and temperature.
- Cold hardiness develops midway through the winter.
- Once growth starts in spring and the buds have extended to approximately 2 inches in length, they are susceptible to damage from temperature below $28^\circ F$.
- South and east facing slopes are desired, very low lying areas should be avoided.

### Soil and water

- Highest growth in soils with good internal drainage but can be grown on a variety of soil types.
- Prior to vineyard establishment, selection of site with drainage considerations is important.
- After selection, soil fertility maintenance is generally more economical.

### Cultivar selection

- Vine hardiness, fruit characteristics, season of ripening, tolerance to disease and insects, vine characteristics.
- American type (slip-skinned): Concord, Natawba and Norton
- Table type:
  - Seeded: Buffalo, Alden, Steuben, Seneca, Golden muscat
  - Seedless: Marquis, Jupiter, Thompson seedless

(...continue)

## Pineapple cultivation

The pineapple ( _Annanas commosos_) is perennial monocarpic herb (i.e. each stem flowers only once and dies after fruiting), xerophytic plant with thick, long sessile and spine leaves. Multiple fruits called synocarp is a parthenocarpic type.

### Origin

The center of origin is believed to be in the Paraguay River drainage basin area in South America where the original seed species still occur wild. Another center of origin is siaid to be along the river banks of the most southern ares in Guyana. The records of the early explorers show that the pineapple was widely distributed throughout most of tropical America and later taken to other tropical areas of the world, e.g. Africa, Asia, the South Pacific and Australia.

### Distribution

In Nepal, pineapple is grown in 45 districts (? when). Major pineapple growing districts are: Jhapa, Morang, Sunsari, Dhanusha, Chitwan, Bara, Nawalparasi, Dhading and Nuwakot.

### Uses

Originally, pineapple was consumed only as a fresh fruit. With the development of processing industry, the fruit is now prepared and consumed in various forms such as pineapple chunks, slices, juices, syrups, jams, crushed pineapple, diced pineapple, etc. Also, the wastes from the processing of the fruit are now further processed into sugar, wines, vinegar, animal feed, etc.

High quality fibers, ideal for the manufacture of luxury clothing, are produced from the leaves of the pineapple varieties. Fibers of the wild species are used for making rope and fishing nets. The fiber is also used for making pulp in the paper industry.

### Soil and climate

Loam, clay loam and clay soil high in organic matter with improved drainage and water control systems are suitable for pineapple cultivation. Soils prone to water logging are unsuitable since they hasten the development of root rot. Best growth is achieved on well-drained, fertile, sandy loam soil with a pH range of 4.5-5.5.

It's habitat is humid tropical region (\SIrange{15.6}{32}{\celsius}). A high temperature at night is deleterious and a difference of at least \SI{4}{\celsius} between day and night temperatures is desirable. 

- Rainfall 1000-1500 mm,
- pH: 5.0 - 6.0,
- Elevation: > 100 masl

Although the pineapple plant is fairly resistant to drought, it requires medium to high evenly distributed rainfall for good growth and the production of healthy fruits. It will grow with an annual rainfall as high as 2500 mm once adequate drainage is provided.

### Varieties

There are 3 main grouping;

- Cayennue: Canning + Table purpose, e.g. Hilo, Chailotte, Kew and Giant kew
- Spinach: Canning + Table purpose, e.g. Red spinach, Singapore, Spinach and Masmerah
- Queen: Queen, Mouritius, Kipley queen and Alexender

Other: 

```{r pineapple-varieties-other}
tribble(
  ~"group", ~"example",
  "Mardilonus", "Perolena, Maipur Random, Mainpure, Milagrena",
  "Permemukulo", "Permanmbuce, Perola and Papelon"
) %>% 
  knitr::kable(booktabs = TRUE, col.names = NULL) %>% 
  kableExtra::kable_styling(font_size = 10) %>% 
  kableExtra::column_spec(column = 2, width = "12em")
```


### Propagation methods

The pineapple is propagated vegetatively. Planting materials obtained from various parts of the plant are identified based on the part they are obtained from.

- **Suckers** (Ratoon or side shoots): Ratoon suckers are produced from ground level and when used will produce shoot in 12-14 months of planting. Side shoots/suckers are produced from above ground level and, when used, bear within 18-20 months of planting.

- **Slips** (Basal suckers): They are the shoots located at the base of the fruit. When used, they bear within 14-16 months of planting.

- **Crown**: Bunch of leaves on top of fruits. When used, it requires > 24 months for them to bear.

- **Stem bits or disc**: After harvesting, the stem is divided into small bits containing at least 2-3 buds. Each bit is known as disc. When used, it requires > 24 months for them to bear.

Sucker and slip are usually preferred for planting.

To avoid rotting of plants after planted, sucker and slips are cured by stripping off the lower leaves followed by drying in sun or in partial shade for about a week before planting.

### Planting

When planting, ensure that the "heart" of the plant is above soil level, otherwise rotting may occur, particularly if the soil is wet. Either the single or double-row system can be used. However, for a more cost-effective land use, the philosophy is to have the largest number of plants per unit area as is practical. This is better achieved by the double row system, which is the recommended system for most conditions.

#### Single row

The rows are spaced 150 cm apart and plants are spaced 60 cm within the row. This allows the population of 11000 plants/hectare.

#### Double row

The spacing of 30 cm, 60 cm and 90 cm between plants, rows and beds respectively in a double row system accommodates upto 43,500 plants/hectare.

### Planting time

Planting time is early monsoon, but after monsoon is preferred time in heavy rainfall areas. Winter planting is possible if water is available.

### Manure and fertilizer

The fertilizer recommendation in pineapple is based on plant density. In general 30t of well decomposed FYM is required for a hectare.

NPK \@ 16g: 5g: 12g per plant per year is recommended, with full dose application of Phosphorus and Potassium and top dressing of half of the nitrogen during May-June on individual basis.

### Weed control

Pineapple plants grow slow and do not cover the ground well enough to suppress weeds from developing. Weed management cost in pineapple may accrue upto 40\% of the total production cost. Manual weeding is difficult, due to presence of spines, and expensive too. Following chemical weedicides are usually applied:

- Bromacil + Diuron \@ 2 kg a.i. per hectare (pre-emergence)
- Glyphosate \@ 2 kg per hectare in \SI{1000}{\litre} of water as directed spray (post-emergence)

### Flowering

Flowering is not uniform and regular. This is due to various ages, growth and health of planting material used. Although it offers a chance to take harvest of fruit spred several months from summer to winter, the later harvest is commercially not good. Young and unhealthy plants should not be induced to flower since they produce small fruits.

Artificial flower-induction permits better scheduling of the harvest as it takes 5 months from the time of induction to full maturity of fruits. Flowering in pineapple may be induced if the stems are kept in horizontal position, due to geotropism. In this way, accumulation of auxins induces bud initiation. The plants in the block to be treated should be homogeneous in size and not less than 12 months old or possess less than 25 leaves. Treatment should take place during cool hours of the day, early in the morning or late afternoon (later preferred).

- The main products used to induce flowering are: Naphthalene Acetic Acid (NAA) and Ethrel. It is done using Planofix (200 ppm), Ethephon (20 ppm) and CaC2 (2\%) for early flowering, uniform flowering, to increase the fruit size and to induce fruiting at anytime of the year.
- Plants raised from suckers and treated with 100 ppm Ethrel + 2\% Urea + 0.004\% CaCO3 at 18 months after planting gave 96\% flowering.
- Ethylene induces 90-100\%  flowering and CaC2 (1g) induces 60-66\% flowering in pineapple.
- NAA induces 60\% flowering, 15-20\% large fruit size
- Increased number of slips delay maturity of fruits, so they are recommended for removal as soon as possible.
- Desuckering can be delayed as much as possible because fruit weight increases with increase in number of suckers/plant.

### Harvesting

As a climacteric fruit, pineapple should be harvested only after the fruit fully ripens. In Nepal, it normally ripens in July-September. Winter ripen fruits are more acidic.

### Maturity indices

- Drying of bracts
- Color changes to yellow
- Less angular near eyes
- TSS: 8-16\%

The pineapple is best harvested with hands by breaking or cutting the stalk with crown a few centimeters below the fruit. Crown removal quickens ripening and reduces shelf life of the fruit.

### Yield

Well managed crop at high density planting normally yields \SIrange{40}{50}{\tonne\per\hectare}. Single row gives \SI{20}{\tonne\per\hectare}.

### Insect pests

1. Mealy bug ( _Dysmiococcus brevipus_)
2. Fruit-eating beetle ( _Carpophilus dimidiatus_)
3. Termites ( _Odontotermes spp_)

### Diseases

1. Mealy bug wilt (Transmitted by mealy bug)

## Strawberry cultivation

- Cultivation as well as production of one of the most popular dessert crop -- Strawberry ( _Fragaria ananassa_) has increased in recent years. Most production occurs in the northern hemisphere.
- Two types are commonly grown: Day neutral and short-day types

### cultivation systems

1. Matted rows

- Employs runner as primary yield component
- Both mother and daughter plants are allowed to runner freely, with periodic training into narrow hills.

2. Hill/plasticulture system
- Relies on crown as primary yield component, any runners that form are removed.

The former system is used to grow short day cultivars in climates with short summers and cold winters (continental europe and North america). The latter system is used primarily in areas having warm winters and hot or moderate summers such as California, Italy and Spain.

- Strawberry is non-climacteric fruit
- Most widly planted cultivar in the world is "Camarosa"
- Other species are _F. chiloensis_ and _F. moschata_.

Stolons: These are the stem that grows horizontally above ground and occur in plants such as strawberry and bermudagrass. Roots may arise at the nodes of the stem as it creeps on the surface. When stolons have long internodes that originate at the base of the crown of the plant (in strawberry), the stolon is called a runner.

### Climate

- Flowering habit of strawberry is directly related to temperature and strongly interacts with photoperiod.
- Cool temperature at summer (\SI{17}{\celsius}) allows flowering induction to occur even under long photoperiod, even in short day varieties.
- Chilling requires at \SIrange{0}{7}{\celsius} for reproductive development (for breaking bud dormancy with vernalization).

### Soil

- Sensitive to salinity and water stress. Soil of pH range 5.5-6.5 is preferred.
- Sub-surface irrigation to root zone is suitable to check fungal growth.

### Varieties

Nyaho, Uno

### Spacing and time of planting

- Spacing: \SI{90}{\centi\meter} x \SIrange{30}{45}{\centi\meter}
- Timing of planting: Late summer or spring
- Plant density: 1500 plants per ropani

### Manuring and fertilization

- FYM: \SI{2.25}{\tonne\per\hectare}
- Urea: \SI{97.83}{\gram\per\hectare\per\year} per plant
- MoP: \SI{140}{\gram\per\hectare\per\year} per plant

### Harvest

- Crop matures usually within the months of Kartik-Chaitra but harvest is signalled when 3/4th of fruit turns red in color.
- Fruits are harvested every other day to make sure berries are not over-ripe.
- Yield: \SIrange{12}{15}{\tonne\per\hectare}

### Insect-pest management

- Around 200 species of pests are known, most damaging of which include slugs, moths, fruitflies, strawberry root weevils, thrips, sap beetles, crown moth, mites, aphids, and mainly lepidopteran caterpillars.
- Diseases:
  - Powdery mildew, leaf spot ( _Sphaerella fragariae_), leaf blight ( _Phomopsis spp._)
  - Verticillium wilt, black rot
  - Gray mould, rhizopus rot
  - Nematodes

## Macadamia nut cultivation

Macadamia is an evergreen genus (family: Proteaceae) that grows 2-12 m (7-40 ft) tall. The leaves are arranged in whorls of three to six, lanceolate to obovate or elliptic in shape, 6-30 cm (2-10 in) long and 3-13 cm (1-5 in) broad, with an entire or spiny-serrated margin. The flowers are produced in a long, slender, simple raceme 5-30 cm (2-10 in) long, the individual flowers 10-15 mm (0.4-0.6 in) long, white to pink or purple, with four tepals. The fruit is a hard, woody, globose follicle with a pointed apex, containing one or two seeds.

- _Macadamia integrifolia_ Maiden & Betche
- _Macadamia jansenii_ C.L.Gross & P.H.Weston
- _Macadamia ternifolia_ F.Muell.
- _Macadamia tetraphylla_ L.A.S.Johnson

Nuts from _M. jansenii_ contain toxic amounts of cyanogenic glycosides, The other three species are cultivated in the commercial production of macadamia nuts for human consumption.

Previously, more species with disjunct distributions were named as members of this genus Macadamia. Genetics and morphological studies published in 2008 show they have separated from the genus Macadamia, correlating less closely than thought from earlier morphological studies. The species previously named in the genus Macadamia may still be referred to overall by the descriptive, non-scientific name of macadamia.

Three of the four species -- _Macadamia integrifolia_, _Macadamia ternifolia_, and _Macadamia tetraphylla_ -- are of commercial importance. The other species, _M. jansenii_, produces poisonous nuts resulting from toxic amounts of cyanogenic glycosides.

### Cultivation

The macadamia tree is usually propagated by grafting, and does not begin to produce commercial quantities of seeds until it is 7-10 years old, but once established, may continue bearing for over 100 years. Macadamias prefer fertile, well-drained soils, a rainfall of 1,000-2,000 mm (40-80 in), and temperatures not falling below $10^\circ C$ ($50^\circ F$) (although once established, they can withstand light frosts), with an optimum temperature of $25^\circ C$ ($80^\circ F$). The roots are shallow and trees can be blown down in storms; they are also susceptible to Phytophthora root disease.

Cultivars: Beaumont, Maroochy, Nelmac II, Renown

- South africa is the leading producer of the Nut.
- Normal yield: \SIrange{18}{20}{\kilo\gram} per tree (8-10 years old)
- Globally 160,000 tons of the nut was produced in 2015.
- The crop is indigenous to Australia
- The nut is considered toxic to dogs.
- Cultivation started in Nepal during 1964 from Trishuli Horticulture Farm, Nuwakot.
- Crop is currently grown in Kaski and Syangja
- In Nepal crop is grown in elevation range of 300 - 1500 masl (optimum: 900-1400 m)
- A sapling of the plant costs Rs 75-150 each.

### Food and nutrition^[https://en.wikipedia.org/wiki/Macadamia]

The seeds (nuts) are a valuable food crop. In a 100-gram amount, macadamia nuts provide 740 Calories and are a rich source (20% or more of the Daily Value, DV) of numerous essential nutrients, including thiamin (104% DV), vitamin B6 (21% DV), manganese (195% DV), iron (28% DV), magnesium (37% DV), and phosphorus (27% DV) (table). Macadamia nuts are 76% fat, 14% carbohydrates, including 9% dietary fiber, and 8% protein.

Compared with other common edible nuts, such as almonds and cashews, macadamias are high in total fat and relatively low in protein. They have a high amount of monounsaturated fats (59% of total content, table) and contain, as 17% of total fat, the monounsaturated fat, omega-7 palmitoleic acid.

## Guava cultivation

- _Psidium guajava_ (Myrtaceae)
- Origin: Tropical america; originated in the warm, lowland tropics
- Contains low energy, low fat and high amounts of vitamin C.
- Commercial trade is smaller in volume compared to other fruits.

### Morphology and Botany

- Photoperiod insensitive
- Can be forced to flower by pruning
- Woody is hardy; resistant to strong winds.
- The rhythmic flushing of vegetative growth and flowering influences when and how pruning is
carried out.
- Guavas are forced to produce year round, and normally the actively growing shoots are tipped once they become too large. This tipping forces regrowth and new flowers to be induced, to keep production high and constant.
- As soon as the harvest is finished, a peripheral pruning of the canopy is performed, in order to force new shoots that will bear flowers.

### Climate and soil

- Temperature lower than $8\ ^\circ C$ is inappropriate for commercial cultivation.
- Moderately tolerant to saline conditions.

### Varieties

- Lucknow-49, Allahabad safeda, Seedless

### Propagation 
- Stem and root cuttings, Air layering and Budding
- Guava trees propagated by grafting have tap roots that provide substantial anchorage. 
- Trees propagated by cuttings have weaker root
systems.

### Planting and orchard establishment

- Spacing with fillers: 3.1 x 7.6 (424 trees/ha)
- Permanent spacing: 6.2 x 7.6 (212 trees/ha), or 4.6 x 7.6 (286 trees/ha)

```{r climac-vs-nonclim}
cl_ncl <- data.frame(climacteric = c("Avocado", "Banana", "Guava", "Mango", "Papaya", "Sapota"), 
                     non_climacteric = c("Litchi", "Mangosteen", "Pineapple", "Mountain apple (Bayer)", NA, NA))

cl_ncl %>% 
  rename_all(function(x)str_replace_all(x, "_", " ") %>% str_to_sentence()) %>% 
  knitr::kable(booktabs = TRUE, caption = "Climacteric and non climacteric fruits commonly cultivated in Nepal") %>% 
  kableExtra::kable_styling(latex_options = "striped", position = "center")
```

### Post-harvest

- High moisture loss occurs
- No dramatic change in taste during storage at higher temperatures (10-15 $\ ^o C$)

## Litchi cultivation

**Taxonomy** 

Family: Nepheleae. Family contains about 150 genera and 2000 species. Related species of the family are Rambuton ( _Nephelium lappaceum_) and pulsan ( _N mutabile_). The genus Litchi has 3 species:

  - Sub species chinensis -- Commercial litchi.
  - Sub species philippinensis -- Developed in Philippines. It has long, oval shaped fruit with inedible flesh and long thorn like protuberances which can be used as root stock.
  - Sub species Javenensis bears fruit similar to Chinensis but the aril is thinner.
  
**Varieties**

Mazzafarpur (Bears profusely, fruits are deep orange to pink, pulp is sweet, tough, moderately juicy, good flavored and pulp:seed ratio is high), Deshi (Bearing is moderate and regular, fruit is oval, oblong or oblong conical in shape), Rose scented (Fruits are delicious rosy flavor in the aril. Trees are vigorous. Fruits are mostly oblong or oblong conical in shape, deep rose pink in color, pulp grayish white, and soft and very sweet), Shahi, Dehradun, Culcuttia, Seedless late, Saharanpur, etc.

Origin: Originated from northern China, Northern Vietnam and Malaysia

Uses: Fruit makes excellent canning. The preservation of fruit can be done in honey.

Botany: Long lived, medium to large, brached, round topped, evergreen tree reaching upto 10 m or more height with short stocky trunk. Leaves are compound, alternate, consists of 4-7 oblong leaflets, glossy dark green. Bark is grayish brown and rough. Vegetative growth is rhythmic and occurs in 3-4 recurrent flushes alternating with period of rest. The inflorescence is compound raceme developing both from terminal and axillary buds. Flowers are small, male, pseudo hermaphrodite (functional male) and hermaphrodite. The edible portion is aril. The seed is dark brown in color.

Climate: The litchi is adapted to the warm sub-tropic, cool, dry, frost free winter, and long, hot summer with high rainfall and high humidity. The litchi usually likes low elevation but can be grown upto 800 masl. The optimum temperature requirement for flowering is 16-20 degree C. High temperatures, more than 30 degree C and low relative humidity (less than 60%) during fruit development stage cause the fruit cracking in litchi. The hot winds in summer cause fruit cracking and subsequently damage the pulp (aril). Sometimes it limits the expansion of litchi cultivation. Wet spring, dry summer and light winter are desirable conditions for fruiting in litchi.

The litchi can grow under wide variety of soild including alluvial soil, loams, and heavy clays with rich source of organic matter. Litchi can withstand water for a considerable period, provided the water does not stagnate, but will die after prolonged immersion. It is suggested that new plant should be grown in soil taken from the vicinity of old trees to introduce the mycorrhiza. The optimum pH range is 5.5-6.5

## Kiwi cultivation (_Actinidia_ spp.)

- Spreading type shrub of family Actinidiaceae. It is a cross pollinated crop.
- There are two major kinds of Kiwi species found in Nepal: _Actinidia strigosa_ and _A. callosa_.
- In Nepal, Kiwi fruit was first planted in 2040 in Charikot, Dolakha. Cultivated mostly in cropping regions of fingermillet and maize and in mountain regions below the line of apple farming, in elevation 1200-2400 masl.
- The crop is not suited to areas with temperature greater than \SI{35}{\celsius}. It tolerates temperature of \SI{10}{\celsius} and that below \SI{5}{\celsius} causes stem and bark of branches to crack and dry.
- According to 2071/72 statistics, a total of 52 hectares of Nepal are planted kiwi with current production of 368 tons.
- Native to china but popularly cultivated in New-zealand since 1910. Hayward Wright first developed the kiwi cultivar called Haywayd in 1924.
- Districts considered suitable for kiwi farming are Illam, Panchthar, Dhankuta, Kavre, Dolakha, Sindhuli, Nuwakot, Lamjung, Kaski, Parbat, Baglung, Myagdi and Salyan.
- South facing slopes with assured irrigation are the most desirable, although, areas with rainfall reaching \SI{1500}{\milli\meter\per\year} are suitable too.
- Flowering: March-April
- Harvesting: November-December
- Popular propagation technique: Splice grafting and bench grafting mostly and less commonly budding and cutting.
- NPK: 170:56:100 per year per hectare.
- Spacing: 6 m x 6 m
- - Four varieties are selected for commercial evaluation in Nepal. Important varieties are:
  - Soyou ( _A. chinensis_): Red fleshed early variety (Bhadra), require less chilling hours and doesn't have hair on fruit skin. Fruits are small in size ($39.3 \pm 9.6 g$). They can be stored in room temperature for about a month.
  - Bruno ( _A. deliciosa_): Green fleshed, quicker maturing, high yielding cultivar with oblong shaped and medium size fruits ($63.9 \pm 12.4 g$). It has better storability than Hayward (can be stored upto 3 months in room temperature). The fruits mature in second week of Kartik (earliest among green fleshed cultivars). The fruit has brown hairs on skin.
  - Hayward round ( _A. spp._): Require 600-800 hours of chilling period. Fruits are larger in size ($81.4 \pm 22.1 g$) and nearly round in shape. It matures from fourth week of kartik and the harvest may be stored for about 3 months.
  - Allison: Grown in altitude of range \SIrange{1200}{2400}{\meter} above sea level. Fruits are brown and hairy, green fleshed, long and have good storability.
  - Abbot
- Golden and red kiwi are expensive and less productive than the rest.
- In temperatures \SIrange{0}{3}{\celsius} and 80\% RH, kiwi can be stored for upto 3 months.

### Planting method and timing

- As a deciduous crop, planting is done in Poush-Magh.
- The crop is dioecious, so nursery should possess both male and female plants in proportion of 1:5 - 1:8.
- Some varieties may exhibit self incompatibility, so planting of a combination of varieties may be beneficial.
- Male plants should be arranged in center to facilitate pollination.
- While planting in pits, which is done best during evenings, planting boards should be used.
- In grafted seedlings, graft association should be established above the ground.
- Irrigation is provided in 3 days interval.
- After planting, bamboo stakes are used to support the plants.
- Mulching prior to irrigation increase efficiency of water use.
- Rootstock varieties: Bruno, Monty, Allison, Ebert.
- Seed matures at Mangsir and this month is suited for seed collection.
- Seed separation: At 12% TSS, fruit pulp is extracted and mixed with sand to facilitate mashing and separation. Thus obtained seed is sun dried.
- Seed is sown after 1-2 months of storage, Germination increases when cold temperature treatment (\SI{5}{\celsius} for 2-3 weeks) is given.
- Seed bed nursery is established at the end of falgun. Seeds are sown at 15 x 5 cm spacing (R-R x S-S) and covered with 1 cm layer of soil.
- Ready for planting seedlings are 4-5 leaved (1-1.5 months). Generally, seedlings are suitable for grafting 10-12 months after sowing.

### Training

- Since the crop grows as a vine, support system is essential for good crop management. It does not produce tendrils but the stem roll over the supporting medium.
- Tip notching/removal of vine should be first done at 7-8 months of planting (\@ 6 feet height).
- Two main systems of training are T-wire method and pergola method.
- Flushes arise from one year old branches in the month of Poush-Magh (Apply 10\% bordeaux paste in the cut surface).

### Disease and insects

- Diseases: Root rot ( _Phytophthora_ spp.), Botrytis fruit rot ( _Botrytis_ spp.), Twig blight ( _Sclerotina_, spp.), Leaf spot ( _Alternaria_ spp., _Phomopsis_ spp., _Glomerella_ spp.)

- Hopper, stem borer and leaf eating caterpillar

## National fruit development directorate (Institutional development of fruits)

The agro-ecological conditions of Nepal are very much suitable for the successful cultivation of large number of fruit species. Growing of fruits in homestead gardens is a traditional practice in Nepal since time immemorial. The systematic fruit development program in Nepal was initiated in a planned manner since 1960.

Agriculture Perspective Plan (APP) stressed development activities for commercial production of two major high value fruit crops such as Apple in high hills & Citrus (especially Mandarin) in mid-hills. The emphasis was given on pocket-package development strategy, which must be carried out as a campaign on a participatory basis from the grass-root level to higher ups. The other fruits of the commercial importance such as Mango, Litchi and Banana etc. are also addressed near/along the highway corridors of Terai and inner Terai belts so as to meet the fruit requirements in the country. However, horticulture sector has not gained the level of expected industrialization/commercialization due to various constrains.

Fruit Development Directorate (FDD) was first established as Horticulture Section in 1955, which evolved as Fruit Development Section in 1966, Fruit Development Division in 1990 and the directorate in 2000. FDD is the central body responsible for the development of fruits, coffee, tea and ornamental crops in the country. National Centre for Fruit Development was established in 2018 after restructure under federal system for Fruit Development Directoriate, National Citrus Development Programme and Coffee and Tea Development Section.
