
# Horticulture

## Plant growth hormones

### Abscissic acid

Physiological effects of ABA

1. Induction of dormancy of plant organs
2. Water balance of plants

- Additionally, ABA plays roles as antagonist of other phytohormones.
- Stratification causes reduction in ABA content and so promotes germination (eg. walnut)
- Inhibition of seed germination by ABA may be reversed by GA application.
- ABA induces formation of storage proteins in seed.
- Bud dormancy is frequently correlated with increase in ABA concentration.
- During vernalization, however, ABA content falls.
- ABA regulates water balance (causes stomatal closure and increase in root growth and hydraulic conductivity rise in water deficit).

## Growth stages of cabbage (9 distinguishible)

1. Cotyledons: No true leaves present
2. Seedling: Upto 5 true leaves
3. 6-8 true leaf
4. 9-12 true leaf: Base of stem still visible from above
5. Precupping: Approximately 13-19 leaves. By the end of this stage, the base of stem and the bases of all leaves are concealed when plant is viewed from above. The innermost heart leaves are growing in an upright fashion and are visible without moving any of the surrounding leaves.
6. Cupping: Approximately 20-26 leaves. The innermost heart leaves, which are still growing in an upright fashion, are concealed by the larger, older leaves surrounding them. All visible leaves will later become the frame leaves (leaves not touching the mature head) of the mature plant.
7. Early head formation: Approximately 2.5-4 inch diameter head. The inner heart leaves, now quickly developing as a ball-like structure of overlapping leaves, are concealed by the surrounding larger leaves. These leaves do not press tightly against the developing head and will later unfold to become frame leaves.
8. Head fill: Approximately 3-8 inch diameter head. A firm round head is visible within the wrapper leaves (the 4 outer loose leaves that touch the mature head). The head has not yet fully developed and thus, is not of harvestable size.
9. Mature: Approximately 6-12 inch diameter head. No new visible leaf production will occur after the head has attained maximum hardness and size. The head is ready for harvest and may split if not harvested in time.

## Citrus cultivation

### Principal rootstocks

Today, five types of rootstock predominate in relatively not cool climates where cold or freezing weather is not probable, especially Florida and southern Europe.

Sour orange rootstock: it is the only rootstock that truly is an orange (the Citrus ? aurantium or bitter orange). It is vigorous and highly drought-resistant.

Poncirus trifoliata: it is a close relative of the Citrus genus, sometimes classified as Citrus trifoliata. It is especially resistant to cold, the tristeza virus, and the fungus Phytophthora parasitica (root rot) and grows well in loam soil. Among its disadvantages are its slow growth-it is the slowest growing rootstock-and its poor resistance to heat and drought. It is primarily used in China, Japan, and areas of California with heavy soils.

Swingle citrumelo: it is tolerant of tristeza virus and Phytophthora parasitica and moderately resistant to salt and freezing. This rootstock selection was hybridized from the Duncan grapefruit (Citrus paradisi Macfadyen) and the Poncirus trifoliata (L.) Raf. by Walter Tennyson Swingle in Eustis, Florida, in 1907. It was released by the US Department of Agriculture to nurserymen in 1974.

Troyer citrange and Carrizo citrange: these reasonably vigorous rootstocks are resistant to Phytophthora parasitica, nematodes, and tristeza virus and show good cold tolerance. They also are highly polyembryonic, so growers can obtain multiple plants from a single seed. Citrange, however, does not do well in clay, calcareous or high-pH soils, and is sensitive to salinity. It is not feasible as rootstock for mandarin scions, as it overgrows them by producing branches of its own in competition with the grafted budwood.[3] Citranges are hybrids of the Washington navel orange and the Poncirus trifoliata. The original crosses, made in the early 1900s by the U.S. Department of Agriculture with the intention of producing cold tolerant scion varieties, were later identified as suitable for use as rootstocks. The commercial use of these rootstocks began in Australia in the 1960s. The Troyer variety generally is found in California, while the Carrizo variety is used in Florida.

Cleopatra mandarin: it is tolerant of salinity and soil alkalinity and also suitable for shallow soils. It is used primarily in Spain, Australia, and Florida. Dade County, for example, has 85% calcareous soil, a typical trait of land that has been under water.[4] The Cleopatra mandarin, originated in India and introduced into Florida from Jamaica in the mid-nineteenth century, has been distributed and tested as a rootstock throughout the world. Nowadays, however, it is considered an inferior rootstock because it is sensitive to many diseases, grows slowly, and is difficult to propagate.

## Guava cultivation

- _Psidium guajava_ (Myrtaceae)
- Origin: Tropical america; originated in the warm, lowland tropics
- Contains low energy, low fat and high amounts of vitamin C.
- Commercial trade is smaller in volume compared to other fruits.

### Morphology and Botany

- Photoperiod insensitive
- Can be forced to flower by pruning
- Woody is hardy; resistant to strong winds.
- The rhythmic flushing of vegetative growth and flowering influences when and how pruning is
carried out.
- Guavas are forced to produce year round, and normally the actively growing shoots are tipped once they become too large. This tipping forces regrowth and new flowers to be induced, to keep production high and constant.
- As soon as the harvest is finished, a peripheral pruning of the canopy is performed, in order to force new shoots that will bear flowers.

### Climate and soil

- Temperature lower than $8\ ^\circ C$ is inappropriate for commercial cultivation.
- Moderately tolerant to saline conditions.

### Varieties

- Lucknow-49, Allahabad safeda, Seedless

### Propagation 
- Stem and root cuttings, Air layering and Budding
- Guava trees propagated by grafting have tap roots that provide substantial anchorage. 
- Trees propagated by cuttings have weaker root
systems.

### Planting and orchard establishment

- Spacing with fillers: 3.1 x 7.6 (424 trees/ha)
- Permanent spacing: 6.2 x 7.6 (212 trees/ha), or 4.6 x 7.6 (286 trees/ha)

```{r climac-vs-nonclim}
cl_ncl <- data.frame(climacteric = c("Avocado", "Banana", "Guava", "Mango", "Papaya", "Sapota"), 
                     non_climacteric = c("Litchi", "Mangosteen", "Pineapple", "Mountain apple (Bayer)", NA, NA))

cl_ncl %>% 
  rename_all(function(x)str_replace_all(x, "_", " ") %>% str_to_sentence()) %>% 
  knitr::kable(booktabs = TRUE, caption = "Climacteric and non climacteric fruits commonly cultivated in Nepal") %>% 
  kableExtra::kable_styling(latex_options = "striped", position = "center")
```

### Post-harvest

- High moisture loss occurs
- No dramatic change in taste during storage at higher temperatures (10-15 $\ ^o C$)

Horticulture development and plans and policies

Q. Government plans and policies have given priority on horticulture. However, the potentiality of fruit sector (an important component of horticulture) has not been exploited properly. In this context, discuss the current problems relating to fruit crop development in Nepal.

A. $\longrightarrow$ Although the long term Agriculture Perspective Plan fared well with the set target in horticulture sector (along with roads and community forest) during the implementation period of 1995-2015, overall history of horticulture sector is less than satisfying in Nepal.

The total area, productive area, production and productivity of fruits in the Fiscal Year (FY) 2014/15 are 150387 ha, 110802 ha, 992703 mt and 8.96 mt/ha respectively (Table \ref{tab:horticulture-fruits-veg}). Recent data shows that contribution of horticulture sector amounts to 15% of the Agriculture GDP. Fruit sub-sector contributes 5.24 ??(when) percent in Agricultural GDP. Besides government engaging in active production of fruit seedlings and provisioning gardening tools, many private firms are working on production of quality planting materials. 

Out of total area occupied by fruits 57% are tropical, 26% are citrus and only 17% are temperate fruit species. Similarly, total area covered by fruits is 4.79% of total cultivated area. It was found that the productivity trend is increasing from 2005/06 (9.47mt/ha) to 2011/12 AD (10.17 mt/ha) and decreasing in 2012/13 (9.25 mt/ha) and in 2013/14 (8.77mt/ha). There is huge gap between exported (24812.17 mt) and imported (2,27,00,266 mt) scenario of different fruits which indicates the immense scope of fruit development in Nepal. Out of 470 fruit nurseries 272 were registered and 198 were non registered.

Currently, more than 90% of the citrus seedlings are annually distributed by private fruit nurseries. Technical support services to the fruit growers are also being provided by government farms in their command areas and to the pocket areas of the targeted districts through the extension service.

For the tabulation of cultivation status of fruits and vegetables in Nepal, refer to \ref{tab:fruits-vegetable-cultivation}.

About 3-4 percent of cultivated land is under fruit acreage. This could be observed from the table below showing the annual cultivation trend since 2005-06 AD. There is an increasing trend in area coverage and fruit production but the yield per hectare is low in comparison with other fruit producing countries. The fruit production needs to be increased to meet the increasing demand for domestic consumption and for exports as well. There are some important issues like infrastructure, physical and environmental, agronomic and technical support post harvest losses, problem oriented research and transfer of technology, which need to be addressed to promote fruit development in the country.

There are 45 species belonging to 37 genera of wild edible fruits. Even seasonal fruits harvested from forests can be seen in several local markets. Despite of increasing area and production and huge potential of fruit cultivation in the country, the quantity and value of foreign import of fruits is increasing in recent years due to higher rate of the increasing internal demand in comparison to rate of increasing domestic production and supply. The per capita availability of fruit in the fiscal year 2014/15 was 34 kg/year.

Major problems relating to development of fruit crops in particular, and horticulture sector in general, are as follows:

1. Despite immense potential of hoticultural crops, upto 80% of the cultivated land in Nepal is used to grow cereal crops. Out of remaining 20% cultivated land, most of which are located in more northern latitudes, suitability of fruit farming in those lands is fringe. Most are rocky terrains with steep slopes and young soil. In addition, hilly regions of Nepal suffer a major setback on soil moisture management, which is critical for a fruit orchard. Without a proper buffer against hailstorm and snowfall, substantial amounts of fruit drops and wastage have been noted.

2. Since a long time past, unavailability of inputs like inorganic fertilizers, progagation materials have hindered progressive farming. Rural farmers are forced to opt for low input management which only give marginal yields and limits the productivity improvement. Furthermore, as fruit crops are given an inferior status relative to cereals, farmhold themselves are negligent about timely provisioning of agricultural inputs for fruit orchard.

3. Irrigation management

4. Postharvest handling

5. Marketing and transport

6. APP was based on a narrow viewpoint of technology -- excessively focused on green revolution perspective that is not appropriate for large parts of Nepalese agroecology. Horticulture sector in particular is largely hurt by this conservative view point of technology adoption. This sector has poorly prospered in terms of access to infrastructure and is one of the least industrially benefitting ones. As hills and rural parts of country are infact the major hubs of vegetable and fruit production, policy emphasis on adoption of more capital intensive technology i.e., that which characterizes green revolution, is just a dream chase afterall. On the other hand, Nepal has areas where specialized farming systems, such as organic citrus and apple production system, could flourish. These regions would have sought mostly technical support for quality assurance, policy support for certification, and basic resource support. All of this were lacking in places such as Mustang, Dolpa and Jumla, where organic apple production could have comparative advantage.

7. With the diverse agroecology, we have southern terai regions suited for cultivation of tropical fruits while mid-hills and high-hills towards north being suitable for sub-tropical to warm and cold temperate fruit and nut species. Research and development programs are not found to have exerted impact on all agroecologies, mostly that of Far Western Development Region of Nepal.

8. There still are several fruit and nut crops that have a greater local demand but are not being integrated into mainstream research and crop development process. Fruit nuts such as chestnut ( _Castanea_ spp.), Sapota, Rubus berry are underutilized while some are at best only identified. Thus, in meeting increasing need of human nutrition and feeding urban population, these minor fruits could play an important role.

Policy level and action oriented strategies that have been or could potentially be taken by Government to uplift horticulture sector include improvement in mainly following aspects:

1. The Decentralization strategy of ADS foresees a decentralized structure of Nepalese Agricultural Research, with establishment of new national research institutes (eg, a National Horticultural Research Institute, National Animal Health Research Institute, and National Aquaculture and Fisheries Research Institute), establishment of a National Agricultural Research Fund (NARF) under NARC, and the establishment of research stations in all regions, including the far western region.

2. Nutritional aspect of food security has a prominent role of hoticultural crops. Because fruits and vegetables supply the bulk of fiber, minerals and vitamins, meeting nutritional need of growing urban cummunity is possible only through empowering of horticulture sector.

3. ADS highlights following key strategies in improvement of fruit sub-sector of horticulture:
  - Formulation and implementation of appropriate land use policy which will significantly help area expansion of horticulture and plantation crops.
  - Rejuvenation of old, unproductive, senile plantations through substitution of old varieties with improved high yielding varieties.
  - Development of demand-driven technology on improved varieties, cultural operations, pest management, harvesting, post-harvest handling, marketing and processing; promotion of value chain approach in extension and development.
  - Public private partnership in developing irrigation schemes including the micro-irrigation, collection centers, wholesale markets, cold storage, processing industries, and encouragement and promotion of packaging and grading centers
  - Planning and prioritization of HHVCs' development in the form of projects on the basis of domestic and external demands, economic viability, comparative advantage, employment and income generation, environmental sustainability and use of local resources.
  - Enforcement of laws and regulations to ensure adherence to safety, hygiene and other standards as per SPS through relevant public institutions and local authorities.
  - Strengthening of present organizations responsible for horticultural and HVC development in the country to be made more accountable to carry out the work responsibilities with new vision, or preferably a separate competent horticulture authority established to coordinate inter and intra ministerial, institutional and departmental agencies and programs.
  - Deployment of technically skilled and capable manpower in adequate numbers on the basis of research, extension and training in relation to area of operation and volume of production; and
  - Formulation of long-term commodities development plans with program wise investment plans and encouraging significant participation and investment of private sectors in promotion, processing and marketing of HHVCs.
  - Increased productivity of fruit proposed for ADS implementation period is from 10 to 15 Mt/ha

## Litchi cultivation

**Taxonomy** 

Family: Nepheleae. Family contains about 150 genera and 2000 species. Related species of the family are Rambuton ( _Nephelium lappaceum_) and pulsan ( _N mutabile_). The genus Litchi has 3 species:

  - Sub species chinensis -- Commercial litchi.
  - Sub species philippinensis -- Developed in Philippines. It has long, oval shaped fruit with inedible flesh and long thorn like protuberances which can be used as root stock.
  - Sub species Javenensis bears fruit similar to Chinensis but the aril is thinner.
  
**Varieties**

Mazzafarpur (Bears profusely, fruits are deep orange to pink, pulp is sweet, tough, moderately juicy, good flavored and pulp:seed ratio is high), Deshi (Bearing is moderate and regular, fruit is oval, oblong or oblong conical in shape), Rose scented (Fruits are delicious rosy flavor in the aril. Trees are vigorous. Fruits are mostly oblong or oblong conical in shape, deep rose pink in color, pulp grayish white, and soft and very sweet), Shahi, Dehradun, Culcuttia, Seedless late, Saharanpur, etc.

Origin: Originated from northern China, Northern Vietnam and Malaysia

Uses: Fruit makes excellent canning. The preservation of fruit can be done in honey.

Botany: Long lived, medium to large, brached, round topped, evergreen tree reaching upto 10 m or more height with short stocky trunk. Leaves are compound, alternate, consists of 4-7 oblong leaflets, glossy dark green. Bark is grayish brown and rough. Vegetative growth is rhythmic and occurs in 3-4 recurrent flushes alternating with period of rest. The inflorescence is compound raceme developing both from terminal and axillary buds. Flowers are small, male, pseudo hermaphrodite (functional male) and hermaphrodite. The edible portion is aril. The seed is dark brown in color.

Climate: The litchi is adapted to the warm sub-tropic, cool, dry, frost free winter, and long, hot summer with high rainfall and high humidity. The litchi usually likes low elevation but can be grown upto 800 masl. The optimum temperature requirement for flowering is 16-20 degree C. High temperatures, more than 30 degree C and low relative humidity (less than 60%) during fruit development stage cause the fruit cracking in litchi. The hot winds in summer cause fruit cracking and subsequently damage the pulp (aril). Sometimes it limits the expansion of litchi cultivation. Wet spring, dry summer and light winter are desirable conditions for fruiting in litchi.

The litchi can grow under wide variety of soild including alluvial soil, loams, and heavy clays with rich source of organic matter. Litchi can withstand water for a considerable period, provided the water does not stagnate, but will die after prolonged immersion. It is suggested that new plant should be grown in soil taken from the vicinity of old trees to introduce the mycorrhiza. The optimum pH range is 5.5-6.5


## National fruit development directorate (Institutional development of fruits)

The agro-ecological conditions of Nepal are very much suitable for the successful cultivation of large number of fruit species. Growing of fruits in homestead gardens is a traditional practice in Nepal since time immemorial. The systematic fruit development program in Nepal was initiated in a planned manner since 1960.

Agriculture Perspective Plan (APP) stressed development activities for commercial production of two major high value fruit crops such as Apple in high hills & Citrus (especially Mandarin) in mid-hills. The emphasis was given on pocket-package development strategy, which must be carried out as a campaign on a participatory basis from the grass-root level to higher ups. The other fruits of the commercial importance such as Mango, Litchi and Banana etc. are also addressed near/along the highway corridors of Terai and inner Terai belts so as to meet the fruit requirements in the country. However, horticulture sector has not gained the level of expected industrialization/commercialization due to various constrains.

Fruit Development Directorate (FDD) was first established as Horticulture Section in 1955, which evolved as Fruit Development Section in 1966, Fruit Development Division in 1990 and the directorate in 2000. FDD is the central body responsible for the development of fruits, coffee, tea and ornamental crops in the country. National Centre for Fruit Development was established in 2018 after restructure under federal system for Fruit Development Directoriate, National Citrus Development Programme and Coffee and Tea Development Section.

## Environmental stress factors in horticulture

Q. What are the important climatic factors that are considered as extreme environmental stress? Suggest their possible control measures.

A. $\longrightarrow$ Climate is a combination of aboveground environmental factors -- temperature, moisture, sunlight and air -- and is characteristic of a region. It determines what crops can be cultivated in a given area. However, alike climate belowground environmental factors -- soil temperature, soil moisture, soil physical characteristics, soil chemical properties, organic matter, etc., also play a role in determining success of a crop in a given agricultural region.

Environmental stresses play crucial role in the productivity, survival and reproductive biology of fruit plants as well as crops. Biotic and abiotic stresses, including drought, extreme temperature, scarcity of water, reducing quality of irrigation water and salinity in soil and water are problems which are becoming really acute. Due to their rapid and unpredictable effects, it is becoming very difficult for horticultural scientists and farmers to respond to challenges posed by biotic and biotic stresses.

Some of the climatic parameters and the stress associated with them, influencing crop development and yield are:

- Water and moisture stress: Required in several stages of crop's life cycle -- germination. Much of the water requirement is met from the precipitation (including the rain and the snow) and evaporation. Lack of moisture in the aboveground environment makes the air less humid, thereby increasing its drying power. The rates of plant processes such as transpiration, diffusion, and evaporation are affected directly by the vapor pressure of the air (the part of the total air pressure attributable to the water molecules present in the air). As previously indicated, if air temperature is increased but the amount of water vapor in the air stays the same, the relative humidity of the air decreases. Excessive moisture in the microclimate of plants predisposes them to disease.

- Temperature and temperature stress: Temperature is the intensity of heat energy and very vital in regulating the plant's biological, chemical and physiological processes, mostly by regulating rate of chemical reactions. Several crops show temperature dependent growth periodicity, i.e. some crops tend to flower in higher temperature of summer season while fruits like apple and ornamental bulbs such as daffodil ( _Narcissus_) and hyacinth ( _Hyacinthus_)require a period of chiling temperature to break the dormancy and initiate flowering. Similarly hardiness of a crop is defined for various temperature scales within a range of $0-50 ^\circ C$. Extremely high temperature kills plants ouright or reduces production severely, when such prevalence coinicides with flowering and fruiting periods. actively growing, succulent tissue in plant parts such as flower buds are more susceptible than dormant tissue to cold damage. Frost damage is critical when flower buds start to open. Warm-season crops are more prone to frost damage.

- Light and light associated stress: The role of light in the growth and development of horticultural plants depends on its quality, quantity, and daily duration. olar radiation is electromagnetic in nature. Radiant energy is described by its wavelength and frequency. The shorter the wavelength and higher the frequency, the higher the energy transmitted. Sunrise and sunset patterns differ from one season to another. Incident light angles are wider in summer than in winter and hence the light intensity is higher in summer months. Even the daily light-dark duration varies over an annual cycle. In winter months dark hours are longer and light hours are shorter, and vice versa for summer months. This gives rise to a physiological behavior in some plants called photoperiodism, which is characterized by the flowering time variation of crop plants based on the light-dark period. One one extreme end, plants exposed continually to very low light intesities exibhit tall statures, yellowing of leaves and lean stem, a condition called etiolation. Likewise light quality affects color development and tillering and branching behavior in grass crops.

Four photoperiodic responses in plants are a basis for classifying horticultural plants:

  - Short-day plants (or long-night plants). Short-day plants will not flower under continuous light. They require a photoperiod of less than a certain critical value within a 24-hour daily cycle. For example, strawberry (Fragaria x ananasia) requires 10 hours of light or less and violet (Viola papilionacea) requires 11 hours. Poinsettia (Euphorbia pulcherrima) requires 12.5 hours of daylight and cocklebur (Xanthium strumarium) requires about 16 hours or less of light. When planted in the field, short-day plants flower in early spring or fall.
  
  - Long-day plants (or short-night plants). Long-day plants are plants that flower only when light periods are longer than a certain critical length. These plants flower mainly in summer and include annuals such as henbane (Hyoscyamus niger), which requires more than 10 hours of light, and spinach (Spinacia oleracea), which requires 13 hours of light. Baby’s breath (Gypsophila paniculata) requires 16 hours or more of daylight in order to flower.
  
  - Day-neutral plants. Day-neutral plants are not responsive to photoperiod and flower according to the developmental stage. Plants in this category include tomato, corn, and cucumber.
  
  - Intermediate-day plants. Certain grasses such as Indian grass do not flower if the days are too short or too long. These plants are said to have two critical photoperiods and are categorized as intermediate-day plants.

Some strategies can be duely adopted to minimize the effects of climate associated stress. There are many approaches like:

- Growers should make use of weather forecast,
- Stubble/plastic mulching to check competative weed growth and conserve moisture.
- Drip and sprinkler irrigation to continually maintain required moisture
- Mist irrigation to maintain appropriate level of atmospheric humidity
- Physiological drought resistance of plants through the use of growth retardants, and anti-transpirants
- Construction of specialized structure such as hot beds, hot caps, polyethylene tunnels, shelterbelts, wind machines, etc. based on necessity.
- Proper orchard geometry and nutrient management, and
- Use of rootstocks tolerant to physiological stresses.

Beside climatic factors themselves, geographical processes such as relief, orientation and exposure of the fruit orchard, presence or absence of water bodies and vegetation in the vicinity, etc. vastly influence growth of fruit crops. As an example, Colder and heavier air occurring at higher altitudes moves down and pushes the warmer and lighter air upward. This air convection leaves the higher band of land warmer. This thermal belt is warm and permits the culture of frost-sensitive crops on certain parts of slopes in areas that are normally too cold for growing crops.

## Tissue culture and significance in Nepal

Q. What do you understand by tissue culture? Describe it's role in fulfiling the need of seed requirement to Nepalese farmers.

A. $\longrightarrow$ The branch of biology in which tissues or cells of higher animals and plants are grown artificially in a controlled environment.

As a method of propagation of horticultural crops, from amongst a range of options available to propagate a plant, tissue culture is one of the major biotechnology activity of public as well as private sectors. Protocols have been developed for in vitro propagation of different species (both animals and plants) elsewhere in the world to be used readily or with some modifications.

The first extracted tissue (also called _explant_) changes into a mass of undifferentiated cells called callus, from which differentiation into shoot and root or embryos may later occur. This method of propagation is commercially adopted in herbaceous (e.g., strawberry, banana, gladiolus, tobacco, carnation, and gloxinia) and woody plants (e.g., apple, rose, and rhododendron).

In several asexually reproducing crops, maintenance of their quality and adequate quanity over period has been a challenge. This amounts to a large number of seed material (propagule) import from neighboring countries for instant use of Nepalese farmers. However, this situation of excessive reliance on import of basic input can be remedied if tissue culture of relevant crops were practiced in Nepal itself.

- Mainly focusing on propagule generation of high value and bulky propagule seems to be a profitable venture as fruit and flower crops like those mentioned above can be successfully regenerated in a simply established tissue culture laboratory. 
- A tissue culture laboratory can host seedling at a very small storage space
- Large number of planting materials can be obtained by establishing cultures and subcultures of different varities/cultivars of crops at the same time, on demand. Typically within a couple of weeks thousands of usable seedlings can be recovered from a single explant.
- Tissue culture laboratory can be operated at a relatively cheap cost if basic infrastructure like culture medium, sterilants, working and housing facility could be assured.
- Tissue culture technology using meristem culture can be used to raise virus and pathogen free seedlings, which is not guarenteed by conventional plant part based propagation.
- Development of hybrid cultivars i.e., that of _Brassica napus_ and _B calabrese_, exploits an use case of tissue culture called embryo rescue to obtain viable hybrids.

## River basin/bed farming and guidelines

Q. If a farmer from a river basin area (600 m) comes to you and request for summer season cucumber cultivation technology what would be your advice to him ?

A. $\longrightarrow$ Riverbed farming can be used to increase household income and to improve the food security of landless and land-poor households in the Terai area of Nepal. It is estimated that about 8,000 hectares of riverbed land would be suitable for agricultural cultivation in the Kailali and Kanchanpur Districts in the Western Terai areas of Nepal. After the river water recedes in the post-monsoon season, vegetables are planted in ditches dug into the seasonal sand banks; the crops are harvested before the onset of the next monsoon. 

General guidelines for performing riverbed farming:

- On average, the water table should not be lower than 1 m; when the water table is lower than this, too much labor is required.
- Ditches are up to 1 m deep and 1 m wide. The length depends on how much land is available.
- A row-to-row spacing of 2-3 m (between the ditches) and plant-to-plant spacing of 0.5-1 m is required depending on the crop.
- The ditches are dug in an east-west orientation to maximize the amount of sunshine they receive and to minimize the collection of sand carried by the prevailing winds.
- Farmers may build shelter close to their plots so that they can be close at hand to fend off thieves and wild animals.
- Apply fertilizer: farmyard manure/compost about 12 tonnes; urea about 100 kg; di-ammonium phosphate (DAP) about 120 kg; and potash about 30 kg per ha.
- Plant seeds/seedlings using the appropriate row-to-row (RXR) and plant-to-plant (PXP) distance for at least one crop. Following planting distances are applicable for the crops mentioned herein:

  - 3 m x 1 : Bottle gourd, pumpkin, and water melon and 
  - 2 m x 0.5: Bitter gourd and cucumber
- Following establishment and maintenance inputs are required (per hectare):

  - Labor to prepare plots, irrigation, and to collect mulching materials (165 person days)
  - Equipment: Sprayers, watering can, spades
  - Materials: Polythene bags, sheets, mulching materials
  - Agricultural seeds, chemical fertilizer, farmyard manure, compost, bio-pesticides, micronutrients.

## Potato cultivation 

Potato ranks 5th in terms of area of cultivation (194115 ha) with the average annual production of 2691037 mt. The major potato growing districts are Kavre, Bara, Jhapa, Solukhumbu, Illam, Khotang, Kailali, Bardiya, Bhojpur, Makawanpur, Nuwakot, Rupandehi, etc.; the largest of all is the Kavrepalanchowk. Currently Kavre also hosts the potato superzone.

Potato is advantageous from the health perspective to patients of high blood pressure, anaemia and gastritis as it contains highly digestible starch, protein and fat.

The crop is native to South America. It is propagated by seeds and tubers (stolon). Eyes on potato are buds that develop on stem and leaves. Leaves are compound with opposite leaflets. Stem are green. Roots are shallow and extend only upto 2 feet. Flowers are 1 inch in diameter, corolla white or purple in color. The fruit type is berry but does not commonly develop. Seeds are similar to Mustard but are highly heterozygous.

Varieties

1. Kufri Jyoti: Medium maturing variety (110-120 days in hills and 100-110 days in mid-hills). It is resistant to late blight. Tubers are large, oval with white smooth skin. The crop has dormancy period of 6 weeks and yields on average 20-25 tons/ha.
2. Kufri Sindhuri: Late maturing variety (110-130 days). Moderately resistant to early blight. Tubers are medium sized and round shape with red smooth skin. The tubers are dormant for about 8 weeks and on average their yield is 25-30 tons/ha.
3. Dejire: Early variety (70-90 days). The plant is susceptible to blight. Tubers are medium sized, oval with smooth red color skin.
4. Janak Dev: The cropping period is about 100-120 days with medium maturity duration. It is suited for all area -- Terai, midhills and highhills. The tuber is medium to large sized, red colored. The dormancy holds for 6-8 weeks and tuber yield is attainable upto 23 tons/ha.
5. Khumal Seto-1: Medium term of maturing (100-120 days). Resistant to viral diseases. Round tubers, small and large size white. 6-8 weeks dormancy. Yield: 25 tons/ha.
6. Khumal Rato-2: Medium maturing (100-120 days). Medium sized, round shape, smooth skin and light red color tuber. Average yield: 20-25 tons/ha. Resistant to Y-virus.
7. Kufri Bahar: Medium maturing, large size, round shaped oval, white skin variety. Susceptible to all major diseases.

Indian varieties: Kufri lamlima, Kufri chandramukhi, Local red round, Magnum bonum, up-to-date.

```{r recommended-varieties}
# potato varieties recommended in nepal "publicdata" repo.
```

**Climate**

- Cool season, crop is moderately tolerant to frost. Germination occurs at $8^\circ C$ ideally.
- Day temperature of $30-35^\circ C$ and night temperature of $20-25^\circ C$ is ideal for growth.
- Tuber formation occurs at $18-20^\circ C$, ideally.
- If temperature is >29 at night, no tuber formation occurs.

**Soil and fertilizer**

- Well drained sandy loam to clay loam soil rich in OM is optimal.
- Ideal pH: 5-7
- Alkaline soil leads to development of Scabies disease
- 20-30 ton FYM per hectare
- 220 kg N2, 140 kg P, 140 kg K, 90 kg CaO and 30 kg MgO
- Potato is heavily fertilized as they have a high nutrient requirement and high gross value per unit area
- Full dose of P, K and 1/2 N2 is applied during land preparation and remaining 1/2 N2 is applied during earthing up by side line or point dressing.

```{r planting-time}
dplyr::tribble(~"Planting season", ~"Harvesting month", 
               "Terai (Sept-Oct)", "Feb", 
               "Lower hills (Nov)", "March", 
               "Mid hills (Aug-Nov)", "Nov-Feb", 
               "High hills (Dec-Feb)", "April-May") %>% 
  knitr::kable(caption = "Planting time of potato in different agro-ecological regions", booktabs = TRUE)
```

**Selection of seed**

- Vegetatively propagated crop. Pure, healthy and disease free tubers are selected.
- Seed treatment: Dip tuber in 0.2% solution of Captan or Dithane M-45 for 20-30 minutes to prevent attack of fungal disease.
- To break dormancy, treat whole seed tubers with ethylene chlorohydrins at 3% for 72 hours. Then cut the tubers and dip in 1% solution of thiourea for one hour. Before planting, tubers are dried in shade.
- Cutting the seed: Cut the large sized tubers with sterilized knife. During cutting, keep 2-3 healthy eyes on the cut piece. Average weight of cut piece should be 30-40 g. Small seed tubers are prone to decay.

**Method of planting**

1. Flat bed planting

- Use of light sandy soil
- Tubers are planted in very shallow furrow. This method requres two earthing up

2. Furrow planting

- Tubers are planted in furrows of 5-7 cm deep and adopted in areas having light sandy soil.

3. Planting on ridges

- Ridges of 30 cm height are made at 60 cm spacing and tubers are planted opposite on the ridges.

**Seed rate and spacing**

- Line to line spacing: 60-70 cm
- Plant to plant distance: 25-30 cm
- 40 g seed with 4 buds are preferred
- 2.5 ton seed tubers are required to plant a hectare.

**Irrigation**: Presowing/pre-emergence irrigation and second irrigation after 30 days (Crop water requirement: 300-636 kg).

**Intercultural operation**

- Weed management: At early stage, when plant is small, noe hand weeding and hoeing is helpful to reduce competition with weeds. When the plant is 10-15 cm, small growing weeds are controlled at the time of earthing.
- Chemically, weeding is done by:
  - Pendimethalin (0.5 kg/ha) as pre-planting or a pre-emergence spray
  - Alachlor (1-1.5 kg/ha)
  - Simazine (0.5 kg/ha) is used
- Earhing up is done when the plant is 15-20 cm tall.

**Harvesting**

- Dehaulming (before 7-10 days of harvest) helps avoid insect vectors carrying virus.
- Stop irrigation before 10-15 days of harvest.
- Yield: 25-35 tons/ha

**Disorder**

- Greening: Sunlight exposure causes production of Solanine. Proper earthing up operation maintains tubers are not exposed to sunlight.
- Black or hollow heart: When stored at high temperature ($35-40^\circ C$) the inner tissue breaks down and becomes dark grey or black which is due to lack of O2, accumulation of CO2 and high temperature. Control: Close spacing, removal of foliage and avoidance of overfertilization; Proper ventillation in storage and transport; Store tubers in cold storage at $2.2-3.3^\circ C$

Q. Describe various types of storage systems used in Nepal to store potato tubers.

A. $\longrightarrow$ By Samita!

Q. Define True Potato Seed (TPS). Do you see its scope in supplementing the requirement of potato seeds in Nepal ? Is so, how can it be improved ?

A. $\longrightarrow$ Potatoes are most often propagated from vegetative tissues, either whole tubers or cut pieces of tubers. Tubers used for propagation are typically called seed tubers, seed potatoes, or seed pieces. Potato plant can be produced from botanical or sexual seeds, but these "true" potato seeds (TPS) are seldom used, because they are genotypically variable and give rise to plants with traits thay may be completely different from those of the parent plants.

A single plant may produce 50-100 berries and single berry contains about 150-200 seeds. TPS is mainly used to develop new potato cultivars and now gaining popularity to raise the commercial crops. About 100-150 g of seed is required to raise the crop in 1 hectare of land. 

_Scope_: 

- Non-availability of good quality seed tuber, 
- High seed cost, 
- Virus infilteration in seed tubers causing degeneration of seed stocks 
- Problem of long distance transport of tuber seeds from seed producing areas 
  
All these constraints have led to the development of TPS technology of crop production. Further benefits TPS technology are:

- Unlike seed tuber, stocking density of TPS seed much more, hence from the same parcel of land, more of TPS can be recovered than seed tuber.
- An additional benefit with the use of TPS is process to obtaining application ready seed is substantially shortened as several post-harvest processess such as curing and tuber treatment can be avoided.

```{r}
readr::read_csv("./data/potato_seed_crop_production_specification.csv", skip = 1) %>% 
  pivot_longer(cols = starts_with(c("Recommended")), 
               names_to = "Region", 
               values_to = "Time of year") %>% 
  mutate(`Region` = str_replace_all(`Region`, "Recommended time_", "")) %>%
  dplyr::select(-1) %>%
  dplyr::select(`Crop`:`Production tons per ropani`, `Region`, `Time of year`, `Remarks`) %>%
  rename_all(function(x)str_replace_all(x, "_", " ")) %>%
  knitr::kable(caption = "Potato production technology for various purposes", booktabs = TRUE) %>% 
  kableExtra::kable_styling(font_size = 8) %>% 
  kableExtra::column_spec(1:13, width = c(rep("4.8em", 10), "5.5em", "5.5em","11em")) %>% 
  kableExtra::landscape(margin = "1.2cm")
```


Q. What is PBS ? Differentiate PBS and TPS. How is PBS is produced in Potato ?

$\longrightarrow$ Pre-basic seed (PBS) potatoes are disease free potato minitubers produced by transplanting pathogen free in vitro potato plantlets under protected condition in aphid-proof glasshouse and/or screen house.

Potato crop variety is very susceptible to degeneration as it is vegetatively propagated, and tends to accumulate virus load over generation of cropping. Therefore, routine maintenace of seed should be carried out to ensure variety stands to the promised useful features. Following set of steps are usually carried out for production of PBS:

- Virus elimination
  - Tubers are allowed to sprout for 2-3 weeks under thermal treatment at  $37^\circ$
  - Shoot tips excised from the sprouts and washed in detergent water
  - Under laminar airflow chamber, shoot surface are sterilized (treated with 70% Ethanol for 30 seconds, washed with sterile distilled water and then sterilized with 2% sodium hypochlorite solution for five minutes and again washed at least three times with sterile distilled water). 
  - The apical meristem (with one or two leaf premordia, about 0.2-0.3 mm in diameter) are excised from the shoot tip by viewing under a stereoscopic microscope, and placed on top of a filter paper bridge on a liquid MS medium (Murashige and Skoog 1962) supplemented with $0.5 mg l^{-1}$ IAA, $0.4 mg l^{-1}$ Kinetin and $0.1 mg l^{-1}$ $GA 3$ (Mellor and Stace-Smith 1977). 
  - The meristem are then cultured in an incubation room under $20\pm 2^\circ C$ with proper illumination (2000 lux) and 16 h photoperiod. 
  - After few weeks of culturing, when green pigmentation and stem and leaf tissues are observable, it is transferred to a solid MS medium for proper rooting and shooting. 
  - These plantlets can be multiplied into several clones by nodal cuttings.
- Virus testing: ELISA or related techniques to screen for virus free clones. Assayed clones are further validated for them being virus free with field tests of standing crops in greenhouse.
- Germplasm maintenance: Virus free clones could be maintained with subsequent subcultures after testing in germplasm laboratory.
- Rapid propagation: Maintained mother plantlets can be propagated by single nodal cutting on modified MS solid media with supplement. To enhance differentiation and production of apical part including leaves, culture should be incubated in growth chamber with culture conditions of 16 h photoperiod, 2000 lux light intensity and $20 \pm 2^\circ C$ temperature.
- PBS production under controlled conditions: Four to six weeks old in-vitro plantlets are transplanted into sterile mixture of 2:1 sand and soil substrate under aphid-proof glasshouse. Special cultivation techniques assure the quality of propagules. They include:
  - 20 x 10 cm spacing
  - UV sterilized water irrigation and proper irrigation schedule
  - Precise dose fertilizer application
  - Earthing up
  - Plant protection

Since the technique used in producing PBS focuses on maintenance of asceptic environment and provides minimum nutrients for early stage plant development, the tuber produce are expected to be low. First generation culture will generally produce small sized tubers. However, a large number of seedlings could be recovered from a tissue cultured plantlets.

Since 1990, National Potato Research Program has been producing about 200,000 pre-basic seeds annually. So far, PBS of 19 different recommended and released potato cultivars has been produced. There cultivars have been tested for major potato viruses like PLRV, PVS, PVX, PVY, PVA and PVM.

**Raising of crops from transplanted seedlings**

TPS is sown in nursery beds to produce seedlings. The best substrate for nursery beds is a mixture of 1:1 FYM and soil. Soak the seeds in water from 24 hours before sowing. TPS germinates well when night temperature drops to about $20^\circ C$ in the plains and shade is provided over nursery bed areas during day time. In plains the best sowing time is October.

The seedlings raised in nursery can either be transplanted in field to produce aware/seed crop or alternatively allowed to tuberize in the nursery beds to produce small seedlings tubers which can be used for field planting next year. The method of raising the crop using small seedlings tubers produce in nursery bed is successful in all the potato growing regions of the country. Seedlings will be ready for transplanting in field in about 20-25 days of germination.

The seedlings are planted at the distance of: 60 x 20, 60 x 15, 60 x 10 and 60 x 5. It is desirable that larger tubers are planted with sparser spacing.


## Tomato production

- Lycopersicon esculentum (2n = 24). It is a warm season crop.
- Presence of lycopene causes development of red color in tomato. Orange color coloration is caused by Prolycopene.
- The fruit is also known as poor man's apple.
- Tomato is native to Tropical America.

**Botany**

- Herbaceous annual; Sexually propagated; Bisexual flower.
- _L pimpinellifolium_ is wild tomato.
- Self pollinated, tap rooted with berry fruit type.
- The crop occurs in determinate and indeterminate variants. Determinate types are dwarf because the plant terminates into a flower bud (self toping/self pruning type). Indeterminate types are tall because the terminal portion of stem is growing and flower is produced at every third internode. Determinate types can be harvested 2-3 times but indeterminate types can be harvested much longer.

**Varieties**

Determinate: These are generally early type in bearing. Roma, NS-2535, NS-815, NS-719, NCL-1 etc.
Indeterminate: They are mostly late bearing type. These varieties are suitable for home consumption or local market. Srijana, Nabin-2000, Marina, etc.

- Srijana variety was developed from the cross of Female HRD-1 and Male HRD-17. It is recommeded for Midhills and Terai region. The plants are 4.5-5 m tall.

## Monthly crop calendar

- Baisakh

  - Intercropping of spring season rice, irrigate it and fertilizer application
  - In order to improve soil fertility status of rice field where previously wheat was grown, raise green manuring crop Dhaincha 50 kg per hectare.
  - For establishment of evergreen fruit and flower orchard, perform layout, pit preparation, fertilizer and soil mixing.
  - Okhle-1 and Kavre kodo-1 nursery bed establishment in upper hills.
  - Interculture, earthing up fertilization and irrigation of maize in hills.
  - Application of botanical pesticides (Neem, Timur, Bojho, Artemisia, Thyme, Mustard oil) for control of storage pests.
  - Planting of rainy season maize in terai and inner terai.
  - Planting of forage grass like Teosinte, sorghum, sudan grass, bajra and dinanath grass.
  - Since this time of the year is conducive to spead of Brucellosis and _Bhyakute rog_, case should be taken to prevent the infection from occuring in cattle shed.

- Jestha

  - Nursery bed raising of rice
  - Transplanting fingermillet as mixed crop in standing maize field in mid hills
  - Desuckering of fruit orchard
  - Spraying of fungicides in evergreen flowers and fruits
  - Broadcast seed of perinnial grassess in terai and mid hills
  - Control of gundhi bug in spring season rice
  - Planting of Hill, Seti, Ransom and Lumle-1 variety of soybean in hils.
  - Citrus orchard sanitation by application of bordeaux paste in Mandarin and Sweet orange.
  
## Offseason onion production

Owing to the lack of Cold storage facilities and due to reigning high temperature during the months of Ashoj and Kartik, the onion harvested during months of Bhadra and Ashoj months is unlikely to meet demands for later Mangsir and later months. This shortcoming can conveniently be avoided if offseason cultivation of Onion could be done. When onion is grown for fresh leaves and bulbs during the Kartik and Mangsir, such seasonal demand meeting produce is often called offseason. Adding to the availability feature, when sold around these months, produce fetches premium price in the market.

Off season onion farming could be done in regions ranging from terai to mid-hills (upto elevation of 1600 masl), however profitable production is being done on lower hills (at 500-1600 masl). Offseason onion farming could basically be done in following ways:

  1. Seed to set and set to bulb
  2. Seed to bulb

Vegetative onion can be obtained directly from seedling cultivation, which is generally established in bulk in nursery bed and disseminated among growers. For simplicity, here is discussed the method of fresh leaf onion and bulb onion production:

  1. Selection of land: Onions will grow satisfactorily on soils with a pH of 6.0-6.8. The ratio of N:P:K applied during seedbed preparation is 1:2:2, although some bulb producers increase the nitrogen ratio according to the soil status. Work by Ahmed (1982) showed that N, P and K applications equivalent to 150 kg/ha produced the largest bulbs and highest total bulb yield at the end of the first year, and that supplementary nitrogen application not exceeding 100 kg/ha in the second year applied during anthesis enhanced seed quality. The higher potassium levels during bulb production were carried over to the second year and also enhanced seed quality.
  Additionally, the land should not permit water stagnation and should drain as early as possible to avoid fungal diseases.
  2. Selection of off season cultivar: The best cultivar for raising crop directly from seed is the Agrifound Dark Red. But Nasik-53 is favored when sets are to be raised to obtain bulbs.
  3. Seedling raising: Seedling is ready for transplanting within 30-40 day of sowing.
  
```{r offseason-onion-seed-bed}
tribble(~"Region", ~"Seedling raising time", ~"Transplanting time", 
        "Lower hills", "Early Ashar", "Early Shrawan", 
        "Mid hills", "Jestha", "Ashar") %>% 
  knitr::kable(booktabs = TRUE, caption = "Timing of seed seed preparation and seedling transplanting for offseason onion production.") %>% 
  kableExtra::kable_styling(font_size = 10)
```
  
  4. Seed requirement: 500 g per ropani
  5. Raising of seedling bed:
    - 10 msq of seed bed is enough to raise seedlings for 1 ropani of land. Each meter square of bed optimally receives 50 g.
    - For healthy seedling raising, bed should be 1 m wide and be raised 30 cm above the ground.
    - Application of 5 kg: 10 g: 10 g: 10 g FYM:Urea:DAP:MoP per 1 msq seed bed.
    - If the seed bed is moisture deficit, irrigate it 2-3 days prior to seed placement.
    - Seed placement in rows of 8-10 cm apart in the bed.
    - Seed should be thoroughly covered with FYM or fine soil after sowing. Bed should be mulched and be moistened to avoid dessication of seeds underneath.
    - Healthy seedlings can be obtained if they are primed in sugar+jathropa solution (5:5 g in 20 ml water) and dried.
    - To prevent seed borned diseases, like seedling rot, spray solution of 2.5 g Dithane-M45, 1 g Bavistin and 1 g Nuvan in 1 ltr water after 1-2 days in seed bed.
    - Seed bed should always be kept moist.
    - Apply Micronutrients for healthy and vigorous seedlings.
    - Seed bed shoudld have provision of drainage to let excess of water.
    - If early bed is to be constructed in rainy season, bed should be protected from spattering rain by sheltering with polythene.
  6. Seedling transplanting
    - Seedlings become ready for transplantation 30-40 DAS. Only healthy seedlings should be selected for planting. This is fulfiled by roguing out unhealthy ones.
    - Prior to seedling uprooting, bed should be softened with water.
    - Based on moisture reign of the land where seedling will be futher grown, provision of irrigation or drainage should be made. Land should be parcelized (into beds) in order to ease operation.
    - R-R: 20 cm x P-P: 10-15 cm

## Offseason tomato production

## Offseason cucumber production

## Offseason production of other vegetable crops

## What do you understand by tissue culture. Describe its principles and techniques suitable to horticultural crops.

## Breeder seed production of Onion

## Describe the process of seed formation with illustration.

## What is fruit? Discuss various causes of poor fruit setting in Mango and suggest suitable measures for overcoming them.

## Discuss in brief post-harvest deterioration of fruits and vegetables and suggest measures to minimize postharvest losses.

Every year horticultural crops suffer significant lossess attributed to storage, warehouse, processing unit, transporation and other post-harvest operations mainly due to insect-pest damage or mechanical injury. An estimate of average post harvest lossess in some of the commonly marketed crops (including cereal crops) is provided in Table \ref{tab:post-harvest-loss-table}.

```{r post-harvest-loss-table}
tribble(
  ~"Crop", ~"Loss (percentage)",
  "Commercial vegetable (average)", "25", 
  "Potato (including seed potato)", "32", 
  "Cereals", "20 (15-20)", 
  "Cauliflower and cabbage", "10-15", 
  "Fruits", "20", 
  "Ginger", "5-10", 
  "Tomato", "22"
  ) %>% 
  knitr::kable(booktabs = TRUE, format = "latex", caption = "Post harvest loss (estimates) in common crops")
```

