
# Agronomy

## Denotified varieties

- For a more complete listing, refer to publicdata repository for the [list of denotified varieties](https://raw.githubusercontent.com/DeependraD/publicdata/master/crop_varieties/denotified_varieties.csv).

- Paddy:
  - IR-8
  - IR-20
  - IR-22
  - IR-24
  - Jaya
  - Khajura-2
  - Barkhe-2
  - Mallika
  - Candina
  - Laxmi
  - Durga
  - Parwanipur-1

- Wheat:
  - NL-30
  - Nepal 251
  - Lerma Roho 64
  - Lumbini
  - S 331
  - Siddartha
  - Pitik 62
  - Vaskar
  - Binayak
  - HD-1982
  - Annapurna 2
  - Kalyan Sona
  - Triveni

- Maize:
  - Janaki Seto
  - Kakani Yellow
  - Rampur Yellow
  - Makalu 2
  - Sarlahi Seto
  - Hetauda Composite
  - Rampur 1

- Chickpea:
  - Trishul
  
- Soybean:
  - Hill

- Mustard:
  - Type-9

## Varietal description of some common rice varieties

```{r rice-varieties}
read_csv("./data/rice_variety_description.csv", skip = 2) %>% 
  knitr::kable(booktabs = TRUE, caption = "Varietal description of recently released rice varieties", escape = FALSE) %>% 
  kableExtra::kable_styling(font_size = 8, latex_options = "striped") %>% 
  kableExtra::column_spec(column = 3:11, c(rep("5em", times = 7), "8em", "12em")) %>% 
  kableExtra::landscape(margin = "1.5cm")
```

```{r maize-varieties}
read_csv("./data/maize_variety_description.csv", skip = 2) %>% 
  knitr::kable(booktabs = TRUE, caption = "Varietal description of recently released maize varieties", escape = FALSE) %>% 
  kableExtra::kable_styling(font_size = 8, latex_options = "striped") %>% 
  kableExtra::column_spec(column = 3:12, rep("5em", times = 10)) %>% 
  kableExtra::landscape(margin = "1.5cm")
```

## Seed technology

1. Breeder seed
  - Genetic purity of the variety: 99.9%
  - Golden yellow tag (as of new seed certification guidelines, it is ?brown) with breeder's signature.
2. Foundation seed
  - Not available for sales/purchase at farmer's level
  - SQCC or specialized organization having granted permit from the SQCC performs the field inspection and certification
  - White colored tag with black letters
2. Certified seed
  - In self pollinated crops two generations may be grown, however only one generation is allowed in cross pollinated species
  - CS-I and CS-II (Blue letters in white tag and blue border, and green letters in white tag with green borders, respectively)
3. Improved seed
  - In cross pollinated and highly CP species, direct production of improved seed from foundation seed is also practiced (Maize, vegetable, pigeonpea).
  - Production takes place in farmer's field or by seed producing organizations.
  - Yellow colored tag

- Normally, genetic purity of FS is 99.5 and that of CS is 99%.
- For longer storage in cereals, vegetables, legumes and oilseeds, and orthodox seeds, optimum moisture content: 12% or less.
- In recalcitrant seeds, moisture content is best kept close to 20%.
- In the moisture regime between 5-14%, for every drop in 1% storage life of the seed is doubled (Jems F Herington)
- Forbidden diseases of crops:
  - Cauliflower: Black rot, Alternaria leaf spot

```{r number-of-seed-per-ten-g}
require(tidyverse)

tribble(~"Crop", ~"Number of seeds", 
        "Carrot", 8280, 
        "Chinese cabbage", 6480,
        "Turnip", 5400, 
        "Tomato", 4140,
        "Onion", 3420, 
        "Pea", 3263,
        "Cabbage, cauliflower", 3240,
        "Brinjal", 2340,
        "Capsicum", 1620,
        "Radish", 1440,
        "Chukandar", 576,
        "Muskmelon", 470,
        "Cucumber", 400,
        "Okra", 180,
        "Watermelon", 75,
        "Luffa gourd", 40
        ) %>% 
  knitr::kable(booktabs = TRUE, caption = "Number of seeds per 10 gram of vegetable crops") %>% 
  kableExtra::kable_styling(latex_options = "striped")
```

## Seed certification system

- Seed act was first enacted in 2045 BS. Since then it's first amendment came into being in 2064.

Two step certification/authentication system is practiced in Nepal:

1. Seed certification: Official body responsible for certifying seed in the SQCC. Three classes of seeds are identified: Foundation, Certified-I, Certified-II and Improved under this system.

Following activities are carried out in various stages of a crop are crucial for inspection by a seed inspector:

  - Standing crop: By licenced inspector, for field inspection check for crop's source and seed class, purity, isolation distance, seed affecting diseases, weed and type of varieties.
  - At harvest: Threshing, processing, transporation and storage.
  - After processing: To verify that minimun quality meets, insect damage (In legume: and maize: 1% and other crops: 0.5%), seed moisture, weed and off type seeds, germination percentage
  - Seed is sampled and taken to lab
  - Lab should return results within 30 days of receiving a sample.
  - If suspected, sample may be subjected to grow out test by the tester.
  - If farmer isn't convinced of the results of the seed testing, s/he may file for recertification.

- Tagging storage container with certification tag, with signature of certifying inspector.
- If the seed is to be stored for more than 1 season, the storage sample must be re-certified for quality standards.
- There are some steps to be followed in prolonging the validity of certification. 

- There are two main types of certification:
  1. Minimum standard certification:
    - Location and land requirement fulfilment certification
      - In terai, a minimum of 1 hectares is required for cereal and cash crops.
      - In terai, for vegetables, a minimum of 0.25 hectares is required.
      - In hills, for cereal, a minimum of 0.25 hectare is the necessary criterion.
      - In hills, for vegetables, a minimum of 0.1 ha land should be cultivated with seed under the question.
      - Maximum distance between plots of same certification lot should not exceed 50 m.
  2. Location and locality of specification
  3. Source of seed verification
  4. Timing of inspector assignment

## Grain legumes

- "Legume" comes from "legere" meaning "to gather".
- Soybean cultivation recorded in China 3000 to 2000 BC
- Globallyh 70.6 million hectares (production: 61.5 million mt) of legumes are planted
- In Nepal, 0.32 million hectares producing 0.26 million mt was grown (when ?)
- Consumption trend of legumes in Nepal is 9 kg per person per year. However, the recommended annual intake is 36 kg.
- Legumes are an important part of natural soil fertility maintainance system -- fixing atmospheric nitrogen.
  - Cowpea: 73-354 kg
  - Chickpea: 103 kg
  - Lentil: 88-114 kg
  - Pigeon pea: 168-280 kg
  - Broad bean: 45-552 kg
- Energy expended per gram of seed (i.e. gram photosynthate):
  - 0.74-0.96 in cereals
  - 2.03-2.09 in pulses/oilseeds

# Breeding

## Why and how can a plant breeder increase genetic variability in crop plants ? Explain.

Breeding objectives are realized by combining the favorable genes of different sources (parents). That is, one cannot breed, for example, disease resistance if the gene conferring resistance of disease of interest does not occur in the base population.

As genetic variation is heritable and results in permanent change in phenotype (trait) of organism, it is the variability of interest of breeder, since breeder can predictably control it through selection. Moreover, genetic variation is detectable at molecular as well as gross morphological level but the variation due to environment cannot be. It is also possible for the source of variation to be channeled through carefully designed mating system.

Primarily there are 3 types of sources of genetic variability:
  1. Gene recombination
  2. Modification of chromosome number
  3. Mutation
  
Plant breeder can increase genetic variability by use of following tools:

1. Hybridization: Crossing un-identical plants to transfer genes or achieve recombination (creating non-parental type through physical exchange of chromosomal segments). In hybridization, effects of segregation and indepenent assortment in crosses between individual heterozygous for given number of gene pairs is evident. If we take 2 alleles for any 'n' of genes, following indicators of variation can be deduced:

  - Different number of gametes produced by $F_1$ heterozygotes: $2^n$
  - Number of combinations of $F_1 \times F_1$ gametes (perfect population size): $4^n$
  - Different kinds of genotypes in $F_2$ = $3^n$
  - Different kinds of $F_2$ genetypes that are homozygous: $2^n$
  - Different kinds of $F_2$ genotypes that are heterozyous: $3^n-2^n$
  - Different types of phenotypes in $F_2$ (complete dominance): $2^n$
  

2. Wide crossing: Crossing of distantly related individuals for desirable gene introgression. Generally wild relatives of native agricultural crops contain such genes for biotic and abiotic stress tolerance. For example in tomato breeding, program resistance to Oidium neolycopersici (powdery mildew of tomato) monogenic resistance was found in _S peruvianum_ accession, _S habrochaites_ accession and polygenic resistance in _S neorikii_ accessions.

Both of these breeding methods depend upon the number of allele variants per loci. In a gene system involving two alleles, 3 genetic variants (genotypes) are possible, while for a triallelic system 6 individuals may arise from hybridization.

3. Polyploidization: It is accomplished through chromosome doubling, which can occur naturally (through modification of chromosome number as a result of hybridization or abnormalities in nuclear division process) or can be induced using spindle disrupter Colchicine. 

4. Mutagenesis: Also known as ultimate source of biological variation, mutation may arise spontaneously in nature of as a result of errors in cellular process such as DNA replication (or duplication) and by chromosomal abberations (deletion, duplication, inversion and translocation). Mutations may too be induced artificially by using mutagenic agents: irradiation and chemicals. Commonly occuring mutations are those for dwarfing and nutritional quality. However, most mutations being deleterious in nature are selected against. Although, recessive mutations may remain hidden until much later, when they are expressed in homozygous combination.

## What are the different factors that affect the amount of natural crossing or selfing ? Explain about the mechanisms that enforce self and cross pollination in field crops.

Species differ in the degrees to which they self pollinate or cross pollinate. Mechanisms promoting self pollination or autogamy are:

1. Cliestomgamy (flowers fail to open), or flowers open only after pollination (chasmogamy)
2. Close proximity of anthers to stigma

Mechanisms that prevent autogamy:

1. Self-incompatibility: In spite of viable and normal pollen and ovule development pollen from a flower is non receptive on the stigma of the same flower, hence incapable of setting seed. This mechanism of reproduction control is conditioned by a single locus "S" with multiple alleles. There are two main type of self incompatibility systems:

  - Heteromorphic: Differences in lengths of stamen and style (heterostyly).
  - Homomorphic:
    1. Gametophytic incompatibility: Ability of the pollen to function is determined by its own genotype and not the plant that produces it.
    2. Sporophytic incompatibility: Incompatibility character of pollen are determined by plant (sporophyte) that produces it.
    
    The self incompatibility mechanism promotes heterozygosity.
    
2. Male sterility: Anther or pollen are non-functional.
  1. Genetic (nuclear/genic): Barley, cotton, soybean, tomato, potato, etc. The system is under control of "ms" gene.
  2. Cytoplasmic male sterility: Controlled by mitochondrial gene. Transmitted only through maternal line.
  3. Cytoplasmic genetic male sterility: Presence of "Rf" gene (nuclear) that overcomes cytoplasmic MS.

3. Dichogamy: Maturing of pistils and stamens of a flower at differnt times. Reduces intensity of self pollination and fertilization. It includes protandry and protogyny.

Factors that affect cross and self pollination are:

1. Environment: 
  - Pollinators: Wind, water, insect, animals (Anemophily, hydrophily, entomophily, zoophily)
  - Temperature: High temperature or low temperature may temporarily overcome SI
  - Light: It induces pollen inviable.
2. Flower appearance: Shape of petal, color radiance, size of flower, olfactory cues, chemical signaling, etc.



