
# Agronomy

## Released, registered and denotified varieties overview

```{r overview-seeds}
readxl::read_xlsx("./data/varietal_notification_status.xlsx") %>% 
  select(SN:Denotified) %>% 
  knitr::kable(caption = "Overview of released, registered and denotified varieties of crops", booktabs = TRUE, longtable = TRUE, col.names = c("SN", "Crop", "Released", "Registration", "Total number (as of 2074-12-12)", "Total number (as of 2075-08-24)", "Denotified")) %>% 
  kableExtra::kable_styling(latex_options = c("striped", "HOLD_position"), font_size = 8) %>% 
  kableExtra::column_spec(column = 2:7, width = c("6em", "5em", "5em", "8em", "8em", "6em"))
```

## Recent addition to notified varieties

## Recently denotified varieties

- For a more complete listing, refer to publicdata repository for the [list of denotified varieties](https://raw.githubusercontent.com/DeependraD/publicdata/master/crop_varieties/denotified_varieties.csv).
- Also the [Seed Quality Control Centre](http://sqcc.gov.np/en) is the authentic government body that routinely updates by enlisting crop variety database and publishing seed related information.

## Varietal description of some common rice varieties

```{r rice-varieties}
read_csv("./data/rice_variety_description.csv", skip = 2) %>% 
  knitr::kable(booktabs = TRUE, caption = "Varietal description of recently released rice varieties", escape = FALSE) %>% 
  kableExtra::kable_styling(font_size = 8, latex_options = "striped") %>% 
  kableExtra::column_spec(column = 3:11, c(rep("5em", times = 7), "8em", "12em")) %>% 
  kableExtra::landscape(margin = "1.5cm")
```

```{r maize-varieties}
read_csv("./data/maize_variety_description.csv", skip = 2) %>% 
  knitr::kable(booktabs = TRUE, caption = "Varietal description of recently released maize varieties", escape = FALSE) %>% 
  kableExtra::kable_styling(font_size = 8, latex_options = "striped") %>% 
  kableExtra::column_spec(column = 3:12, rep("5em", times = 10)) %>% 
  kableExtra::landscape(margin = "1.5cm")
```

## Crop production technology

```{r crop-cultivation-cereal-commercial-remarks}
read_csv("./data/crop_production_technology_cereal_commercial.csv") %>% 
  knitr::kable(booktabs = TRUE, caption = "Summary of production remarks on cultivated crops (Cereal and commercial)", longtable = TRUE) %>% 
  kableExtra::kable_styling(font_size = 5, position = "center", latex_options = "striped") %>% 
  kableExtra::column_spec(column = 1:12, width = c("12em", "12em", rep("10em", 10))) %>% 
  kableExtra::landscape(margin = "1cm")
```

```{r crop-cultivation-legume-oilseed-remarks}
oil_joined <- tibble(`Production remark` = c("Time of sowing", "Seed rate (kg/ha)", 
                                             "Spacing (cm)", "NPK (kg/ha)", "Seed yield (tons/ha)"), 
                     `Mustard/rapeseed/sarsoon` = c("Mid September - Mid October", 
                                                    "5-8", "30 x 7.5 - 10", "60:40:20", "1.5-2.0"), 
                     Groundnut = c("Mid June - 1st Week of July", 
                                   "Spreading type: 60-80; Bunchy type: 80-100", "45-60 x 10-15; 30 x 10 - 15", 
                                   "20-40:50-60:20-40", "1.5-2.0"), 
                     Sesamum = c("First week of June - First week of July", 
                                 "2.5-5.5", "30-45 x 15-22", "30:60:30", "0.2-0.6"), 
                     Sunflower = c("Rainy, 1-15 July; Spring, 1-15 March", 
                                   "8-10", "60 x 20", "60-80:40-60:20-40", "1.5"), 
                     Niger = c("Mid July (June-August)", 
                               "6-8", "30 x 15", "20:20:10", "0.3-0.4"), 
                     Linseed = c("First week of October - Mid November", 
                                 "30-35", "22-30 x 5-6", "50:40:30", "1.0-1.9"))

leg_oilseed <- read_csv("./data/crop_production_technology_legume_oilseed_other.csv")

leg_oilseed %>%
  left_join(oil_joined) %>%
  rows_upsert(oil_joined %>% # update the values in x from y and and also insert missing rows from y.
                left_join(leg_oilseed)) %>%
  select(`Production remark`, Lentil, Blackgram, Moongbean, Pigeonpea, Cowpea, Chickpea, Pea, Soybean) %>% 
  knitr::kable(booktabs = TRUE, caption = "Summary of production remarks on cultivated crops (Legume)", longtable = TRUE) %>% 
  kableExtra::kable_styling(font_size = 6, latex_options = "striped", position = "center") %>% 
  kableExtra::column_spec(column = 1:9, width = c("14em", rep("11em", 8))) %>% 
  kableExtra::landscape(margin = "1cm")

leg_oilseed %>%
  left_join(oil_joined) %>%
  rows_upsert(oil_joined %>% # update the values in x from y and and also insert missing rows from y.
                left_join(leg_oilseed)) %>%
  select(`Production remark`, `Mustard/rapeseed/sarsoon`, Cotton, Sunflower, Groundnut, Linseed, Sesamum, Niger) %>% 
  knitr::kable(booktabs = TRUE, caption = "Summary of production remarks on cultivated crops (Oilseeds)", longtable = TRUE) %>% 
  kableExtra::kable_styling(font_size = 6, latex_options = "striped", position = "center") %>% 
  kableExtra::column_spec(column = 1:8, width = c("14em", rep("12em", 7))) %>% 
  kableExtra::landscape(margin = "1cm")
  
read_csv("./data/crop_production_technology_legume2.csv") %>% 
  knitr::kable(booktabs = TRUE, longtable = TRUE) %>% 
  kableExtra::kable_styling(font_size = 6, latex_options = "striped", position = "center") %>% 
  kableExtra::column_spec(column = 1:9, width = c("10em", rep("12em", 8))) %>% 
  kableExtra::landscape(margin = "1cm")
```

### Optimum pH range for some crops

```{r crop-cultivation-ph}
readr::read_csv("./data/crop_production_ph_optimum.csv") %>% 
  mutate(half_sep = ifelse(row_number() < ceiling(nrow(.)/2), 1, 2)) %>% 
  group_by(half_sep) %>% 
  mutate(row_id = 1:n()) %>% 
  ungroup() %>% 
  pivot_wider(names_from = half_sep, names_glue = "{.value}_{half_sep}", values_from = c(pH, crop), values_fill = list(value = 0)) %>% 
  select(-row_id) %>% 
  relocate(pH_1, .after = crop_1) %>% 
  relocate(pH_2, .after = crop_2) %>% 
  knitr::kable(booktabs = TRUE, col.names = rep(c("Crop", "pH"), times = 2), longtable = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped", font_size = 8)
```


### Notable varieties of some crops

```{r crop-notable-varieties}
readr::read_csv("./data/crop_notable_varieties.csv") %>% 
  rename_all(function(x)str_to_sentence(str_replace_all(x, "_", " "))) %>% 
  select(-1) %>% 
  knitr::kable(booktabs = TRUE, longtable = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped", font_size = 8) %>% 
  kableExtra::collapse_rows(1) %>% 
  kableExtra::column_spec(column = 1:3, width = c("8em", "10em", "24em"))
```


### Potential crop varieties for import substitution and export promotion (Seed vision, 2013-25)

```{r}
tribble(~"Crop", ~"Varieties", ~"Trade potential", 
        "Rice", "Khumal-4, Shambha mahasuri sub-1, Jethobudho, Sunaulosugandha", "Import substitution", 
        "Lentil", "Shishir, simrik", "Export promotion", 
        "Tomatoes", "Srijana hybrid, Lapsigede", "Import substitution", 
        "Radishes", "Minnow early, 40 days, Tokineshe, Pyuthane, Rocky-45 (Bangladeshi)", "Import substitution and export promotion", 
        "Broccoli", "Green sprouting and calabrese", "Import substitution", 
        "Peas", "Arkel, Sikkim, Azad", "Import substitution and export promotion", 
        "Cowpea", "Prakash, akash, khumal tane, some selected lines of Chinese origin", "Import substitution", 
        "Onion", "Red creole, Agrifound dark red", "Import substiution and export promotion", 
        "Garlic", "Chinese and local", "Import substitution", 
        "Carrot", "Nantes and newkuroda", "Import substitution and export promotion", 
        "Egg plant", "Nurki, Sarlahi green, Pokhara lurke", "Import substitution", 
        "Cucumber", "Bhaktapur local and kushle", "Import substitution and export promotion", 
        "Squash", "Gray zuchhini", "Import substitution", 
        "Bean", "Trishuli, four season, contender, some lines of Chinese origin", "Export promotion", 
        "Mung bean", "Pusabaisakhi, Kalyan, Pratiksha", "Import substitution", 
        "Pigeon pea", "Bageshwori and Rampur arhar-1", "Import substitution", 
        "Potato", "Khumal laxmi, Kufri jyoti, Khumal rato", "Import substitution"
        ) %>% 
  knitr::kable(booktabs = TRUE, longtable = TRUE) %>% 
  kableExtra::kable_styling(position = "center") %>% 
  kableExtra::column_spec(column = 1:3, c("6em", "18em", "12em"))

```


## Seed technology

1. Breeder seed
  - Genetic purity of the variety: 99.9%
  - Golden yellow tag (as of new seed certification guidelines, it is ?brown) with breeder's signature.
2. Foundation seed
  - Not available for sales/purchase at farmer's level
  - SQCC or specialized organization having granted permit from the SQCC performs the field inspection and certification
  - White colored tag with black letters
2. Certified seed
  - In self pollinated crops two generations may be grown, however only one generation is allowed in cross pollinated species
  - CS-I and CS-II (Blue letters in white tag and blue border, and green letters in white tag with green borders, respectively)
3. Improved seed
  - In cross pollinated and highly CP species, direct production of improved seed from foundation seed is also practiced (Maize, vegetable, pigeonpea).
  - Production takes place in farmer's field or by seed producing organizations.
  - Yellow colored tag

- Normally, genetic purity of FS is 99.5 and that of CS is 99%.
- For longer storage in cereals, vegetables, legumes and oilseeds, and orthodox seeds, optimum moisture content: 12% or less.
- In recalcitrant seeds, moisture content is best kept close to 20%.
- In the moisture regime between 5-14%, for every drop in 1% storage life of the seed is doubled (Jems F Herington)
- Forbidden diseases of crops:
  - Cauliflower: Black rot, Alternaria leaf spot

```{r number-of-seed-per-ten-g}
tibble::tribble(~"Crop", ~"Number of seeds", 
        "Carrot", 8280, 
        "Chinese cabbage", 6480,
        "Turnip", 5400, 
        "Tomato", 4140,
        "Onion", 3420, 
        "Pea", 3263,
        "Cabbage, cauliflower", 3240,
        "Brinjal", 2340,
        "Capsicum", 1620,
        "Radish", 1440,
        "Chukandar", 576,
        "Muskmelon", 470,
        "Cucumber", 400,
        "Okra", 180,
        "Watermelon", 75,
        "Luffa gourd", 40
        ) %>% 
  knitr::kable(booktabs = TRUE, caption = "Number of seeds per 10 gram of vegetable crops", longtable = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped")
```

### Seed certification system

- Seed act was first enacted in 2045 BS. Since then it's first amendment came into being in 2064.
- Organized seed production program intitiated in Nepal in 1973 (Wheat)
- 90\% of the seeds are supplied through informal sector (SQCC, 2011).
- Seed replacement rate of cereals: 9.75\% (SQCC, 2011)
- SRR for vegetable crops was as high as 66% in 2009 (NSV)
- Seed vision, 2013-2025 published by SQCC, Nepal has envisioned improvement in SRR during the implementation period, with
  - Cereals: 25% SRR
  - Vegetables: 90% SRR

```{r seed-replacement-rate-crops}
tribble(~"Crop", ~"SRR (percent, 2009)", ~"SRR (percent, 2010)", 
        "Paddy", 9.14, 11.38,
        "Maize", 6.54, 9.03,
        "Wheat", 9.4, 11.0,
        "Millet", 1.97, 2.35,
        "Lentil", 3.08, 3.26,
        "Oilseeds", 2.62, 2.86,
        "Vegetables", 66.3, 68.3,
        "Potato", 5, 5.5,
        "Average (cereals)", 8.78, 10.68) %>% 
  knitr::kable(booktabs = TRUE, caption = "Seed replacement rate of different crops (SQCC, 2010).") %>% 
  kableExtra::kable_styling(latex_options = "striped") %>% 
  kableExtra::add_footnote(label = c("Source: SQCC, 2011 as cited in Seed vision, 2013-2025."))
```


- Establishment of agriculture input corporation: 1974
- In 1960, High yielding variety of wheat Lerma-52 was released.
- In 1962, Seed testing laboratory was established under Agronomy Division, Khumaltar.
- In 1964, Central Seed Lab of Nepal was designated membership with International Seed Testing Association (ISTA).
- In 1966, Seed testing laboratory moved to Agri-Botany Division.
- In 1966, Rice variety CH-45 was released.
- In 1974, AIC established under Corporation Act, 1965.
- In 1980, Seed production and input storated project (SPISP, USAID) launched
- In 1982, Seed technology and improvement programme
- In 1984, Central seed science and technology division established
- In 1988, Seed act enacted
- In 1997, National seed policy approved
- In 1998, Seed production guidelines published
- In 2001, Seed quality control center established
- In 2002, National seed company limited established
- In 2010, Seed science and technology division revived as central disciplinary division of seed in NARC
- In 2013, Promulgation of seed regulation, 2013
- In 2004, First hybrid of maize -- Gaurav, was officially released.
- In 2010, First hybrid of tomato -- Srijana, was registered
- Upto 2010, 231 hybrids were registered for cultivation in Nepal.


Two step certification/authentication system is practiced in Nepal:

1. Seed certification: Official body responsible for certifying seed in the SQCC. Three classes of seeds are identified: Foundation, Certified-I, Certified-II and Improved under this system.

Following activities are carried out in various stages of a crop are crucial for inspection by a seed inspector:

  - Standing crop: By licenced inspector, for field inspection check for crop's source and seed class, purity, isolation distance, seed affecting diseases, weed and type of varieties.
  - At harvest: Threshing, processing, transporation and storage.
  - After processing: To verify that minimun quality meets, insect damage (In legume: and maize: 1% and other crops: 0.5%), seed moisture, weed and off type seeds, germination percentage
  - Seed is sampled and taken to lab
  - Lab should return results within 30 days of receiving a sample.
  - If suspected, sample may be subjected to grow out test by the tester.
  - If farmer isn't convinced of the results of the seed testing, s/he may file for recertification.

- Tagging storage container with certification tag, with signature of certifying inspector.
- If the seed is to be stored for more than 1 season, the storage sample must be re-certified for quality standards.
- There are some steps to be followed in prolonging the validity of certification. 

- There are two main types of certification:
  1. Minimum standard certification:
    - Location and land requirement fulfilment certification
      - In terai, a minimum of 1 hectares is required for cereal and cash crops.
      - In terai, for vegetables, a minimum of 0.25 hectares is required.
      - In hills, for cereal, a minimum of 0.25 hectare is the necessary criterion.
      - In hills, for vegetables, a minimum of 0.1 ha land should be cultivated with seed under the question.
      - Maximum distance between plots of same certification lot should not exceed 50 m.
  2. Location and locality of specification
  3. Source of seed verification
  4. Timing of inspector assignment

\blandscape

### Certification standards

```{r certification-standards-field}
readxl::read_xlsx("./data/seed_certification_requirements.xlsx", sheet = "field_standards") %>% 
  rename_all(function(x)str_replace_all(x, "_", " ")) %>% 
  knitr::kable(caption = "Field standards of standing crop for certification", booktabs = TRUE, longtable = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped", font_size = 8) %>% 
  kableExtra::column_spec(column = 1:9, width = c("8em", rep("6em", times = 8)))
```

```{r certification-standards-seed}
readxl::read_xlsx("./data/seed_certification_requirements.xlsx", sheet = "seed_standards") %>%
  select(1:9) %>% 
  rename_all(function(x)str_replace_all(x, "_", " ")) %>% 
  knitr::kable(caption = "Minimum seed standards for certification", booktabs = TRUE, longtable = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped") %>% 
  kableExtra::column_spec(column = 1:9, width = c("4em", rep("5em", times = 14)))

# kableExtra::landscape(margin = c("1cm", "1cm", "1cm", "1cm"))

readxl::read_xlsx("./data/seed_certification_requirements.xlsx", sheet = "seed_standards") %>% 
  select(1,10:15) %>% 
  rename_all(function(x)str_replace_all(x, "_", " ")) %>% 
  knitr::kable(caption = "Minimum seed standards for certification", booktabs = TRUE, longtable = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped") %>% 
  kableExtra::column_spec(column = 1:9, width = c("4em", rep("6em", times = 14)))
```

\elandscape

### Seed production: Constraints

1. Climatic and edhapic factors:
- Subtropical type climate range ($26^\circ 22'' - 30^\circ 27''$ N).
- Mean daily temperature: $27-28^\circ C$ during summer and $17-18^\circ C$ during winter.
- 750-2250 mm annual rainfall, large interannual variability, frequent drought, unimodal rainfall distribution
- fragile ecosystem, remoteness, extensive soil mining

2. Weeds
3. Market and trade
- Indian seed market
- Price instability

4. Agronomical
- Seed treatment, seed availability, knowledge of good production package, irrigation facility, fertilizer insufficiency/unavailability, haphazard use of chemical fertilizer.

5. Socio-economic
- Government policy does not prioritize local production.
- Devalued currency and price instability
- Tragedy of free trade policy
- Infrastructure drudgery
- Poor technology adoption
- Low extension workers: farmer ratio
- Government budget allocation and planning for agriculture showing gaps

6. Pests and diseases
- Upto 80% loss due downy mildew in Maize.

#### Insects and diseases of major crops

```{r insect-pests-major}
tibble::tribble(
  ~"Crop", ~"Major insects", ~"Major disease",
  "Rice", "Rice bug, rice hispa, yellow stem borer, stripped stem borer, rice gall midge, mole cricket, plant hopper", "Bacterial blight (Xanthomonas oryzae), Blast (Pyricularia oryzae), False smut (Ustilaginoides virens), Brown leaf spot (Helminthosporium oryzae)", 
  "Wheat", "Armyworm, cutworm, shoot fly, stem borer, termites", "Leaf spots (Helminthosporium spp), rust, leaf streak (Xanthomonas spp), loose smut", 
  "Maize", "Stalk borer, shoot fly, cutworm, jassid, armyworm", "Rust, leaf blight (Helminthosporium maydis), smut (Specealothica reliana)", 
  "Barley", "Green bug, corn sawfly, fruitfly, wheat bulb fly", "Barley yellow dwarf virus, powdery mildew (Erysiphe graminis sp. hordii), Net blotch (Helminthosporium sativum)"
) %>% 
  knitr::kable(caption = "Major disease and insect pests of major cultivated crops in Nepal", booktabs = TRUE) %>% 
  kableExtra::kable_styling(latex_options = c("striped", "HOLD_position")) %>% 
  kableExtra::column_spec(column = 2:3, width = c("12em", "12em"))
```

7. Other factors
- Poor education and low literacy rate
- Customary use of conventional tools; lack of mechanization
- Improper finance
- Storage facilities

### Scope of seed production in Nepal

- Climate variability
- More demand and less supply of seed and seed materials
- Rapid expansion of road and urban infrastructure
- Extensive reach of irrigation projects
- Development of seed enterprises
- Employment opportunities
- Seed exporting opportunities
- Increaseing productivity of cereal and legume crops

### Expect results of Seed vision 2013-25

1. 10 lakh farm families will have access to quality seeds upon demand prior to planting season. The country will be self-sufficient in food crop seeds.
2. 88 mt breeder, 2,978mt foundation and 92,527 mt improved seeds will be produced through formal system by 2025. 
3. 750mt Nepal produced high quality seeds will have improved access to export by 2025.
4. Seed replacement rates will be increased up to 25 percent for cereals and over 90 percent for vegetables.
5. Good quality seeds will be available in market through quality assurance with genuine label.
6. Four hundred twenty-three open pollinated varieties and 60 hybrids will be released by 2025.
7. 40,000 seed samples will be tested and analyzed annually by seed laboratories. 
8. Seed production and marketing will be done through structured and efficient seed system.
9. Yield of rice and vegetable crops will be increased up to 3.8 mt/ha and 19 mt/ha respectively. 
10. Enhanced participation of private sector will increase availability of quality seeds in the market.Private sector will establish or strengthen four big seed companies.
11. 293 highly skilled seed specialists will be developed in private and public sectors.
12. Farmers' rights will be protected and breeders will get incentive for developing better varieties.
13. Seed import and export regulations will be harmonized in line with WTO and SAARC.
14. All the stakeholders will be accountable to farmers and be responsible participants of Nepalese seed system.
15. 255 thousand people will get additional full time employment upon the implementation of Seed Vision 2013 -2025.
16. Seed Vision will contribute to food security. Edible food availability by 2025 will reach 8 million mt,worth around 200billion rupees at current price. 17. Agro-based industries will have adequate raw materials from increased production.
18. Nepal's seed sector will be able to share its experience and knowledge to other countries.

## Seed production

- Sampling of the variety to btian nucleus seed
- Not more than 15 new varieties in one crop at station in one year

- Seeding breeder's stock: 1.2 ha for Wheat, 3 ha for Rice (transplanted)
- Roughing should be done before flowering.
- Purity: 99.9% for breeder's seed.

### Oilseed crops

For seed and crop production practice of oilseed crops, refer to Table \ref{tab:crop-cultivation-legume-oilseed-remarks}.

### Sugarcane

- Area cultivated: 58,101 ha
- Production: 2354412 tons
- Productivity: 40.52 tons/ha

**Isolation**

- 5 m all around from sugarcane field.
- Time of planting;
  - Mainly October and Spring planting
  - Seed crop harvest: 8-10 months in tropics, 10-12 months in sub-tropics

**Field inspection**

- 1st inspection: 2 months of planting
- 2nd inspection: 2 months before harvest

**Harvest and seed yield**: 40-60 tons/ha

### Cotton

- Area cultivated: 100 ha
- Production: 59 tons (2008/09)
- Productivity: 0.59 tons/ha
- Variety: Tamcot-SP-37 released upto 2010

**Isolation requirements**

- Foundation seed: 50 m
- Certified seed: 25 m
- It is commonly self pollinated crop, but natural cross-pollination occurs.
  - 10-50% in _G. hirsutum_
  - 1-2% in _G. arborum_
  - 1-5% in _G. herbaceum_
  - 5-10% in _G. barbadense_
  
- Time of sowing: May-June
- Seed rate: 15-25 kg/ha (American type), 12-16 kg/ha (Desi type)
- Fertilizer: 100:50:50 kg NPK kg/ha
- Inspection: Before anthesis and at full flowering
- Seed yield: 300-600 kg/ha

### Radish/carrot

1. Seed to seed method

- This method is also called in-site seed production technique because seed is produced in crop production plot itself. Although good production can be obtained from this method, high quality seed cannot be ensured. This method allows for quick harvest in expense of purity of the seed produced.

2. Root to seed method

- Well developed, uniform tap root are harvested and transferred to seed production plots.
- A small portion of above ground part and some of the below ground portion is trimmed off while maintaining only 8 cm long root stubs.
- The stubs are then strored for 8-10 days in a cold storage facility maintaining $0^\circ$ C and 90-95% relative humidity.
- The root stubs are called stalks and the process of preparing those is called stalking.
- Stalks are planted in seed production plots at a spacing of 15 x 15 cm R-R x P-P distance.

### Onion

1. Seed to seed method (similar to radish/carrot)
2. Bulb to seed method

- Bulbs are dug/pulled out when 75\% of the top portion of onion topples.
- Those harvested bulbs are cured for 10-15 days in natural condition
- 2-3 cm upper portion of the bulb is removed.
- Desirable bulb size: 50-80 kg
- The bulb is then planted at 10-15 x 10-15 cm distance in seed production plot.
- With this method, high quality seed material with desirable purity can be maintained.

### Cabbage

1. Seed to seed method (similar to radish/carrot)
2. Head to seed method

- Once healthy fully mature upper portion of it is given a cross shaped out (i.e., cut to center from all directions).
- Heads are stored in shade for 5-7 days
- Thenafter, heads are planted at 15-20 x 15-20 cm distance on seed production plot
- Seed production of quantity 100-200 seed/ha could realized.

### Cauliflower

1. Seed to seed method (similar to radish/carrot)
2. Curd to seed method

Central part of the curd is cut and removed giving it a both 3-4 cm deep and wide cut -- a practice called scooping. Scooping helps in the earlier emergence of the flower stalk. On 5-7 days of scooping the curds are transferred to seed production plot and planted at a spacing of 15-20 x 15-20 cm.
- Seed production of quantity 200-600 kg/ha could be realized in this way.




## Grain legumes

- "Legume" comes from "legere" meaning "to gather".
- Soybean cultivation recorded in China 3000 to 2000 BC
- Globallyh 70.6 million hectares (production: 61.5 million mt) of legumes are planted
- In Nepal, 0.32 million hectares producing 0.26 million mt was grown (when ?)
- Consumption trend of legumes in Nepal is 9 kg per person per year. However, the recommended annual intake is 36 kg.
- Legumes are an important part of natural soil fertility maintainance system -- fixing atmospheric nitrogen.
  - Cowpea: 73-354 kg
  - Chickpea: 103 kg
  - Lentil: 88-114 kg
  - Pigeon pea: 168-280 kg
  - Broad bean: 45-552 kg
- Energy expended per gram of seed (i.e. gram photosynthate):
  - 0.74-0.96 in cereals
  - 2.03-2.09 in pulses/oilseeds

## Biological nitrogen fixation

```{r biological-nitrogen-fixation}
tribble(~"crop", ~"nitrogen_kg_ha", 
        "Pea", "52-77", 
        "Cowpea", "73-345", 
        "Soybean", "60-168", 
        "Bean", "40-70", 
        "Chickpea", "103", 
        "Lentil", "88-114", 
        "Pigeonpea", "168-280", 
        "Forage soybean", "168", 
        "Stylo", "200", 
        "Lucerne", "290", 
        "White clover", "128", 
        "Vetch", "110", 
        "Centro", "125-400", 
        "Horsegram", "45-52") %>% 
  rename_all(function(x)str_replace_all(x, "_", " ") %>% str_to_sentence()) %>% 
  knitr::kable(booktabs = TRUE, caption = "Amount of nitrogen (in kilogram) fixed by different leguminous crops", longtable = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped")
```


# Breeding

## Center of origin and estimated time of cultivation

```{r}
tribble(
  ~"Crop",	~"Length of time domesticated (years)",	~"Possible region of origin",
  "Maize, Zea mays",	"7000",	"Mexico, Central America",
  "Rice, Oryza sativa",	"4500",	"Thailand, Southern China",
  "Wheat, Triticum spp.",	"8500",	"Syria, Jordan, Israel, Iraq",
  "Barley, Hordeum vulgare",	"9000",	"Syria, Jordan, Israel, Iraq",
  "Sorghum, Sorghum bicolor",	"8000",	"Equatorial Africa",
  "Soybean, Glycine max",	"2000",	"North China",
  "Oil palm, Elaeis guineensis",	"9000", "Central Africa",
  "Coconut palm, Cocos nucifera",	"100", "Southern Asia",
  "Rapeseed, Brassica napus",	"500",	"Mediterranean Europe",
  "Sunflower, Helianthus annus",	"3000",	"Western United States",
  "Beans, Phaseolus spp",	"7000",	"Centra America, Mexico",
  "Lentil, Lens culinaris",	"7000",	"Syria, Jordan, Israel, Iraq",
  "Peas, Pisum sativum",	"9000",	"Syria, Jordan, Israel, Iraq",
  "Potato, Solanum tuberosum",	"7000",	"Peru",
  "Cassava, Manihot esculenta",	"5000",	"Brazil, Mexico",
  "Sweet potato, Ipomoea batatas",	"6000",	"South Central America",
  "Sugar beet, Beta vulgaris",	"300",	"Mediterranean Europe",
  "Tomato, Lycopersicum esculentum",	"3000",	"Western South America",
  "Cabbage, Brassica oleracea",	"3000",	"Mediterranean Europe",
  "Onion, Allium spp.",	"4500",	"Iran, Afganistan, Pakistan",
  "Orange, Citrus sinensis",	"9000",	"South-east Asia",
  "Apple, Malus spp.",	"3000",	"Asia Minor, Central Asia",
  "Grape, Vitis spp.",	"7000",	"Eastern Asia",
  "Banana, Musa acuminata, M. balbisiana",	"4500",	"South-east Asia",
  "Cotton, Gossypium spp.",	"4500",	"Centra America, Brazil",
  "Coffee, Coffea spp.",	"500",	"West Ethiopia",
  "Rubber, Hevea brasiliensis",	"200",	"Brazil, Bolivia, Paraguay",
  "Alfalfa, Medicago sativa",	"4000",	"Iran, Northern Pakistan",
) %>% 
  knitr::kable(booktabs = TRUE, longtable = TRUE) %>% 
  kableExtra::kable_styling(position = "center", latex_options = "striped") %>% 
  kableExtra::column_spec(column = 1:3, width = c("8em", "12em", "12em"))

```


## Modes of reproduction

### Gametophyte

- Fertilization requires Nicking to occur.
- Perfect, complete, monoecious/dioecious flower types.
- Male gametophyte:
  - Each anther is bilobed (has two pollen sacs)
  - Pollen sac (microsporangium) is filled with number of large sized cells called sporogenous cells or microsporocytes.
  - Microsporocytes produce microspore mother cells by mitotic division.
  - Microspore mother cell or pollen mother cells undergo meiotic division to produce microspores at the end of second division.
  - Anther contains exine, intine, generative nucleus and vegetative nucleus.
- Female gametophyte:
  - Within ovary, starts as small outgrowth (tiny knob) within nucellus.
  - Nucellus is multilayered main body of ovule which encloses one or two protective layers called integuments, except for a small pore at one end. Integuments are outer and inner, outer layer eventually becomes testa.
  - Developing ovule attached to placenta by funiculus, scar called hilium forms after detachment.
  
**Agents of pollination**

- Wind/anemophily: Maize, oats, coconut palm, cannabis, etc.
- Water/hydrophily: Hydrilla, zostera marina, etc.
- Insect/entomophily: Mustard
- Bird/ornithophily: Red silk cotton, bottle brush, etc.
- Bat pollination/chiropterophily: Adansonia

**Endosperm development**

## Miscellaneous

Q. Why and how can a plant breeder increase genetic variability in crop plants ? Explain.

$\longrightarrow$ Breeding objectives are realized by combining the favorable genes of different sources (parents). That is, one cannot breed, for example, disease resistance if the gene conferring resistance of disease of interest does not occur in the base population.

As genetic variation is heritable and results in permanent change in phenotype (trait) of organism, it is the variability of interest of breeder, since breeder can predictably control it through selection. Moreover, genetic variation is detectable at molecular as well as gross morphological level but the variation due to environment cannot be. It is also possible for the source of variation to be channeled through carefully designed mating system.

Primarily there are 3 types of sources of genetic variability:
  1. Gene recombination
  2. Modification of chromosome number
  3. Mutation
  
Plant breeder can increase genetic variability by use of following tools:

1. Hybridization: Crossing un-identical plants to transfer genes or achieve recombination (creating non-parental type through physical exchange of chromosomal segments). In hybridization, effects of segregation and indepenent assortment in crosses between individual heterozygous for given number of gene pairs is evident. If we take 2 alleles for any 'n' of genes, following indicators of variation can be deduced:

  - Different number of gametes produced by $F_1$ heterozygotes: $2^n$
  - Number of combinations of $F_1 \times F_1$ gametes (perfect population size): $4^n$
  - Different kinds of genotypes in $F_2$ = $3^n$
  - Different kinds of $F_2$ genetypes that are homozygous: $2^n$
  - Different kinds of $F_2$ genotypes that are heterozyous: $3^n-2^n$
  - Different types of phenotypes in $F_2$ (complete dominance): $2^n$
  

2. Wide crossing: Crossing of distantly related individuals for desirable gene introgression. Generally wild relatives of native agricultural crops contain such genes for biotic and abiotic stress tolerance. For example in tomato breeding, program resistance to Oidium neolycopersici (powdery mildew of tomato) monogenic resistance was found in _S peruvianum_ accession, _S habrochaites_ accession and polygenic resistance in _S neorikii_ accessions.

Both of these breeding methods depend upon the number of allele variants per loci. In a gene system involving two alleles, 3 genetic variants (genotypes) are possible, while for a triallelic system 6 individuals may arise from hybridization.

3. Polyploidization: It is accomplished through chromosome doubling, which can occur naturally (through modification of chromosome number as a result of hybridization or abnormalities in nuclear division process) or can be induced using spindle disrupter Colchicine. 

4. Mutagenesis: Also known as ultimate source of biological variation, mutation may arise spontaneously in nature of as a result of errors in cellular process such as DNA replication (or duplication) and by chromosomal abberations (deletion, duplication, inversion and translocation). Mutations may too be induced artificially by using mutagenic agents: irradiation and chemicals. Commonly occuring mutations are those for dwarfing and nutritional quality. However, most mutations being deleterious in nature are selected against. Although, recessive mutations may remain hidden until much later, when they are expressed in homozygous combination.

Q. What are the different factors that affect the amount of natural crossing or selfing ? Explain about the mechanisms that enforce self and cross pollination in field crops.

$\longrightarrow$ Species differ in the degrees to which they self pollinate or cross pollinate. Mechanisms promoting self pollination or autogamy are:

1. Cliestomgamy (flowers fail to open), or flowers open only after pollination (chasmogamy)
2. Close proximity of anthers to stigma

Mechanisms that prevent autogamy:

1. Self-incompatibility: In spite of viable and normal pollen and ovule development pollen from a flower is non receptive on the stigma of the same flower, hence incapable of setting seed. This mechanism of reproduction control is conditioned by a single locus "S" with multiple alleles. There are two main type of self incompatibility systems:

  - Heteromorphic: Differences in lengths of stamen and style (heterostyly).
  - Homomorphic:
    1. Gametophytic incompatibility: Ability of the pollen to function is determined by its own genotype and not the plant that produces it.
    2. Sporophytic incompatibility: Incompatibility character of pollen are determined by plant (sporophyte) that produces it.
    
    The self incompatibility mechanism promotes heterozygosity.
    
2. Male sterility: Anther or pollen are non-functional.
  1. Genetic (nuclear/genic): Barley, cotton, soybean, tomato, potato, etc. The system is under control of "ms" gene.
  2. Cytoplasmic male sterility: Controlled by mitochondrial gene. Transmitted only through maternal line.
  3. Cytoplasmic genetic male sterility: Presence of "Rf" gene (nuclear) that overcomes cytoplasmic MS.

3. Dichogamy: Maturing of pistils and stamens of a flower at differnt times. Reduces intensity of self pollination and fertilization. It includes protandry and protogyny.

Factors that affect cross and self pollination are:

1. Environment: 
  - Pollinators: Wind, water, insect, animals (Anemophily, hydrophily, entomophily, zoophily)
  - Temperature: High temperature or low temperature may temporarily overcome SI
  - Light: It induces pollen inviable.
2. Flower appearance: Shape of petal, color radiance, size of flower, olfactory cues, chemical signaling, etc.


## Gene action and methods of breeding

- Self pollinated species: When additive gene action predominates in a self-pollinated species, breeders should consider using selection methods such as pure line selection, mass selection, progeny selection and hybridization. However, when non-additive gene action predominates, effective methods of breeding are the exploitation of heterosis in breeding hybrid cultivars.

- Cross pollinated species: When additive gene action predominates in a cross-pollinated species, recurrent selection may be used to achieve general combining ability (GCA). Specific breeding products to pursue include synthetic varieties and composites. In the case of non-additive gene action, heterosis breeding just like in self pollinated species is recommended for breeding hybrid cultivars. Alternatively, breeders may consider recurrent selection for specific combining ability (SCA) for poulation improvement. Where both additive and non-additive gene action occur together, reciprocal recurrent selection may be used for population improvement.
