
# Agronomy

## Denotified varieties

- For a more complete listing, refer to publicdata repository for the [list of denotified varieties](https://raw.githubusercontent.com/DeependraD/publicdata/master/crop_varieties/denotified_varieties.csv).

- Paddy:
  - IR-8
  - IR-20
  - IR-22
  - IR-24
  - Jaya
  - Khajura-2
  - Barkhe-2
  - Mallika
  - Candina
  - Laxmi
  - Durga
  - Parwanipur-1

- Wheat:
  - NL-30
  - Nepal 251
  - Lerma Roho 64
  - Lumbini
  - S 331
  - Siddartha
  - Pitik 62
  - Vaskar
  - Binayak
  - HD-1982
  - Annapurna 2
  - Kalyan Sona
  - Triveni

- Maize:
  - Janaki Seto
  - Kakani Yellow
  - Rampur Yellow
  - Makalu 2
  - Sarlahi Seto
  - Hetauda Composite
  - Rampur 1

- Chickpea:
  - Trishul
  
- Soybean:
  - Hill

- Mustard:
  - Type-9

## Varietal description of some common rice varieties

```{r rice-varieties}
read_csv("./data/rice_variety_description.csv", skip = 2) %>% 
  knitr::kable(booktabs = TRUE, caption = "Varietal description of recently released rice varieties", escape = FALSE) %>% 
  kableExtra::kable_styling(font_size = 8, latex_options = "striped") %>% 
  kableExtra::column_spec(column = 3:11, c(rep("5em", times = 7), "8em", "12em")) %>% 
  kableExtra::landscape(margin = "1.5cm")
```

```{r maize-varieties}
read_csv("./data/maize_variety_description.csv", skip = 2) %>% 
  knitr::kable(booktabs = TRUE, caption = "Varietal description of recently released maize varieties", escape = FALSE) %>% 
  kableExtra::kable_styling(font_size = 8, latex_options = "striped") %>% 
  kableExtra::column_spec(column = 3:12, rep("5em", times = 10)) %>% 
  kableExtra::landscape(margin = "1.5cm")
```

## Seed technology

1. Breeder seed
  - Genetic purity of the variety: 99.9%
  - Golden yellow tag (as of new seed certification guidelines, it is ?brown) with breeder's signature.
2. Foundation seed
  - Not available for sales/purchase at farmer's level
  - SQCC or specialized organization having granted permit from the SQCC performs the field inspection and certification
  - White colored tag with black letters
2. Certified seed
  - In self pollinated crops two generations may be grown, however only one generation is allowed in cross pollinated species
  - CS-I and CS-II (Blue letters in white tag and blue border, and green letters in white tag with green borders, respectively)
3. Improved seed
  - In cross pollinated and highly CP species, direct production of improved seed from foundation seed is also practiced (Maize, vegetable, pigeonpea).
  - Production takes place in farmer's field or by seed producing organizations.
  - Yellow colored tag

- Normally, genetic purity of FS is 99.5 and that of CS is 99%.
- For longer storage in cereals, vegetables, legumes and oilseeds, and orthodox seeds, optimum moisture content: 12% or less.
- In recalcitrant seeds, moisture content is best kept close to 20%.
- In the moisture regime between 5-14%, for every drop in 1% storage life of the seed is doubled (Jems F Herington)
- Forbidden diseases of crops:
  - Cauliflower: Black rot, Alternaria leaf spot

```{r number-of-seed-per-ten-g}
require(tidyverse)

tribble(~"Crop", ~"Number of seeds", 
        "Carrot", 8280, 
        "Chinese cabbage", 6480,
        "Turnip", 5400, 
        "Tomato", 4140,
        "Onion", 3420, 
        "Pea", 3263,
        "Cabbage, cauliflower", 3240,
        "Brinjal", 2340,
        "Capsicum", 1620,
        "Radish", 1440,
        "Chukandar", 576,
        "Muskmelon", 470,
        "Cucumber", 400,
        "Okra", 180,
        "Watermelon", 75,
        "Luffa gourd", 40
        ) %>% 
  knitr::kable(booktabs = TRUE, caption = "Number of seeds per 10 gram of vegetable crops") %>% 
  kableExtra::kable_styling(latex_options = "striped")
```

## Seed certification system

- Seed act was first enacted in 2045 BS. Since then it's first amendment came into being in 2064.

Two step certification/authentication system is practiced in Nepal:

1. Seed certification: Official body responsible for certifying seed in the SQCC. Three classes of seeds are identified: Foundation, Certified-I, Certified-II and Improved under this system.

Following activities are carried out in various stages of a crop are crucial for inspection by a seed inspector:

  - Standing crop: By licenced inspector, for field inspection check for crop's source and seed class, purity, isolation distance, seed affecting diseases, weed and type of varieties.
  - At harvest: Threshing, processing, transporation and storage.
  - After processing: To verify that minimun quality meets, insect damage (In legume: and maize: 1% and other crops: 0.5%), seed moisture, weed and off type seeds, germination percentage
  - Seed is sampled and taken to lab
  - Lab should return results within 30 days of receiving a sample.
  - If suspected, sample may be subjected to grow out test by the tester.
  - If farmer isn't convinced of the results of the seed testing, s/he may file for recertification.

- Tagging storage container with certification tag, with signature of certifying inspector.
- If the seed is to be stored for more than 1 season, the storage sample must be re-certified for quality standards.
- There are some steps to be followed in prolonging the validity of certification. 

- There are two main types of certification:
  1. Minimum standard certification:
    - Location and land requirement fulfilment certification
      - In terai, a minimum of 1 hectares is required for cereal and cash crops.
      - In terai, for vegetables, a minimum of 0.25 hectares is required.
      - In hills, for cereal, a minimum of 0.25 hectare is the necessary criterion.
      - In hills, for vegetables, a minimum of 0.1 ha land should be cultivated with seed under the question.
      - Maximum distance between plots of same certification lot should not exceed 50 m.
  2. Location and locality of specification
  3. Source of seed verification
  4. Timing of inspector assignment

## Grain legumes

- "Legume" comes from "legere" meaning "to gather".
- Soybean cultivation recorded in China 3000 to 2000 BC
- Globallyh 70.6 million hectares (production: 61.5 million mt) of legumes are planted
- In Nepal, 0.32 million hectares producing 0.26 million mt was grown (when ?)
- Consumption trend of legumes in Nepal is 9 kg per person per year. However, the recommended annual intake is 36 kg.
- Legumes are an important part of natural soil fertility maintainance system -- fixing atmospheric nitrogen.
  - Cowpea: 73-354 kg
  - Chickpea: 103 kg
  - Lentil: 88-114 kg
  - Pigeon pea: 168-280 kg
  - Broad bean: 45-552 kg
- Energy expended per gram of seed (i.e. gram photosynthate):
  - 0.74-0.96 in cereals
  - 2.03-2.09 in pulses/oilseeds
