
# Soil and irrigation

- Study of soil is known as pedology
- Water holding capacity is highest in clayey soil but better aeration in sandy soil.
- Soil = 5% of OM(organic matter) + 25% soil water + 25% soil air + 45% mineral/inorganic matter
- Study of soil: pedology (soil genesis)/ edaphology (soil in relation to plant production)
- Clayey soil is the bank of plant nutrient.
- Naturally formed soil aggregate is ped and artificially formed soil aggregate is clod.
- Red soil is due to presence of Fe.

**Soil textural classes**

## Fertiliers and/or extracts and their chemical composition

**Nutrient content of Green manuring crops**

```{r nutrient-green-manure}
tribble(~"Crop", ~"Nitrogen (percent)", ~"Phosphorus (percent)", ~"Postassium (percent)", 
        "Dhaincha", 3.5, 0.6, 1.2, 
        "Sirish", 2.9, 0.7, 2.6, 
        "Asuro", 4.3, 0.9, 4.5, 
        "Titepati", 2.4, 0.4, 4.1, 
        "Sanai", 3.2, 0.8, 1.8, 
        "Taramandal", 5.0, 0.9, 4.2, 
        "Khirro", 2.8, 0.7, 2.9, 
        "Banmara", 2.4, 0.7, 4.0) %>% 
  knitr::kable(booktabs = TRUE)
```


**Nutrient composition of vermicompost**

```{r nutrient-vermicompost}
tribble(~"Chemical", ~"Content", ~"Unit",
        "Nitrogen", 5.3, "percent",
        "Phosphorus", 3.8, "percent",
        "Potassium", 4.3, "percent",
        "Manganese", 645, "ppm",
        "Zinc", 580, "ppm",
        "Copper", 106, "ppm",
        "Sulfur", 0.4, "percent", 
        "Calcium", 1.5, "percent", 
        "Magnesium", 0.4, "percent", 
        "Iron", 0.7, "percent", 
        "Born", 23, "ppm", 
        "Molybdenum", 47, "ppm") %>% 
  mutate(Content = paste0(Content, " ", Unit)) %>% 
  select(-Unit) %>% 
  knitr::kable(booktabs = TRUE) %>% 
  kableExtra::kable_styling(position = "center")
```

**Nitrogen fixation capacity of different leguminous crops**

```{r n-fixation-legume}
tribble(~"Crop", ~"N-fixation (kg/ha)", 
        "Horsegram", "45-52", 
        "Pea", "52-77", 
        "Soybean", "60-168", 
        "Chickpea", "103", 
        "Cowpea", "73-354", 
        "Bean", "40-70", 
        "Lentil", "88-114", 
        "Pigeonpea", "168-280"
        ) %>% 
  knitr::kable(booktabs = TRUE) %>% 
  kableExtra::kable_styling(position = "center")
```


## Crop water requirement

```{r crop-water-requirement}
read_csv("./data/crop_water_requirements.csv", skip = 2) %>% 
  dplyr::rename_all(list(~str_replace_all(., "_", " "))) %>% 
  knitr::kable(booktabs = TRUE, caption = "Crop water requirement and growth duration of major crops of Nepal", longtable = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped", font_size = 8) %>% 
  kableExtra::column_spec(column = 1:5, width = c("6em", "6em", "8em", "8em", "20em"))
```


## Crop coefficients

These are the properties of plants used in predicting evapotranspiration (ET). The most basic crop coefficient, $k_c$, is simply the ratio of ET observed for the crop studied over that observed for the well calibrated reference crop under the same conditions.

$$
\mathrm{PET} = K_c \times \text{RET}
$$

Potential evapotranspiration (PET) is the evaporation and transpiration that potentially could occur in a field of the crop had an ideal unlimited water supply. RET is the reference ET often denoted as $ET_0$.

Even in agricultural corps, where ideal conditions are approximated as much as is practical, plants are not always growing (and therefore transpiring) at their theoretical potential. Plants have growth stages and stages of health induced by a variety of environmental conditions.

RET usually represents the PET of the reference crop's most active growth. $K_c$ then becomes a function or series of values specific to the crop of interest through its growing season. These can be quite elaborate in the case of certain maize varieties, but tend to use a trapezoidal leaf area index (LAI) curve for common crop or vegetation canopies.

Stress coefficients, $K_s$, account for diminished ET due to specific stress factors. These are often assumed to combine by multiplication.

$$
\mathrm{ET_{estimate}} = K_w \times K_{s_1} \times K_{s_2} \times K_c \times ET_0
$$

Water stress is the most ubiquitous stress factor, often denoted as $K_w$. Stress coefficients tend to be functions ranging between 0 and 1. The simplest are linear, but thresholds are appropriate for some toxicity responses. Crop coefficients can exceed 1 when the crop evapotranspiration exceeds that of RET. 

Below are given estimates of the crop coefficient for multiple crops commonly cultivated in Nepal

```{r krop-coefficient}
read_csv("./data/crop_coefficient.csv", skip = 2) %>%
  tidyr::pivot_longer(cols = `Jan_first half`:`December_second half`, 
                      names_to = c("Month", "Month half"), 
                      names_pattern = "(.*)_(.*)", 
                      values_to = "Crop coefficient") %>%
  dplyr::mutate(`Month half` = str_to_sentence(`Month half`)) %>% 
  tidyr::pivot_wider(names_from = Month, values_from = `Crop coefficient`) %>% 
  knitr::kable(caption = "Crop coefficients", escape = FALSE, booktabs = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped") %>%
  kableExtra::column_spec(column = 1:15, width = c("10em", "5em", "5em", rep("2em", times = 7), rep("3em", 5))) %>% 
  kableExtra::landscape(margin = "1.25cm")
```


## Estimates of potential evaporation

**Thornthwaite equation (1948)**

$$
PET = 16 \left(\frac{L}{12}\right) \left(\frac{N}{30}\right) \left(\frac{10 T_d}{I}\right)^\alpha
$$

Where,

  - PET is the estimated potential evapotranspiration (mm/month)
  - $T_d$ is the average daily temperature (degree C; if this is negative, use 0) of the month being calculated
  - N is the number of days in the month being calculated
  - L is the average day length (hours) of the month being calculated
  - $\alpha = (6.75 \times 10^{-7})I^3 - (7.71 \times 10^{-5})I^2 + (1.792 \times 10^{-2})I + 0.49239$
  - $I = \sum^{12}_{i = 1} {\left(\frac{T_{m_i}}{5}\right)}^{1.514}$ is the heat index which depends on the 12 monthly mean temperatures $T_{m_i}$.
  
Somewhat modified forms of this equation appear in later publications (1955 and 1957) by Thornthwaite and Mather.

**Penman equation (1948)**

The penman equation describes evaporation (E) from an open water surface, and was developed by Howard Penman in 1948. Penman's equation requires daily mean temperature, wind speed, air pressure, and solar radiation to predict E. Simpler hyrdometeorological equations continue to be used where obtaining such data is impractical, to give comparable results within specific contexts, e.g. humid vs arid climates.

**Penman-Monteith equation (1965)**

## Soil erosion

Soil erosion is the displacement of upper layer of soil, one form of soil degradation. The agents of soil erosion are water ice, and wind, each contributing a significant amount of soil loss each year.

### Process

1. Detachment

Process depends upon type of soil, OM, moisture, nature of detaching agents (energy). Rainsplash is the most important detaching agent. As a result of raindrops striking a bare soil surface, soil particles may be thrown through the air over distances of several centimeters. Continuous exposure to intense rainstorms considerably weakens the soil.

Decrease in organic matter lowers the binding action causing the low aggregates formation which can be easily eroded by water, wind and ice movement. 

Detachment of soil particles by flowing water in rill erosion (D) varies with square of velocity of flowing water (V2). The soil is also broken up by weathering processes, both mechanical, by alternate wetting and drying, freezing and thawing and frost action, and biochemical.

The transporting agents comprise those that act a really and contribute to the removal of a relatively uniform thickness of soil, and those that concentrate their action in channels. The first group consists of rainsplash, surface runoff in the form of shallow flows of infinite width, sometimes termed sheet flow but more correctly called overland flow, and wind The second group covers water in small channels, known as rills, which can be obliterated by weathering and ploughing, or in the larger more permanent features of gullies and rivers. 

Process depends upon size, and shape of detached materials and velocity of the transporting agent. The higher the size of the particles the higher will be the energy required and higher will be the depth required. Ability of flowing water to transport soil particles (T) varies with fifth power to the velocity of flowing water (V5). The severity of erosion depends upon the quantity of material supplied by detachment over time and the capacity of the eroding agents to transport it.

## Irrigation development in Nepal

```{r irrigation-development-nepal}
read_csv("./data/irrigation_development.csv", skip = 2) %>% 
  dplyr::rename_all(list(~str_replace_all(., "_", " "))) %>% 
  dplyr::mutate(`Plan period` = str_replace_all(`Plan period`, "_", " ")) %>% 
  knitr::kable(caption = "Irrigation development in Nepal during different period plans", 
               booktabs = TRUE, escape = FALSE) %>% 
  kableExtra::kable_styling(font_size = 8, latex_options = "striped") %>% 
  kableExtra::column_spec(column = 1:9, width = c("18em", rep("7em", 4), "5em", "5em", "5em", "5em")) %>% 
  kableExtra::landscape(margin = "1.5cm")
```
