
# Crop protection

- A popular fungicide, generally used for seed treatment, called Carbendazim is available in commercial formualtion as KI-BESTIN (Carbendazim 50% WP).
  - The commercial seed treatment fungicide is composed of:
    - 51% Carbendazim 98% (at minimum) a.i.
    - 2% Surface acting agent
    - 2% Dispersing agent
    - 2% Sticking agent (Glue powder)
    - 43% Inert carrier (China clay)
  - In case of carbendazim poisoning medical charcoal preparation 6-10 times is recommended.
  - It has green colored warning level.
  - It is manufactured by Kisan Agro Chemicals, Parsa, Birgunj, Nepal.
  - KI-BESTIN is a broad spectrum systemic fungicide useful as both spray and wetted powder form.

## Biopesticides

- _Nisarga_
  - Active ingredient: _Trichoderma viridae_
  - Effectiveness: Stem rot, Root rot, Sett rot, Damping off, Ganoderma etc. Against _Fusarium_, _Sclerotium_, _Phytopthora_ and _Ganoderma_.
  - Utility crops: Potato, tomato, sweet pepper, garlic, cauliflower, onion, tea, coffee and pulses.
  - Dosage: Spray 5 gm Nisarga per liter of water solution. While applying in soil, 500 gm Nisarga is mixed with 2.5 kg of mature FYM or compost. This suffices for 1 ropani of land.

- _Pseudomonas_
  - Active ingredient: _Pseudomonas fluorescence_
  - Effectiveness: Onion smut, Paddy blast, Bacterial wilt of pepper and Dieback of tomato.
  - Useful against soil borne, seed borne and air borne pathogens.
  - Secondary metabolites, i.e. Auxin, Gibberelic acid and Cytokinins promote plant health.
  - Dosage: Spray 5 g of Pseudomonas commercial formula in 1 liter of water. While applying in soil, 500 gm Nisarga is mixed with 2.5 kg of mature FYM or compost. This suffices for 1 ropani of land.

- Verticillium fungicide is effective against sucking insects and nematodes.
  - In a ropani of land, use 500 gm of verticillium preparation with 2.5 kg of FYM/compost.

- Neem Baan
  - Contains Azadirachtin.
  - Effectiveness: Against phytophagous insects for deterrence. It inhibits oviposition and is ovicidal (kills larvae if hatched)
  - Most effective against sap sucking type insects (Aphid, mealy bug, white fly, thrips, etc.) and chewing type insects (Stem and fruit borer larvae)
  - Has contact and systemic property
  - Dosage: 2-5 ml liquid in 1 ltr of water is sprayed in 12 days interval, 2-3 times.
  - Composition: 0.03%, 0.15%, 1%, etc.

- Nuclear polyhedrosis virus (NPV)
- Granulosis virus (GVs)

## Pathogens

### Nematodes

- Nematode is derived from the Greek words, "Nema" = thread/fibre, "toda" = worm.
- In germany, there is a separate University of Nematology.
- Nematodes can be defined as unsegmented, bilaterally symmetrical, tryploblastic, pseudocoelomate, invertebrate, and thread like worms.
- So far 50000 nematode species are recorded worldwide. 10000 are found in fresh water and soil. 300 species are known to be plant parasites.
- Molya disease ( _Heterodera avenae_) causes 6-7 crore/year loss in Rajasthan and ear cockle ( _Anguina tritici_) causes 8 crore loss in India.
- _Radopholus similis_ was found associated with citrus decline in Florida, USA.
- In India, _Tylenchulus semipenetrans_ was associated with citrus decline.

- Plant parasitic nematodes are triploblastic, bilaterally symmetrical, unsegmented, pseudocelomate and vermiform animals.
- The body of the nematode may be elongated, spindle shaped, fusiform tapering towards the end but the cross section is always circular.
- Smallest nematode is 10mm long ( _Paralongidorus_).
- Female nematodes are more virulent and agressive than male in attacking and parasitizing the plants.
- Plant parasitic nematode possess spear or stylet.
- Nematodes are known to transmit viruses:
  - Two single stranded RNA virus genera, Nepovirus (NEPO) and Tobravirus (TOBRA).
  - 11 species of Xiphinema transmit 13 NEPO virus (Grapevine fan leaf virus)
  - 11 species of Longidorus transmit 10 NEPO virus
  - 14 species of Trichodorus transmit various strains of TOBRA virus: tobacco rattle and pea early browning

Insect transmitted viruses

```{r insect-transmissed-viruses}

tribble(~"Virus", ~"Nematode", 
        "Rice dwarf virus", "Nephotettix cincticeps", 
        "Rice tungro virus", "Nephotettix impicticeps", 
        "Rice grassy stunt virus", "Nilaparvata lugens", 
        "Tomato spotted wilt virus", "Thrips tabaci, Frankliniella spp.", 
        "Tomato yellow leaf curl virus", "Bemisia tabaci", 
        "Tomato yellow mosaic virus", "Bemisia tabaci", 
        "Soybean yellow mosaic", "Bemisia tabaci", 
        "Grapevine virus A", "Pseudococcus longispinus (Mealybugs)", 
        "Cowpea mosaic virus", "Epilachna varivestis", 
        "Potato virus X virus", "Melanoplus differentialis (Grasshopper)", 
        "Tobacco mosaic virus", "Liriomyza langei (Leafminer); Mechanical transmission", 
        "Onion mosaic virus", "Eriophyses tulipae (Mites)", 
        "Soybean mosaic virus", "Aphids", 
        "Potato leaf roll virus", "Myzus persicae") %>% 
  knitr::kable(booktabs = TRUE, caption = "Insect mediated virus transmission")
```


## Crop diseases

### Ergot (Wheat, barley, oats, rye, triticale)

Hosts: 
All grasses, particularly, blackgrass ( _Alopecurus myosuroides_)

Symptoms:
- Causal fungus only attacks ears of flowering, replacing the grain in a few spikelets by a hard, purple black sclerotium, known as ergot.
- Such ergots can be very large, upto 2 cm in length, and very obvious in the standing crop in contaminated grain samples.

Life cycle:
- Ergot is not truely a seed borne disease, however it can be spread by ergots in contaminated seeds.
- At or near harvest, ergots fall to the ground where they remain untill the following summer, when they germinate to produce club-shaped spore bearing structure (stroma). These ascospores are spread by the wind to nearly open flowers of grasses/cereals. The spores germinate in flower, infecting the ovaries. This infection leads to the production of secondary spores (condia) encased in sticky secretion commonly referred to as honeydew. This attracts insects which carry the spores to other flowers, where further infection can occur.
- Wheat and other cereals are less severely affected than rye although, occassionally more open-flowerd wheat variety can be badly affected.
- Disease is favored by cool, wet conditions during flowering which facilitate spore production and prolong the flowering period, making infection more likely.

Importance:
- Very little direct effect on yield.
- Affects stocks which when fed to flour made with cereals with large amount of toxic alkaloid containing ergot, possess health risks.

### Fusarium

Fusariusm head blight/ear blight, foot rot, seedling blight
Pathogen: _Fusarium spp._ and _Microdochium nivale_
Hosts: Wheat, barley, oats, rye triticale and grasses.

Symptoms:
- Form a complex of diseases on seeds, seedlings and adult plants.
- _Microdochium nivale_ (formerly known as _Fusarium nivale_) is seed-borne pathogen and causes seedling blight resulting in seedling death and thinning of plant stand.
- _M. spp_ (other than _M. nivale_) cause a range of symptoms including brown lesions on stem bases, often restricted to outer leaf sheath.
- _Fusarium lesions_ often begin in the leaf sheath at the stem base where crown roots split the leaf sheath when emerging.
- This infection can spread up the leaf sheath causing long dark brown streaks at the stem base. The other symptom in cooler regions is brown staining of lower nodes.
- In older plants, fusarium infection can produce a true foot rot, where the stem base becomes brown and rotten, resulting in lodging and white heads.
- Symptoms are prevalent in very dry seasons as well.
- Ear blight causing fungus: _F culmorum_ and _F graminearum_ are common. Other are, _F avenaceum_, _F poae_ and _F langsethiae_.
- Infection frequently results in the whole or part of the ear becoming bleached.
- Symptoms seen when ears become infected during the early flowering stages, later infection may result in infection of grain but without obvious bleaching of the ears.
- Important due to its mycotoxin that gets accumulated in grains.

Life cycle:
- Most important source is seed but fungus survives on debris in soil also.
- Spores are splashed in canopy causing ear blights and seed borne infection, in wet seasons, especially during flowering and grain formation.
- Most fusarium species have competative saprophytic abilities which allow them to colonize debris and stubble in soil.

Importance:
- When wet season coincides with flowering high levels of ear blight can occur.
- Due to seed borne nature of pathogen, seed treatment plays role in preventing seedling loss in wheat.

### Major diseases of rice

1. Blast

- Bavistin, Dorosal 2-3 g per kg seed treatment
- Tricyclazole 75% WP 0.75 g per ltr spray at 15 days interval
- Kasugamycin 3% SL 1.5 ml per ltr at 15 days interval

2. Bacterial leaf blight
  
- Use Agromycin-100 0.25 g per ltr for seed soaking for 30 minutes

3. Brown leaf spot disease

- Bavistin, Dorosal
- Apply Mancozeb 75% WP (Dithane M-45) 3 g per ltr water, Propineb 70% WP 3 g per ltr water at 15 days interval for 3 times.

4. Foot rot

- Carbendazim 50% WP seed treatment

5. Sheath blight

- Maintain spacing
- Validamycin 3% L 3 g per ltr water; Pencycuron 22.9 SC 1.5 ml per ltr; Carbendazim 70% WP 1.5 g per ltr spray at 10-12 days interval for two times.

6. Khaira disease

- 20 g $\mathrm{ZnSO_4}$ + 12% $\mathrm{CaCO_3}$ in 50 ltr water per ropani at 10 days interval for 2 times.

### Major diseases of Wheat

1. Leaf blight

- Small brown dots on leaves
- Later on the dots coalesce to cause wilting or blighted appearance
- Use Vitavex-200 2 gm per kg seed as presowing treatment
- Increase potassium fertilizer dosage

2. Brown rust

- Orange color spots on upper surface of leaves.
- Spots do not coalesce or merge
- Mancozeb (Dithane M-45 45 WP) 1.5-2 kg in 750 ltr water spray at interval of 15 days for 2-3 times.

3. Yellow rust

- Yelow colored spots, elongated and jointed to form stripes
- Cultivation of resistant varieties: WK-1204, Pasang Lhamu.

4. Loose smut

- Instead of grains black mass of fungal hyphae fills the panicle.
- Use of healhty seeds, Vitavex-200 2 g per kg seed treatment
- Bury the sick panicles in initial stage of disease appearance.
- Annapurna variety is relatively tolerant to disease.

5. Stinking smut/hill smut

- Diseased grains are rounded, black colored spores filled
- Spores only released after grain is crushed
- Smell of fish
- Crop rotaion for 2-3 years, Vitavex-200 2 g per kg seed treatment.


## Crop insects

### Major insects of rice

1. Seed bed bettle, mole cricket, field cricket
2. Borer
3. Rice hispa
4. Hoppers
5. Rice bug
6. Leaf roller
7. Mealy bug

### Major insects of wheat

1. Larvae of wireworm

- Similar to cutworm in Maize (damages the crop at night)
- Use Bt for control
- Malathion 5% DP 2 g per kg with wheat bran 1/2 kg per ropani, during evening
- Chlorpyrifos 10% Granule or Malathion 5% DP 1 kg per ropani for soil treatment

2. Aphid

- Lady bird beetle is its natural enemy
- Dimethoate 30% EC 1 ml per liter water

3. Pink stem borer

- Same as that for control of Maize stem borer

## Pesticide toxicity

- A pesticide is any substance used to control pests. Pests may be target insects, vegetation, fungi, etc. Most control the pests by poisoning them. Unfortunately, pesticides can be poisonous to humans as well.

- Toxicity: The toxicity of a substance is its capacity to cause injury to a living system. A living system can be things such as a human body, parts of the body (lungs or respiratory system), a pond, a forest and those creatures that live in there. Toxicity represents the kind and extent of damange that can be done by chemical. In other words, if you know the toxicity of a pesticide, you know how poisonous it is.

- Dose-time relationship of pesticide toxicity
  - Dose is the quantity of a substance that a surface, plant or animal is exposed to.
  - Time means how often the exposure occurs.
  - This relationship gives rise to two types of toxicity.
  1. Acute toxicity: This refers to how poisonous a pesticide is to a human, animal or plant after a single-term exposure. It generally implies the effect that occurs within 24 hours of exposure.
  2. Chronic toxicity: This refers to delayed poisonous effects from exposure to substance.
  
- Routes of entry:
  1. Local: local effect refers to those that take place at the site of contact with material. e.g. skin irritation/inflammation on th hand in response to hand contact, irritation of mucous membrane lining the lungs due to inhalation of toxic fumes.
  2. Systemic: Effect that occur away from the original point of contact. These pesticides are distributed throughout the body once they enter. They function by blocking or stimulating a chemical signal, generally that of the nervous system (Cholinesterase).

- Pesticides may have following actions:
 - Additive, antagonistic or synergistic
 - Immediate or delayed
 - Reversible or irreversible action
 
- Exposure may result in following effects:
  - Reproductive effects
  - Teratogenic effects: Effect on unborn offspring, such a birth defects.
  - Carcinogenic effects: Cancer in living animal tissues.
  - Oncogenic effects: Tumor forming effect (not necessarliy cancerous)
  - Mutagenic effects: Permanent effect on genetic material that can be inherited
  - Neurotoxicity: Poisoning of nervous system, including the brains.
  - Immunosupression
  
Acute toxicity measures

To figure out how acutely toxic a pesticide is, scientists give laboratory animals short-exposure to does of pesticide being tested. Experimental doses are given orally, as well as put on eyes, skin, and in the air that test animals breathe. These animals are then carefully observed for the changes.

**LD50**

Amount of a pesticide that has killed half of the animals in a laboratory test. LD50 values are effective for both oral and dermal routes of exposure. But they do not tell us about how the chemical acts, nor about how sensitive different organs within an animal or human might be. LD50 for different chemicals can be compared if the same test animial was used. The LD50 values are measured in unit of weight called mg per kg (or interchangeably, parts per million).

**LC50**

This measure of toxicity gives the acute inhalation toxicity.

Chronic toxicity measures

There is no standard measure like LD50 for chronic toxicity studies. Often the length of the experiment is in days, months or years and the amount of each dose is stated. For e.g., a study of chronic oral toxicity might look like, "8 mg of pesticide to rats daily for two years. No symptoms of poisoning appeared."

- Two classes of pesticides, organophosphates and carbamates can slowly poison by attacking an essential body chemical called "cholinesterase". The chronic exposure to Organophosphate pesticides can be measured by monitoring changes in blood cholinesterase levels. In humans, decrease in cholinesterase levels are sure sign that exposure to these types of pesticides should be avoided untill the level is measured as being normal again.

Categories of pesticide toxicity

```{r pesticide-hazard-category}
tribble(~"Toxicity class", ~"Toxicity label", ~"Oral LD50 (mg/kg)", ~"Dermal LD50 (mg/kg)", ~"Inhalation LC50 (mg/L)", 
        "Highly toxic", "Danger", "0-50", "0-200", "0-0.2", 
        "Moderately toxic", "Warning!", "50-500", "200-2000", "0.2-2", 
        "Slightly toxic", "Caution!!", "500-5000", "2000-20000", "2-20", 
        "Relatively non-toxic", "Caution!!", ">5000", ">20000", ">20") %>% 
  kable(booktabs = TRUE, caption = "Categories of pesticide toxicity")
```

Status of pesticide use in Nepal

- Initially, DDT was imported in 1952 AD for control of Maleria.
- For the same purpose, DDT was reimported in 1955 AD
- For use in crops, DDT was imported in 1956 AD.
- According to Thapa, 2003, average pesticide use in Nepal is 142 gm/ha.
- In general, cropwise analysis of pesticide use signals alarming levels of residues, hence their current state of use being haphazard.
  - Tea: 2100 gm/ha
  - Cotton: 2560 gm/ha
  - Vegetables: 1400 gm/ha
  
- On environmental perspective, pesticides are of following types, based on bio-degradation:
  1. Environmentally degradable/non-persistent:
    - Dimethoate (Nuger, Roger, Dimet)
    - Dichlorovos (Dum, Vapon)
    - Fenitrothion (Folithion)
    - Malathion
  2. Environmentally non-degradable/persistent:
    - PoPs: Aldrin, chlordane, DDT, Dialdrin, Eldrin, Heptachlor, Mirex, Toxaphene, HCB, PCB, Dioxyn, Furan, etc.
    - These pesticides require special treatment facility for disposal.
