
# Plans and policies

## General

- MOA has issued an "Implementation Guidelines on PPS for Agriculture Development, 2055 (1998)" to ensure common framework for implementation of pocket package strategy in order to carry out APP at field level. The guidelines have presented the pocket package program under two categories namely Crop production and Livestock production.
- Agriculture fair is based on the concept of "Seeing is believing".
- Department of Agriculture first broadcasted improved production technology in agriculture through radio for the first time in 2016 B.S.
- Agriculture program broadcast over radio was stopped for a few years and later resumed back in 2023 Mangsir 20.
- In group approach of extension is used in convincing a farmer to use a technology when s/he has been made aware of it through mass media approach.

## Millennium development goals (MDG)

Millennium Development Goals (MDG) Progress Report (NPC, 2010) stated that the poverty rate has decreased, a reduction in the population suffering chronic food insecurity, and reduced unemployment rates, and there is a positive trend towards gender equality as indicated by a balanced enrollment of girls and boys in primary schools. The Report notes there has been increased allocation of public resources in favor of marginalized groups in remote areas, and greater attention has been given to environmental conservation and adaptation to climate change. The policy environment for achieving most of the MDG targets seems favorable with the Interim Constitution of Nepal 2063 (2007) and its subsequent laws emphasizing inclusive, participatory and decentralized governance.

```{r}
# readxl::read_xlsx("./data/millineum_development_goals.xlsx") # this needs to be filled
```


## Periodic planning

- The nineth plan considered poverty reduction and other development goals. The main target of this plan was to reduce the headcount poverty in the headcount poverty index from as high as 42% to 32%, 23%, 15% and 10 by 2002, 2007, 2012 and 2017, respectively. However, it could not be realized as the poverty rate in 2011 was 25%. 
- The tenth plan (2002-2007) is considered as a comprehensive plan designed for the reduction of poverty to a desired level. Designed as the poverty reduction strategy paper (PRSP), the tenth plan incorporated the involvement of the private sector in the market economy with the facilitation from the government side. Because of the bad political situation and existing conflict inside the country, reduction of poverty was a daunting task and this was reflected in the level of poverty that country was facing by the end of this plan.
- Instead of eleventh plan, interim plan for 3 years was launched from 2007/08-2009/10. The main objective of this plan was to generate an experience of a direct feeling of change in the lives of the general public by supporting in the establishment of peace and reducing the existing unemployment, poverty and inequality. This plan was designed in a critical state of transition where the country was formulating a new constitution to restore peace and reconstruction and reintegration among people suffering from the conflict.
- The twelve three year plans (2010/11-2012/13) has kept poverty alleviation as its major agenda and inclusive employment generation and equitable economic growth are considered two important pillars for realizing poverty alleviation. The purpose of this plan is to incorporate every community from all around the country to make sure that every underprivilaged and marginalized group is included. This plan has mobilized volunteers through National Development Volunteer Services (NDVS). NDVS has been implemented by district development committees. The purpose of this program is to mobilize volunteers as moderators in disadvantaged communities such as poor, ethnic, Dalit communities so that they will help them as facilitator in education, health, infrastructure, social mobilization and other areas.

## National planning commission

Introduction and History

The National Planning Commission (NPC) is the specialised and apex advisory body of the Government of Nepal for formulating a national vision, development policy, periodic plans and sectoral policies for overall development of the nation.
 
- It is headed by the Right Honorable Prime Minister. The NPC assesses resource needs, identifies sources of funding, and allocates budget for socio-economic development. 
- It serves as a central agency for monitoring and evaluating development policy, plans and programs. 
- The NPC also serves as an intellectual hub for the exchange of new development ideas and proposals from scholars, private sector, civil society, and development partners.

Membership

At present, the NPC has 
	- one full-time Vice-Chairman, 
	- six members, and 
	- one Member-Secretary who also heads a fully staffed secretariat.

The Chief-Secretary and the Finance Secretary are ex-officio members of the Commission. The national statistical organization of Nepal, the Central Bureau of Statistics (CBS), functions as a specialized entity of the NPC, headed by a Director-General.

The Commission consists of four functional divisions each headed by a Joint-Secretary: 

i. Economic Management; 
ii. Infrastructure and Production;
iii. Research and Evaluation; and 
iv. Good Governance and Social Development. 

Each division is further divided into Sections headed by an Under-Secretary.

Historical Background

The Planning Commission was first created in Nepal in 1956. It was soon renamed in accordance with the Yojana Mandal Act of 1957. Following the introduction of the partyless Panchayat system in 1961, the National Planning Council was formed under the then King. In 1963, the Council was dissolved and a new planning body, with an identical name, was constituted under the Chairman of the Council of Ministers. All the Ministers became ex-officio members of the Council; and the Ministry of Economic Affairs was renamed the Ministry of Economic Planning.

In 1968, all tasks related to development budget and foreign aid hitherto carried out by the Ministry of Economic Planning were assigned to the Ministry of Finance. The National Planning Council then morphed into the National Planning Commission (NPC) under the Chairmanship of the Rt. Hon'ble Prime Minister. A 1972 study on the functions and responsibilities of the central planning agency resulted in the reconstitution of the NPC.

After the historic restoration of multiparty democracy in 1990, the newly elected government reconstituted the NPC again with the Prime Minister as Chair, a full-time Vice-Chairman, five Members, and a Member-Secretary. The current organization and functions of the NPC draw on the Executive Order issued by the cabinet in 2018.


## National tea policy, 2000

- GoN approved and implemented National Tea Policy, 2000 as per intention of National Tea and Coffee Development Board Act, 1992 for:
  - Income generation (enhancing employment, earning foreign currency)
  - Participation in private sector in production, processing and commercial transaction through systematic and sustainable utilization of resources in country.
- Working policies:
  A. Production and processing
    - Banks to provide priority credit
    - Upto 80% loan of total project cost to registered tea plantation industry
    - 7 years grace for orthodox tea and 5 years for green tea on loan in hills and terai respectively.
    - Interest on loan not capitalized in grace period
    - Income tax not levied within grace period
    - Interest and principal to be fully paid up within 10 years from the end of grace period
    - Exemption of 75% on land registration fee
    - Land revenue exemption
  B. Market and trade promotion
  C. Institutional arrangement
  D. Manpower development
  E. Development and promotion of auxillary industries

## National agricultural policy, 2004

Vision

- To bring about an improvement in the standard of living through a sustainable agricultural development to be achived by transforming the current subsistence oriented farming system into a commercial and competative farming system.

Objectives

Broad

- To ensure food security and poverty alleviaion by achieving a high and sustainable economic growth through commercial and competative farming system.

Specific

- Agricultural production and productivity shall be increased
- The bases of a commercial and competative farming system shall be developed and made competative in regional and world markets
- Natural resources, as well as the environment and biodiversity shall be conserved, promoted and properly utilized

Policies

Objective 1

- Increasing of agricultural production and productivity
  1. Utilize local potentialities, comparative advantage and special opportunities; Ensure development, extension and utilization of agricultural technology; Commercialization and diversification of agriculture
  2. Scientific land use system
  3. Irrigation, agricultural roads, rural electrification and appropriate agriculture technology development and epansion
  4. Pocket development of HVAP
  5. Entrust local body for formulating, implementing and monitoring agricultural plans suitable to local needs and priorities. Conditional basis of grant.
  6. Multi-district projets to promote agricultural produciton and enterprises operated and suppored at implementation level through central departments and directorates.
  7. Farmer's group appraoch for on-site extension service
    - Agriculture and frest colleges to extend agriculture technologies through package programs
    - IT and mass communication development
  8. National agricultural resource centres for high quality inputs
    - based on development regions and geo-graphical subdivisions
    - NARC to be converted to integrated centre
  9. Participatory and competative agricultral research and development system promotion
    - Technology exchange encouraged between organizations
  10. Private and foreign investment in agriculture research and development encouraged
  11. Input supply monitoring and regulating guarenteed
  12. Surveillance system to assess impact of excessive rains, droughts and calamities.
  13. Emphasis on farmer's training programs
  14. Linkage of agricultural production and enterprises' return to ensure agriculture credit flow.
  15. Establishment of agriculture and forestry university. Quality of agriculture human resource to be increased by arranging cooperation and exchange of technicians and experts among universities/colleges, agriculture research centre and national agriculture resource center.
  16. System development for collecting, analyzing and projecting data required for formulation of plans, determining policies and carrying out monitoring and evaluation of activities related to agriculture sector shall be strengthened.
  17. Women's participation in all fields of agriculture upto 50%
  18. Resource poor farmers (with less than 4 hectares of land) will be identified and classified.

Objective 2

- Special facilities for target groups: For farmer's having less than 0.5 hectares and unirrigated lands, dalit and utpidit classes of farmers and other marginal farmers
  1. Opportunity of gaining access to lands
    - Legal ceilings on landholdings and exemptions
    - Progressive taxation
    - Legal provision for contractual farm lands.
  2. Land bank establishment
    - Local body as information provider
    - Concessional loans
  3. Forest upgradation and other land will be handed to the traget communities under lease.
  4. Special facilities to target groups to build and install infrastructures: small irrigation (pedal pumps, power pumps, sprinklers, drips and water harvesting ponds)
  5. Utilization of means to increase production and income to mitigate food deficit. Network for food mobilization development.
  6. Government services based on priority from food security viewpoint.
  7. Food safety nets development for farmers with less than 0.5 hectares of land.
  
Objective 3

- Development of commercial and competative farming system
  1. Large production pockets, priority for comparative advantage products. Technology and technical sevices as well as infrastructure mobilized in integrated manner.
  2. Local production of food grains will be encouraged under food supply programme
  3. Double track management system will be adopted in government farms and centres
  4. Livestock insurance programmes extension. Poultry insurance and HVAP and seed crops insurance.
  5. Organic farming will be encouraged; certification support
  6. Production of hybrid seeds and improved seeds will be encouraged, GMOs will be regulated.
  7. Traditional, local original agriculture products and technologies will be registered and promoted.
  8. Agriculture training classification
    - Capacity improvement training for agriculture workers and farmers
    - Enterprise promotion training
  9. Training educated but unemployed youths
  10. Local production, sale and distribution of improved agriculture resource inputs (seeds, plants, saplings, breeds, fingerlings, etc.) and manure, insecticides, pesticides, regulated and quality control.
    - Private agriculture lab services will be regulated and accreditated
    - High quality product processing facilites and services will be offered
  11. Agriculture and livestock quarentine services will be systematized and strengthened in order to ensure the production of high quality agriculture products and raise their credibility in local and external markets.
  12. Participation of local bodies will be strengthened in determining, controlling, certifying, and regulating standards of food stuffs.
  13. Regulatory services will be upgraded as per international treaties and agreements and the national requirements.
  14. Promotion of cooperative based agriculture industries and enterprises
  15. Mobilize agriculture industry and enterprise promotion board
    - To analyze and provide outlet to complaints and suggestions
  16. Commodity and subject specific policies equipped with incentives developed in order to attract cooperative and private sectors to make investments in commercial production, processing and marketing
  17. Agricultural industry development policy formulation
    - Interlink agriculture research + production + processing industries + internal and external export markets
  18. Free-based agricultural technology extension services in areas of commercial agricultural production
  19. Private sectors engaged to operate suitable farms/centers through contract lease agreements.
  20. Developing and extending market information system and disseminating such information shall be carried out in private + cooperative + local bodies partnership
  21. Collection centres close to production centres (Hat bazzars, well equipped wholesale and seasonal markets as well as private cooperatives)
  22. Agriculture enterprise promotion board to provide capital and incentives/facilities to industries and enterpreneurs for:
    - import substitution
    - export promotion
  23. Cooperative promotion of potential farmer groups and enterprenuers by mobilizing and promoting local small capital and other resources
    - Rural cooperatives as local delivery points
  
Objective 4

- Conservation, promotion and utilization of natural resources and environment
  1. Impact minimization of agro-chemicals
  2. Organic fertilizer promotion
  3. Gene banks and insitu conservation. Participatory biodiversity parks
  4. Biodiversity conservation, promotion and utilization to improve condition of degraded forests and natural reservoirs.
  5. Conservation oriented farming system
    - Watershed management through local participation
    - Soil erosion control through local participation
  6. Checking cultivable land fragmentation and ensure scientific mangement
  
Objective 5

- Implementation and monitoring arrangement
  1. Participatory involvement of stakeholders in process of formulating, monitoring and evaluating plans connected with agriculture sector from local to central level.
  2. Role of national agricultural development board (Central agriculture development committee and Regional agriculture development committee) devolution for policy implementation to local bodies (VDCs and DDCs) to ensure formulation, implementation, monitoring and evaluation of plans in accordance to LSGA, to:
    - DADC (District agriculture development committee) with technical feedbacks to CADC and RADC
    - ADC (Village level ADCs) with technical feedbacks to CADC and RADC.
  3. Concerned ministries shall implement this policy on their related sectors
    - NADC to monitor the implementation on national level
    - Strategies, programs and responsible bodies of related matters may be implemented by concerned ministries upon approval.

## Medium term expenditure framework (MTEF)

Recognizing the fiscal deterioration and the need to revise the Tenth Plan accordingly, the Government initiated the preparation in parallel of a Medium Term Expenditure Framework (MTEF) in late 2001. The key objective of the MTEF was to begin to implement the Tenth Plan from the beginning of the fiscal year 2002/03, without waiting for the formal finalization of the Plan. Accordingly, the key projects/programs and activities that are considered essential for achieving the Tenth Plan's poverty reduction goals and their resource requirements for the next three years were to be identified, and were to be given priority in terms of budget allocations in the 2002/03 budget. Additional considerations guiding the MTEF, in view of the fragility of the fiscal situation, were: (i) How to streamline the budget which was already overextended well beyond resource availability, with over 600 ongoing activities/budget lines ? (ii) How to protect the Tenth Plan's key priorities in the event of further shortfalls in resource availability ? and (iii) How to ensure that the resources allocated and released for priority activities would be effectively utilized? The MTEF was expected to be the principal instrument for operationalizing the Tenth Plan, by prioritizing the proposed Plan activities according to the changing resource situation, and firmly linking the annual budget process and the Plan.

Thus, a three year MTEF was prepared in early 2002 by virtually the same line ministry teams responsible for Tenth Plan preparation, using similar methodology for ranking and prioritizing activities, (but with a sharper focus on implementability, short term financing needs and results). On the basis of the MTEF, (i) the 2002/03 development budget was initially set at a significantly lower level -- Rs. 38 billion, compared to Rs. 50 billion in the 2001/02 budget; and the number of budget lines were reduced from over 600 to 430. (ii) Even more importantly, all activities were classified into three groups according to their priority (P1being the highest priority and P3 being the lowest); and (iii) The principle was established that the priority classification would be strictly followed in releasing funds for development activities. Thus, P1s will receive first priority in budget releases; and P2 and P3 will get funded only if funds are still available after providing for P1s. This has established clear criteria and methodology for adjusting resource allocations on the basis of Tenth Plan's priorities, in line with changes in the resource situation. (iv) Finally, resource allocation was linked to expenditure reporting and performance, in order to ensure satisfactory results. Thus, under the new system that has been evolved, activities which do not provide statements of expenditures (SOEs) for the preceding trimester will not be given additional funds, until they provide the SOE's.


## 10th plan (FY: 2059/60-2063/64)

Four pillars of poverty reduction strategy:

1. Broad based economic growth; 
2. Social sector development including human development;
3. Targeted programs including social inclusion, in order to bring the poor and marginalized groups into the mainstream of development, together with targeted programs for the ultra poor, vulnerable and deprived groups (who may not adequately benefit from the first two pillars); and 
4. Good governance.

The plan stresses strategic cross-cutting approaches with regard to: 

(a) redefining the role of the State, and limiting public interventions; 
(b) enlisting the private sector to play a leading role in employment and income generation and together with NGOs, INGOs and CBOs, in complementing government efforts in service delivery functions in key areas, as well as in implementing key activities; 
(c) promoting community participation in and management of activities at the local levels; and 
(d) accelerating the decentralization process, which is also a key element under good governance

Implementation strategy of four-pillars and their implications for sectoral programs and activities are:

1. Macroeconomic framework
2. Structural reform agenda
3. Broad based economic growth
4. Social sector development
5. Social inclusion and targeted programs
6. Good governance
7. Improving financial management and accountability
8. Decentralization

```{r tenth-periodic-plan}
read_csv("./data/periodic_plan_10th.csv", skip = 1) %>% 
  rename_all(function(x)str_replace_all(x, "_", " ") %>% str_to_sentence()) %>% 
  knitr::kable(booktabs = TRUE, caption = "Achievements of 9th and goals for 10th periodic plan under normal and restrained cases.") %>% 
  kableExtra::kable_styling(latex_options = "striped", font_size = 8) %>% 
  kableExtra::collapse_rows(columns = 1, valign = "middle") %>% 
  kableExtra::column_spec(column = 1:5, width = c("10em", "10em", "8em", "8em", "8em"))
```


**Decentralization as implementation strategy of 10th plan**

Decentralization is an important means of bringing development closer to the rural poor—by involving local communities in developing appropriate programs which are best suited to their needs and in implementing them, ensuring greater accountability for use of public resources, and mainstreaming the poor and deprived groups. The local Self Governance Act (LSGA, 1999) provides a framework for decentralization, which the government has agreed to implement in a phased manner. The Local Bodies Finance Commission (LBFC) has also recently made recommendations for fiscal devolution. A Decentralization Implementation and Monitoring Committee was also set up to oversee effective implementation. But progress so far has been hampered by institutional capacity and fiscal constraints, and more recently by the dissolution of elected local bodies.

The main objectives of the Tenth Plan in this area are to ensure greater participation of people in the governance process to accelerate the development process by implementing fiscal devolution in a phase-wise manner as envisaged in the LSGA; and creating necessary institutional mechanisms including the formation of a Local Service Commission. The main strategies being adopted to achieve these objectives include: the devolution of basic service delivery functions, capacity building of local bodies and decentralization of certain revenue mobilization functions to local bodies.The major activities being undertaken to carry out this strategy include the transfer of the management and operation of DIDO, schools, health posts, postal service, agriculture and small irrigation projects, as well as rural roads, to local bodies. The government will constitute local service cadres and enhance the planning, management and evaluation capacities of the local bodies by providing the necessary training and equipment. Revenue collection and fiscal management of local bodies will be improved and auditing and accounting system refined. As a result of these initiatives, it is expected that over the Plan period, 25,000 primary schools, all sub-health posts and agriculture extension services will be transferred to local bodies. The share of local revenue in the local bodies' budget will rise, transparency and accountability will increase and service delivery to communities at village levels will improve considerably.

## 14th plan (FY: 2073/74-2075/76)

- Recently proposed projects: 13
- 21.6% of the population is below poverty line by the end of 13th plan

```{r forteenth-periodic-plan}
read_csv("./data/periodic_plan_13th.csv", skip = 1) %>% 
  knitr::kable(booktabs = TRUE, caption = "Goals and achievements of 13th periodic plan (2073/74-75/76) for selected indicators") %>% 
  kableExtra::kable_styling(latex_options = "striped") %>% 
  kableExtra::column_spec(column = 1, width = "18em")
```

- Gross national income in 13th periodic plan was expected to rise by 6.0 percentage, however by 2070/71 it was 5.72%, in 2071.72 2.32% and in 2072/73 it remained 0.77%. Hence resulting in average economic growth of 2.92%.
- There was a projection for an increase in 4.5% in agriculture sector, however, in years 2070/71, 71/72 and 72/73 respectively, 4.55%, 0.81% and 1.33% growth were achieved. Hence resulting in average growth of 2.22%
  - Food grain production was projected to reach 1,08,81,000 mt annually. However at the second year of the plan only 92,66,000 mt were realized.
  - There was an estimation of 12,60,000 mt production of fruits. But only 11,86,000 mt were produced.
  - There was an estimation of 36,00,000 mt production of vegetables. However, 36,23,000 mt were produced of it, surpassing the set target.

## 15th plan
    
\begin{questions}

\question Describe the 15th plan approach paper:
% \makeemptybox{1.5in} % use this to provide a box in question

\begin{solution}

The 15th plan approach paper is a piece of shit.

\end{solution}

\end{questions}
