## Civic society

### Civil Society and Political Parties

Political party is a part of society, not whole like polity and the state. It claims the right to defend certain ideas and interest, represents citizens and fractionalises the society to expand its vote base. Party has emerged in opposition to despotism and politicised citizens to fight for democracy. As an agent of social change, the shared dharma of political party and civil society is to modernise the general society and work for its reform and renewal to make democracy a way of life. Conscious of freedom and autonomy Nepalis can articulate their voice for a fulfilling destiny.

Rationality

Genuine civil societies stand above partisan prejudice for worthy initiatives and build bridges across various empirical social, economic and political divides. But if they emerge from the incubus of parties and geopolitical forces, they stoke social and political cleavages, incarcerate the public sphere and indulge in instrumental actions. For average Nepalis, it increases the cost of politics and stifles the birth of civility and national will. Acute party-mindedness of leaders has turned Nepali parliament feeble to advance public interests. Several verdicts of the Supreme Court against their actions stand as a hostile witness. The names of Nepali parties do not match with their ideologies. Leaders' impulse for durability in power made many compromises and less faithful to their own parties' vision. The rationality of civil society rooted in altruistic human essence seeks to improve leaders' utilitarian instinct.

Genuine civil societies offer democratic impulse to families, communities, societies, political parties, business and bureaucracy for cooperation. They champion satto guna based on enlightenment ethos - freedom, justice, solidarity, ecological ethics and peace. The instinct of political party is rajo guna, a craze for power calculation and animus dominandi, instinct for domination. The virtues of civil society propel them to perform niskam karma (selfless service) while parties are the domain of partisan selfishness. Their fierce competition with rivals for a control of rule, authority and legitimacy flags the purpose of political power what Alexander Solzhenitsyn says, "A call to service." Civil societies bear cultural and cosmological values while political parties shape tools of power. Their domain of action is confined to the state's realm. Political parties are indulged in ideological indoctrination what Karl Marx calls "false consciousness," partisan attachment and political mobilisation forking paths for the left and the right but a common lust for power. They turn wild when power shift in coalition swings the perception of each other. Only leaders of democratic parties maintain a balance between their drive for power, sense of personality and upholding of conscience to act in response to vox populi.

The fundamental questions in building interface between the two are: How can Nepali civil society and political parties mediate the polity and social classes to remove the evils that divide them in sharing constitutional goals? Can both work for participatory culture aiming to eliminate all forms of biases and privileges from public life and public policy and open up possibility for Nepalis to become cultural and creative citizens? Is Nepali society capable of sustaining its history and tradition of tolerance when civic education, interest articulation and communication performed by parties and civil society nourish pre-political style based on biological and social differentiation? When citizens face authoritarian decision-making from family to the halls of power? How can civil society-political party interface contribute to the canon of a just constitutional order which can bring the redeeming power of politics to settle contesting issues? Can Nepali leaders achieve good governance when they have to operate under pre-modern polity devoid of real power separation, autonomy, capacity to implement the constitution and functional specialisation of roles? Ordinary Nepalis' support for transparency, however, is opening actors and institutions to vital reason and inquiry.

Nepal's major political parties - Communist Party of Nepal, Nepali Congress, Federal Socialist Forum, Nepal and Rastriya Janata Party-Nepal, dozens of smaller parties and myriad of civil society operate in multiple directions. Some of them lack constitutional will as they are liberated from the ethical norms of politics. Synergy of party-civil society interface is not enough unless they stand on the reason of state, shape a shared view on democracy, human rights and justice and coordinate their policies for a common future. The stability of Nepali polity rests on institutional equilibrium of power and moral strength of its leaders. Nepali people are citizens of Nepali state, not polity, government and political parties, which entitles them with equal constitutional and human rights.

Constant engagement of both in educating citizens about civic knowledge, life-skills and wealth-creation helps them to realise their potential and perform public duties. An elevation of human condition calls for opportunity in building this nation and bridging the fundamentalist gaps that drown the voice of reason infecting political polarisation. Nepali citizens' claim of state sovereignty can help political parties to contextualise policies and laws and set a common ground for the resolution of symbolic, cultural and distributive conflicts symbolising regression of human values. Habit of oiling legal tradition of politics can stop interventionist mismanagement of statecraft by means of several strategies:

First, civil society as an epithet of reason and feeling and capable of achieving self-awareness, must instil historical awareness among party leaders of Nepal to transmit the authority of polity in society, respond to changing aspirations of citizens and curb the affinity of politics to personal, family, territorial and privileged interests. Social inclusion can fortify the social base of politics and clog extra-constitutional engagement, extra-parliamentary formation of caucus politics, identity-crazy forces, interest and pressure groups, lobby networks and anti-institutional social movements engaged in a vicious battle for micro-narratives and privileges, not real social transformation.

Second, civil society of Nepal needs to widen the old binary frame of friend and enemy politics for a modern politics of mutual action based on constitutional spirit. This cuts the conditioning of politics by caste, class, ethnic, gender, territorial and ideological factors. It subverts democracy's optimal values of national integration opening an avenue for protest politics. Leaders' forage on deliberation, discipline and feedback in the party committees can simply rear personality cult rising from authority distinction than performance and dilutes inner-party democracy.

Third, the historical crisis in Nepal's reformist politics reflects the flaws of parties to uphold a middle path, act as a mediating agency of society and infuse reforms in each generation of citizens for the rational idiom of political life. Civil societies have to assume duty to limit them to the structure of public law, democratise their ties with their auxiliary bodies and citizens and fortify inter-party dialogue for broad-based consensus on issues of public goods. It revitalises the hopes of moderates in politics.

Fourth, political parties and civil society need to collaborate on justice. It is related to the vision of Nepal's constitution and global humanitarian obligations to the poor. Both need to revive old solidarity with the marginalised. Nepal's parties are poorly institutionalised in terms of subsidiarity, autonomy and complexity of functions. Top leaders loath creative ideas and inconsistent in organisational direction but deftly bargain for power with factions. Civil society need to introduce code-based values to make them sensitive to human rights.

Fifth, modern legitimacy stems from the elections of multi-level governance. Nepali leaders need to link the top with the bottom of society to cope with what Aristotle calls "mean state between self-indulgent and being insensible." Coherence between the initial ideals of parties and the justification of current path ease them to keep historical identity and settle fissiparous inclinations. Their natural formation can free both from borrowed life, open debate in the public sphere about the duty of public power and institutions of modernity and embolden leaders to execute the constitution opening path to political stability, prosperity and happy life. The removal of vacuum in security can expedite democracy, development and peace processes. The civic culture imagines a qualitative change in the cognition, values and attitudes of Nepali leaders to inspire a stake of Nepalis in the constitution.

Sixth, decision-making sphere of Nepal entails rational politics to open up leaders' mind to social learning of the changing nature of citizens' rights in order to inspire their orderly civic participation and initiate social and economic reforms. Deeper institutional transformation leaps from alternative vision and leadership provided by grassroots organisations, local bodies and civil society. A real democratisation of Nepal's political parties, however, requires arresting the collective diversion underway from constitutional goals and behaviour and synergising all the connective forces and resources of social capital for a rational construction of future order. Federations of many civil society groups are moderating the hierarchical system of production, appropriation and control through intermediary associations- free cooperatives, small enterprises, self-help and economic institutions as they do not disregard nature and labour's pain.

But the key propellers of modernity such as education, economy, health, technology, institutions and leadership are not yet fully rationalised in Nepal to achieve the integrity of rule of law. In this context, the civil society-party interface requires getting rid of their own self-ironies and enforcing democratic accountability attached to the public expectation of their duties. Protection of the integrity of Nepali politics from pre-rational actors entails the state to create a just public order above power tussle.

Constellation

Democracy is a national system where the state, economy and citizenship find constellation. It is opened to its citizens but maintains certain closure as democracy demands self-rule, immigration control and nationality. Nepali civil society and political parties must acquire ability for self-control and add to the rejuvenation of polity and society. This liberates their tendency to indulge in pre-modern adversarial practices that creates a society of winners and losers. Modern politics of collective action demands Nepali leaders to overcome inner deficiencies, attain the outer constitutional and universal goals of SDGs and enabling the nation to secure public goods, progress and peace.

### Power of Civic Education

Life's progress rests on continuous learning. Civic education answers the aim of life-enlightenment. It allows attaining self-illumination to combat all kinds of subordination- familial, economic, political, religious, institutional and communicative one and gain guts for freewill. It unchains Nepali citizens from the absurdity of indoctrination, thought control, consent manufacturing and false consciousness and promotes critical faculty of judgment. Conscious citizens exercise dignity and autonomy in public affairs and build good character. Nepalis of all generations need civic education to widen their democratic desire. It transforms passive people into active citizens, aids them to exercise constitutional rights and duties, bridges the gap between knowledge and wisdom, makes social contract implementable and amplifies valid voting turn out in the elections. Nepali state promises a good life to citizens by means of instilling enduring positive values of liberty, justice, tolerance, solidarity and cooperation. The legal equality of Nepalis has created a basis for active citizenship to fight for factual equality vital for the level playing field for civic participation at villages and towns. Both nurture a perspective to envision Nepal from the native viewpoint. It frees the captive minds that deem all ills of the nation to natives by enabling them to learn the utility and limits of universal knowledge to national condition.

Reflective Consciousness: Constitutional knowledge bears positive impacts on the behaviour of citizens enabling them to act constitutionally for ordered life and restyle civic culture worthy of oneness of humanity. Nepali leaders' duty lies in educating the union of the sovereignty of the state and popular sovereignty enabling them to use it. The legitimacy to rule is derived from free and fair election which allows the circulation of new leadership in power inducing rational change. Constitutional ideals provide Nepalis an orientation to reform their condition, assist acculturation to public morale, prepare for civic duties and build their patriotic character without being biased to humanitarian norms. In a multi-cultural society like Nepal, citizens' active engagement in civic associations that dot the free landscape is crucial for social and national integration. Experience about civic virtues animates a balance of the leaders' use of democratic concepts and ordinary citizens' richness in the nation's philosophical, historical, cultural and spatial life. The praxis of civic education in Nepal entails building dynamic relation of politics to citizen's life and cultivating civic sense to beat petty infantilism, eternal paternalism and arrogant juvenility of undemocratic leaders. It critically informs the authority of parents, teachers, leaders and priests instilling a belief that they are neither divine nor finer in wisdom. The use of despotic doctrine of necessity and too much personalisation of authority dehumanise the citizens and reduce them into an object of manipulation like instrumental approach adopted by a few donors to foment social fissures. Both do not train people into citizenship and reveal common ground for conflict mitigation. The noise of foreign meddling without naming the country reflects leaders' fear while handy to bend Nepali citizens' right to know. Public security is vital to vitalise citizens to critique on authorities' dilemma and venture far and wide across the cultures and national geography to symbolically bind all Nepalis in their common identity. It nullifies the challenge to the unity of state by internal forces of class, caste, ethnicity, gender, religion and region and external meddling.

Leaders' reflection on the Nepali society's diversity opens the possibility of feedback from many mini publics and facilitates them to learn from citizens about the way to resolve the practical problems of personal and national life. This can contribute to perfect their inter-subjective skills, broaden cross-cultural understanding and strengthen citizens' attachment with the Nepali state. Improved rationality of political practices can build a national consensus for constitutional, political and policy coherence on the basis of changing popular mandate. The concepts underlined in the structural and normative details of the Constitution, such as directive principles, separation of power, checks and balances, constitutional and human rights, rule of law, autonomy of constitutional bodies and cultural industries, social justice, welfare state, human security for all the citizens etc, animate self-consciousness for ample civic actions—cleaning house, road, toilets, school compound, controlling domestic violence, girl's education, safety of disables, creating green parks, health awareness, deference to elders, education about local government, etc.

Civic education is both a top-down and bottom-up socialization of human nature. This means learning contents should be mutually decided by citizens, teachers and leaders. It is process of mutual liberation, not domestication of citizens and denationalization of experts devoid of national feeling and insight of history. Democratic space exists within the nation-state. If Nepali state loses control over its citizens owing to migration of youth, immigration of outsiders and economic, social and political matters to global forces, the order of democracy shrinks from the shortage of critical masses of change agents of society. This brings an authority crisis for the government and loyalty pattern of citizens to the state, political parties, parliament and the actors of governance imposing problems for the institutionalisation of democracy. Institutions of governance fail to remain inclusive if the vision of building Nepali nation cannot address key challenges: alleviating poverty, job-creation, ending social exclusion, combating political alienation and the ecological, social, gender and inter-generational justice as they are related to the realisation of subjective and objective rights and escape from the fear of natural selection of life.

Civic Competence: A live discourse enables both leaders and citizens to gain self-confidence and know as to how their "negative rights" have inspired the democratic struggle, the state protected the heritage of the nation's syncretic culture and equipped the citizens with the muscle to shape and reminisce the screen of national history that they and their institutions, such as political parties, media, civil society and a host of associations, often reconstituted. Democracy becomes tedious if its discourses freeze in the public mind leaving the leadership free to interpret democracy, progress and peace in their own way, like in Nepal, and foster empirical divisions of society for identity politics at the cost of Nepali identity. A culture of silence, fatalism and apathy of citizens is the lethal sign of democratic debit while eternal stir radicalizes the society and erodes the pivot of democratic centre.

The basic values of citizenship lie in paying tax, becoming law-abiding, remaining vigil and assuming duty in community action. Citizenship, as a member of state, begins with commitment to and respect for state sovereignty and territorial integrity, national language, anthem, national flag and social harmony in a spirit of solidarity among all the members of Nepali state irrespective of religious, racial, linguistic, class, caste and gender diversities. Deep internalisation of the values of citizenship by leaders desists from commercialising public goods under the gyration of post-modernism and market and dialectical materialism. They detach citizens from their legendry bond to land, culture and nature. An ideology-bound education blinds one to native reality. Civic education is a program with sovereign citizens, not passive patients, clients, ignorant, aliens and objects. Democratic construction of citizenship in Nepal is vital to make political system functional. It empowers them to select workable choices that the polity offers and constantly feeds their power of feeling, thinking, working and creating the lot the Nepali nation lives by.

Political Acculturation: Nepalis since old days have created and transmitted their syncretic cultures and values across generations. Now, they seek a solution to the problem of a clash among political rationality of democracy, the economic rationality of the market and ideological rationality of social engineering that deconstructs its culture, national heritage, history and society and makes citizens rootless without any harbour to its ancient roots of enlightenment. If ethical and cultural values that bond Nepalis are torn, all that is left to power fails to represent citizens and the power of leaders becomes disproportional to their representation. In this situation, Nepali state has to ensure regulatory frame that fits to all citizens. Cognitive service to citizens enables them to judge the action of their leaders on the basis of value criteria of democracy, human rights, social justice and peace. They also need a mass of stable civic bodies to ensure a balance in the polity and bind them all, even the minorities, on the policies of Nepali state. Mutual trust and concern for fellow citizens help those who are lacking a bit deep for their dignified life mainly the weakest members. If rights of Nepali citizens remain unenforceable, like party manifestoes, the rational vision of Nepali society spins absolute impossibility. The vital task ahead for Nepali polity is to bridge the rising hiatus in the words and deeds of national preachers, teachers and leaders and saving democracy from its performance crisis. This presupposes the utilisation of the art of civic education to communicate the common heritage of the nation with a larger spectrum of people, detribalise them and enable them to exercise the power of the public for synergy of good governance.

### Civic Role in Development

Nepali society is in a steady process of change. The wind of change impelled by science is altering its social structures and rules of civic engagement. Nepal's Constitution has rewarded an enabling milieu for citizens to engage in development. But the vitality for this needs knowledge, vision, desire and will for mutual aid. Institutional explosion in the nation has offered a non-party option of associational solidarity. It emerged as a revulsion against the burden of swollen political classes and their penchant for patronage. The guiding aim for citizens is to realise constitutional goal of an egalitarian society through distributive justice and detribalisation of social life.
Nepalis across the diverse social and economic strata engage in collective action in matters of public good facilitated by their solidarity. Other approach to engagement is induced from above or catalysed by civic bodies from grassroots as per value change. Human rights recognise personhood of each Nepali beyond the rationality of collective greed. Social and cultural forces exert influence for a new national construction on the basis of their expanded rights and welfare state which can build civic capacity for poverty alleviation, reduce income difference, spur public ventures and nurture full citizenship. Diasporas go beyond national relativism of Nepaliness to widen development view and ignore the roots of native wisdom. In a democracy, self-awakening and action raise adaptive learning to change and reduce the cost of development. Rule-based civic stirrings across the party lines now stand for a common life of citizenship.

The seething cauldron of discontent of revolutionaries is subversive disposed to deconstruct the construction of state-citizens ties for a new way of life. Many constructions of activism in Nepal through the justification of contending development theories and ideologies are unsettling to autonomous development and civic culture built on bonding and bridging trust, common experience and mutual respect. Still, its syncretic culture offers social capital central to share the cost of development. Nepal has notably improved its education, communication, health, banking, market expansion, urbanisation and migration but remains slow in energy, production, trade and social progress vital to graduate to developing nation by 2022 and attain the SDGs.

Anticipation of a robust spell of high and inclusive economic growth needs a firm political course, massive scale of cooperation of private sector and international community and conversion of needs, interest and opinion of ordinary Nepalis into public policy. Mobilisation of local economic and human resources, diversified revenue and tax base to support the levelling up of the poor through distributive justice and voluntary public action against the raid over local common can propel self-directed prosperity, the leitmotif of development.

The Constitution provides the local bodies powers of self-governing polity and its vital elements- inclusiveness, transparency and accountability, if not responsiveness, to ensure good life. Still, the centre defines the rules, maintains vertical control over the provinces and local authorities on legal, financial and technical aspects and holds critical leverage despite peddling autonomy. In a diverse society only by up-scaling civic engagement can cut costs of progress. The right-based discourse in Nepal is politicising them and increasing their engagement in a myriad of institutions and activities of diverse scales- voting, writing letters to the editor, public hearing, deliberation, rallies, protests and social movements to realise their rights. But this discourse has prized lawyers and judges. The claim of rights expresses less concern for public good.

Similarly, the feeling for life has recognised a symbiosis between ecology and economy and production and exchange to meet survival needs. Political parties, civil society, NGOs and social media have improved the feedback between leaders and citizens. But the huge costs of their dependency, debt, delay and skewed or non-action in matters of projects are fading the national immune system. Auditor-General has suggested that INGOs need state regulation as they are running a "shadow economy" violating the laws, hiding huge money in banks and enriching kleptocratic networks thus reducing development indicators.

Though federal parliament has taken holiday from policy making and adopted all the policies originated from international institutions since the onset of modernity and democracy including the latest one the SDGs, local bodies are entrusted with the duty to execute most of them. They are, therefore, engaged in crafting policies, programmes and activities as per their needs, address their problems and steer development gear. Compared to States which act as intermediary, citizens' direct engagement in local bodies is growing. Social inclusion of women, Dalits, minorities and functional interest groups of society in the elected bodies provide scope to participate, interact with authorities, influence policy outcome in matters of selection of projects, allocation of resources, infrastructure development, management of environment and supply of public goods.

Active civic engagement is culturally rooted practices of Nepal. What is innovated is the use of emancipatory democratic values-freedom, scientific knowledge, human rights, civic autonomy, political voice and visibility, solidarity and peace. They tend to ground Nepali institutions in universalistic ethics and unleash social, economic and political energies for democratic development. It prevents the citizens' fall off of economic scale.

Assuming the public good as uppermost priority, development process can bring the state, polity, government, political parties, economic institutions and local bodies closer to citizens, increase their access to resources, institutions and agencies and enforce leaders' accountability. In Nepal, public good and services are supplied by the state's coercive means, private sector market exchange, co-sharing of cooperatives, voluntary mode of NGOs and community based bodies, donors' food for work and direct engagement of citizens which is cost-effective. It bypasses bureaucratic rut as citizens are producers, consumers, stakeholders and beneficiaries of development and, therefore, better state-citizen ties can liberate them from being passive recipients of welfare benefits common in a paternalistic regime.

But the centre-periphery tensions over their jurisdictions limp local bodies as the later are facing barriers to positive civic engagement owing to bigger sizes of Gaon Palikas and municipalities, geographic isolation of certain parts, reluctance of district officers to shift their works to local bodies, under supply of public servants, poor infrastructures, convoluted laws, loss of production dynamism owing to the migration of youths abroad, inexpert local authorities in matters of mediation of interests, conflict between elected and non-elected leaders and corruption of power. In some localities weak market institutions and strewn hamlets hinder the supplies while in remote areas where the presence of government is limited to administration, NGOs, civil society and market fail to meet the hope of citizens for development dividends.

In some localities, returnee migrant workers are doing innovative production works, such as poultry business, bee-keeping, commercial farming, horticulture, milk production, meat supply, organic farming, vegetable production, cash crops like tea, coffee, etc. fulfilling demand-supply equilibrium. Exposure has provided them development insights of other nations. Nepal has also offered training on vocational and technical education to enhance the stock of human resource. But it is insufficient to optimally exploit the nation's endowment. Citizens' practical engagement in labour, production, trade, finance, investment, technology and distribution is primal entailing modern tools to meet competitive standards. Similarly, various projects such as education, health, drinking water, sanitation, agriculture, cooperatives, infrastructures, farming techniques and small scale projects also require upgrading of technology.

Nepali leadership often flips to new theory of development to solve the nation's problems while citizens are socialised in unfilled promises. Each theory proved defunct to capture the diversity and complexity of Nepali society. Development experts, detached from social learning of the empirical reality, ignored the culture, knowledge and experience of citizens and remained insulated from the disturbing effects of their failures. Development rests not only on the use of capital, technology, market and materials but also human ingenuity to utilise what Elinor Ostrom calls "institutional incentive" without crippling the nature's gift for renewal. It optimises the sustainability of development.

Constitutional checks are vital to control obstreperous actors who spoil these conditions and get private benefits at public costs. The leaders need to revamp the failing institutions owing to the corrosive influence of interest groups, partisan politics and weak management, share the burden and benefits of common good with all Nepalis, craft dense associations to cultivate public spirit and common life of citizenship beyond private profit. Dispersal of power is vital but not enough unless they cultivate the virtue and skill, acquire wealth, power and specialisation for engagement in the public sphere of local commons and raise the scale of self-governance.

A strong doze of civic education for Nepalis is vital to liberate them from the Hobbesian trap, engage them in the civilisation process to pacify the wild human nature and uplift the quality of life by building citizens' civic capacity, mobilising local resource and cultivating human scale of community, society, organisation and national affinity crucial to cut the costs of development. 

Source: Dev Raj Dahal, \url{http://therisingnepal.org.np/news/33812}

### Usefulness of Civic Education

The idea of civic education embraces liberation of people from their infantile thinking, belief and behaviour. It transforms them from biological origin to social and then to cosmological spheres awakening them to the fact that different people are not too dissimilar to learn from each other's experience. Civic consciousness, like glow of light, razes all fallacies revealing citizen's life beyond the needs of body and mind. Practically, it publicises them with democratic theories, laws, institutions and rights to match the variation of material condition and duties holding each other accountable for their action. A national curriculum and praxis of education can enable Nepalis to acquire high quality of knowledge, virtues and skill to perform their tasks and build character which is crucial to peacefully engage in the nation's life. Nepalis live in a plurality of life-world who share the common state with the natives, core values of shared humanity with other humans and a common Earth with other living species that define their duties beyond self and attain universal emancipation. Innovated by sages, its past tradition of enlightenment was not elitist in nature. It sought the perfection of good life unblinked by the siren call of fundamentalism, fatalism, unsocialised animal spirit and populism of today.

Enlightenment values, shaped by the knowledge of light, were based on punya karma (good action), not only human reason and will for self-promotion. They enabled Nepalis to engage in improving life, acquire a worldview and gain ability for self-decision. Those devoid of them are dictated by others and cannot be said free citizens able to solve their personal and social problems and contribute to democracy consolidation in Nepal. Astavakra Gita reveals the nature of Atma Gyan (internal vigilance measured as a precondition to liberty) which has the drive to transcend unsocial human nature. Social virtue is essential to bridge the gap between knowledge and wisdom and order and free will and build human affinity enriching good character. For Gautam Buddha, the path to dharma is a way to the bliss of nirvana, a way that frees humans from the karmaic cycle of suffering, impermanence of life and impersonality of phenomenon and link inner self to a single universe. It helped to detribalise Nepali society. Production of good citizens is the core of civic education affirmed by Socrates in the West and Buddha in the East as a source of polity fit for peace. Nepal's Constitution and policy document too stress on civic education for every citizen though the modern path to it is lost in the din of contesting ideologies, strategies and manoeuvres. As a result, Nepali families prefer successful children, no matter good or bad. The ironies of too much rights, few duties and poor material resources have kept Nepali politics into a combustible gear. This shows that modern Nepali leaders need to confront the nation's future with the ideals of the past defined in the spirit of middle path. Science has linked diverse citizens but did not cultivate the civility to achieve common good.

Immanuel Kant says, "Enlightenment is man's emergence from his self-incurred immaturity. Immaturity is the inability to use one's own understanding without the guidance of another." He says that this inability arises from the lack of courage to exercise freedom. He speaks of three-fold tutelage or immaturity: domestic socialisation where children are dependent on the advice of parents and elders every time and follow their judgments in every aspect of life rather than becoming capable of using their intellect and creativity; civil tutelage of law, leaders and teachers whereby adults uncritically imbibe and faithfully obey their orders and authority without questioning their rationality; and life-long stage of pious tutelage of religious scripts and priests that dominate their belief and behaviour. Each sort of tutelage depoliticises citizens, fosters conformist political culture and kills their sovereign power to uphold inquiring mind to solve their problems created by political stasis and inexorable tides of globalisation. Nepalis cannot exercise their sovereignty so long as they are divorced from the native tradition of enlightenment which looks timelessly modern but now vulnerable to the glitter of modernity's materialism which has reduced media, schools, culture, health, nature, security and law into consumer goods undermining the civic health of the nation. The enchantment with the instrumental nature of reason has undermined the emancipatory potential of enlightenment and punished the weaker part of Nepali society.

The praxis of civic education evokes practical ends of human sensibilities. Its reflective awareness on lived experience of citizens enable them to self-legislate, build personal integrity and national character and become stakeholders of democracy. Inability of Nepali leaders to spread civic virtues has converted citizens into an unreflective mass who impulsively jubilate every regime change, democratic or not as they judge it from existential possibility. The critique of this praxis is vital to transform Nepalis of diverse social, economic and geographic origins into equal citizens and hone their talent of judgment. Praxis cannot be governed by only scientific or rationalistic knowledge claims which is deterministic in nature but by choice-based normative, ethical and moral belief. Nepalis have wisdom of 'common sense' judgment of every day practices. On the basis of this, they participate in community, local assemblies, organise public action, engage in the production, circulation and exchange of vital goods and fulfil their needs and rights. A society that displays ample civic virtues can make their milieu clean and do not look down to cleaners as socially inferior in caste, gender or status. It breaks the barriers to work together in pursuit of the nation's overall progress.

Liberty exists in the gift of critical judgment about ideological blindness, demagogic illusion, irrationality and determinism of class, caste, gender, ethnicity and locality. Education divorced from real life experience of citizens disables them to reflect, engage and influence societies, politics, law and economic planning. In this sense, the rebellious spirit of Nepalis against their subordination to organised power is vital to realize the Constitutional vision of "an egalitarian society." Civic competence unfolds the choices to Nepalis to confront, not just during election but in every turn of life and influence vital policies that affect their lives. It fosters a sense of social equality, tolerance of opposing voice, builds trusts and gears up the spirit of public action. Frank Bruni says, "It is dangerous to forget that in a democracy, college is not just about making better engineers but about making better citizens, ones whose eyes have been opened to the sweep of history and spectrum of civilisation." Those suffering from historical amnesia are hardly inspired by the example of ancestors, their temper and spirit of national freedom. John Dewey says that human nature is not fixed. Civic education and associated life determine human conduct. They nurture civic virtues of individuals that underpin justice based social order. De-culturation, by contrast, is alienating. In this context, social modernisation in Nepal supposes freedom for all. Civic culture bears three effects: growth of individuality, freedom of choice and personal accountability in the entire cycle of life. The last one enables Nepalis to cultivate social virtues of trust, tolerance and teamwork to fight endless causes of the nation's marginality.

Nepalis, as public citizens, are members of the state. They are free to take any initiatives not constrained by habit, tradition and arbitrary actions of authorities. Civic learning means not the passive absorption of lifeless facts or passage of text books written in a totally alien context. It is about engaging in active reflection and inquiry on the life-context of Nepalis, their everyday issues, actors and rules to resolve problems. A vibrant democracy requires active citizenship. It brings cognitive and affective attachment to Nepali democracy, nourishes their openness and engages them in the philosophical traditions for the renewal of national order without smothering the capacity of society for reunion and resilience to the changing times. A common citizenship can transcend the primitive differences shaped and institutionalised by political parties, their liberating fronts and the Constitution. It can create a shared condition for judgment about the art and outcome of polity. Rational choice has failed to control egoism entailing the rediscovery of its enlightenment to rebuild the state from the societal muscle.

Enlightened Nepalis have the guts to defy myth and capricious practices and articulate in public sphere. It frees them from inferiority complex denying their human potential and constitutional propensity. The inversion of Nepal's enlightenment tradition through radical political and social engineering has bred gutless elites. Inorganic intellectuals and leaders are disembodied from the episteme of national culture, history and life-world. They hardly learn from their experience. They are weak to indigenise universal knowledge to fit to local conditions. Ideology-driven education creates ignorant person. Consent manufacturing by unfair media while thought-and-body control of legislators during crucial parliamentary voting denied them from using conscience. Artificial awareness makes it easy to abuse mini identities against the Nepali state and wane social cohesion. Undemocratic leaders fear that civic education fades their superior status and, therefore, seek loyal clients and cadres, not free citizens.

Nepali state has made great strides in the domain of social inclusion and increased women's rights—reproduction, equal parental property and authority, non-exploitation, non-discrimination, social inclusion in politics and administration and positive discrimination in education, health, income-generation, security and peace-promotion activities marking towards participatory democracy. Equal rights to grant citizenship to their children in their name, gender budgeting, appreciation of their family work and zero tolerance to violence can widen the civic circle of life and remove the remaining impediments. This means learning of universal standards and laws can change attitude, capacitate Nepali women for leadership and close the gender gaps in vital sphere of life. Promotion of cardinal civic virtues such as moderation, patience and wisdom is vital for a plural society like Nepal for reconciliation of various identities through the delight of new lens. The transmission of these virtues to many generations of citizens can give impetus to those native civic values of giving, charity, healing and safeguarding social peace. It renews social ties, engineers democracy from below and gives Nepalis a chance to converse with power elites in equal terms. This is vital to create a common ground for public action and foster normative consensus over democratic values across the political parties of various hues to repair state-citizenship ties.

Source: Dev Raj Dahal, \url{http://therisingnepal.org.np/news/27235}

## Bureaucracy

German sociologist Max Weber, who used an "ideal type" approach to extrapolate from the real world the central core of features that would characterize the most fully developed bureaucratic form of organization. Weber's ideal type of bureaucracy possesses the following characteristics:

1. The bureaucrates must be free as individuals; they can only be bossed around with respect to the impersonal duties of their offices.
2. The bureaucrats are arranged in a clearly defined hierarchy of offices, the traditional scalar chain wherein every bureaucrat has an unambigious place -- and knows his or her place!
3. The functions of each office are clearly specified in writing.
4. The bureaucrats accept and maintain their appointments freely -- without duress. Slave bureaucrats, while once fashainable in the Ottoman empire and Imperial china, are an inherent contradiction except within military or prison organizations.
5. Appointments to office are made on the basis of technical qualifications, which ideally are substantiated by examinations administered by the appointing authority, a university, or both.
6. The bureaucratis receive money salaries and pensions rights, which reflect the varying levels of the hierarchy. While the bureaucrats are free to leave the organization, they can be removed from their offices only under previously stated, specific circumstances.
7. The office must be the bureaucrat's sole or at least major occupation.
8. A career system is essential; while promotion may be the result of either seniority or merit, it must be premised on the judgment of hierarchial superiors.
9. The bureaucrats do not have property rights to their office nor any personal claim to the resources that go with it.
10. The bureaucrat's conduct must be subject to systematic control and strict discipline.

## New Public Management (NPM)

This concept was put forward in 1990s and aims to overcome the drawbacks facing traditional system of administration. NPM is best understood as a conceptual framework that draws on private-sector notions of disaggregated operations with market-based controls to produce efficient and effective outcomes in the public sector. In the briefest of quips, proponents of NPM contend that government should be run like a business. All of this aims to transformation of public administration to public management, improvement in public services and administration, improvement of administrative capacity and competitiveness.

According to Christopher Pollitt (1993), "the new public management" has four main aspects:

1. A much bolder and larger-scale use of market-like mechanisms for those
parts of the public sector that could not be transferred directly into private
ownership (quasi-markets).

2. Intensified organizational and spatial decentralization of the management
and production of services. 

3. A constant rhetorical emphasis on the need to improve service "quality".

4. An equally relentless insistence that greater attention be given to the wishes of the individual service user/"consumer."

In the final quarter of the twentieth century NPM reforms swept across developed democracies in countries such as New Zealand, Australia, Great Britain and the US. The political prominance of these reforms, combined with resulting changes in management practice makes it tempting to view NPM as the way the work of state gets done today. In practice, public management remains highly varied and much of the old structures of centralizaiton and bureaucratic hierarchy remain firmly entrenched.

Two impetuses that gave rise to a new concept of public management are bureaucratic distrust and public choice theory. Shortcomings of the public management in early days were picked as being grown to unimaginable proportions, far from politically neutral, relied upon formal rules and internal hierarchy to function. An NPM framework moves away from large, discernable public agencies, preferring to engage a host of agencies - public and private - in the delivery of public goods and services. Further, an NPM framework divests itself of the democratically derived mandate for political neutrality by expressing democracy differently, either through the provision of service options (i.e., the people "vote" with their choice) or public participation (i.e., the people "vote" through participation). Finally, an NPM framework leaves behind formality and hierarchy to stress managerial dexterity and entrepreneurialism.

The public choice theory has a long history stretching back more than two centuries. In brief, public choice theory -- or more simply, public choice -- looks to apply economic concepts to political ideas. One of the more influential works on the subject was offered by James M Buchanan and Gordon Tullock in their 1962 The calculus of consent: Logical foundations of constitutional democracy. Two themes undergird the authors' work: methodological individualism and decentralized organizational arrangements. 

For the welfare of public, there should exist alternatives in public sector as well. Only then people can have their choice, much alike the idea private sector thrieves on.

Principal agent theory


It is also termed as principal agent problem. This occurs when one person or entity (the "agent"), is able to make decisions and/or take actions on behalf of, or that impact, another person or entity: the "principal". This dilemma exists in circumstances where agents are motivated to act in their own best interests, which are contrary to those of their principals, and is an example of moral hazard. The principal and agent theory emerged in the 1970s from the combined disciplines of economics and institutional theory.

When it comes to fighting corruption, the 'Principal-Agency Theory' is helpful in discerning the good or the bad, which assumes that corruption is an individual aberration where due to information asymmetry, inadequate monitoring and incentive and the punishment system, the principal cannot detect corruption by his/her agent(s).The solution here is to get rid of one or two rotten apples in order to save the whole lot. There is no problem with this theory as long as its assumptions are correct. 

https://kathmandupost.com/opinion/2016/06/19/corruption-vs-anti-corruption

## Organizational behavior

### Theory X and Theory Y

Psychologists have long been noted for their studies of human and animal motivation. The operative discipline in organizational motivation is philosophy. The sincerity and rigor of the motivation of employees toward their duties is a direct reflection of management's philosophy toward them. That an operative philosophy is neither written down nor formally stated is irrelevant to its existence. Many organizations have commendable formal managerial philosophies. Ream upon ream of paper have been dedicated to espousing an official view of what an idyllic place the jurisdiction is in which to work. The only problem with these fine-sounding philosophies is that they are seldom operational, tending to exist only on paper. 

All managerial philosophies are premised on a set of assumptions about human behavior. Douglas McGregor, through his 1960 book The Human Side of Enterprise, popularized the contending concepts of managerial philosophy with his now famous Theory X and Theory Y sets of assumptions. McGregor hypothesized that a manager's assumptions about human behavior predetermined his administrative style. Because of the dominance of traditional theory in managerial thought, many managers had long accepted and acted on a set of assumptions that are at best true of only a minority of the population. McGregor labeled as Theory X the following assumptions:

1. The average human being has an inherent dislike of work.
2. Most people must be coerced or threatened with punishment to get them to put forth adequate effort.
3. People prefer to be directed and wish to avoid responsibility.

While McGregor's portrait of the modern industrial citizen can be criticized for implying greater pessimism concerning human nature on the part of managers than is perhaps warranted, Theory X is all the more valuable as a memorable theoretical construct because it serves as such a polar opposite of Theory Y, which assumes the following:

The expenditure of physical and mental effort in work is as natural as play or rest.
2. A person will exercise self-direction and self-control in the service of objectives to which he or she is committed.
3. Avoidance of responsibility, lack of ambition, and emphasis on security are generally consequences of experience, not inherent human characteristics.
4. The capacity to exercise a relatively high degree of imagination, ingenuity, and creativity in the solution of organizational problems is widely, not narrowly, distributed in the population. 

Of course, these differing philosophic orientations are extremes for purposes of example. Most work situations would require a mix rather than a simplistic acceptance of either construct. Nevertheless, assumptions shape destiny; they tend to create self-fulfilling prophecies.

## Civil service

The civil service is an institutional mechanism to support the government in executing its tasks and realising the goals of the nation state that have been defined under the Directive Principles and Policies of the State in the Constitution of Nepal, 2072. Creation of a public welfare state is the main political objective, and a public service is to be set up and operated for fulfilling this objective. It specifies that the state shall be oriented towards promoting public welfare by maintaining peace and order in society, making arrangements for the just distribution of economic resources available in the country, and running the development activities for overall socio-economic development in collaboration with the governmental, cooperative and private sectors. The government's major role is to deliver the services to the people according to their demands and needs in a speedy, efficient and effective manner. 

But despite many repeated efforts the civil service still has many weaknesses, including:
- The low level of motivation of civil servants;
- The administrative machinery becoming too much bloated due to the government assuming the whole range of functions affecting people's lives;
- Unnecessary expansion in the number of government agencies and its employees;
- Too many layers in the decision-making process;
- Weak mechanism to make individual officials responsible;
- Adopting new technology effectively remains a challenge;
- Lack of responsiveness towards citizens; and a lack of decentralization of necessary authority to the officials down to the field offices.

Significant structural, managerial and behavioural challenges remain to be addressed, but some positive lessons have already been learned, including:

- Decentralization brings governance closer to the people and helps service delivery and also empowers the people;
- Strengthening local governance has brought many positive changes;
- Local hiring helps to retain the personnel locally and improve a sense of ownership;
- Many local level agencies feel more empowered; and
- Limiting rights to transfer has helped to retain people at decentralised sites

Following features define a civil service:

1. Recruitment based on merit system
2. Legal basis of service offering
3. Permanency and continuity
4. Non-military
5. Careerism and development
6. Pre-defined benefits and facilities
7. Professionalism
8. Job protection and honor
9. Political neutrality and impartiality
10. Mission based
11. Based on hierarchy
12. Dedication to public service

Government's success and people's prosperity relies significantly on effectiveness, efficiency and accountability of the public service sector.

Following points constitute duties of public service sector:

1. Support people elected government
2. Implement programs and policies approved by government
3. Perform as a representative of government
4. Provide assurance of service on behalf of government: Service prioritization; categorization, distribution; Proper and judicious service distribution; Timely service; Monitoring and evaluation.

**Activities under civil services**

In order to establish of peace, harmony and the rule of law in the country, tax collection from the public is ages old tradition that has been as primary mandate of the public service. With progress of time, several other activities have been assigned status of public services. In recent years, with changes in attitude of government towards governance system, reforms have also propagated to public service sphere. Some of the activities currently performed by the service are:

1. Assist government in legislation of people oriented policies.
2. Establish peace and harmony in the country
3. Conduct developmental activities
4. Make use of and utilize properly resources and inputs available in limited amounts in the country.
5. Provide fundamental services necessary to the public in coherent way
6. Promote activities that improve national dignity at international level
7. Perform activities that ensure social justice and social security
8. Launch activities for upliftment of socially backwarded and marginalized groups
9. Collaborate and work jointly with private and non-government sector.

Principles of civil services

1. Principle of equality: Non discrimination, public welfare is all that matters.
2. Principle of neutrality: Politically neutral, only dedicated to clients' welfare.
3. Principle of legal recognition: Implemented as provisioned in law, provisions of punishment upon violation.
4. Principle of continuity: Acts as a form of permanent goverment.
5. Principle of accountability: Towards government, public, programme, goals, colleague.
6. Principle of transparency

Characteristics of Sound Civil Service

- Adjusted to the level of social and economic development in the country;
- Efficient and effective in the delivery of services;
- Highly professional and capable of offering the best technical advice to the democratically elected government;
- Operated according to merit-based principles, combined with culturally sensitive management practices;
- Loyal in the execution of the policies of the acting government, while operating in accordance with the Constitution and the laws;
- Strongly committed to the public interest;
- Disciplined and intolerant of unproductive or unethical behavior;
- Honest and devoted to serving the population in an unbiased and impartial manner;
- Broadly representative of society;
- Upholding fair administrative practices, transparent in its operations and accountable to the citizens and their representatives;
- Willing to hire the best people available at each level, based on a fair and transparent recruitment process and to maintain competitive pay practices that will foster a motivational climate for state employees;
- Strongly in support of proper training and development at all levels;
- Capable and willing to develop partnerships with various groups and organizations in civil society.

### Managing the Civil Service

Central Personnel Agency: The MoGA is the central civil service personnel management agency and is responsible for formulating, implementing and reviewing plans, policies and programs for recruitment, placement, promotion, disciplinary action and retirement of civil servants. It also functions as the Salary and Allowance Review Committee and the Scholarship Committee. It is the central keeper of civil service personnel records through its line department, Nijamati Kitabkhana (DCPR). All ministries, departments, commissions, and central level offices need MoGA approval before taking any action related to civil service management. The ministry is mandated to conduct management audits of government agencies to ensure compliance with the Civil Service Act 1992. The reports are submitted to the OPMCM, which can decide to implement the recommendations. However, in practice, there have been just a few audits, and the MoGA did not have a dedicated management audit section when this study was done in 2010. Further, there have been hardly any decisions on the reports that MoGA has submitted to OPMCM.

Job descriptions: The Civil Service Act 1992 mandates job descriptions stating the specific functions, duties, responsibilities, and authority of each civil service position. Newly appointed employees must receive job descriptions with criteria for performance evaluation along with the letter assigning them to their positions. However, this requirement is rarely observed. Job descriptions that do exist do not have specific and quantifiable performance criteria. These remain a low priority, because the civil service has not traditionally had a performance-based management culture.31Promotions: Civil servants are promoted on the basis of competency evaluations, internal competitive examinations, work performance and experience. Competency is measured through performance evaluations, seniority, educational attainment, service in different geographical regions and training. Separate promotion committees recommend advancement for professional and support level civil servants respectively. There is also a provision for promotion to professional class for employees who have held the same position for at least 15 years. However, despite clear and specific criteria, the promotion system has a number of shortcomings, and in practice, performance evaluations are primarily based on personal judgments rather than objective indicators. Code of conduct: There is a code of ethics that civil servants are required to uphold. The law contains procedures and sanctions to deal with cases of misconduct, including indiscipline and breach of instructions. It restricts employees from being involved in politics, criticizing the government and its policies, receiving gifts, presents, donations and loans, and agitating or participating in strikes. Further, all civil servants are required to accord decent treatment to service recipients and work in an impartial, fair and efficient manner.

Job security: The civil service law is very protective of the rights of government employees. No civil servant can be removed or dismissed from service without a reasonable opportunity for self-defense. No case may be filed without government approval against any civil servant for any act committed while discharging his or her official duty. Where there has been approval, the government defends that civil servant even after the termination of service. The conditions of service are also fully secured: the salary, pension and other benefits allocated to any civil servant at the time of appointment cannot be altered without his or her consent. Any order for departmental action requires prior consultation with the PSC, and there is an Administrative Court to hear appeals against departmental actions.Civil servants retire at 58 years of age, or after three years of service in the case of Chief Secretary and five years in the case of Secretary. Trade unions: Civil servants up to the class three professional level have trade union rights, and authorized unions have the right of collective bargaining. In principle, trade unions are expected to provide constructive suggestions and support the government in making the civil service more effective while safeguarding and promoting employees' rights and interests. The Civil Service Act envisions one authorized trade union. However, there are six different trade unions in the civil service, and only one of them is unaffiliated with any political party.32 All of these unions enjoy trade union rights. The government may consult the authorized trade unions and seek suggestions while preparing bills, rules and regulations. But since the trade unions are allied to political parties, they often focus on parochial interests rather than the interest of the civil service as a whole. Since 2006, employee unions have been accused of undue influences in promotions and transfers, among other things.Individual ministries are responsible for preparing and enforcing job descriptions and managing employees, including maintaining records of disciplinary action. They have the authority to transfer staff within the ministry. They also review organizational structures and positions, and are required to inform the PSC about vacant positions. All ministries must have their salary reports approved by the Nijamati Kitabkhana (DCPR). Likewise, all ministries are required to evaluate the performance of civil servants. While this is the case, human resource management decisions of the ministry are routinely contested by the unions

### Pros

1. Stong judicial basis, consitutional
2. Predefined areas of work
3. Meritocracy
4. Oversight (by Parliamentary committee, Auditor general, CIAA)
5. Permanency
6. Professionalism (groups and hierarchies)
7. Deciplined (code of conduct guidelines)
8. Trade union
9. Social honor
10. Careerism
11. Integrated civil service
12. Inclusive civil service

### Cons

1. Lacking modernizaiton and technology adoption
2. Lack of motivation and low morale
3. Non predictable career path
4. Poor information management system
5. Low confidence among personnels
6. Tendency to seek rights and refrain duties
7. Lack of redress facility
8. Boss oriented culture
9. Vision and norms have not been institutionalized and widely recognized
10. Corruption
11. Lacking human resource development
12. Excessive centralization
13. Impunity and lacking discipline
14. Resistant to change
15. Lack of a comprehensive system of work reporting
16. Work performance is not closely tied to reward and punishment
17. Innovation and inventiveness is not encouraged
18. Reward and punishment as expressed is narrowly associated with transfer and promotion.
19. Malpractices from trade union

### Future direction

1. Competitive civil service

- Privatization
- Service marketization
- Result oriented management
- Agency based and autonomous approach
- Collaboration

2. Enhancement of effectiveness of civil service

- Gender sensitive
- Responsive
- Efficient
- Accountable
- Transparent
- Participative and predictable
- Adequate facility
- Lean structure

3. Working strategy should integrate following

- Challenging
- Comparison with tax payers
- Consult with tax payers
- Competition with private sector
- Contracting out
- Collaboration with people

4. Reform culture

- Responsibility
- Direct accountability
- Continuous reform
- Work directory
- Tele working
- Use of ICT
- Unified civil service
- Out sourcing
- Virtual organization
- Ombudsman (An official appointed to recieve and investigate complaints made by individuals against abuses or capricious acts of public officials)
- Portfolio management
- Information system management

5. Transitional management

- Enforcement of Law and order
- Control of political interference
- Social legitimacy
- Documentation and implementation of inclusive policies and laws

### Integration of civil service in federal system of governance

1. Generate human resource policy.
2. Merit based federal service and inclusive provincial service.
3. Defining of standards for integration.
4. Protection of service rules and facility.
5. Retirement should be voluntary and not be imposed.
6. Promotion of employee while posting from federal employee to local or provincial tier.
7. Transfer or posting should be more on voluntary basis.
8. Increase age limit to provincial personals as opposed to central level personals.
9. Provisioning of voluntary retirement service (VRS) and CRS with appropriate facility
10. Encourage personnel to be motivated to work at provincial levels rather than at central level.
11. While appointing personnel for rural locations, they should be offered higher benefits
12. Proper integration of career builds while having personnel transfer between federal and provincial levels.
13. Gathering of wide national consensus.

### Fixing problems with civil services

1. Legal problems

- Unclear administrative policies
- Poor implementation of existing policies
- Upgradation and continuity of extant policies is poor
- Unstable government policies
- Policies for personal gains

2. Structural problems

- Large organizational structure
- Redundancy of work
- Cumbersome organizational make-up bearing permanent bodies for tenured jobs
- Structures generated without thorough analysis are defunct.
- Excessive centralization
- Lack of power delegation to implementation body
- Unclear directions and regulation mechanism.

3. Implementation problems

- Process orientation, and unclear ways to activities
- Redundancy in execution
- Poor record and book-keeping system
- Complaint management (grievance handling) system is lacking
- Opaque processes to execution
- Lack of proper communication

4. Behavioural problems

- Power orientation
- Rent seeking
- Buck passing (blame game)
- Lack of professionalism
- System does not always ensure right man in the right place
- Status quo
- Self-centric
- Nepotism, cronyism, sycophany
- Negligence and defence approach
- Non-competance and systematic drag down of competent individuals
- Ruler conscience
- Violation of norms and values
- Lack of team spirit and collaboration

```{r}
tribble(
  ~"Roles and functions", ~"Sector",
  "Macro economic Management", "Economic sector",
  "Education", "Social sector",
  "Energy (electricity, alternative energy and others) and drinking water", "Infrastructure sector",
  "Secretarial services to the Council of Ministers, supervision and coordination", "Governance sector",
  "Transportation (road, air, water and railways) construction and management ", "Infrastructure sector",
  "Civil administration selection management", "Governance sector",
  "Agriculture, food security and irrigation", "Economic sector",
  "land reform and management", "Economic sector",
  "Industry", "Economic sector",
  "Commerce and supply management", "Economic sector",
  "Tourism and culture", "Economic sector",
  "Forest and soil-conservation", "Economic sector",
  "Auditing", "Economic sector",
  "Health, nutrition and sanitation", "Social sector",
  "Social security and protection", "Social sector",
  "Inclusiveness (gender, ethnic, religious, regional)", "Social sector",
  "Mobilization of youth force, capacity development, labour and employment, sports", "Social sector",
  "Decentralisation, local self-rule and local development", "Social sector",
  "Information and communication", "Infrastructure sector",
  "Science and technology (Information technology)", "Infrastructure sector",
  "Environment and climate change", "Infrastructure sector",
  "Urban development and rural infrastructure", "Infrastructure sector",
  "Peace and security", "Governance sector",
  "Defence", "Governance sector",
  "Foreign affairs and economic diplomacy", "Governance sector",
  "Law, justice and parliamentary affairs", "Governance sector",
  "Peace, reconstruction and rehabilitation", "Governance sector",
  "Corruption control", "Governance sector",
  "Statistics and planning", "Governance sector") %>% 
  knitr::kable(booktabs = TRUE)
```

### Provision in place for regulation of civil service

1. Constitution of Nepal, 2072 (Through different constitutional organs and committees)
2. Civil service act, 2049 and regulations, 2050
  - Efficient, competent, service oriented and accountable civil service
  - Provisions regarding conduct, discipline and punishments
3. Prevention of corruption act, 2059
4. Good governance act, 2064 and regulations, 2065
  - Delineates the monitoring and evaluation guidelines for leadership of both political and administrative spheres.
5. Right to information act, 2064 and regulations, 3065
  - Transparency
6. Administrative court rules, 2051 [link](http://www.lawcommission.gov.np/en/wp-content/uploads/2018/09/administrative-court-rules-2051-1995.pdf)
7. Legislature parliament, _rajya byabasta tatha susasan samiti_ (State affairs and good governance), Public accounts committee (For a complete listing of various committes under legislative parliament refer to section on Governance)
8. CIAA, Office of auditor general, Public service commission
9. Ministry of federal affairs and general administration
10. Judicial units of various levels
11. Department of Revenue Investigation
12. Administrative court
13. Administrative Restructuring Commission (ARC) (formed 2008)
14. Speaker ( _prabakta_) of respective units
15. Orientation toward New Public Management (NPM) during policy implementation
16. Awareness has improved among service takers
17. Media and civil society have empowered
18. Concept of inclusive administration
19. Adherence to several international laws and policies
20. Employees are encouraged to be service motive while delivering their services

### Public Service in Asia

The Asian identity mirrors a rainbow civilisation, both with respect to its internal diversity and cultural particularity. It is, therefore, difficult to weave different sets of identifications of Asians into a common regional awareness and set policy coherence. The common pattern of similarity in Asia is its political culture. Despite the sign of various modes of governance - traditional, democratic, praetorian and authoritarian - cultural attitudes play a key role in shaping institutions and human conduct. Hindu, Buddhist, Confucian, Islamic and Christian civilisations shape this attitude and give primacy on group interests over ideas of individual rights.

Much of philanthropic giving finds roots in religious feeling of compassion. Pre-religious Vedic scriptures of South Asia elaborate four different types of human conduct: bad conduct hurts others; normal conduct is meant to regulate personal and family lives; virtuous conduct aims to help others, such as establishing resting places, inns, schools, hospitals, well, water spouts and uplifting the poor; and supreme conduct governs Yagna, protecting the earth, heavens and the cosmos for tranquility and peace.

It is binding for the adherents of Islam to give 2.5 per cent of their net yearly earning to religious and social institutions called Zakat (alms), which is distributed to the needy. Buddhism seeks to stop the root of violence within every individual so that better relations with family, neighbours, community and states can be built. Stable peace requires developing compassion and giving support to the needy thamtaan. In both Hindu-Buddhist traditions, property obliges the people to perform paropakar, charity work and social service. Confucian thought, like those of Hindus, confers respect to authority, morality, unity and a sense of duty to serve people within the familial, social and political institutions. The highest ethical standard of Confucian tenet is jen. It means affective concern for the well-being of humanity. Confucianism, like Buddhism, is pacifist and attaches value to the notion of datong, implying that the whole world is one big family, akin to the ancient Hindu wisdom—Basudaiba Kuttumbakam.

The current Chinese leadership feels that Confucianism can fill its spiritual void and compete against Western individualism and consumerism for the creation a harmonious society. The defence of Easternisation containing certain Asian values is a defensive response to a one-sided Westernisation. The base of state-society harmony is precisely the familial values of mutual help which has cut the autonomous organisation of civil society. In Eastern Asia officials monopolise the public realm allowing people liberty to chase self-interests in the private sphere. The Christian practice is better institutionalised in charity. Every devout Christian is obliged to give 10 per cent of his or her income to the Church which is spent on public welfare. The Confucian doctrine and the family-oriented Hindu-Buddhist values oblige people to help relatives in crisis times.

Asian businesspersons value the morals of these traditions and assume corporate social responsibility by establishing trusts, charity society and foundations to fund relief projects and social initiatives. These non-profit associations have contributed to the development of specific laws, regulations and several codes of conduct to govern their lives, fulfil diverse needs of citizens and strengthen civil society's engagement in social responsibilities. Despite the persistence of high religions embedded within the institutions often lending legitimacy to state authority and in the religious practices of everyday life of citizens, even in the modernising countries like Singapore, Thailand and Malaysia capitalism is pervading every sphere of life and rapidly individualising citizens eroding the muscle of society for public action.

All over Asia exchange ties determine power and authority. Even in modern states leaders arbitrate the use of rules so that interpersonal relations can run easily, unhindered by laws. This has limped the growth of strong institutions. Those who occupy the state authorities and served by them remain satisfied with weak institutions and fear the depersonalised laws, institutions and processes. And when leaders fail to perform duties citizens replace the leaders, not changing the rules of the game and institutions. It is the strength of personal and professional ties that enable the growth of social capital and moderation of conflict. The general Asian family values make the social change process elite-driven, trickle-down and incremental overlapping tradition and modernity.

On the one hand citizens are loyal to authority, order and hierarchy and on the other hand there is fast erosion of this culture thus marking a paradigm shift from ritualised, clan-based, rural and customary traditional society to detribalised, de-ruralised, class-based, urban and modern society required by rapidly changing economic fundamentals, rationality and political modernisation. This shift has been propelled by information, education and science, growth of modern industrial hubs and cities and globalisation. As a result, one can see a new public sphere in the making entrenched in the democratic processes. A basic tension exists between consensus and conflict in the operation of Asian politics. Educated citizens are mobile, skilled and cosmopolitan and exercise more choices in public life. The globalisation is exposing the Asian public not only into each other's culture but also to world cultures. Enlightened leaders have, therefore, a general willingness to accommodate sub-groups in a common bond of the state.

Many of the Asian leaders' style can be characterised by patrimonialism. Highly personalised leadership coexists with impersonal laws and political institutions bearing the legacy of the traditional polity in which government is personalised and public administration is the extension of the ruler's patronage. Powerful leader controls the political economy and the personal ties increasing people's personal fortune. It implies a lack of strong social discipline—where cultural norms allow the flouting of rules and laws and indulging in corruption--thus a culture of impunity indicating the problem in the institutionalisation of political parties.

The social and cultural context shapes the performance of political institutions. The competition over institutions among various groups is more about power sharing than reshaping lawful institutions. In the weak states of Asia a fatal gap exists in the political order. Decay of democratic institutions mirrors public apathy in politics ripping cross-cutting bonds of solidarity across political institutions with new forms of identity, ideological, ethnic, regional and communal conflicts and even eroding the civility of society. One area of progress is the filling of gender gap in public life. Here, the works of civil society have proved vital in the transformation of politics based on freedom, gender equality, identity and social justice and release modern politics from primitive clutches.

In Asia, the tradition of public service differs from one society to another depending on the level of modernization and reconciliation between traditional and modern ways to adopt the aspirations of diverse peoples. The Asian style democracy represents personality cult, communitarian culture, patronage and valorisation of the state. In the well-off states like Japan, South Korea, Malaysia and Singapore, a basic harmony exists between the capacity of civil society to create public interests and the ability of states to fulfill them. In poorer parts, political institutions are far behind in social, economic and technological change and political promises of leaders outstrip the state's capacity to deliver, inflicting an atrophy of polity.

In South Asia, argumentative culture upholds a pluralistic worldviews animating civic life despite gross inequality. As culture shapes political development, Asian leaders, confronted with the task of setting up modern states, respond to this by shaping patrimonial forms of power that please their deep mental longing for security, stability and status. This may seem authoritarian but it is a natural response to people's needs. It ensures community solidarity, group loyalties and public service. Asia's accelerating transformation may breed new version of civic culture to avoid class conflict, if not geopolitical.

The modernity's wave enabled the Asian leaders to set up the centrality of bureaucracy, dominant party system and traditional politics. Now, its erosion now gave the birth to coalition politics where diverse interests are represented in the polity offering the mobility of class, caste and gender. Old Asian texts treasure precious ideas for protecting civilized life, liberty and identity beyond social contract. They offer liberal thought on learning, leadership, virtue, just and unjust wars, civility, nobility and common good. In urban nodes the birth of civic coalitions seeks to prevent the fading egalitarian culture while their rural counterparts detest clintalistic nature of winner-takes all political economy.

The East Asian style of trust building is typified by informal, leader-centric and non-interfering in other's affairs. Religions also played a role in it. The Buddhist concept of Sangha and Hindu's Sat Sanga played their role in incubating social capital like the art of association defended by Tocqueville. Each civilisation has produced a distinctive pattern of ties between the state and society and tied people's compliance to order, authority and discipline. In some regions secular nationalism is still the dominant force at the elite level. In others, religious and ethno-nationalism as well as tribal instinct are challenging the state's writ.

The denial of democratic rights to sub-groups of society has forced them to resort to distributional, identity and rights-based struggle as a reaction to the imposition of Western model of state in multi-cultural societies. Persistence of weak mediating structures in society framed in the image of golden mean involving the satisfaction of basic needs, individual freedom in thought, speech and action, decentralised society, avoidance of a complete separation of the sacred and secular in ordinary life and articulation of sustainable progress explain this. This golden means is vital in every sphere of life to establish the condition of middle ground between liberalism and relativism, globalism and localism and cosmopolitan citizenship and nationalism. 

Source: Dev Raj Dahal, \url{http://therisingnepal.org.np/news/28797}

## Public administration in Nepal

Modern public administration has a relatively short history in Nepal. It began taking shape in the early 1950s and has continued to evolve alongside the political changes and upheavals that have marked the period between 1950 and 2010. Politics shape government structure and functions, and the state administration cannot remain insulated from changes in the political landscape. Nepal has had two administrative reform commissions, one in the 1950s and another in the early 1990s. There was no such commission after the political changes of 2006, but it is likely there will be a new one after the new constitution is prepared, because the prospective new federal system will require restructuring public administration.

Nepal has had an essentially parliamentary form of government since 1990, and state structures have been designed to match the needs of the polity. An analysis of government structures and functions suggests that they resemble those in countries with similar forms of government, but with idiosyncrasies and inefficiencies unique to Nepal. For instance, there are over 150 development boards and committees created for various purposes (for expediting projects, for example), and many of them are not even functional. While such boards and committees can often be effective, the fact that so many are inactive suggests that they are redundant, unnecessary, or have outlived their purpose.

The governance act (GGMOA) clearly lays down the standards for government business at all levels. These include taking into account the larger interests of the people and nation, equity and inclusion, rule of law, human rights, transparency, objectivity, accountability and integrity, financial discipline, fairness, efficiency, accessibility, decentralization and devolution, utilization of local resources and participation. These principles, however, are not always evident in practice, as is clear from the reports of different accountability agencies such as the CIAA, NVC, and parliamentary committees, as well as the mass media and other civil society groups. 

In principle, governments exist to serve the people. This ideal informs public policies like the provisions of the GGMOA just enumerated. But ideals expressed in broad terms like 'national interest' are more easily interpreted for political convenience than concrete benchmarks like the *Rashtriya Mool Niti* that existed before the 1990s. Clear and concrete language would better guide decision makers in interpreting the provisions of the GGMOA, and better serve the public interest.

Openness and transparency are vital for accountability, because only an informed public can demand and make use of effective public services. While both openness and transparency are acknowledged public policy objectives, the lack of adequate and timely information remains a major obstacle that continues to prevent people from accessing services effectively or holding public officials accountable for non-performance.

Further, while accountability is a stated government objective, the fact that Cabinet deliberations are still protected by law from public scrutiny or court review suggests that the desire to 'protect' government information is a strong one, and that thoroughgoing accountability will remain hard to achieve. The GNBTR and GNBAR are not public documents in Nepal, while similar laws are readily available on the websites of other parliamentary democracies such as India. While there may be good reasons for protecting the confidentiality of government deliberations until a decision has been reached (though this has not been the practice, particularly since 2006), keeping Cabinet deliberations secret forever represents a mismatch between the ideal of accountability and the actual practice of government in Nepal.

Nepal has constitutional checks and balances on the executive, with bodies that get their mandate from Interim Constitution. Again, in principle, these bodies are independent — for example, appointees cannot be members of political parties — but their independence is questionable, given that political parties nominate the candidates, often based on affiliation rather than competence, a particularly common practice since 2006. Since political parties are not transparent about their membership, and there is no way to check membership roles of parties that do not release them, there is no way to ascertain the party affiliation of appointees. The clause requiring non-partisan nominees is meaningless in the absence of a mechanism to check if nominees hold party membership or not.

## Public participation and participatory governance

Participation is life of democracy and foundation for sustainable development.

Simply, participation indicates involvement. Getting involved in any activities, development action, plans and project in some means is participating. But still, participation is a very vague element in development and vital compound in good Governance. Following definitions are more constructive in defining participation.

"People's Participation is conceptually integrated with re-distribution of economic and political power in the process of development and is therefore subject to value judgment." -- M.R. Ghonemy

Participation is more than simply voting. Participatory processes must entail open dialog and broadly active civic engagement. It requires that individuals have a voice in the decisions that affect them. Processes, not just outcomes are key to this broader interpretation of participation. 

Participation means many things to many people. Perception differs as well: some see participation as a technique to implement better projects, while others see it as a development goal or part of a rights-based development approach. Participatory development means involving people in diagnosing the situation and problems, planning and deciding courses of action, implementing agreed upon tasks, monitoring, evaluating and sharing the benefits as well as responsibilities of joint action.

Participatory development is a process through which stakeholders can influence and share control over development initiatives, and over the decisions and resources that affect themselves. Participation is not something that can be conjured up or created artificially. Participation is a feeling on the part of the people, not just the mechanical act of being called in to take part in discussions.

Participation possess following potentialities:

- Joining together the interest groups both direct and indirect to a convincible point
- Management for sustainability
- Good governance
- Equitable benefit sharing

In order for development projects to reflect positively, following aspects of participation are necessary:

- Bottom Up approaches of planning including well communicated ideas and demand based perception.
- Capacities, culture and empowerment factors relatively adjusted with community and identified stakeholder.
- Equity based benefit sharing mechanism.
- Identified role and responsibilities clearly defining the output and its impact to participating agency.
- Conducive environment and comfortable provisioning.

As participation is founded by democratic and people oriented ruling in the same time it needs a good political culture and conductive environment. "Participation for development and vice versa." Nepalese development process lacks resonance to this particular effect. Supplied participation on blue print approaches has fostered pseudo participation. Pseudoparticipation has resulted in failure of institutions to convey the outputs.

Pseudo-participation is the outcomes to implementation failure of policy, unethical doing from leaders and institutions. These are due to transitional society dominated by bureaucratic inefficiency and corrupted political culture. This arises in small community based multi-sectored project, where actors fail to adopt team work, sharing of common benefit and recognition of value for each other as a outcomes are not realized. 

Poverty is still extreme. Access to capacities building and its use in enterprise and services are less possible to majority of rural people. Although one of the strategic approaches to poverty reduction by the government is "improvement of livelihoods focusing on people centered participatory development", we have still not gotten rid from trickle down development approaches and blue print tendencies. Social taboos are so strong that the effort, indigenous skills attained and people's positive aspirations are consumed out before resonance. Often blamed statement are; People sell votes for foods and cloths, politician are corrupted and make bias policy, bureaucracy holds ruling attitude, community user groups are elite dominated absorbing resources and misused, industry and commerce are pointed to benefiting through quality and quantity difference/curtailing etc. These factors are interdependent and needs some kind of coordinated orchestration for irrational job to commence.

Leaders were never aware and committed for crime free, corruption less and cooperative society, instead they have frequently committed for so in media to gain cheap popularity. Every news paper, radio television and informant talks about power, corruption and praise for money. Continuous exposure to these events has a brainwashing effect on public sentiments.

Participation as a key element in Nepal's Sustainable Development: "Substantial participation creates a more rational basis for local decision-making, increases the transparency and accountability of the administration, and allows better monitoring of resource management." The vital task to be compiled for winning the Nepal struggle for development are,

a) democracy with political culture and stability 
b) Poverty reduction with social and economical reform 
c) National vision orientation with unified and identified outputs. 

Participatory processes (like 'voice', openness and transparency) promote truly successful long term development. These processes do not guarantee success. But an understanding of the centrality of open, transparent and participatory processes in sustainable development helps us to design policies - strategies and processes - that are more likely to lead to long term economic growth, and that reinforce the strengths of the processes themselves. Evidence from several reports clearly depicts that the benefits of participation outweigh the costs. 

To sustain the rule of law, the 'checks' on abuses of the power and influence participation have to be strengthened. There are three elements to achieve this: 

1. Strengthen civil society as a source of countervailing power (political parties, unions, consumer groups, think tanks and other NGOs).
2. Increase transparency by protecting 'right to know' among the citizens
3. Extend citizens' rights to legal recourse, to sue the non-action by the government.

Development requires change. Change is threatening. Participatory processes ensure that these concerns are not only heard, but also addressed; as a result, these processes dissipate much of the resistance to change. Participation is thus essential to effect the systemic change in mindset associated with the development transformation, and to engender policies that make change more acceptable.

At the same time, we need experts to think of various solutions, to present alternatives, and then to discuss them with the poor. We also need to have experts to help understand the larger socioeconomic and historical contexts or the systemic causes of problems, such as poverty, which are not immediately visible to the people The views of the poor and experts must be seen as complementary for development to succeed.

Therefore, open, transparent, and participatory processes are important ingredients in the development transformation - important both for sustainable economic development and for social development that should be viewed as an end in it and as a means to more rapid economic growth. Beside as participation is culture which also transmits; leaders, professionals and institutions must promote ethical and good governed practices to demonstrate the culture of participation to local institutions and community. Mechanism must ensure equity based benefit sharing, ownership among the stakeholder and to best the demand led projects. Meaningful participation can only be expected where people have access to basic needs with quality education and health and for this participation. It is fairly a competitive pursuing of people oriented participation to integrate the strength to improving cycle moving toward goal of sustainable development.
 
Nepal development realm suffers from pseudo-participation. This has affected achieving equity based empowered local communities. Experiences achieved and lessons learned are, so far, least put into trial. In spite of frequent talking about participation of marginalized and good governance, our action are not unlike the three monkey who refuse to see, hear and speak on behalf of misunderstanding, misuse and corruption in participatory mechanism. Great emphasis and highlights on reports has been given to mechanical participation, but had we ever revitalized and reinvented participatory approaches for its effectiveness and efficiency? Fraction on perception and expectation on participation from ground communities and power practitioner are different, practiced in two different ways. Gradually participation in development has now been a show-piece and it continuously loosing value based on outputs. Finally, for participation to succeed, it should not be conceived in terms of technique or even as a goal, but rather in terms of its authenticity.

### The Participatory Governance

Nepal has come along an important and far-reaching process of political development and transformation over the last few years. It has undergone change from a centralised unitary state structures to federal democratic republic to redeem pledge to deliver justice, prosperity and welfare to citizens. The Constitution enacted by the popularly elected Constituent Assembly in 2015 has restructured Nepal into three tiers of government - federal, province and local. And these governments have been assigned with important functional competencies and mandates.

Roles

In the federal architecture of functional assignment, the local governments - rural and urban municipalities - are entrusted with important roles and functions that are mainly related with local policy making, development planning and implementation, school education and health service delivery and local dispute resolution and management. The Local Government Operation Act, 2017 elaborates the functional responsibilities of the local government that range from governing, regulating to management of local development planning and service delivery functions at the local level. At the core of these functional allocations among the three-tier of the government lies the imperatives of the representation and participation of citizens in the government decision making process. This aspect is very much spelt out in the preamble, directive principles of state policy, fundamental rights chapter of the constitution, acts and guidelines issued by the government time and again.

Moreover, the Local Government Operation Act 2017 intends to promote collaborative governance through citizen participation at the local level while the Good Governance Guidelines 2018 issued for enhancing transparency practices at local government posits citizens and CSOs as the key actors to engage with local governments to strengthen social accountability process and practices.

Needless to repeat, citizen participation is indeed a key ingredient in the recipe for democratisation and development. It contributes to increase transparency and ownership of citizens in decision making process. If citizens are involved in the policy development, they will be able to make government officials more accountable and responsive for their actions and decisions. In fact, local and participatory democracy can be institutionalised properly only when a constructive collaboration and engagement between local government institutions and citizens is duly facilitated, fostered and strengthened. It is also the duty of citizen to participate in decision making process at local level because, in one way or another, he or she will be influenced by the policies implemented by the government.

Moreover, citizen participation should not be undertaken under any special circumstances, but should be a part of the everyday life of citizens. It is not a seasonal or occasional event. It is each citizen's right and responsibility to participate not only during the election period but also in times between elections. Citizens need time to learn about an issue and also know how they can influence the decision making process. The government should inform the citizens about policy development as well as their roles. Media plays an important role in citizen participation. Information can be disseminated to the population at large in a very easy and efficient matter through use of the media. Citizen participation in local government decision making process not only enhances legitimacy of local government but also promotes and institutionalises practices on accountability and transparency.

It is also realistic and less transactional for citizens to oversee and monitor the governing and local development planning and implementation practices at the local level. This can help ensure that local resources are properly utilised consistent to local needs and preferences. In addition to it, citizen stakeholders can participate and feed their inputs into local policy formulation which can ensure local authorship and ownership of local laws and bylaws. Though enabling policy and legal framework exist for citizen participation at the local level as enshrined in the constitution of Nepal, and local government law, the proposed national integrity policy and directives issued by the Home Ministry to rein in on the activities of the INGOs and NGOs have, however, given rise to concerns and apprehensions on the intent of the new government in dealing with civil society. Moreover, concerns have been raised in the media regarding the partisan orientation of the civil society from local to the national level that has allegedly diminished their independent role and objectivity.

Moreover, restructuring of local government, carried out prior to local elections in accordance with the recommendations of the Local Level Restructuring Commission (LLRC) held in 2016, has slashed the number of municipalities from around 3,500 to 753 which has increased size of the local governments both in terms of territory and demography. It has also impacted the political density with significant variation in population - representative ratio. Moreover, the local government offices have gone removed from the walking distance of local communities presenting difficulties in contacting the representatives and officials apart from accessing services.

Collaboration

Since the local level elections held last for the rural municipalities (Gaupalikas) and Municipalities (Nagarpalikas), the elected representatives have formulated policies, plans, programmes and projects for the last fiscal year and this year in succession. It is generally found that the local projects are selected and formulated without consultation with the local citizens which can be labelled as pork barrel projects. The projects are generally found being allocated based on the partisan consideration not on the basis of the local needs. Moreover, local tax rates and service fees/tariffs have been revised allegedly without consultation with local community and concerned stakeholders. It is time multi-tiered governments especially the local government should interact with citizens on every issue that impact them in their day to day life. This can give a sense of ownership and motivate citizens to cooperate and collaborate with the activities of the government institutions.

### Virtues of Participatory Democracy

Participatory democracy is rooted in the active engagement of citizens in the public interest associations to attain self-governing goal, improve the condition of life and resolve common problems in the framework of social justice. Its defining values are popular sovereignty, right-based culture, power devolution, justice, human rights and an admission of citizens to the realm of public life. These values create popular space for citizens to build choice, organisation, leadership, knowledge and solidarity beyond the classical frame of party politics. The right to participate is altering the nature of class-based party politics of industrial era socialised by mass media to multi-channel communication and representation of citizens' diverse interests in political power to shape public agenda.

Citizens' conscious participation, based on their own experience and capacity for self-reflection on their living condition, itself is the source of learning about freedom, legitimacy, authority, ownership and mitigation of the problem of collective choice. Participation in multi-functional spheres maximises their power, secures essential needs, rights and legitimate aspirations and enables them to control the effects of their action. The linking of economy to ecological and social policies helps to overcome the pitfalls of representative democracy where rule and domination become prerogative of majority party exercised from top down and less caring for common good of those at the bottom. It has fed grievance, political instability and conflict infecting good life.

Participatory democracy marks citizens' voluntary and innovative zeal in politics and public affairs and its favorable outcome for a just political order. It has built their confidence in cutting the distance between constitutional vision and uneven conditions and preventing the return of demosclerosis. Joshua Cohen speaks to ensure "participation rights" in constitutional dispensation to "guarantee a fair value of these rights" and "produce legislation that encourages a fair distribution according to the difference principle."

The crisis of representative democracy is caused by its inability to fit economy's race to become global. There are other factors: the decay of industrial civilization with the onset of information revolution and its class based format of politics based on capital and labor, friend and foe and their free collective bargaining; surge of new issues such as climate change, interdependence, globalization, technological acceleration, migration etc entailing beyond state-centric basis of cooperation; birth of multi-classes of electorates with the participation of previously marginalized groups such as women, poor, labor and minorities struggling for recognition and renegotiation of a new social contract; decline of efficacy of ideological representation of social classes of multi-cultural society; rise of anti-party, anti-institutional social struggles and global associational revolution enabling Nepalis to reclaim decision making power to determine their destiny. It has offered opportunity for their participation in areas affecting their life, liberty and choice and halt a race to the bottom. The ideological crisis of representative democracy has created a transformative moment for participatory democracy driven by pluralistic politics, knowledge and information revolution. Integration of women, Dalits, Janajatis, Aadibasis, Madhesis, labor, minorities, human rights, ecologists and peace advocates in the representation of knowledge, policy, power and resource sharing has created a new constitutional basis for level playing field for organized groups. Nepali constitution ensures 33 percent representation of women in national and sub-national parliaments and 40 percent in local bodies.

Participatory democracy has enlarged the rights of citizens including self-determination in politics, ecology, culture, social and economic spheres seeking to prevent outside bullying and address the concerns of diverse society. It marked a qualitative shift of citizenship from "original position" of inequality to equal constitutional entitlements of inclusive citizenship and proportional representation. It is shifting the power of citizens previously governed by what John Rawls calls "the veil of ignorance" to entitlement-based enlightenment now. In Nepal, this is supported by inclusive procedures in civic, political and state institutions, change of electoral system, shared and self-rule with legislative, executive, judicial and financial power, social welfare, formation of inclusive commissions for many groups of citizens, etc despite their unequal virtues, intelligence and abilities.

Another shift lies in collective bargaining of interest between the capital and the labor to a spirit of cooperation of the government, private business and social enterprises of Nepali society to support public goals common in pluralistic politics though the distribution of its profits is skewed. Still, another shift is in the culture of bureaucracy from administration to communication, coordination and delivery of public goods and services to Nepalis to make multi-level governance functional. Participatory democracy says David Held "fosters human development, enhances a sense of political efficacy, reduces a sense of estrangement from power centers, nurtures a concern for collective problems and contributes to the formation of an active and knowledgeable citizenry capable of taking a more acute interest in governmental affairs." Active citizens equipped with a sense of duties do not escape from social responsibilities but creatively participate in public affairs, the market, political parties, civil society, the state and global public sphere seeking fulfilling life. The praxis of participatory democracy evolves where citizens enrich their cognitive resources and experience, apply their practical knowledge and exercise their reason in problem solution through public action as per Local Government Operation Act.

Political emancipation of citizens generates civic competence and innovative social capital while human emancipation enables to acquire humanity based on the protection and promotion of human rights as an intrinsic value of participatory democracy. Information revolution has pluralised the governing powers, civic association and activism and entailed the involvement of affected citizens in decision making turning it transparent and accountable. An awareness about mutual vulnerability of singular institutional life is vital for the birth of solidarity of larger public to participate in collective action. The social struggles of non-class Guthis, journalists, human rights, conflict-victims, educational and health institutions, sub-national units, recently freed bonded labor and civil society demanding stakeholders' consultation in law making beyond the partisan bias of parliament reflect the assertion of popular sovereignty assuming moral, democratic and cultural dimensions of politics and a revolt against Nepali legislature for alienating self from policy production, confinement to public resources through patronage networks and treating Nepalis as an object of development intervention.

The right to shape consultative outcome can flatten vertical integration of party structures giving cadres a sense of freedom, equality and autonomy. Nepali leaders can directly communicate and receive feedback from their voters now. Those not in the loops are facing social, gender and intergenerational tensions in the party committees. The ideological solidarity has shifted to value and issues-based networks and bottom up collective action. Democratization of the inner life of Nepali political parties can break the vices of clientalism, paternalism, factionalism and leadership cult bolstering the base of participatory democracy where experiences of citizens enables in transforming their condition.

Nepali citizens have cognitive ability and skill derived from contextual learning, will and opinion forming institutions about participation in political agitation, hunger strike, protests, demonstration, electoral campaigns, social movements and community engagements. They are vital to improve participatory quality and exert democratic control on decision expecting to spur higher scale of freedom, equity and justice. The political efficacy of Nepalis' engagement in many self-chosen, self-managed and solidaristic organizations such as consumer committee, ward assembly, community governance, civil society, NGOs, cooperatives, school, community forestry, irrigation and local bodies' federations etc is self-evident. They acquire experience and resources to influence policies. Local laws oblige citizens, elected leaders and government officials codetermine planning, manage resources, build infrastructures, execute and monitor various scales of projects. Local leaders and civil servants alone cannot self-determine vital matters without public hearing. Still, in a number of cases where civic competence of citizens is low cooptation of influential persons provide them a semblance of ownership in the initiatives and civic action. In other cases, citizens become mere implementers of decisions which mark the continuity of paternalistic, bureaucratic and hierarchic working style antithetical to participatory democracy.

Increased voters' participation in Nepali elections to above 70 per cent is sufficient to legitimise the polity. Those not voting were either abroad, harbored political apathy or lacked civic virtues. Invalid voting turnout of about 3 percent can be attributed to a lack of voters' education on the election system. It is doubtful whether the Nepali voters have full knowledge of the issues, leadership and organisations debated by the media or attend local meetings with leaders vital foWr evolving participant culture. Their final decision to cast ballot is determined more by social characteristics and leader affiliation than rational calculation to hone the virtues of institutional life of democracy. Rational participation requires a strong doze of civic enlightenment in Nepal. It helps to address the rationality of governance performance and nurture the virtues of participatory democracy.

Source: Dev Raj Dahal, \url{http://therisingnepal.org.np/news/32199}

### Active Citzenship in Democracy

Democratic state demands attentive and free citizens. Regular dialogues on basic values, actors, norms and issues of democracy can politicise Nepali citizens and rears active citizenship that can influence public policy. In this sphere, they can shape political will and preference with co-nationals to attain national vision. Citizenship is the membership of an enclosed political community, the state, governed by legitimate monopoly on power, loyalty to which surpasses loyalty to sub-national identities. Active citizens can assume duty to community, society and the state which is crucial to gain equal security, identity and welfare benefits for good life.

Civic virtues

The Constitution of Nepal has laid down 31 rights for citizens - freedom of speech and organization, right to work, social justice, social security, education, health, etc. It ensures their representation and participation in governance and fortifies the belonging to the state. It has stipulated four duties - protection of national sovereignty by being loyal to the state, obedience to law, compulsory service needed by the state and safety of public property. The ability of citizens to pay taxes defines their legal and economic status which is a mark of civic duty. If citizens are around their own selfish interest, they end up tearing the web of rights and duties, taking and giving and privileges and charity and pass rational judgment on social contract. Tax-evasion, corruption, illicit capital flight and national betrayal are crimes that make democracy empty, without funds to fulfill citizens' rights. The maker of modern Nepal, P. N. Shah said, "Bribe givers and takers are the enemies of the nation." His universe, united by a common garland, echoes the Nepali national anthem now. Periodic renewal of citizens' active consent through fair elections refills the civic virtues.

The boundary of Nepali state does not correspond to society which is large. Nepali citizens have moral duties to aliens as the nation is a member of many global regimes and bearer of human rights. It offers them the prospect of post-national citizenship and the resolution of antinomies between nationality and humanity. The Constitution stipulates self-perpetuating unitary idea of sovereignty. Nepalis are nationals, not reduced to determinism of provincial and local domains or vicious hatred of the 'Other'. Their duty lies in creating a national identity out of differences of individual interests, ideologies and biological origins. It creates a shared background for the mitigation of fissures.

Constitutional awareness of Nepalis instills in them with patriotism which is vital to bridge the chasm of loyalty at multi-actor governance even to the global spheres where many citizens earn their living. Hope of progress in Nepal can return if group-differentiated citizenship does not cut the ability of the nation's minorities to fulfill their special needs, become equal citizen, protect their cultural identity and nurse national feeling. Group-differentiated citizenship has the potential to re-tribalise the nation. It is averse to modernity couched in terms of individual and human rights and tramples its moral ethos of respecting outsiders and giving them justice like a post-Westphalian state. The Nepaliness is defined in terms of their stake in the state.

Active citizenship is a condition for a vibrant democracy. Citizenship, as a political character, is linked with the public good. Progress, aided by Nepal's wisdom and way of life, evolved national character and dignified freedom. Democracy has granted rights to get passports to join in the global labour market, some even in unsafe jobs. But it did not build a worldwide public sphere where Nepalis can control decisions though they hugely contribute to national and global economic progress. The participation of ordinary citizens in the internal economic opportunity is waning. Monopoly, syndicate and powerful organisational leadership, less accountable to Nepalis, stymie the national productivity and rescue them for free will. Only monied classes, diaspora and globally mobile elites reap the benefits from the new economy. Globalisation has spurred internal democratic deficits because decisions do not require the "consent of Nepali people" making mockery of popular sovereignty, global ethics and universal human rights culture that marked a transition from the state of nature to civil society, constitutional rule, international law and cosmopolitan ethics. Growth of civic virtues rooted in constitutionalism can enable Nepali citizens to strategise for the future through active learning about tolerant politics and actively participate in rebuilding of the community, society and the state.

Social Learning: Social learning enables one to know that they are citizens of a state, universal humans and species beings and have parallel duties at the level of constitutional state, jus gentium (inter-state law) and cosmopolitan norms and solidarity. Nepal's citizenship rights are rooted in both jus sanguine, hereditary and jus soli, birth origin. Democratic spirit transforms passive Nepalis into active citizens at various spheres. It equips them with cognition, skill, disposition and full maturity, jagrit manushya. Political education opens a new possibility for cooperative action across the nation's populace of 125 caste and ethnic groups, 123 languages and 10 religions each with its own values and interests. It eggs on them to reflect seriously on their living condition, acquire wisdom and repair the ethical capital for societal cohesion.

Better ethical capital means less chance of misuse of inhuman ethnic, class, caste, gender, or tribal determinism. It greases shared norms, values and institutions that make common life of citizenship possible. Active citizenship is governed by personal integrity and responsibility about common good. Praxis of socialisation on public education, works and resolution of collision between public demands and budget deficits enable them to harness and use land, labour, capital, technology and knowledge depending on their relative benefits. Passive citizens are satisfied with their representation in politics while active ones are participants of the polity. In Nepal, increased political awareness, social inclusion, proportional representation and subsidiarity opened new domain of citizen activism. Social and economic citizenship through membership in self-chosen association, union, civil society, cooperative, movement, network and federation in development infrastructures, school management, drinking water, consumer groups, health and sanitation, irrigation, cooperatives, clubs, etc. have increased their civic competence. Nepal's local governance has introduced public hearing in project selection and budget allocation, social audit, information and public consultation making the governance responsive.

Citizen-Driven Politics: Nepal's constitution has reconciled popular and state sovereignty and set their authority over other actors. In this context personalisation and centralisation of authority by leadership is immoral. Leaders need to strengthen the parliament and its committees and a culture of deliberation in the parties, before taking any decision. It can set up democratic procedures. Recent national, provincial and local elections have ignited hope for citizen-driven politics keeping the spirit of democracy alive. It seeks to address the needs of citizens from public security, rule of law, health, education, production, market, voice, etc. and bring the state closer to communities. Citizens' active participation in planning, budgeting, monitoring, evaluation and social mobilisation enables the governance actors to fulfil institutional duties and meet the needs of social justice groups. It brings the private citizens into political sphere and creates an interface between them and the state. Citizen's Charter refers to their rights and duties, aims to improve public service delivery and fosters responsive governance. Nepal has enlarged the domain of women's rights in many areas. The right to grant citizenship in mother's name, effective gender budgeting, appreciation of their family work, zero tolerance to violence, 33 per cent representation of women in parliament, speaker or deputy speaker of the House, president or vice-president and chairman or vice chair of local bodies are important steps to liberate them from patriarchal culture, its discipline and rationalisation.

Innovators

Supply roles of the Nepali polity enable citizens to gain access to public goods. The state receives necessary feedback to sustain democratic dynamic. The new civil society with the feeling of "we" are providing hope for citizens and spurring the need for institutional reforms. Emerging from diverse geographical, social and economic backgrounds, the youth, workers, ordinary citizens, local entrepreneurs and cooperatives are giving support to citizens' needs. These democratic innovators aim to change Nepali politics, economy and society. Without active citizenship, Nepalis will be clientalised to families, lineage groups, friends and leaders. They cannot combat misery and fatalism by transforming passive people into active Nepali citizen. The international community can add by showing relevant experience, financial support, resources, tools and technical skills for citizenship-building beyond instrumental and empirical approach. But the local society must learn to update archaic infrastructure, apply practical standards and ensure that scarce public resources are visibly utilised to contribute to a self-regulating, adaptable and resilient democratic communities to combat natural risks and social vices rejuvenating democratic political culture and harnessing the potential of national economy.

Source: Dev Raj Dahal, \url{http://therisingnepal.org.np/news/22420}

### Role of private sector in nation building/Nature of private sector

MESI ko CC PAPER

- Market oriented
- Efficiency
- Strategic
- Innovative
- Competitive
- Customer driven
- Profit oriented
- Accountability
- Performance management
- Economy
- Result oriented

Private sector in 15th plan

- Play a major role in overall development process
- Contribute to national development in the way envisaged by the constitution of Nepal.
- Play a role as supportive actor in achievement of the objectives related to government's vision, policy, goals and plan.
- Create investment friendly environment
- Collaborate with government and public to form PPP
- Enhance cometitiveness of production as well as service sector
- Creation of jobs and entreprenurial opportunities
- Build human capital and physical infrastructure
- Generate revenue for government

The role of the private sector in accelerating the pace of economic growth in Nepal started afterthe newly formed democratic government adopted liberalized economic policies in 1990. Earlier, the government itself operated major large industries in the country and patronized few business houses from the private sector. Over the last 25 years, the private sector has thrived in Nepal; the GDP has increased to US\$ 19. 40 billion in 2013 from US\$ 3.6 billion in 1991; the dependence on agriculture for employment has come down to 74\% in 2008 from 83\% in 1991.

The unemployment rate of Nepal is 46\% in 2012. The employment in Nepal is 54\% with the contribution of private sector in only about 25\%. For the reference, 70\% of employment is from Agriculture sector. As per World Bank report - 2011, the private sector contributes 16.34\% of GDP of Nepal. Agriculture remains Nepal's principal economic activity, employing 70\% of the population and providing 37\% of GDP. Agriculture in Nepal is mostly at the subsistence level. However, there is a vast potentiality to commercialise it through modernization and commercialization of comparative advantage products. Government of Nepal (GoN) is committed to promote private sector for national economic development. In this context, GoN has promulgated Industrial Policy 2010 with the purpose of facilitating investment by private sector in the industrial sector. For the first time GoN has given high priority in the policy to Micro-Enterprise Sector as the means of poverty alleviation targeting the Women, Poor and Excluded in rural areas and later linking them with small, medium and large enterprises for their scale up. GoN has also drafted the Industrial Enterprise Act and its By Laws that is in the process of approval by the newly formed parliament. Likewise, other several policies and acts are drafted and in the process of approval such as Special Economic Zones Act, Foreign Direct Investment Act, Investment Policy, Intellectual Property Act, revision in Company Act, Nepal Standard Act, National Trade Integrated Strategy, etc. Likewise, GoN has created several funds such as Micro, Cottage and Small Industry Fund, Women Entrepreneurs Development Fund and Technology Transfer Fund to provide direct support to Micro, Cottage and Small Enterprises (MSEs). All these initiatives taken by the government are for the purpose of creating conducive environment for economic development through enterprise development. In spite of having tremendous potentialities of enterprise development based on locally available resources (raw materials) from community forest, agriculture, mines, water, etc. the enterprise sector development in Nepal has been sluggish in the absence of low and mid level trained manpower in the specialised area. Realising the fact GoN has established Council for Technical Education and Vocational Training (CTEVT) to produce such manpower needed in different enterprise sector. CTEVT runs several courses in about 21 Training Centres and more than 500 private training centres are affiliated to CTEVT that are running different vocational training course.

After adopting economic liberalization policy and entry into World Trade Organisation (WTO), a fast and significant development has been observed in financial service sector. Before adopting economic liberalization, GoN had only three government owned Banks but now about 247financial institutions2 such as Private Banks, Finance Companies, Micro-Finance Institutes, Financial NGOs (FINGOs) and Co-Operatives which are affiliated to the central Bank and operating financial transaction all over the country. Therefore, there is no dearth of access to finance opportunities for micro and small enterprises. However, still there are policy issues to be addressed to resolve the problems of easy access of credits by micro and small enterprises.

Apart from Government and donor agencies many NGOs and INGOs are also working in Nepal for the promotion of micro and small enterprises focusing to strengthen the private sector for employment creation and for economic growth of Nepal. In order to do so, I/N/GOs have focused on upgrading the occupational skills of the youth of Nepal based on curriculums of Council of Technical Education and Vocational Training (CTEVT) and others. About 3% women and men have been upgraded skills in last one decade. To accelerate private sector, the government has also adopted policies to provide facilities to the financial institutions in the rural areas and has created a number of initiatives to support such initiatives. Numerous of them are youth self-employment fund, Rural Self Reliance Fund, Fund for Women Entrepreneurs, etc. The government has formed Investment Board to facilitate the large scale projects in the country. Recent government efforts and policies have increasingly recognized the role of private sector for employment generation and to accelerate economic growth.

Nepal is one of the poorest countries in the world, but the demand for infrastructure and capital investment is enormously high. 143 out of 144, second last from 144 countries surveyed, this is the rank of Nepal in Infrastructure as given by the Global Competitiveness Report 2012-2013 produced by the World Economic forum. Therefore, Government of Nepal (GON) has given high priority on PPPs for infrastructure development in all sectors, for instance road, transport, energy, communication, agriculture, tourism industry, production industries, etc. In order to strengthen and sustain the PPPs strategy and lure investment in Nepal, Government established 'Investment Board Nepal' in November, 2011as a one window solution for foreign investors.

Public-Private Partnerships (PPPs) is an innovative option used for delivering a public service. It is basically a contractual agreement between a public entity and a private entity, whereby the private entity performs part of a government organization's service delivery functions, and assumes the associated risks for a specified period of time. In return, the private entity receives a benefit/financial remuneration according to predefined performance criteria. The government enters into PPPs to attract private capital investment; to increase efficiency and use available resources more effectively; and to reform sectors through reallocation of roles, incentives, and accountability.The Government of Nepal has recognized PPPs as an important approach for development but country has not been able to implement PPP infrastructure project in a substantial way. Lack of clear-cut vision, political transitions, unstable government and social transformations have been the current and future constraints for the PPP approaches to be executed.

## Accountability

Accountability is assured when the jobs and responsibilities of public officials are clearly defined and the public is informed. The specification of duties and responsibilities is different at the political and bureaucratic levels.

### Political accountability

Generally most political jobs lack the kind of performance requirements that allow for strict, formal accountability. The constitution says that Prime Minister and the ministers are collectively accountable to parliament, and ministers are accountable to parliament and the Prime Minister. In practice, however, the term 'accountable' seems to imply winning the support of parliament rather than satisfying certain performance requirements, because the Prime Minister can keep his job (all prime ministers so far have been male) as long as he enjoys the confidence or political support of parliament, and ministers remain in office as long as the Prime Minister wants them in office. A general election is the only time when political appointees have to face the public for a verdict on their performance. Accountability remains elusive in Nepal's parliamentary system, where the separation of legislative and executive powers is not as clear as it generally is in presidential systems. In a parliamentary system, the executive provides political leadership to parliament and also serves as its agent in executing decisions. This arrangement makes it difficult for either side to clearly delineate its tasks. It is important, therefore, for the government to develop parliamentary practices and conventions to ensure accountability in the conduct of government business, particularly the implementation of policies and priorities. The accountability of the political leadership that has been evolved is based on the following conventions:

1. The government announces new policies and programs in the President's anuual address to parliament, and seeks their approval.
2. The government reviews its activities, income and expenditure for the year in the annual budget speech, and seeks parliamentary approval.
3. Ministers respond to enquiries from parliamentary committees.
4. It is the government's duty to fulfill the committments ministers make to the parliament.
5. Ministers are obligated to give sincere and honest answers to questions from memebers of parliament.
6. The government is expected to respond to the issues raised in the parliamentary debates.
7. The government must respect parliamentary resolutions.

### Administrative accoutability

Accountability is more straightforward in the bureaucracy, as there are job descriptions (albeit not uniformy used), performance benchmarks and formal procedures for delegating the authority, duties and responsibilities of the Chief Secretary, secretaries, and heads of departments. In some cases, the law also defines the duties and responsibilities of lower level officials such as regional administrators, chief of province level and local level offices.

### Mechanisms for ensuring accoutability

Vertical and horizontal accountability: There are built-in mechanisms for both vertical and horizontal accountability. Vertically, regional administrators, CDOs and LDOs are responsible for supervising agencies within their authority. Heads of agencies are responsible for supervising their subordinates. At the district, ministry and chief secretary levels, there are committees to monitor service delivery.Horizontally, the Auditor General's office can send auditors to check accounts and review physical documentation. The PSC audits personnel functions, and the Financial Comptroller General's Office (FCGO) sends its accounts officers to examine the books at different offices. Various institutions are mandated to promote accountability. The National Vigilance Center (NVC), established in 2002 under the direct supervision and control of the Prime Minister in accordance with the Control of Corruption Act 2002, is vigilant about integrity and good governance. The CIAA has authority to make inquiries or investigate matters where it believes there are shortcomings. The GGMOA requires public service providers to hold public hearings once every six months to make the public servants sensitive to public interests, choices and conveniences.Generally, accountability can be breached in three ways. First is failure to meet job requirements in terms of targets, standards, quality, efficiency, etc. Evaluation is usually done by the supervisor, and sanctions are usually departmental actions as provided by the civil service law and regulations. At the political level, steps may be taken by the head of government, the political party, or parliament, depending on the seriousness and political implications of the breach. However, formal actions taken against politicians are rare. The second type of breach includes violation of laws or procedures, financial irregularities, unlawful gains, and unfair dealings. These are addressed in the same manner for both political and administrative personnel under the applicable laws (the Financial Procedure Act 1998 and Rules 2007, the Public Procurement Act 2006, and the Control of Corruption Act 2002). The third type of sanction stems from citizens' right to redress of grievances and compensation for damages, as provided by the GGMOA. An errant official can be individually liable for damages and compensation. The Right to Information Act 2007 and Regulations allow citizens to seek information from the government to build cases for compensation.

### Oversight institutions

The National Vigilance Center (NVC): The NVC was established in 2002 under the direct supervision and control of the Prime Minister in accordance with the Control of Corruption Act 2002. The task of the NVC is to promote integrity and good governance. It is empowered to monitor the incomes and individual property statements of public servants, and it has the authority to accept information on the shortcomings of government agencies and officials, and suspicions of irregularity or corruption. Commission for Investigation of Abuse of Authority (CIAA): The CIAA is a constitutional body with the authority to investigate abuse of authority and unlawful gains by public officials, including political appointees. If the CIAA finds that a person is abusing their public office, it can file a case in court, or recommend departmental action by the relevant authority. Another very important function of the CIAA is to advise the government on policies and laws to curb corruption. CIAA commissioners, including the chief commissioner, are appointed by the President on the recommendation of the Constitutional Council for a period of six years. They retire at the end of their term, or at 65 years of age, whichever comes first. The CIAA submits an annual report to the President, who sends it to the legislature.State Management Committee (SMC) and Public Accounts Committee (PAC): Parliamentary committees such as the SMC and the PAC, constituted under the Rules of Procedures of the Legislature-Parliament, can enquire into any issue of financial or management irregularity, unfairness or inefficiency. The committees can summon ministers or civil servants to committee hearings for questioning.

Auditor General: The Auditor General audits the finances of all government offices, including the judiciary and the legislature. These audits are required for all public agencies. The Auditor General is a constitutional appointee recommended by the Constitutional Council, who serves for a period of six years, or until reaching the retirement age of 65. The Auditor General submits its annual report to the President, who sends it to the legislature. The PAC discusses the report on behalf of the parliament.

#### Office of the Auditor General

The Office of the Auditor General (OAG) was established with the appointment of the first Auditor General as per the Constitution of the Kingdom of Nepal, 1958. Prior to establishment of OAG, the institution named Kumari Chowk Adda used to review the government accounts. It is assumed that Kumari Chowk was established in the year 1769 with the aim of strengthening administrative system after the unification of Nepal. It is also anticipated that there was also existence of audit institution prior to unification.

Vision, Mission and Core Values

Vision: "An Independent, efficient and effective audit institution to promote good governance."
Mission: "Provide quality audit service to the nation for the efficient management of public resources."
Core Values: "Independence, Integrity, Credibility, Professionalism and Serving the Public Interest."

Constitutional Council

According to the Interim Constitution 2063(2007), on the recommendation of the Constitutional Council*, the President appoints the Auditor General. The functions, duties and powers of the Auditor General are specified in the Interim Constitution. (Part12).

Constitutional Council consists of the following Chairman and Member:

```{r}
tribble(
  ~"Designation", ~"Membership", 
  "The Prime Minister", "Chairman", 
  "The Chief Justice", "Member",
  "The Speaker of the Legislative Parliament", "Member",
  "Three Ministers designated by the Prime Minister", "Member"
) %>% 
  knitr::kable(booktabs = TRUE)
```


The Constitution of the Kingdom of Nepal, 2015 (1959 A.D.) clearly described about appointment and scope of functions and duties of an auditor general. Part-8, Article-61 endows the explicit authority to the King to appoint an Auditor-General. Although, the official shall not, except on the same grounds and in the same manner as a judge of the Supreme Court, be removed from his office prior to the expiry of his term of office. Likewise, the Article-62 informs that accounts and ledgers of all Departments and Offices of the Government and of all courts as well as the offices of the Cabinet, the Secretary of the General Assembly, the Secretary of the House of Representative, the Supreme Court and the Public Service Commission, be kept in such a format as may be made with the approval of His Majesty's Government and in such manner as may be prescribed by the Auditor-General and shall be audited by the Auditor-General and shall be audited by the Auditor-General; and the Auditor-General and his deputies shall, at all times, be entitled to have access to all documents relating to such accounts.

In the Constitution of Nepal, 2015, provisions regarding appointment, qualification, term of office, jurisdiction and functions, duties and powers are defined in the Part 22.

Audit objectives and policies of the Office of the Auditor General are:

- To promote public accountability and transparency
- To bring forth economy in the mobilization of public resources and enhance efficiency
- To enhance effectiveness of public entities
- To improve and assure clean practices in the working system of administrative, financial and managerial system;
- To assist in compliance of the existing laws;
- To encourage abandonment of the discretionary work-style
- To recommend practical suggestions for improvement by identifying weaknesses and lacunae in the existing approaches, processes, practices and legal provisions; and
- To encourage the practice of taking actions against delinquents and rewarding the excellent performers.
