## Write an essay in about 800 to 1000 words in following issues:

1. We live in a world besieged by media and technology. Every sector of our life is impacted by the presence of modern forms of communication system. Write an essay discusses the influence of modern forms of communications on everyday life.

\begin{solution}

The very nature of human being as a social creature underlines the importance of communication in one's day-to-day life. Human beings interact, possibly much effectively and efficiently than any other life forms in this planet. Much of the achievements we are relishing today, from kitchen utensils to high performance computers are all, in a way, fruits of reliable communication. This is to say that innovations are triggered in several ways through communication. It is also true that modern day communication, we use in the form of media and technology, have been made possible by the feats of human innovation -- once again guided by state of the art communication systems.

Early developments in communication technology consisted of invention of telegram and telephone, which extensively relied on electrical signals on copper wire. As the search for better communication infrastructure continued, data (including video and audio) sharing over wireless frequencies and optical cables are on upsurge today. To recapitulate the role of communication channel, we can refer to an event that is purported to have put an end to the World War II. Invention of cipher (a tool that decrypts hidden messages) on those days became a landmark in intercepting German radio communication channels and gaining the information about overhanging threat military assault. In the retrospect, communication development during World War might not seem as much of a feat when we consider high performing broadband facilites that are at our disposal today.

Communication system today pervades all spheres of life be it physical or socio-economic. It wasn't uncommon, just a few decades earlier, for people to commute great deal of distance just for the sake delivering information through postal services. Proliferation of telecommunication services in every major inhabitated part of world has made it non-essential for the paper mail services to physically transport tremendous amounts of exchanges people are making. This trend in frequent exchange of ideas and even general information among peoples -- families, friend, scholars, researchers, etc., is expected to only expand in future. The imperative is that, as means of communication gets cheaper and accessible, people need not be constrained to utilize information wealth due to their physical distance.

More often than not, means of communication exert their importance at times of crisis. Remote areas are remote for more than a reason but one of those consistently includes them not being able to communicate with authorities on time. This is a situation that typically results in greater marginalization, as individuals from distant localities lacking proper communication methods are increasingly unheard of their problems and scarcities. For example, when a disease outbreak occurs in such an area, relief package and medical interventions become delayed, resulting in significant losses of life. In contrast, well-connected areas where government can quickly initiate a rescue operation will have more lives saved.

When looked at for side-effects on an individual, communication affects even the mundane aspects of life. Citizens of 21 century take toolsets of communication for granted. There is little to how every day engagements of families are ran without crediting to modern communication technologies. Grocery stores and dairy shops have the communication network set to have knowledge of market; A dairy handler has to know who are his customers for what products. Daily updates on traffic and weather to work are all televised through various news channels so that commuters can be punctual to their work. Furthermore, in the workplace, all commercial startups communicate over telephone or email to their clients. This, in comparison to early days approach of door to door visit for sales promotion and order acquistion, is a stark gain in efficiency and scale. A supply order could be made over a phone call while also provisioning for accumulated order the previous day for the same household equipment to areas nearby thereby significantly cutting down the cost of repeated transportation.

Our social thread forms a web of close circle of friends and families along with tendencies to keep-on-touch with them. This emotional and social sphere of our lives too is strung more tightly together than ever before with the advent of new communication technologies. Rashid can now ask what the weather is like in Australia to one of his friends from Nepal in a video call, while also hoping some days to see them there in person. His mother can make a quick appointment with a doctor in a phone call after knowing that Birat is ill. In addition to surfing a personalized news-stand online right before the meeting, one is now also able to address a meeting which includes a mass of 50 people. This is made possible due to wide projection and affordable cinema technology which allow people to conduct presentations as well as entertain with movie.

Besides trivial occurances of everyday life, various channels of communication are used by government organizations to address public and state their vision for country. Likewise, several government insitutions have made social media formum also a platform for obtaining public feedback and outcries, while also attempting to address them. For example, if suspected for a criminal activity, goverment can call for a surveillance of an individual or a group through the use of closed circuit cameras (CCTV) or the GPS system of suspected's smartphone. Although for the case it is clear that criminal threat outweights the expectation that government action hurts individual privacy, there can be scenarios where this application of technology can be subjected to debate. Meanwhile, there is always a chance that political groups may misuse media to propagandize their statements to only gain popularity, while masking up facts that give different flavor.

With all the merits it has brought to individual and societal lives, modern communication technologies can well be thought of as integral part of development. While it has connected people for good, indiscriminate retrieval and exchange of some information can have detrimental consequences -- With great power, comes a greater responsibility.

\end{solution}

2. Write an essay that discusses how different cultures maintain their distinctive tradition in your community. In what ways does this cultural diversity have a positive impact on your community ?

3. Diplomacy has been defined as "the management of relations between independent states by the process of negotiation". Write an essay citing some major events of modern international diplomacy that have resolved conflict through the process of negotiation and have shaped the world in which we live.

## International relations

1. Nepal and United Kingdom have celebrated 200 years of their relationship. How do you view this relationship and what recommendations, if you have, you wish to offer to take the subsisting relationship to the next level ?

\begin{solution}

Nepal established diplomatic relations with the United Kingdom in 1816. Treaty of Friendship between Great Britain and Nepal was signed in 1923 which further formalized relations between the two countries. Ever since the establishment of their diplomatic relations, friendship, mutual understanding, cordiality, cooperation and respect for each other's national interests and aspirations have characterized relationship between the two countries. The UK is also the first country in the world with which Nepal had established diplomatic relations and also the first to have established its Embassy in Kathmandu, the capital of Nepal. This is the country where Nepal had established its first diplomatic mission (Legation) back in 1934. It was elevated to the Ambassador level in 1947 AD.

The UK remained one of the top development partners of Nepal with the annual British aid on an increasing trend. Tourism, trade, education, and British Gurkha connection remained the key dimensions of the bilateral relations.

Since then, relations between the two countries have continued to grow, with a new Treay of Perpetual Peace and Friendship signed in 1950 which expanded areas of cooperation and an exchange of State Visits. Amicable relations continue today; Nepal continues to be the source of recruitment of Gurkha soldiers into the British army -- a tradition dating back to the nineteeth century but still an essential part of Britain's modern army -- and the UK remains one of the most significant providers of development assistance to Nepal.

As one of Great Britain's allies during the two world wars, Nepalese soldiers in hundreds of thousands fought and sacrificed their lives in many battlefields of the world, thus leaving a heritage of deep and sincere friendship in the history of two countries. The Gurkha soldiers are the most visible bridge between Nepal and the United Kingdom. It is to state that the Gurkha's service in the British army started on April 24, 1815, which is continuing till the date spanning over more than two hundred years. The solid foundation of the relation is built on the history of service, sacrifice and bravery of these Gurkha soldiers.

Nepal and United Kingdom are celebrating the bicentenary of bilateral relations between two countries. To mark the 200th year of the beginning of official contacts between the two countries in 1816, a number of programs and events from both sides have been organized in 2016-2017. There have been intermittent to and fro high level visits between Nepal and United Kindom since as early as 1960.

\textbf{Treaties/Memorandum of Understandings/Agreements between the UK and Nepal}

Bilateral consultation mechanism

With a view to enhancing the friendly cooperation and further promoting the bilateral relations Nepal and the United Kingdom have signed a MoU on 7th January, 2014. On behalf of Nepal and UK, Foreign secretary Mr. Arjun Bahadur Thapa and UK's Permanent Under Secretary Sir Simon Frazer signed the MoU for establishing the Consultation Mechanism at the Foreign and Commonwealth Office, London respectively. As per the MoU, both the parties will hold consultations at jointly decided intervals alternately in the capital of either country at an appropriate level as may be decided upon and hold discussion on all aspects of bilateral relations as well as on international or regional issues of mutual interest.

Finalized Agreements/MoU

- TIFA Agreement signed in 2011.
- An agreement on promotion and protection of investment between Nepal and UK signed on March 2, 1993
- Development Partnership Agreement between Nepal and UK, signed on December 2013. It supersedes the Technical cooperation agreement desiring to strengthen the traditional cooperation and cordial relation between Nepal and the UK signed on 31 May 1994.
- Trade Agreement signed in 1965.

Agreements/MoU in Progress

- Bilateral Agreement on Avoidance of Double Taxation

\textbf{Economic and Technical Cooperation (Development Cooperation)}

The UK aid to Nepal in various fields of activities started in 1961. The United Kingdom has been offering fellowships to the Government of Nepal since 1950s. British volunteers are engaged in Nepal since 1964. These programmes have contributed to Nepal's need for specialised and trained manpower and also have developed important links between the people of the two countries. In order to better understand and address the issues of poverty in Nepal, in April 1999 Department for International Development (DFID) established an in-country office staffed by a multi-disciplinary team of both UK nationals and staff appointed in country. DFID presence in Nepal has greatly assisted in working with Government of Nepal in the development of understanding and establishment of networks that will promote opportunities for change.

Recently Development Partnership Arrangement to provide a transparent and mutual accountability framework between Nepal and the United Kingdom on development assistance was signed on December 2013. The assistance of British Government generally comes through the Umbrella Agreement. The first such agreement of Pound Sterling 12 million was signed on 23 September 1979 for the implementation of various projects. The UK Government’s Department for International Development (DFID) prepared an Operational Plan for Nepal during 2011-2016. This plan commits up to Pound Sterling 413 million of UK official development assistance during the period 2011-2016. All bilateral aid with Nepal is on a grant basis.

As per Development Cooperation Report 2014-2015 by Ministry of Finance, UK remained the top ODA provider (based on disbursement) with a total assistance of US dollars 168.07 million in the Fiscal Year 2014/2015.

DFID Technical Support has been instrumental in promoting foreign direct investment especially in large hydropower project through Investment Board Nepal for the last few years. This very programme has provisioned of supporting IBN to promote private investment in the country. Likewise there are provisions of continuing support to the second and third phases of Financial Sector Reform Programme of the Government initiated few years back. The program will last for six years from 2015 to 2020.

The DFID Nepal's operational Plan is divided into four main areas: governance and security, inclusive wealth creation, human development (basic services including education and health), and climate change/disaster management. The UK has assisted Nepal in the areas of livelihoods, e.g. agriculture, forestry, transport and communications, local development; basic services, e.g. education, health, water supply and sanitation; good governance, human rights and peace building efforts.

\textbf{Earthquake, 2015}

Nepal received generous support and assistance from the Government of the United Kingdom in the aftermath of the earthquake of 25 April 2015. It responded to the Nepal earthquakes immediately with the deployment of Eight Disaster Response Specialists under Rapid Response Facility and with the release of GBP 5 million. The United Kingdom Government pledged USD 110 million for the reconstruction and rebuilding of Nepal. 

\textbf{Trade}

Nepal and UK signed trade agreement in 1965. Despite our extensive engagements and linkages, both at the level of government and, people in the backdrop of our long-standing friendly and cooperative relations, trade and investment relations between the two countries have largely remained way below its actual potential.

With high imbalance and elements of unpredictability, this is, of course, a typical feature of a trade relation between a developed and least developed country. Nepal faced a trade deficit of 142,182,856 Rupees in the first eight month of 2016. The total export to UK stands at 1,256,018,669 Rupees while the total import stands at 1,398,201,525. It is for this reason that Nepal places high importance to directing foreign development cooperation in areas which help enhancing our productive capacity. Aid for trade is, therefore, another preferred area for us in our development cooperation. 

The major items of import from UK are: Motor car, vehicle, parts of aeroplane and helicopter, Whiskies, Malt not roasted, Sweets biscuits, Chocolate in blocks, slab or bar etc. Major Nepalese exports to UK are Pashmina shawls, goatskin, leather goods, Nepalese paper and paper products, woollen carpets, handicrafts, ready-made garments, silverware and jewellery. Likewise, major imports from UK are copper scrap, hard drinks, cosmetics, medicine and medical equipment, textiles, copper wire rod, machinery and parts, aircraft and spare parts, scientific research equipment, office equipment and stationery.

Exchange of visits by trade delegations from Nepal and Britain has added a new dimension to the commercial relations between Nepal and Britain. Nepal-Britain Chamber of Commerce and Industry is active in promoting trade and investment between Nepal and UK.

\textbf{Tourism}

A sizeable number of British tourists come to Nepal every year for trekking and mountaineering and other leisurely activities. Given our historical linkages and increasing people-to-people contacts, the prospects for increasing the tourist arrival from UK in Nepal are quite promising.

UK's support and cooperation in line with our broader policy priorities will be much appreciated, especially in construction and upgradation of tourism infrastructure with focus on tourist safety. DIFID’s recent involvement and interest in tourist security, trekkers tracking, and maintenance of damaged trekking trails, including the Great Himalayan Trail, is most welcome.

Embassy of Nepal in London is planning to organize the many tourism promotional programme to increase the number of tourists from UK.

\textbf{Consular activities}

Embassy of Nepal in London involves in various consular service ranging from passports, visas, Travel Document, NRN Card, Verification, Attestation and Power of Attorney. During Fiscal Year 2075/76, Embassy of Nepal in London generated GBP 542048.72 (NRS 7,73,71,510.41) revenue.

\textbf{Investment}

Foreign investment commitment from UK stands at 1663 million rupees till 2013/14. The major investment is in the areas of banking, tourism, education, technology. An agreement for the promotion and protection of investment between Nepal and UK was signed on March 2, 1993.

There are some British joint ventures in the areas of hotel, travel and trekking, tea production, education, garment, bio-technology and consultancy. These are significant figures, but do not reflect the true potential. There are many British and Non Resident Nepalese (NRN) entrepreneurs who are still making trade and investment successes in Nepal. Standard Chartered Bank and Unilever are two major British companies who have made a huge return from their investment in Nepal in the last 20 years.

An agreement on promotion and protection of investment between Nepal and UK signed on 02 March 1993 provides a framework to further expand cooperation in this field. In this context, the proposal for signing another agreement on avoidance of double taxation will further help boost the flow of FDI between the two countries.

\textbf{Foreign employment and British Gurkha}

Thousands of Nepalese are working at service sectors like hospitals, university, hotels and restaurants in UK. About 200 Nepalese are recruited as Maritime Security Guards in UK. About 150,000 Nepalese are living in UK. Many Nepalese belong to Ex-Gurkha Servicemen and their families.

British Gurkhas is a fully integrated part of the British Armed Forces. It constitutes an important element in Nepal-Britain relations.  Over 160,000 Gurkhas were enlisted in the Gurkha Brigade and other units of the Indian Army during the First and Second World Wars. In recognition of their distinguished service, the British Gorkhas servicemen from Nepal have won 13 Victoria Crosses (VC), the highest British gallantry honour.

\textbf{Cultural Cooperation}

British Council in Nepal is working to extend cultural relations between UK and Nepal. Similarly, Nepalese diaspora and Gurkha soldiers are visible bond who have interlinked the Nepalese culture and traditions in UK.

\textbf{Education cooperation}

UK is supporting US dollars 260,012 for the education of marginalised girl in Kailali district of Nepal. DFID was the core education funding agency in Nepal, supporting the foundations of education values. Active as a School Sector Reform Plan (SSRP) pooling Development Partner. With Education for All overall goals, the SSRP objectives were to 
\begin{enumerate}
\item ensure access and equity in primary education; 
\item improve the efficiency and institutional capacity of primary education; and 
\item enhance the quality and relevance of basic primary education for children and illiterate adults.
\end{enumerate}

In 2013, DFID committed around GBP 2 million to establish 810 additional girls' toilet and water supplies/WASH facilities across 18 districts selected based on low indicators in hygiene and sanitation with an implementation date which was extended to July 2015.

The British Government has been regularly providing scholarships in different areas for the development of human resource in Nepal since 1950s. Britain is offering Chevening Award to Nepalese from many years. A lot of Nepalese students are pursuing higher level studies in the UK. 

\textbf{Security cooperation}

UK has provided support for upgrading of Nepal Army Birendra Peace Operations Training Centre (BPOTC) located at Panchkhal (45 km East of Kathmandu). It has also provided Explosive Ordinance Device (EOD) related Improvised Explosive Device Disposal (IEDD) related Special Training to Nepal Army. UK is continually providing the special training courses to Nepal Army officials.

High level visit from Nepal Army and British Army are regularly taking place in UK and Nepal respectively.

\textbf{Nepalese students in UK}

\end{solution}

2. Discuss about the bilateral US relations with Nepal

**Solution:**

The United States recognized the Kingdom of Nepal on April 27, 1947, when Joseph C. Satterthwaite presented a letter from President Harry S. Truman to then king Tribhubana of Nepal. Satterthwaite was on a special diplomatic mission to Nepal as personal representative of the President. The two countries established diplomatic relations in 1948. Bilateral relations are friendly, and U.S. policy objectives toward Nepal center on helping it build a peaceful, prosperous, resilient and democratic society. Primary U.S. objectives in Nepal include supporting a stable, democratic Nepal that respects the rule of law; promoting investor-friendly economic development; and improving disaster risk management systems.

The United States enjoys a strong and positive relationship with Nepal. Years of diplomatic, development, and military engagement have advanced U.S. interests in transitioning Nepal into a more peaceful, stable democracy with significant economic potential. Since the end of its 10-year civil war in 2006 and the devastating earthquakes of 2015, Nepal has successfully transitioned into a full-fledged constitutional federal republic grounded in a constitution promulgated in 2015. With the recent formation of a new government, Nepal may now be on the cusp of a period of much-needed political stability.

Nepal has sought a bilateral consultation mechanism with the United States and has signed similar agreements with other countries like Australia in summer 2017. Such a forum might include as topics: Millenium Challenge Cooperation (MCC) coordination and implementation; trade (TPS, technical assistance, etc.); security and defense cooperation; and humanitarian assistance and disaster response.

To improve Nepal's economic situation, the MCC signed, in 2017, a \$500 million Compact with Nepal to expand Nepal’s electricity transmission infrastructure and improve its road maintenance regime. The Nepali government has committed another \$130 million for a program total of \$630 million. The Compact will build 300 kilometers (km) of high-voltage electric transmission lines, three substations, perform enhanced road maintenance on 305 km of strategic highways, and provide technical assistance to the national electric utility, the new electricity regulator, and the Department of Roads. The Compact is entered into force after the end of FY 2019.

U.S. assistance has been critical to helping Nepal rebuild after the devastating 2015 earthquake and remains committed to building Nepal's resilience in the event of any future disaster. The United States has provided over \$190 million for earthquake relief, recovery, and reconstruction. While these commitments have exceeded our initial pledge of \$130 million made at the International Conference on Nepal's Reconstruction in June 2015, significant recovery needs remain. To date, the United States has built 36 schools and hospitals; has directly helped rebuild over 16,000 homes; trained 260,000 people in safer construction; and developed policies, systems, and controls to ensure that $8.6 billion in reconstruction results in safer structures for all. In addition to rebuilding a safer Nepal, we are working to help Nepal institutionalize all lessons learned from the 2015 earthquake. We are also helping Nepal to implement its new disaster management law and stand up a new National Disaster Management Authority. We are further supporting Nepal as it introduces federalism by working with newly elected local governments to implement their own disaster management plans—thus helping local authorities meet commitments made to their constituencies. These efforts will help Nepal on its journey to self-reliance.

The United States has also committed security assistance to Nepal, working with the Nepali Army to strengthen their peacekeeping and disaster response capabilities.

The United States and Nepal have signed a trade and investment framework agreement, providing a forum for bilateral talks to enhance trade and investment, discuss specific trade issues, and promote more comprehensive trade agreements between the two countries. In 2016, Nepal became one of few countries in the world with a single-country trade preference program with the United States. Principal U.S. exports to Nepal include agricultural products, aircraft parts, optic and medical instruments and machinery. U.S. imports from Nepal include carpets, apparel and jewelry.

### Impacts of climate change

What are the impacts of climate change ? Discuss on detail about the initiative taken by Nepal Government in tackling the impacts of climate change.

Climate change is now affecting every country in every continent. It is disrupting national economies and affecting lives, of individuals, communities and countries today and is likely to do so in the ensuing future. Weather pattern are changing, snowline are melting due to global warming and sea level are rising. 

**Negative impacts**

1. Numerous mountains are seeing accelerated the melting of snow due to global warming.
2. Changes in monsoon patterns, such as erratic rainfall with alternate drought periods are the serious impacts of climate change in Nepal which hampers the agricultural production. The changes in monsoon pattern in Nepal greatly exacerbate the situation of poverty and food insecurity in the country.
3. Flood, landslides, forest fires, snow melt are the disasters that occur as a result of climate change without any prewarning and causes heavy loss.
4. High hills and mountains are more vulnerable to climate change compared to terai, thus climate change make large scale economic and ecolgical disruption in upper latitudes.
5. Loss of plant and animal biodiversity.
6. Increase in number of insects, pest and diseases and sporadic outbreaks becoming much frequent.

The positive impact of global warming, a characteristic feature of climate change, is that the crops that could in the past be only growth in certain low altitudes, requiring prolonged exposure to heat can now be growth in higher regions such as Manang and Mustang districts of Nepal. The is evidently due to increase average temperature of the region, thus allowing the crop to perform well. However, as noted earlier, there are more prominent negative impacts than the benifits we can derive from the climate change.

Nepal became signatory of UN Framework Convention on Climate Change (UNFCC) in 1994. Ever since then the issue of climate change has been routinely incorporated into policy and strategic frameworks by Government of Nepal. Following policies and strategies are currently in place to help mitigate the impacts of climate change in Nepal:

1. Constitutional approach

Article 30(1) of the constitution of Nepal guarantees a "clean environment" as fundamental right and elaborates that "every citizen shall have the right to live in a clean and healthy environment." Article 30(2) of the constitution also encourages the state to formulate necessary legal framework to balance environment and development.

2. Approaches in periodic plans
3. Climate change policy, 2067 (2011)
4. National climate change policy, 2076 (2019)
6. National adaptation plan of action (NAPA, 2010)
7. Framework for local adaptation plan for action (LAPA, 2011)

### Highlights of NCCP, 2076

Reasons for Nepal's vulnerability to climate change are Topographical fragility, fragile geological structure, sensitive ecosystems and diversity of climate and micro-climate zones prevalence in Nepal. Additionally, poverty, illiteracy, social disparity and high dependence of communities on natural resources for livelihood.

- Nepal National Red Strategy implemented to check greenhouse gases emmission from deforestation and forest destruction.
- Provision of climate change coordination committee in Chairmanship of Prime minister.
- Establishment of Climate change management section under Ministry of Forest and Environment as focal point for implementation of international treaty.
- Provision of Climate budget code by National planning commission.
- Alternative energy promotion center has established climate and carbon unit for executing Clean Development Mechanism (CDM) related activities.

**Goal**

To contribute to socio-economic prosperity of the nation by building a climate resilient society.

**Objectives**

a. To enhance climate change adaptation capacity of persons, families, groups and communities vulnerable to, and at risk of, climate change;
b. To build resilience of ecosystems that are at risk of adverse impacts of climate change;
c. To promote green economy by adopting the concept of low carbon emmission development;
d. To mobilize national and international financial resources for climate change mitigation and adaptation in just manner;
e. To conduct research, make effective technology development and information service delivery related to climate change;
f. To mainstream or integrate climate change issues into policies, strategies, plans and programs at all levels of State and sectoral areas;
g. To mainstream gender equality and social inclusion (GESI) into climate change mitigation and adaptation programs.

**Policies** (Sectoral policy, strategies and working policies)

1. Agriculture and food security
2. Forest, biodiversity and watershed conservation
3. Water resources and energy
4. Rural and urban habitats
5. Industry, transport and physical infrastructure
6. Tourism and natural and cultural heritage
7. Health and drinking water and sanitation
8. Disaster risk reduction and management
9. Gender equality and social inclusion, livelihoods and good governance
10. Awareness raising and capacity development
11. Research, technology development and expansion
12. Climate finance management

Policy number 9-12 are inter-thematic areas.

### Population census, 2068

```{r}
dplyr::tribble(~"Indicator", "Unit", ~"Value", 
               "Male", "count", 12849041,
               "Female", "count", 13645463,
               "Annual population growth rate", "percentage", 1.35,
               "Population density", "individuals per msq", 180,
               "Total engagement in agriculture", "percentage", 65.6,
               "Male engagement in agriculture", "percentage", 60.2,
               "Female engagement in agriculture", "percentage", 72.8, 
               "Crude birth rate", "per thousand", 21.8,
               "Crude death rate", "per thousand", 7.3,
               "Child death rate", "per thousand", 40.5,
               "Net fertility rate", "percentage", 2.5, 
               "Average life expectancy", "years", 66.6) %>% 
  knitr::kable(caption = "Population census, 2068 summary", booktabs = TRUE, longtable = T)

```

### Hulaki rajmarga (Hulaki highway)

http://www.karma99.com/
http://www.karma99.com/2015/09/hulaki-rajmarga.html

### Demographics of Nepal

- Refer to wikipedia link \url{https://en.wikipedia.org/wiki/Demographics_of_Nepal}, for demographic statistics of Nepal. This includes
  - Population growth
  - Vital statistics: Live births per year, Deaths per year, Natural change per year, CBR, CDR, NC, TFR, IMR
  - Population structure: Based on age group
  - Ethnicity statisics
  - Languages
  - CIA factbook summary
  - Religion, etc.

### Madan puraskar

Madan puraskar is a literary honor in Nepal which Madan Puraskar Guthi confers annually for an outstanding book in Nepali language published within the calendar year. It is considered the most prestigious literature award in Nepal.

It was established by daughter in law of Prime Minister Chandra Shumsher JBR Queen Jagadamba Kumari Devi, in memory of her husband late General Madan Shumsher JBR, on September 26, 1955. Since then it has been presented every year, except in 1974 and 1977. At the time of its establishment, the award prize was NRS 4000. Since the Golden Jubilee year 2005 the award prize has been increased to NRs. 400,000.

### List of winners

```{r winners-list}
# madan_puraskar <- read_html("https://en.wikipedia.org/wiki/Madan_Puraskar") %>% 
#   html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[2]')
# 
# madan_puraskar %>% 
#   .[[1]] %>%
#   html_table() %>% 
#   as_tibble() %>% 
#   write_csv("./madan_puraskar_awardees.csv", na = "")

madan_puraskar <- read_csv("./data/madan_puraskar_awardees.csv")

madan_puraskar %>% 
  knitr::kable(booktabs = TRUE, caption = "List of Madan Puraskar winners till 2075", longtable = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped", font_size = 8) %>% 
  kableExtra::column_spec(column = 2:3, width = c("14em", "24em"))
```
