## Essay writing on issue (800 to 1000 words)

1. We live in a world besieged by media and technology. Every sector of our life is impacted by the presence of modern forms of communication system. Write an essay discusses the influence of modern forms of communications on everyday life.

\begin{solution}

The very nature of human being as a social creature underlines the importance of communication in one's day-to-day life. Human beings interact, possibly much effectively and efficiently than any other life forms in this planet. Much of the achievements we are relishing today, from kitchen utensils to high performance computers are all, in a way, fruits of reliable communication. This is to say that innovations are triggered in several ways through communication. It is also true that modern day communication, we use in the form of media and technology, have been made possible by the feats of human innovation -- once again guided by state of the art communication systems.

Early developments in communication technology consisted of invention of telegram and telephone, which extensively relied on electrical signals on copper wire. As the search for better communication infrastructure continued, data (including video and audio) sharing over wireless frequencies and optical cables are on upsurge today. To recapitulate the role of communication channel, we can refer to an event that is purported to have put an end to the World War II. Invention of cipher (a tool that decrypts hidden messages) on those days became a landmark in intercepting German radio communication channels and gaining the information about overhanging threat military assault. In the retrospect, communication development during World War might not seem as much of a feat when we consider high performing broadband facilites that are at our disposal today.

Communication system today pervades all spheres of life be it physical or socio-economic. It wasn't uncommon, just a few decades earlier, for people to commute great deal of distance just for the sake delivering information through postal services. Proliferation of telecommunication services in every major inhabitated part of world has made it non-essential for the paper mail services to physically transport tremendous amounts of exchanges people are making. This trend in frequent exchange of ideas and even general information among peoples -- families, friend, scholars, researchers, etc., is expected to only expand in future. The imperative is that, as means of communication gets cheaper and accessible, people need not be constrained to utilize information wealth due to their physical distance.

More often than not, means of communication exert their importance at times of crisis. Remote areas are remote for more than a reason but one of those consistently includes them not being able to communicate with authorities on time. This is a situation that typically results in greater marginalization, as individuals from distant localities lacking proper communication methods are increasingly unheard of their problems and scarcities. For example, when a disease outbreak occurs in such an area, relief package and medical interventions become delayed, resulting in significant losses of life. In contrast, well-connected areas where government can quickly initiate a rescue operation will have more lives saved.

When looked at for side-effects on an individual, communication affects even the mundane aspects of life. Citizens of 21 century take toolsets of communication for granted. There is little to how every day engagements of families are ran without crediting to modern communication technologies. Grocery stores and dairy shops have the communication network set to have knowledge of market; A dairy handler has to know who are his customers for what products. Daily updates on traffic and weather to work are all televised through various news channels so that commuters can be punctual to their work. Furthermore, in the workplace, all commercial startups communicate over telephone or email to their clients. This, in comparison to early days approach of door to door visit for sales promotion and order acquistion, is a stark gain in efficiency and scale. A supply order could be made over a phone call while also provisioning for accumulated order the previous day for the same household equipment to areas nearby thereby significantly cutting down the cost of repeated transportation.

Our social thread forms a web of close circle of friends and families along with tendencies to keep-on-touch with them. This emotional and social sphere of our lives too is strung more tightly together than ever before with the advent of new communication technologies. Rashid can now ask what the weather is like in Australia to one of his friends from Nepal in a video call, while also hoping some days to see them there in person. His mother can make a quick appointment with a doctor in a phone call after knowing that Birat is ill. In addition to surfing a personalized news-stand online right before the meeting, one is now also able to address a meeting which includes a mass of 50 people. This is made possible due to wide projection and affordable cinema technology which allow people to conduct presentations as well as entertain with movie.

Besides trivial occurances of everyday life, various channels of communication are used by government organizations to address public and state their vision for country. Likewise, several government insitutions have made social media formum also a platform for obtaining public feedback and outcries, while also attempting to address them. For example, if suspected for a criminal activity, goverment can call for a surveillance of an individual or a group through the use of closed circuit cameras (CCTV) or the GPS system of suspected's smartphone. Although for the case it is clear that criminal threat outweights the expectation that government action hurts individual privacy, there can be scenarios where this application of technology can be subjected to debate. Meanwhile, there is always a chance that political groups may misuse media to propagandize their statements to only gain popularity, while masking up facts that give different flavor.

With all the merits it has brought to individual and societal lives, modern communication technologies can well be thought of as integral part of development. While it has connected people for good, indiscriminate retrieval and exchange of some information can have detrimental consequences -- With great power, comes a greater responsibility.

\end{solution}

2. Write an essay that discusses how different cultures maintain their distinctive tradition in your community. In what ways does this cultural diversity have a positive impact on your community ?

3. Diplomacy has been defined as "the management of relations between independent states by the process of negotiation". Write an essay citing some major events of modern international diplomacy that have resolved conflict through the process of negotiation and have shaped the world in which we live.

## Miscellaneous

\begin{questions}

\question Discuss in brief the political history of Nepal.

\begin{solution}
Records mention the Gopalas and Mahishapalas believed to have been the earliest rulers with their capital at Matatirtha, the south-west corner of the Kathmandu Valley. From the 7th or 8th Century B.C. the Kirantis are said to have ruled the valley. Their famous King Yalumber is even mentioned in the epic, ‘Mahabharat’. Around 300 A.D. the Lichhavis arrived from northern India and overthrew the Kirantis. One of the legacies of the Lichhavis is the Changu Narayan Temple near Bhaktapur, a UNESCO World Heritage Site (Culture), which dates back to the 5th Century. In the early 7th Century, Amshuvarma, the first Thakuri king took over the throne from his father-in-law who was a Lichhavi. He married off his daughter Bhrikuti to the famous Tibetan King Tsong Tsen Gampo thus establishing good relations with Tibet. The Lichhavis brought art and architecture to the valley but the golden age of creativity arrived in 1200 A.D with the Mallas.

During their 550 year rule, the Mallas built numerous temples and splendid palaces with picturesque squares. It was also during their rule that society and the cities became well organized; religious festivals were introduced and literature, music and art were encouraged. After the death of Yaksha Malla, the valley was divided into three kingdoms: Kathmandu (Kantipur), Bhaktapur (Bhadgaon) and Patan (Lalitpur). Around this time, the Nepal as we know it today was divided into about 46 independent principalities. One among these was the kingdom of Gorkha with a Shah ruler. Much of Kathmandu Valley’s history around this time was recorded by Capuchin friars who lived in the valley on their way in and out of Tibet.

An ambitious Gorkha King named Prithvi Narayan Shah embarked on a conquering mission that led to the defeat of all the kingdoms in the valley (including Kirtipur which was an independent state) by 1769. Instead of annexing the newly acquired states to his kingdom of Gorkha, Prithvi Narayan decided to move his capital to Kathmandu establishing the Shah dynasty which ruled unified Nepal from 1769 to 2008.

The history of the Gorkha state goes back to 1559 when Dravya Shah established a kingdom in an area chiefly inhabited by Magars. During the 17th and early 18thcenturies, Gorkha continued a slow expansion, conquering various states while forging alliances with others. Prithvi Narayan dedicated himself at an early age to the conquest of the Kathmandu Valley. Recognizing the threat of the British Raj in India, he dismissed European missionaries from the country and for more than a century, Nepal remained in isolation.

During the mid-19th Century Jung Bahadur Rana became Nepal’s first prime minister to wield absolute power relegating the Shah king to mere figureheads. He started a hereditary reign of the Rana Prime Ministers that lasted for 104 years. The Ranas were overthrown in a democracy movement of the early 1950s with support from the-then monarch of Nepal, King Tribhuvan. Soon after the overthrow of the Ranas, King Tribhuvan was reinstated as the Head of the State. In early 1959, Tribhuvan’s son King Mahendra issued a new constitution, and the first democratic elections for a national assembly were held. The Nepali Congress Party was victorious and their leader, Bishweshwar Prasad Koirala formed a government and served as prime minister. But by 1960, King Mahendra had changed his mind and dissolved Parliament, dismissing the first democratic government.

After many years of struggle when the political parties were banned, they finally mustered enough courage to start a People’s Movement in 1990. Paving way for democracy, the then-King Birendra accepted constitutional reforms and established a multiparty parliament with King as the Head of State and an executive Prime Minister. In May 1991, Nepal held its first parliamentary elections. In February 1996, the Maoist parties declared People’s War against monarchy and the elected government.

Then on 1st June 2001, a horrific tragedy wiped out the entire royal family including King Birendra and Queen Aishwarya with many of their closest relatives. With only King Birendra’s brother, Gyanendra and his family surviving, he was crowned the king. King Gyanendra abided by the elected government for some time and then dismissed the elected Parliament to wield absolute power. In April 2006, another People’s Movement was launched jointly by the democratic parties focusing most energy in Kathmandu which led to a 19-day curfew. Eventually, King Gyanendra relinquished his power and reinstated the Parliament. On November 21, 2006, Prime Minister Girija Prasad Koirala and Maoist chairman Prachanda signed the Comprehensive Peace Agreement (CPA) 2006, committing to democracy and peace for the progress of the country and people. A Constituent Assembly election was held on April 10, 2008. On May 28, 2008, the newly elected Constituent Assembly declared Nepal a Federal Democratic Republic, abolishing the 240 year-old monarchy. Nepal today has a President as Head of State and a Prime Minister heading the Government.

The Constituent Assembly made significant progress to accomplish the mandate of writing a new democratic constitution of Nepal during its first 4 years term. The country also had an extensive democratic exercise in that direction including collection of public inputs on the contents of the new constitution and intense deliberations in the Assembly. However, due to political disagreements on some of the contentious issues like federal provinces and form of government, the first CA could not accomplish the historic task and there was natural termination of its mandate in 2012. The election of CA II was held in November 2013 and in its first meeting, leaders of political parties set the timeline of 1 year to complete the task of writing the new constitution.

Devastating earthquake of 7.8 magnitude hit Nepal in April 2015 followed by several powerful aftershocks causing loss of life, infrastructure and property in an unimaginable scale. Most mid hill districts of Nepal including Kathmandu valley saw massive devastation. This terrible experience created a sense of urgency among political parties to expedite the constitution writing so that a political process would come to a meaningful conclusion and country can divert all its focus on post disaster reconstruction.

After weeks of zeroing in on most contentious issues, political parties sorted them out paving the way to finalize the constitution. The new constitution of Nepal was promulgated through an overwhelming majority of the votes of CA members on September 20, 2015. With this historic achievement, the decades-long dream of Nepali people to have a constitution made through an elected representative body has now been realized. As per the provisions of the new constitution, elections of the new President, Prime Ministers and some other State positions have been successfully held.

source: \url{https://mofa.gov.np/about-nepal/history-of-nepal/}
\end{solution}

\question Law and ethics are considered to be the two tools for controlling human conduct so as to make it conducive to civilized social existence. Discuss how they achieve this objective. Giving examples, show how the two differ in their approaches.

\begin{solution}

Law are the system of rules which a particular country or community recognizes as regulating the actions of its members and which it may enforce by the imposition of penalties.
Ethics seeks to resolve questions of human morality by defining concepts such as good and evil, right and wrong, virtue and vice, justice and crime.

How law and ethics help in conductive social existence

By moulding human character

Ethics tries to mould human character into morally sound individual so that he/she follows set of rules that uphold good values. These values see to it that the individual does not indulge in unethical and non acceptable activities.

Fear of punishment

Laws are designed on the principles of generating fear in an individual for violating certain regulations imposed by the state. These regulations have to be followed in all conditions so that society is free from chaos. The laws have the ability to create virtual boundaries for human behaviour.

Clarity of thought process

The ultimate idea of law and ethics is to create a clarity in thought process of individuals so that they can understand what is good and bad. This will enable them to take the right decision, which will be beneficial for civilised society.

How ethics and law differ from each other

Explicit and implicit nature

Laws are basically codified by either the state or relevant authorities and explicitly defined, making its interpretation and understanding easier. Ethics do not have a fixed version and is basically case to case basis when it comes to interpretation. Ethical in one case may not be ethical in another.

Ex: Murder is a crime that is explicitly given in law and does not vary on case to case basis.
    Lying is unethical in some cases whereas it is acceptable in some cases.
    
General vs Specific

Ethics is based on general idea of good and bad, correct and wrong and is thusvague when it comes to exact relevance. Law is usually specific for a specific case and applies for that case only.

Ex: When a particular crime occurs, ethics tries to understand the underlying circumstances to say whether it is ethical or unethical. In law there is no focus on circumstances but instead only sees whether the act was crime under the law.

Thus, laws and ethics may have different reasons for its existence but the underlying idea of both is to maintain socially conductive practices.

source: \url{https://www.upsciq.com/law-and-ethics-considered-two-tools-controlling-human/}
\end{solution}

\question What is water-use efficiency? Describe the role of micro-irrigation in increasing the water-use efficiency.

\begin{solution}

Water use efficiency is the ratio of water used for plant metabolism to the water lost by plant through transpiration. It can be understood as a measure to reduce water wastage from plants by providing only the necessary amount.

Role of micro-irrigation in increasing water-use efficiency

Reduces excess water accumulation
    
In conventional irrigation system water is used in bulk without proper understanding of the requirement of the crop. This will result in accumulation of water without proper utilisation leading to wastage. In micro-irrigation, the water needed for the plant to grow is given in right amount so as to reduce wastage.

Recycling of expelled water

In micro-irrigation system such as greenhouse farming, the plant canopy is covered in plastic to prevent transpiration water from escaping. The water is then diverting back to the plant. This will ensure that water wastage from plant will be lower in comparison to the input.

Reduces run-off

Water irrigation in conventional methods like canals and tube well will result in water runoff from the fields without getting utilised. This will contribute towards wastage especially in areas where there is shortage of water. In micro-irrigation techniques, water will be targeted at the root of the crops so that wastage is as low as possible.

Way forward
It is necessary to device methods to make water use efficiency high and address shortage of water across the globe.

Thus, micro-irrigation is a very important idea when it comes to sustainable use of water resources. The initial investment once recovered will make agriculture a profitable venture.

source: \url{https://www.upsciq.com/what-is-water-use-efficiency/}
\end{solution}

\question Anger is a harmful negative emotion. It is injurious to both personal life and work life. Discuss how it leads to negative emotions and undesirable behaviours. How can it be managed and controlled?

\begin{solution}
Anger can be termed as a extreme behaviour, an intense emotional state involving a strong uncomfortable and hostile response to a perceived provocation, hurt or threat. This is considered a negative emotion because it has potential to hurt someone or cause any physical or mental damage to ourselves or others.

How anger leads to negative emotions and undesirable behaviours

By blocking our emotional intelligence

Anger has the power of blocking the right thinking process that usually allows us to make right judgement. This state will influence us to make certain decision that would not have taken place if we would have calmly handled the situation. This will result in actions that are both undesirable and unworthy.

Ex: Slapping a friend in public for a small action that was accidental or unintentional.
    Creating sense of disappointment

Negative emotions are developed in an individual if a sense of disappointment prevails due to anger. This disappointment will create a strong sense of negative emotions such as disgust and hate, which will force us to display undesirable behaviour.

Ex: when you find out that your colleague is indulging in unethical action, you sense a feeling of anger. This anger will make you develop negative emotions like hate and disgust towards the colleague without hearing the other side of the argument.

How to control and manage anger

Mind calming techniques

The best method to control anger is using techniques to introduce calmness in thinking process. This can be achieved by practicing yoga, meditation and other forms of natural healing. This will provide that required moment of calmness that prevents our anger to overshoot, which otherwise leads to negative emotions and unruly behaviour.

Physical activities

Physical activities such as exercise, jogging etc will help in reducing the frustration that builds up and is responsible for anger. By getting rid of underlying and hidden feeling, we can atleast prevent ourselves from overreacting to a situation through unruly behaviour.

Way forward

Anger in a certain amount is beneficial as it keeps the spirits of an individual high. But exceeding the limit will usually create chances of negative emotions.

Thus, anger management is an important component in the life of an individual. It not only keeps negative emotions from emerging, it also ensures that energy of an individual is used productively and  not spent on emotional angry outbursts.

source: \url{https://www.upsciq.com/anger-harmful-negative-emotion-injurious-both-personal-life/}
\end{solution}

\question The global lockdown imposed due to COVID-19 pandemic has turned out to be beneficial for environment on short term. Validate.

\begin{solution}
The major countries of the world, including India have self imposed a lockdown to tackle spread of COVID-19 pandemic. This has impacted economy of the world in an adverse manner causing slowdown. But the silver lining in the situation is improvement in environmental conditions that we have observed.

Environmental benefits due to global lockdown

Reduction in air pollution levels

The major environmental concern in this century is climate change, that has been caused due to air pollution across major human settlements. Due to the lockdown the pollutant level in air such as CO2, NO3, SO4, ozone, PM 10, PM 2.5 have seen a drastic drop. This has made the air cleaner for breathing and reducing health concerns.

Reduction in water pollution levels

Water pollution in developing countries is mainly due to factories dumping their waste in water bodies. The lockdown has forced the factories to shutdown and cut production. This has reduced or completely cut off waste discharge from factories.

Reduction in solid municipal waste

With low outdoor activities and travel, the household waste produced in cities have seen reduction. The reduced level makes it easier to dispose off the waste more effectively. Dumping grounds have a chance to recover if this trend continues.

Increase in biodiversity

Withdrawal of humans from outdoors have opened the spaces for biodiversity to recover their lost territories. Sea mammals like dolphins have returned back to the shores due to lack of disruption from humans in their territories.

Reduction in sound and light pollution

Cities have turned silent due to stoppage of vehicular movement and industrial activities. This has made the environment suitable for calmness. Excessive usage of light has been reduced too due to the same reasons.

Way forward

This restriction has to be converted into a habit for people to follow in future so that environmental pollution is kept in check.

So, the global lockdown has helped the environment to recover upto certain extent. What needs to be seen is the impact that return of normality is expected to bring.

source: \url{https://www.upsciq.com/the-global-lockdown-imposed/}
\end{solution}

\question What is the better form of government -- Parliamentary, Presidential and Semi-presidential ?

\begin{solution}

Parliamentary systems, such as of Nepal, United Kingdom and India, combines the lawmaking and executive functions, which allows the national legislature to select and remove the head of government and cabinet.

Presidential systems, such as of Mexico and Nigeria, feature a cabinet that is mostly responsible to the elected executive, with a legislature that can only remove cabinet members through impeachment. An example of a presidential system would be a country in which the president holds the roles of head of the state and head of government, selects the executive cabinet, and is elected in a direct, popular election while members of the lower and upper houses are elected by voters in the respective districts and regions. These systems have separate fixed-term, popular elections for the national legislature, and a top executive leader serving as both head of state and head of government.

Semi-presidential systems, such as Russia, feature separate popular elections for the president and for the national legislature, allowing the president to nominate a prime minister (who must be approved by the legislature); members of the cabinet are held accountable by both the president and legislature.

Although parliamentary systems have fewer institutional obstacles to enact policy than presidential systems (presidential systems have divided branch powers), parliamentary systems have their own checks on the executive branch. Parliaments may censure cabinet ministers, refuse to pass executive proposed legislation, question the executive and cabinet ministers, and impose time deadlines on calling new elections.

Presidential and semi-presidential systems have divided powers among governmental branches which allows the legislature to more commonly change or reject executive policy proposals. For instance, Nigeria's senate rejected the president's proposed budget in 2016 that relied too heavily on borrowing from foreign sources.

source: \#APCompGov2020
\end{solution}

\question Discuss about the advantages and disadvantages of Unitary and Federal government systems.

\begin{solution}

Unitary government advantages

Unitary government do have some significant strengths. Most common of those over federal system are:

\begin{enumerate}
\item National direction is clear; Policies can be made by a single government without the need for negotiation or cooperation with subnational states.
\item There can be no confusion as to accountability. It is clear to voters which level of government is responsible for a particular problem or function. (Legislators who wish to spend money must raise it; it is not possible for legislators to seek to shift the blame for failure to another level of government.)
\item Duplication of legislatures, bureaucracies, and programs is avoided with significant savings in direct costs -- and the more subtle but no less real costs of needless duplication.
\item Issues of fairness in raising and spending money between levels of government (vertical fiscal imbalance) simply do not arise.
\end{enumerate}

Federal government advantages

Unitary government also have major drawbacks, which in turn correspond to the major strengths of federal systems. These include:

\begin{enumerate}
\item A federal system has greater scope of diversity and experimentation in policy.
\item The need to debate issues rather than enact them instantly may provide a more considered and viable policy outcome. This tendency toward incrementalism is seen as integral to democracy.
\item A federal system must consider the different ethnic or cultural groupings that may predominate in a particular state and wish to pursue a distinct cultural policy -- such as the French Canadians in Quebec. (As the cases of Quebec and Bosnia show, membership in a federation may still fail to fulfill the nationalist aspirations of many people.)
\item The danger always exist in a large country that a unitary government may be too remote for appropriate democratic participation by regional centers located away from the capital; a federal system encourages -- indeed demands -- regional participation in governance.
\item The danger exists in a unitary government that the stronger regions, the larger racial groups, or more powerful interests will provide insufficient allowance for the needs of minorities or weaker groups.
\end{enumerate}
\end{solution}

\question What are the provisions in Nepal relating on the "Right to Information" ? Discuss in detail.

\begin{solution}
Government of Nepal enacted the Right to Information Act, 2064 (2007)

\textbf{Preamble}

1. To make the functions of the state open and transparent in accordance with the democratic system and to make responsible and accountable to the citizen
2. To make the access of citizens simple and easy to the information of public importance held in public bodies
3. To protect sensitive information that could make adverse impact on the interest of the nation and citizen, and for the necessity to have legal provisions to protect the right of the citizen to be well-informed and to bring it into practice.

\textbf{Right to information}

1. Every citizen shall have the right to information subject to RTI act.
2. Every citizen shall have access to the information held in the public bodies.

\textbf{Provision regarding commission}

1. An independent national information commission will be established for the protection, promotion and practice of right to information.
2. A Chief Information Commissioner and two other information commissioners will be in the commission.
3. In order to appoint Chief Information Commissioner and Information Commissioners, there will be a committee comprised as follows:
  a. The speaker -- Chairman
  b. Minister of state minister for information and commission -- Member
  c. President, federation of Nepalese journalists -- Member
4. The Government of Nepal on the recommendation of the Committee under Sub-Section (3), will appoint a Chief Information Commissioner and Information Commissioners; while appointing in this way, at least one female will have to be included.
5. The Committee in accordance with Sub-Section (3) will follow inclusive principles as much as possible with recommending for appointment of Chief Information Commissioner and Information Commissioners.
6. The procedures of recommendation in accordance with Sub-Section (3) will be as determined by the Recommendation Committee itself.

Why RTI

RTI Act 2007 gives you the right to ask for and receive information about any activity from a public body. Public bodies include the following entities:

1. Bodies formed under the constitution
2. Bodies established by the Act
3. Bodies constituted by the government of Nepal
4. Public service provider organizations or foundations established as per the law
5. Political party and organization registered under prevalent laws
6. Organized institutions under the full or partial ownership or under control of the GoN, or organized bodies receiving grants from the GoN
7. Organized institutions formed by the Bodies established by the GoN or any other agencies lawfully formed through an agreement
8. NGOs/Institutions directly or indirectly funded by the GoN or foreign government or international organizations/institutions
9. Other bodies or institutions defined as public bodies by the GoN through a notice published in Nepal gazette.

The seeker of information has to submit a written application to the concerned information officer. It is important for the applicant to clearly specify the information sought. Unclear requests make things difficult to provide correct information. Hence, requests for information should be as clear and specific as possible.

No fees will be charged for the first half hour of monitoring in public bodies or getting information from work place, getting records, or undertaking records inspection and observation. After that the charge in Rs 50 per person, per hour. Likewise, there are no fees for inspection in public libraries and public places where information is made available.

Public bodies will nominate a staff as an information officer for the public's ease of access to information. The head of the office should provide information to the PIO to disseminate it in accordance to sub-section (I) of RTI act. An information section will be set up in all public bodies to disseminate information. The PIOs regularly update their information and keeps records regarding their work plans, expenditure, financial transactions, rules, regulations, laws, and directives. The RTI act anticipates that public bodies store (archive) their information going as far back as 20 years from the commencement of the RTI act. Unless there are resonable and sufficient grounds to validate those exceptions, public bodies cannot evade the responsibility to disclose information. Following information, however, shall not be obliged to be made available:

1. Information harmful to soverignty, integrity, national security public security, law and order of the country; and information that can jeopardize international relations;

2. Information that could have direct bearing on research, investigation and prosecution of a criminal case

3. Information harmful to economic, financial, monetary wellbeing or posing threat to someone's intellectual property rights, as also to banking and financial secrecy.

4. Information that could jeopardize the harmonious realations subsisting between various castes and communities.

5. Information posing threat to a person's right to privacy, personal security, health and property.


In case decision rendered by the Chief of Public body is not satisfactory, one can file the appeal (within 35 days of decision received) to the commission either by directly approaching the commission office or getting it registered in the concerned office or through the concerned District Administration Office. Alternatively, one can also mail the appeal directly to NIC, via any postal services.

Centre for Law and Democracy, Canada has published a comparative study report on existing RTI laws of 111 countries in the world, in which Nepal has been ranked 27th in the world with the score of 104 out of 150. The worldwide experience shows that RTI supports good governance and helps to reduce corruption. The National Information Commission had completed its first RTI Audit in 2016 through ex-secretaries of the government. The second RTI Audit was completed in 2017 through senior journalists.

After analysing all aspects, it was found that 11 ministries (35\%) were found working as per law. Similarly, 20 ministries (65\%) were partially satisfactory. The commission has made 10 suggestions, including adding "imparting information" in Article 27 of the constitution, taking out of the word "secret" from the employees' oath, amongst other, to the government on the basis of the results of this RTI audit.

\end{solution}

\question According to Army Act, 2063 what are the punishment provisioned through Court Marshall ?

\begin{solution}
According to the Army Act, 2063 (Chapter-11; Section (dafa) 101), 

1. Following penalties may be imposed by court marshall:

\begin{itemize}
\item Life imprisonment along with confiscation of entire property,
\item Confiscation of his or her entire share of ancestral property,
\item Imprisonment of upto Fourteen years,
\item Removing from service without prejudicing eligibility for any governmental service in the future, or dismissing with ineligibility for any governmental service in the future,
\item Demotion to a lower rank,
\item Freezing of promotion for Five years,
\item Deduction of salary and allowances for Three months in case the offence has been committed in the course of military action,
\item Freezing of salary and allowances until the recovery of compensation,
\item Freezing of salary increments for a period not  exceeding Two years,
\item Warning (Nasihat).
\end{itemize}

2. The following penalties shall be imposed for the following offences:

\begin{enumerate}
  \item Life imprisonment along with confiscation of entire property for the commission of offences as referred to in Section 38, 39 or Clause (a) of Sub-section (1) of Section 42.
  \item A maximum Fourteen years of imprisonment shall be imposed for the commission of the following offences:
  \begin{enumerate}
    \item Offences provided in Section 40,
    \item Offences provided in Section 41 and Clause (a) of Section 45 during a military operations,
    \item Offences provided in Section 43 or 46 if committed intentionally,
    \item Offences provided in Clause (a) or (b) of Section 44 committed against a person while discharging his/her duties during a military operation,
    \item Offences provided in Section 47.
  \end{enumerate}
  \item A maximum Ten years of imprisonment shall be imposed for the commission of the following offences:
  \begin{enumerate}
    \item Offences provided in Clauses (a) or (b) of Section 44 committed against a person not during the military operation,
    \item Offences provided in Section 48, 49 or 50, (3) Offences provided in Clause (a) of Section 51 or Clause (a) of Section 52.
  \end{enumerate}
  \item The penalty of confiscation of entire property or maximum of Seven years of imprisonment shall be imposed for the commission of offences provided in Clause (b) of Sub-section (1) of Section 42.
  \item A maximum of Seven years of imprisonment shall be imposed for the commission of the following offences:
  \begin{enumerate}
    \item Offences provided in Section 41 committed not during military operation,
    \item Offences provided in Section 46 that are committed unintentionally,
    \item Offences provided in Clauses (b), (c), (d), (e), (f) or (g) of Section 52,
    \item Offences provided in Section 43, 54 or 63.
  \end{enumerate}
  \item A maximum of Five years of imprisonment shall be imposed for the commission of the following offences:
  \begin{enumerate}
    \item Offences provided in Clause (c) of Section 44, (2) Offences provided in Clause (a) of Section 45 committed not during the period of a military operation,
    \item Offences provided in Clauses (b) or (c) of Section 51 or Sections 45, 56 or 57.
  \end{enumerate}
  \item A maximum Three years of imprisonment shall be imposed for the commission of offences provided in Clauses (c) or (d) of Sub-section (1) of Section 42 or Sections 58, 59 or 60.
  \item A maximum of Two years of imprisonment shall be imposed for the commission of the following offences:
  \begin{enumerate}
    \item Offences provided in Clauses (e), (f), (g) or (h) of Sub-section (1) of Section 42,
    \item Offences provided in Clause (b) of Section 45,
    \item Offences provided in Clause (h) of Section 52 committed by personnel below the rank of officer while discharging duties during a military operation,
    \item Offences provided in Section 43 that are committed unintentionally,
    \item Offences provided in Section 61.
  \end{enumerate}
  \item A maximum of Six months imprisonment shall be imposed if a personnel below the rank of officer commits offences provided in Clause (h) of Section 52 except in the period of military operation or while on duty.
  \item A person who commits offences provided in Clauses (i) and (j) of Section 52 shall be removed from the service.
  \item A person who commits offences provided in Clause (k) of Section 62 shall be barred from Two salary increments or promotion for Two years.
  \item A person who attempts to commit such offences as referred to in Chapter-7 or induces another person to commit such offences shall, unless there is specific provision in this Act, be imposed the following penalties:
  \begin{enumerate}
    \item A maximum of Fourteen years of imprisonment for attempting to commit an offence for which a penalty of life imprisonment along with confiscation of entire property may be imposed, and half of the maximum (range) imprisonment for other offences which have been attempted,
    \item Equal penalty for inducement (incitement) to commit an offence as imposed for committing an offence,
    \item A maximum of Fourteen years of imprisonment for inducement to commit an offence which has not been committed (completed), and half of the maximum (range) punishment for inducement to commit any other offence which had not actually been committed (completed).
    \item If a person who committed an offence provided in Clause (b) of Sub-section (1) of Section 42 has absconded and is not found (traced) does not produce him/herself within Thirty Five days, his/her partition share to ancestral property shall be frozen from being transacted. If such person does not appear him/herself or could be found through making a search within Three years, his/her partition share to ancestral property shall be confiscated and auctioned, and the money therefrom shall be deposited to the Government Treasury.
    \item After confiscation of partition share to ancestral property, if the  offender is found, he/she may be imposed life imprisonment along with confiscation of the entire property.
    \item A person who is dismissed from service and disqualified for any governmental job in future pursuant to Clause (j) of Sub-section (2), shall not be paid his/her remaining salary or allowances and pension which he/she is entitled and other facilities except the money accumulated in his/her provident fund.
  \end{enumerate}
\end{enumerate}

Besides these there is Battlefield penalty (Section 104) and Departmental action (Section 105) and Collective fine (Section 107).

\end{solution}

\question What provisions and guidelines have Forest act, 2049 set for utilization and management of forest in Nepal? Mention.

\begin{solution}

The Forest Act, 1992, defines national forest, state-managed forest, protected forest, community forest, leasehold forest, religious forest and private forest as well as ownership of different types of forest, and their management. The Act encourages people to own and to use forest, rather than depend on national or state-owned forest. The Forest Regulation, 1994, fixes users' price rates for timber, fuel-wood, medicinal plants and herbs, flowers and furs, fruits and seeds, bark, logs, and allied forest products. Likewise, charges for animal and cattle grazingare also fixed by the regulation.

Similarly, separate regulations for national parks such as the Royal Chitwan National Park and Royal Bardia National Park make it possible to charge different fees for entrance and recreational activities such as safari, hunting game species and fishing. Another milestone in the forestry sector is the Buffer Zone Management Regulation, 1995, which aims to harmonize relation between national parks and the people around those parks. This regulation also encourages the communities around parks to form users' groups so that they can mange and use the buffer zone resources.

\end{solution}

\question What are the current acts and laws in Nepal ?

\begin{solution}
\begin{enumerate}
\item Nepali Language Publication Corporation Act, 2021 (1964)
\item Nepal Ship (Registration and Logbook) Act, 2027 (1971)
\item Nepal Special Service Act, 2042 (1985)
\item Non-resident Nepali Act, 2064
\item Passports Act, 2024 (1967)
\item The Consumer Protection Act, 2075 (2018)
\item Work Operation Fund Act, 2043 (1986)
\item Water Tax Act, 2023 (1966)
\item Nepal Treaty Act, 2047 (1990)
\item Supreme Court Act, 2048 (1991)
\item Regional Development Plans (Implementation) Act,2013 (1956)
\item Rajya Rajauta (Petty States and Rulers) Act, 2017 (1961)
\item Periodic Tax Recovery Act, 2012 (1955)
\item The University Grants Commission Act, 2050 (1993 A.D.)
\item The Stamp Duty Act, 2019 (1963)
\item The Pesticides Act, 2048 (1991)
\item Radio Act, 2014 (1957)
\item Act Restricting Investment Abroad, 2021(1964)
\item Act to Provide for Enhancement of the Circulation of Nepalese Currency, 2014(1957)
\item The Administrative Procedures (Regulation) Act, 2013 (1956)
\item Animal Health and Livestock Services Act, 2055 (1999)
\item Animal Slaughterhouse and Meat Inspection Act, 2055(1999)
\item Anti Defection Act, 2054 (1997)
\item Contingency Fund Act, 2016 (1959)
\item Drugs Act, 2035, (1978)
\item Nepal Tourism Board Act, 2053
\item Public Security Act, 2046 (1989)
\item Associations Registration Act, 2034 (1977)
\item AUDIT ACT, 2048 (1991)
\item banking-offence-and-punishment-act-2064-2008
\item Child Labor (Prohibition and Regulation) Act, 2056 (2000)
\item Commissions of Inquiry Act, 2026 (1969)
\item Infectious Disease Act, 2020(1964)
\item Immovable Property Requisition Act, 2013 (1956)
\item Human Trafficking and Transportation (Control) Act, 2064
\item The Copyright Act, 2059 (2002)
\item Development Board Act, 2013 (1956)
\item Domestic Postal Goods (Insurance) Act, 2019 (1962)
\item Foreign Investment Tax Act, 2019 (1963)
\item Judicial Fund Act, 2043 (1986)
\item Nepal Citizenship Act, 2063 (2006)
\item Essential Services Operation Act, 2014 (1957)
\item Export and Import (Control) Act, 2013 (1957)
\item Essential Goods Protection Act, 2012 (1955)
\item Extradition Act, 2045 (1988)
\item Legal Aid Act, 2054 (1997)
\item Nepal Agency Act 2014
\item International Financial Transactions Act, 2054 (1998)
\item National Judicial Academy Act, 2063 (2006)
\item Mother's Milk Substitutes (Control of Sale and Distribution) Act, 2049 (1992)
\item Compensation Relating to Torture Act, 2053 (1996)
\item Multimodal Transportation of Goods Act, 2063(2006).
\item National Trust for Nature Conservation Act, 2039 (1982)
\item Electricity Act, 2049 (1992)
\item Interpretation of Laws Act, 2010(1954)
\item Lottery Act, 2025 (1968)
\item Nepal Electricity Authority Act, 2041 (1984)
\item Constitutional Council (Functions, Duties, Powers and Procedures) Act, 2066 (2010)
\item Feed Act, 2033 (1976)
\item Jhora Area Land Act, 2028 (1971)
\item National Parks and Wildlife Conservation Act, 2029 (1973)
\item National Guidance Act, 2018 (1961)
\item Loan and Guarantee Act, 2025(1968)
\item Narcotic Drugs (Control) Act, 2033 (1976)
\item Motor Vehicle Tax Act, 2031 (1974)
\item Nepal Health Research Council Act, 2047 (1991)
\item Competition Promotion and Market Protection Act, 2063 (2007)
\item Evidence Act, 2031
\item Money Laundering Prevention Act, 2063 (2008)
\item Motion Picture (Production, Exhibition and Distribution) Act, 2026 (1969)
\item Arbitration Act, 2055 (1999)
\item Domestic Violence (Offence and Punishment) Act, 2066 (2009)
\item Donations Act, 2030 (1973)
\item National Cooperatives Development Board Act, 2049 (1992)
\item Insurance Act, 2049 (1992)
\item Aquatic Animal Protection Act, 2017(1960)
\item National Dairy Development Board Act, 2048(1992)
\item Explosives Act, 2018 (1961)
\item Essential Commodities Control (Authorization) Act, 2017 (1961)
\item Impeachment (Regulation of Procedure) Act, 2059 (2002)
\item Birth, Death and Other Personal Events (Registration) Act, 2033 (1976)
\item The Foreign Investment and Technology Transfer Act, 2019 (2075)
\item House and Land Tax Act, 2019 (1962)
\item The Industrial Enterprises Act, 2049 (1992)
\item Nepal Standards (Certification Mark) Act, 2037 (1980)
\item National Tea and Coffee Development Board Act, 2049 (1993)
\item Bonded Labour ( Prohibition )Act, 2058 (2002)
\item Diplomatic Privilege and Immunities of the Foreign States and Representatives Act, 2027 (1970)
\item The National Broadcasting Act, 1993
\item Industrial Enterprises Development Institute Act, 2053 (1996)
\item Lumbini Development Trust Act, 2042 (1985)
\item Act Prohibiting the Charging and Taking of Amount (Bhatti Charsa etc.) by Birta-holder on Birta, 2015 (1958)
\item Act Relating to Institutions Acting as Financial Intermediary, 2055 (1999)
\item Act Relating to Land of Jhora Area, 2028 (1971)
\item Nepal Agricultural Research Council Act, 2048(1992)
\item Agriculture and Forestry University Act, 2067 (2010)
\item Nepal Water Supply Corporation Act, 2046 (1989)
\item Nepal Trust Act, 2064 (2008)
\item Nepal Pharmacy Council Act, 2057 (2000)
\item Nepal Nursing Council Act, 2052 (1996)
\item Nepal Rastra Bank Act, 2058 (2002)
\item Nepal Medical Council Act, 2020 (1964)
\item Nepal Health Professional Council Act, 2053
\item Nepal Bar Council Act, 2050 (1993)
\item Nepal Law Commission Act, 2063 (2007 )
\item Nepal Airlines Corporation Act, 2019 (1963)
\item National Human Rights Commission Act, 2068 (2012)
\item National Insurance Corporation Act, 2025 (1968)
\item Nepal Administrative Staff College Act, 2020 (1982)
\item National News Agency Act, 2019 (1962)
\item National Foundation for Upliftment of Aadibasi/Janjati Act, 2058 (2002)
\item Nepal Music and Dance Academy Act, 2064 (2007)
\item Mines and Minerals Act, 2042 (1985)
\item Mediation Act, 2068 (2011)
\item Archives Preservation Act, 2046 (1989)
\item International Finance Corporations Membership Acquisition Act, 2022 (1965)
\item Corporations Act, 2021 (1964)
\item Black Marketing and Some Other Social Offenses and Punishment Act, 2032 (1975)
\item Caste Based Discrimination and Untouchability Offence and Punishment Act, 2068 (2011)
\item Forest Act, 2049 (1993)
\item Construction Business Act, 2055 (1999)
\item Bretton Woods Agreement Act, 2018 (1961)
\item Far-western University Act, 2067 (2010)
\item Negotiable Instruments Act, 2034 (1977)
\item Mid-western University Act, 2067 (2010)
\item Gorkhapatra Corporation Act, 2019 (1963)
\item Hotel Management and Sale and Distribution of Liquors (Control) Act, 2023 (1966)
\item The B. P. Koirala Institute of Health Sciences Act, 2049 (1993)
\item The Greater Janakpur Area Development Council Act, 2055 (1998)
\item Iodized Salt (Production, Sale and Distribution) Act, 2055(1998)
\item Amount and Governmental Contract Arrangement Act, 2020(1963)
\item Insolvency Act, 2063 (2006)
\item Land Acquisition Act, 2034 (1977)
\item Kathmandu University Act, 2048 (1991)
\item International Development Association’s Membership Acquisition Act, 2019 (1962)
\item Land Revenue Act, 2034 (1978)
\item Martyr Gangalal National Heart Disease Centre Act, 2057 (2000)
\item Lumbini Bouddha University Act, 2067 (2006)
\item Land (Survey and Measurement) Act, 2019 (1963)
\item Local Administration Act, 2028 (1971)
\item Local Development Training Academy Act, 2049 (1993)
\item International Centre for Integrated Mountain Development Act, 2040 (1983)
\item The Human Body Organ Transplantation (Regulation and Prohibition) Act, 2055 (1998)
\item The Guthi Corporation Act, 2033 (1976)
\item Ancient Monument Preservation Act, 2013 (1956)
\item The Electronic Transactions Act, 2063 (2008)
\item Armed Police Force Act, 2058 (2001)
\item Arms and Ammunition Act, 2019 (1962)
\item Ayurveda Medical Council Act, 2045 (1988)
\item The B.P. Koirala Memorial Cancer Hospital Act, 2053(1997)
\item Birta Abolition Act, 2016 (1959)
\item The Building Act, 2055
\item Nepal Chartered Accountants Act, 2053 (1997)
\item Citizen Investment Trust Act, 2047 (1991)
\item Commission for the Investigation of Abuse of Authority Act, 2048 (1991)
\item Communication Corporation Act, 2028 (1972)
\item Compensation Act, 2019 (1963)
\item The Lands Act, 2021(1964)
\item Construction Business Act, 2055 (1999)
\item Corporations Act, 2021 (1964)
\item Customs Act, 2064 (2007)
\item Decorations Act, 2064 (2007)
\item Education Act, 2028 (1971)
\item Electricity Theft Control Act, 2058 (2002)
\item Excise Duty Act, 2058 (2002)
\item Nepal Fine Arts Academy Act, 2064 (2007)
\item Privatization Act, 2050 (1994)
\item Water Resources Act, 2049 (1992)
\item Private Forests Nationalization Act, 2013 (1957)
\item The Social Welfare Act, 2049 (1992)
\item Notary Public Act, 2063
\item Right to Information Act, 2064 (2007 )
\item Trade Union Act, 2049 (1992)
\item Tax Settlement Commission Act, 2033 (1976)
\item Telecommunications Act, 2053 (1997)
\item Statistics Act, 2015 (1958)
\item Working Journalists Act, 2051(1993)
\item Wealth Tax Act, 2047 (1990)
\item Water Tariff Fixation Commission Act, 2063 (2006)
\item Water Supply Management Board Act, 2063 (2006)
\item Nepal Veterinary Council Act, 2055 (1999)
\item Ukhada Act, 2021 (1964)
\item Tribhuvan University Act, 2049(1992)
\item Town Development Fund Act, 2053 (1997)
\item Town Development Act, 2045 (1998)
\item Road Board Act, 2058 (2002)
\item Postal Act, 2019 (1963)
\item Partnership Act, 2020 (1964
\item Police Act, 2012 (1955)
\item The Patent, Design and Trade Mark Act, 2022 (1965)
\item Special Court Act, 2059 (2002)
\item [The] Kathmandu Valley Development Authority Act, 2045 B.S. (1988)
\item Technical Education and Vocational Training Council Act, 2045 (1988)
\item Summary Procedures Act, 2028 (1972)
\item Standard Measurement and Weight Act, 2025 (1968)
\item Sports Development Act, 2048 (1992)
\item Some Public Documents Authentication (Procedures) Act, 2063 (2006)
\item Solid Waste Management Act, 2068 (2011)
\item Soil and Watershed Conservation Act, 2039(1982)
\item Social Practices (Reform) Act, 2033 (1976)
\item Ship Registration Act, 2027 (1971)
\item Senior Citizens Act, 2063 (2006)
\item Seeds Act, 2045 (1988)
\item Tourism Act, 2035 (1978)
\item Tobacco Products (Control and Regulatory) Act, 2068 (2011)
\item [The] Ownership of Joint Housing Act, 2054 B.S. (1997)
\item Security of the Health Workers and Health Organizations Act, 2066 (2010)
\item Secured Transaction Act, 2063 (2006)
\item Scholarships Act, 2021 (1964)
\item Revenue Tribunal Act, 2031 (1974)
\item Securities Act, 2063 (2007)
\item Press Council Act 2048 (1992) 
\item Patan Academy of Health Sciences Act, 2064 (2008)
\item The Prevention of Corruption Act, 2059 (2002 A.D)
\item Purwanchal University Act, 2050 (1994)
\item Public Service Commission Act, 2066 (2010)
\item Public Roads Act, 2031 (1974)
\item The Public Procurement Act, 2063 (2007)
\item Public Debt Act, 2059 (2002)
\item Prisons Act, 2019 (1963)
\item Remuneration, Conditions of Service and Facilities of the Authorities of the Constitutional Bodies Act, 2053 (1997)
\item Press and Publication Act, 2048 (1991)
\item Poverty Alleviation Fund Act, 2063 (2006)
\item Pokhara University Act, 2053 (1996)
\item Plant Protection Act, 2064 (2007)
\item Nepal Petroleum Act, 2040 (1983)
\item Pasture Land Nationalization Act, 2031 (1974)
\item Mutual Legal Assistance Act, 2070 (2014)
\item Judicial Council Act, 2073 (2016)
\item Employee Provident Fund Act, 2019 (1962)
\item An Act to Regulate and Control International Trade in Endangered Wild Fauna and Flora
\item Administration of Justice Act, 2073 (2016)
\item Pashupati Area Development Trust Act, 2044 (1987)
\item Karnali Academy of Health Sciences Act, 2068 (2011)
\item Nepal Health Service Act, 2053 (1997)
\item Army Act, 2063 (2006 )
\item Civil Service Act, 2049 (1993)
\item Motor Vehicles and Transport Management Act, 2049 (1993)
\item Passports Act, 2024 (1967)
\item The Crime Victim Protection Act, 2075 (2018)
\item The Privacy Act, 2075 (2018)
\item The Public Health Service Act, 2075 (2018)
\item The Right to Employment Act, 2075 (2018)
\item The Act Relating to Rights of Persons with Disabilities, 2074 (2017)
\item The Right to Food and Food Sovereignty Act, 2075 (2018)
\item The Right to Housing Act, 2075 (2018)
\item The Right to Safe Motherhood and Reproductive Health Act, 2075 (2018)
\item The Social Security Act, 2075 (2018)
\item The Act Relating to Children, 2075 (2018)
\item The Act Relating to Compulsory and Free Education, 2075 (2018)
\item Bank and Financial Institution Act, 2073 (2017)
\end{enumerate}
\end{solution}

\question Write about the Inland Revenue Department of Nepal.

\begin{solution}

Background

Organizational Information

The Inland Revenue Department (IRD) is currently responsible for the enforcement of Tax Laws and administration of the following taxes: Income Tax, Value Added Tax, Excise Duty and duties like Entertainment fee (Film Development Fee). Value Added Tax Department was merged with Department of Taxation on April 15, 2001. Later it was renamed as Inland Revenue Department since July 16, 2001.
 
IRD is centrally located in Kathmandu. There are 59 field offices throughout Nepal including 1 Large Taxpayers Office, 1 Medium Level Taxpayer Office,  38 Inland Revenue Offices and 19 Taxpayer Service Offices. IRD carries out the following functions:
1. Tax Administration
2. Tax Policy
3. Tax Payer Services
4. Registration, Revenue Collection
5. Tax audit
6. Tax Enforcement and investigation
7. Review & Appeal
8. Tax Refund
9. Advance Ruling
10. Tax Treaty and International Taxation
11. Excises and Liquor Administration

TAXPAYERS RIGHTS AND ASSURANCES

In addition to the specific rights and assurances provided in the concerned Acts, taxpayers in general are assured of the following rights. These assurances will be strictly adhered to. The Taxpayers rights, among others, include:

A. Each taxpayer is treated with due respect and honor.

B. Each taxpayer will get an immediate receipt of return submission whilst a receipt of other correspondences is to be provided in ONE hour.

C. Each taxpayer willing to seek an advance ruling may apply either in the Department or through concerned Inland Revenue Office. There is a legal provision that a decision should be made within 45 days after the request is made.

D. Each taxpayer can represent or make any correspondence through an authorized person including legal assistants and the accounts specialist or auditors just in case if he/she is unable to represent or make any correspondence.

E. Each taxpayer is to get copies of the documents related with him/her or submitted to the IRO or documents or decisions affecting her/him generally within TWENTY FOUR hours of his/her request in the concerned IRO.

F. Each taxpayer is to get refund of his excess input tax (VAT) and other taxes. The exporter will get refund generally within 15 days in IROs and 30 days in IRD upon receiving application for refund along with required documents.

G. Each taxpayer is to get certificate/renewal of Excise/Liquor business generally within TWENTY FOUR hour after the request is made.

H. Each taxpayer is to get sales and purchase books certified by the Tax Officer generally within TWENTY FOUR hours after the request is made.

I. Each taxpayer is to get initials of the Tax Officer on stock transfer or sales to the non-registrants generally within TWENTY FOUR hours after the request is made.

J. Each taxpayer is to get Permanent Accounts Number (PAN) or the Certificate generally within 3 days after the request is made, and

K. Each taxpayer is to get Tax Clearance Certificate generally within TWENTY FOUR hours after the request is made.

\end{solution}

\end{questions}

## Events and programmes

### Nepal Investment Summit

Nepal hosted 3rd edition of Nepal Investment Summit on March 29-30, 2019 in the Kathmandu in an effort to further enhance Nepal's standing as a promising investment destination and to build on the recent encouraing developments. The 2nd summit was held in 2017, while the first one was held in 1992. 

Major Outcomes of Investment Summit, 2019

- Nepal Investment Summit 2019 has reportedly witnessed signing of 15 new deals and MoUs for new projects in Nepal
- Besides, the country has also received proposals, 17 applications for 10 other projects presented as part of the Summit
- Transport, hydropower, hospital, airport, convention center and agriculture infrastructure were among areas that received investment proposals
- India, Vietnam, United States, South Korea, China, Britain were among countries that represented participating investors
- International Finance Corporation, World Bank Group, Asian Development Bank, Multilateral Investment Guarantee Agency, were among the key multilateral donor agencies entered into separate investment and cooperation deals during the summit
- One of the world's largest solar development company SkyPower announced a new partnership with Nepal's Chaudhary Group for a 600-MW capacity renewable solar energy projects.
- PM and Finance Minister Khatiwada held 30 side-line bilateral meetings with investors and delegates as part of the Summit
- More than 1300 participants and around 735 foreign delegates from more 300 companies attended the Summit


The government, in an effort to attract foreign investors, has introduced three vital legislations—Foreign Investment and Technology Transfer Act, Public-Private Partnership and Investment Act, and Special Economic Zone Act (First Amendment).

### Nepal Infrastructure Summit

Nepal Infrastructure Summit (NIS) 2014, organized by CNI in collaboration with Government of Nepal and YCNC, with the theme "Accelerating Investment in Infrastructure" was 1st ever summit organized of its kind in Nepal. The Summit was successful as it was able to deliver high attendance, important and high-rated keynote speech, excellent speakers and panelists who knew their subject, well facilitated dialogue sessions, a great venue for networking and the presence of aspiring young participants.

The Summit provided a platform for exchange of relevant experience in selected topics including:

\begin{enumerate}
\item Outlook and Opportunities
\item Sustainable Infrastructure Development for Socio-Economic Growth
\item Investment Climate in Nepal
\item Infrastructure Financing Mechanisms
\item Enabling Policy Initiatives
\end{enumerate}

Accordingly, CNI has also been focusing on the resilience of the infrastructure of the country as the result of which CNI came out with a policy document of "Construction of Eco-Friendly Linear Infrastructure" with the assistance WWF Nepal. In nutshell, after the NIS 2014, the activities performed were:

- Formation and effective operationalization of Infrastructure Cell in CNI.
- Follow up workshops/seminars.
- Workshop on Infrastructure Financing - Opportunities and Challenges.
- Series of workshop on "Public-Private Partnerships".
- Workshop on financing of storage-based hydroelectricity projects
- Lobby on various issues:
  - Establishment of Infrastructure Development Bank.
  - Establishment of Project banks/Project Readiness.
  - Public-Private Partnership (PPP) Centre.
  - Other legal and institutional issues.

The 2nd Nepal Infrastructure Summit hosted by the Confederation of Nepalese Industries (CNI) facilitated discourse on the theme "Private Sector as indispensable Partner" for infrastructure development. The primary objective of the summit was to attract large scale private investment in areas of Nepal’s core needs and strengths; forge stronger networks and alliances in the region for seamless connectivity; and raise awareness on why investment in infrastructure needs an urgent push in the country.

The summit focused on the strategic vision of the infrastructure, value proposition, competitiveness, policy particularly on land acquisition and environmental clearance; and regulatory enablers and business environment. CNI in collaboration with Idea Studio will collect various infrastructure ideas from different sectors of the communities to come up with a creative and an innovative **infrastructure** (transportation, irrigation, energy, housing, digital, etc) **ideas** to solve the existing epic infrastructure deficit and address to the possible future need for our country’s development.

The Summit provided a platform for exchange of relevant experience in selected topics including:

- Infrastructure: The Big Picture
- Barriers and Constraints in attracting Private and Public Investment
- Expediting Public Private Partnership (PPP)
- Financing Resilient Infrastructure
- Successful PPP Model: Global and regional Perspective

### Relationship between climate change and energy use

- Energy use has close ties with climate change. India and china are among the greatest emitters of greenhouse gases and consumers of fossil fuel, and it is predicted that their consumption of coal as major source of energy will continue to rise for at least a decade (IEA, 2019).
- Into 2016 and we still have 80% of our total energy requirement met from fossil fuels (mostly, oil, coal and natural gases). International energy agency's report says that 30% of all energy-related carbon dioxide emission is related to coal as fuel, making it the biggest single contributor (2019).
- An individual effort to parsimoniously treat energy resources available to us determines individual footprint. We could choose to reuse a battery/cell, with minor power drops, in alternative electronic items compatible with it or just treat at a refuse at our home. The latter activity with unforeseen consequences might lead to the battery being disposed improperly and polluting an ecosystem due the leakage of its electrolyte chemicals.
- While electricity is lighting up most houses, its generation has been a markedly different process worldwide. The process basically depends with the resource endowment of a nation in their approach to choosing ways to explore it. Coal, for example, is used largely in China and the US despite its much tauted big contribution in greenhouse emission.


## Impacts of climate change

What are the impacts of climate change ? Discuss on detail about the initiative taken by Nepal Government in tackling the impacts of climate change.

Climate change is now affecting every country in every continent. It is disrupting national economies and affecting lives, of individuals, communities and countries today and is likely to do so in the ensuing future. Weather pattern are changing, snowline are melting due to global warming and sea level are rising. 

**Negative impacts**

1. Numerous mountains are seeing accelerated the melting of snow due to global warming.
2. Changes in monsoon patterns, such as erratic rainfall with alternate drought periods are the serious impacts of climate change in Nepal which hampers the agricultural production. The changes in monsoon pattern in Nepal greatly exacerbate the situation of poverty and food insecurity in the country.
3. Flood, landslides, forest fires, snow melt are the disasters that occur as a result of climate change without any prewarning and causes heavy loss.
4. High hills and mountains are more vulnerable to climate change compared to terai, thus climate change make large scale economic and ecolgical disruption in upper latitudes.
5. Loss of plant and animal biodiversity.
6. Increase in number of insects, pest and diseases and sporadic outbreaks becoming much frequent.

The positive impact of global warming, a characteristic feature of climate change, is that the crops that could in the past be only growth in certain low altitudes, requiring prolonged exposure to heat can now be growth in higher regions such as Manang and Mustang districts of Nepal. The is evidently due to increase average temperature of the region, thus allowing the crop to perform well. However, as noted earlier, there are more prominent negative impacts than the benifits we can derive from the climate change.

Nepal became signatory of UN Framework Convention on Climate Change (UNFCC) in 1994. Ever since then the issue of climate change has been routinely incorporated into policy and strategic frameworks by Government of Nepal. Following policies and strategies are currently in place to help mitigate the impacts of climate change in Nepal:

1. Constitutional approach

Article 30(1) of the constitution of Nepal guarantees a "clean environment" as fundamental right and elaborates that "every citizen shall have the right to live in a clean and healthy environment." Article 30(2) of the constitution also encourages the state to formulate necessary legal framework to balance environment and development.

2. Approaches in periodic plans
3. Climate change policy, 2067 (2011)
4. National climate change policy, 2076 (2019)
6. National adaptation plan of action (NAPA, 2010)
7. Framework for local adaptation plan for action (LAPA, 2011)

## Highlights of NCCP, 2076

Reasons for Nepal's vulnerability to climate change are Topographical fragility, fragile geological structure, sensitive ecosystems and diversity of climate and micro-climate zones prevalence in Nepal. Additionally, poverty, illiteracy, social disparity and high dependence of communities on natural resources for livelihood.

- Nepal National Red Strategy implemented to check greenhouse gases emmission from deforestation and forest destruction.
- Provision of climate change coordination committee in Chairmanship of Prime minister.
- Establishment of Climate change management section under Ministry of Forest and Environment as focal point for implementation of international treaty.
- Provision of Climate budget code by National planning commission.
- Alternative energy promotion center has established climate and carbon unit for executing Clean Development Mechanism (CDM) related activities.

**Goal**

To contribute to socio-economic prosperity of the nation by building a climate resilient society.

**Objectives**

a. To enhance climate change adaptation capacity of persons, families, groups and communities vulnerable to, and at risk of, climate change;
b. To build resilience of ecosystems that are at risk of adverse impacts of climate change;
c. To promote green economy by adopting the concept of low carbon emmission development;
d. To mobilize national and international financial resources for climate change mitigation and adaptation in just manner;
e. To conduct research, make effective technology development and information service delivery related to climate change;
f. To mainstream or integrate climate change issues into policies, strategies, plans and programs at all levels of State and sectoral areas;
g. To mainstream gender equality and social inclusion (GESI) into climate change mitigation and adaptation programs.

**Policies** (Sectoral policy, strategies and working policies)

1. Agriculture and food security
2. Forest, biodiversity and watershed conservation
3. Water resources and energy
4. Rural and urban habitats
5. Industry, transport and physical infrastructure
6. Tourism and natural and cultural heritage
7. Health and drinking water and sanitation
8. Disaster risk reduction and management
9. Gender equality and social inclusion, livelihoods and good governance
10. Awareness raising and capacity development
11. Research, technology development and expansion
12. Climate finance management

Policy number 9-12 are inter-thematic areas.

## Population census, 2068

```{r}
dplyr::tribble(~"Indicator", ~"Unit", ~"Value", 
               "Male", "count", 12849041,
               "Female", "count", 13645463,
               "Annual population growth rate", "percentage", 1.35,
               "Population density", "individuals per msq", 180,
               "Total engagement in agriculture", "percentage", 65.6,
               "Male engagement in agriculture", "percentage", 60.2,
               "Female engagement in agriculture", "percentage", 72.8, 
               "Crude birth rate", "per thousand", 21.8,
               "Crude death rate", "per thousand", 7.3,
               "Child death rate", "per thousand", 40.5,
               "Net fertility rate", "percentage", 2.5, 
               "Average life expectancy", "years", 66.6) %>% 
  knitr::kable(caption = "Population census, 2068 summary", booktabs = TRUE, longtable = T) %>% 
  kableExtra::kable_styling(latex_options = "striped")

```

## Hulaki rajmarga (Hulaki highway)

http://www.karma99.com/
http://www.karma99.com/2015/09/hulaki-rajmarga.html

## Demographics of Nepal

- Refer to wikipedia link \url{https://en.wikipedia.org/wiki/Demographics_of_Nepal}, for demographic statistics of Nepal. This includes
  - Population growth
  - Vital statistics: Live births per year, Deaths per year, Natural change per year, CBR, CDR, NC, TFR, IMR
  - Population structure: Based on age group
  - Ethnicity statisics
  - Languages
  - CIA factbook summary
  - Religion, etc.

## Madan puraskar

Madan puraskar is a literary honor in Nepal which Madan Puraskar Guthi confers annually for an outstanding book in Nepali language published within the calendar year. It is considered the most prestigious literature award in Nepal.

It was established by daughter in law of Prime Minister Chandra Shumsher JBR Queen Jagadamba Kumari Devi, in memory of her husband late General Madan Shumsher JBR, on September 26, 1955. Since then it has been presented every year, except in 1974 and 1977. At the time of its establishment, the award prize was NRS 4000. Since the Golden Jubilee year 2005 the award prize has been increased to NRs. 400,000.

List of winners

```{r winners-list}
# madan_puraskar <- read_html("https://en.wikipedia.org/wiki/Madan_Puraskar") %>% 
#   html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[2]')
# 
# madan_puraskar %>% 
#   .[[1]] %>%
#   html_table() %>% 
#   as_tibble() %>% 
#   write_csv("./madan_puraskar_awardees.csv", na = "")

madan_puraskar <- read_csv("./data/madan_puraskar_awardees.csv")

madan_puraskar %>% 
  knitr::kable(booktabs = TRUE, caption = "List of Madan Puraskar winners till 2075", longtable = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped", font_size = 8) %>% 
  kableExtra::column_spec(column = 2:3, width = c("14em", "24em"))
```


\begin{questions}

\question How does the management information system help in decision making process ?

Management information system (MIS) is an information system used for decision-making, and for the coordination, control, analysis, and visualization of information in an organization.

In public governance, MIS are the tools to promote transparency, accountability and participation. More specifically, recent approaches to delivering bureaucratic services are expeceted to meet their goals of public orientation and result based performace evaluation aided through information system. Establishment of computerised Integrated Financial Management Information System (IFMIS) is being mulled over to implement single platform central accouting system as a part of a public finance management reform initiative by Government of Nepal with support of World Bank. For example, under this system information about expenditure made by the Department of Roads and the revenue collected by the Inland Revenue Department will be disseminated via the same platform. Also the concept of digital National Identity Card based is a part of broader vison of government, currently under delegation of the National ID Management Center (NIDMC) under aegis of Ministry of Home Affairs, to aggregate vital information database of the citizens.

In a corporate setting, the ultimate goal of the use of a management information system is to increase the value and profits of the business. This is done by providing managers with timely and appropriate information allowing them to make effective decisions within a shorter period of time.

While management information systems can be used by any and every level of management, the decision of which systems to implement generally falls upon the chief information officers (CIO) and chief technology officers (CTO). These officers are generally responsible for the overall technology strategy of an organization including evaluating how new technology can help their organization. Once decisions have been made, IT directors, including MIS directors, are in charge of the technical implementation of the system. Upon implementation, the assigned users will have the appropriate access to relevant information. It is important to note that not everyone inputting data into MIS need necessarily be management level.

The role of MIS in public and private governance is further invigorated by widespread networking facilities made possible through internet. There are also multiple cases of MIS platform sharing across multiple business centres to benefit from cross enterprise information exchange.

The MIS represents the electronic automation of several different kinds of counting, tallying, record-keeping, and accounting techniques of which the by far oldest, of course, was the ledger on which the business owner kept track of his or her business. Automation emerged in the 1880s in the form of tabulating cards which could be sorted and counted. These were the punch-cards still remembered by many: they captured elements of information keyed in on punch-card machines; the cards were then processed by other machines some of which could print out results of tallies. 

Waves of innovation spread the fundamental virtues of coherent information systems across all corporate functions and to all sizes of businesses in the 1970s, 80s, and 90s. Within companies major functional areas developed their own MIS capabilities; often these were not yet connected: engineering, manufacturing, and inventory systems developed side by side sometimes running on specialized hardware. Personal computers ("micros," PCs) appeared in the 70s and spread widely in the 80s. Some of these were used as free-standing "seeds" of MIS systems serving sales, marketing, and personnel systems, with summarized data from them transferred to the "mainframe." In the 1980s networked PCs appeared and developed into powerful systems in their own right in the 1990s in many companies displacing midsized and small computers. Equipped with powerful database engines, such networks were in turn organized for MIS purposes. Simultaneously, in the 90s, the World Wide Web came of age, morphed into the Internet with a visual interface, connecting all sorts of systems to one another.

Midway through the first decade of the 21st century the narrowly conceived idea of the MIS has become somewhat fuzzy. Management information systems, of course, are still doing their jobs, but their function is now one among many others that feed information to people in business to help them manage. Systems are available for computer assisted design and manufacturing (CAD-CAM); computers supervise industrial processes in power, chemicals, petrochemicals, pipelines, transport systems, etc. Systems manage and transfer money worldwide and communicate worldwide. Virtually all major administrative functions are supported by automated system. Many people now file their taxes over the Internet and have their refunds credited (or money owning deducted) from bank accounts automatically. MIS was thus the first major system of the Information Age. 

Advantages and disadvantages

The following are some of the benefits that can be attained using MIS:

\begin{itemize}
\item Improve an organization's operational efficiency, add value to existing products, engender innovation and new product development, and help managers make better decisions.
\item Companies are able to identify their strengths and weaknesses due to the presence of revenue reports, employee performance records etc. Identifying these aspects can help a company improve its business processes and operations.
\item Giving an overall picture of the company.
\item Acting as a communication and planning tool.
\item The availability of customer data and feedback can help the company to align its business processes according to the needs of its customers. The effective management of customer data can help the company to perform direct marketing and promotion activities.
\item MIS can help a company gain a competitive advantage.
\item MIS reports can help with decision-making as well as reduce downtime for actionable items.
\end{itemize}

Some of the disadvantages of MIS systems:
\begin{itemize}
\item Retrieval and dissemination are dependent on technology hardware and software.
\item Potential for inaccurate information.
\end{itemize}

source: \url{https://www.inc.com/encyclopedia/management-information-systems-mis.html}; \url{https://en.wikipedia.org/wiki/Management_information_system}

\question What do you mean by inventory management ("Jinsi byabasthapan") ? Highlight about its importance in Nepalese public administration.

\question Write short notes on:
\begin{enumerate}[label*=\alph*., leftmargin=1cm]
\item Tax reduction at source
\item Civil service 
\end{enumerate}

source: \url{https://kathmandupost.com/money/2017/06/07/integrated-financial-info-system-mulled}

\question What are the geographical features of Nepal ? Mention in brief.


\question (Long answer type) What is energy crisis ? What are the measures that can be taken to solve the crisis ? What are the different sectors affected by energy crisis ?

\question Explain briefly on the prospect of hydroelectricity development in Nepal.

\question What is human development index (HDI) ? Explain in brief.

\question Discuss in brief the role of private sector in economic development of Nepal. 

\question What is poverty ? What is the difference between absolute and relative poverty ?

\question How are indigeneous and local ethnic groups prioritized in Nepal ? Mention what strategies are taken by government to that end.

\question (Long answer type) "Foreign employment significantly contibutes to customs earning of Nepal". Briefly elaborate the statement.

\question What are the different types of climates occuring in Nepal based on the geographical features ?

\question Write in brief about the importance of water resources of Nepal.

\question What are the major causes of deforestation in Nepal. Discuss.

\question What is the status of environmental pollution in Nepal ? Discuss about the efforts made by Government of Nepal in reducing effects of environmental pollution and preparing for climate change ?

\question What is meant by standard time ? Mention how is standard time of Nepal determined ?

\question Justify: Unplanned urbanization negatively impacts development of nation.

\end{questions}
