## Essay writing on issue (800 to 1000 words)

1. We live in a world besieged by media and technology. Every sector of our life is impacted by the presence of modern forms of communication system. Write an essay discusses the influence of modern forms of communications on everyday life.

\begin{solution}

The very nature of human being as a social creature underlines the importance of communication in one's day-to-day life. Human beings interact, possibly much effectively and efficiently than any other life forms in this planet. Much of the achievements we are relishing today, from kitchen utensils to high performance computers are all, in a way, fruits of reliable communication. This is to say that innovations are triggered in several ways through communication. It is also true that modern day communication, we use in the form of media and technology, have been made possible by the feats of human innovation -- once again guided by state of the art communication systems.

Early developments in communication technology consisted of invention of telegram and telephone, which extensively relied on electrical signals on copper wire. As the search for better communication infrastructure continued, data (including video and audio) sharing over wireless frequencies and optical cables are on upsurge today. To recapitulate the role of communication channel, we can refer to an event that is purported to have put an end to the World War II. Invention of cipher (a tool that decrypts hidden messages) on those days became a landmark in intercepting German radio communication channels and gaining the information about overhanging threat military assault. In the retrospect, communication development during World War might not seem as much of a feat when we consider high performing broadband facilites that are at our disposal today.

Communication system today pervades all spheres of life be it physical or socio-economic. It wasn't uncommon, just a few decades earlier, for people to commute great deal of distance just for the sake delivering information through postal services. Proliferation of telecommunication services in every major inhabitated part of world has made it non-essential for the paper mail services to physically transport tremendous amounts of exchanges people are making. This trend in frequent exchange of ideas and even general information among peoples -- families, friend, scholars, researchers, etc., is expected to only expand in future. The imperative is that, as means of communication gets cheaper and accessible, people need not be constrained to utilize information wealth due to their physical distance.

More often than not, means of communication exert their importance at times of crisis. Remote areas are remote for more than a reason but one of those consistently includes them not being able to communicate with authorities on time. This is a situation that typically results in greater marginalization, as individuals from distant localities lacking proper communication methods are increasingly unheard of their problems and scarcities. For example, when a disease outbreak occurs in such an area, relief package and medical interventions become delayed, resulting in significant losses of life. In contrast, well-connected areas where government can quickly initiate a rescue operation will have more lives saved.

When looked at for side-effects on an individual, communication affects even the mundane aspects of life. Citizens of 21 century take toolsets of communication for granted. There is little to how every day engagements of families are ran without crediting to modern communication technologies. Grocery stores and dairy shops have the communication network set to have knowledge of market; A dairy handler has to know who are his customers for what products. Daily updates on traffic and weather to work are all televised through various news channels so that commuters can be punctual to their work. Furthermore, in the workplace, all commercial startups communicate over telephone or email to their clients. This, in comparison to early days approach of door to door visit for sales promotion and order acquistion, is a stark gain in efficiency and scale. A supply order could be made over a phone call while also provisioning for accumulated order the previous day for the same household equipment to areas nearby thereby significantly cutting down the cost of repeated transportation.

Our social thread forms a web of close circle of friends and families along with tendencies to keep-on-touch with them. This emotional and social sphere of our lives too is strung more tightly together than ever before with the advent of new communication technologies. Rashid can now ask what the weather is like in Australia to one of his friends from Nepal in a video call, while also hoping some days to see them there in person. His mother can make a quick appointment with a doctor in a phone call after knowing that Birat is ill. In addition to surfing a personalized news-stand online right before the meeting, one is now also able to address a meeting which includes a mass of 50 people. This is made possible due to wide projection and affordable cinema technology which allow people to conduct presentations as well as entertain with movie.

Besides trivial occurances of everyday life, various channels of communication are used by government organizations to address public and state their vision for country. Likewise, several government insitutions have made social media formum also a platform for obtaining public feedback and outcries, while also attempting to address them. For example, if suspected for a criminal activity, goverment can call for a surveillance of an individual or a group through the use of closed circuit cameras (CCTV) or the GPS system of suspected's smartphone. Although for the case it is clear that criminal threat outweights the expectation that government action hurts individual privacy, there can be scenarios where this application of technology can be subjected to debate. Meanwhile, there is always a chance that political groups may misuse media to propagandize their statements to only gain popularity, while masking up facts that give different flavor.

With all the merits it has brought to individual and societal lives, modern communication technologies can well be thought of as integral part of development. While it has connected people for good, indiscriminate retrieval and exchange of some information can have detrimental consequences -- With great power, comes a greater responsibility.

\end{solution}

2. Write an essay that discusses how different cultures maintain their distinctive tradition in your community. In what ways does this cultural diversity have a positive impact on your community ?

3. Diplomacy has been defined as "the management of relations between independent states by the process of negotiation". Write an essay citing some major events of modern international diplomacy that have resolved conflict through the process of negotiation and have shaped the world in which we live.

## International relations

1. Nepal and United Kingdom have celebrated 200 years of their relationship. How do you view this relationship and what recommendations, if you have, you wish to offer to take the subsisting relationship to the next level ?

\begin{solution}

Nepal established diplomatic relations with the United Kingdom in 1816. Treaty of Friendship between Great Britain and Nepal was signed in 1923 which further formalized relations between the two countries. Ever since the establishment of their diplomatic relations, friendship, mutual understanding, cordiality, cooperation and respect for each other's national interests and aspirations have characterized relationship between the two countries. The UK is also the first country in the world with which Nepal had established diplomatic relations and also the first to have established its Embassy in Kathmandu, the capital of Nepal. This is the country where Nepal had established its first diplomatic mission (Legation) back in 1934. It was elevated to the Ambassador level in 1947 AD.

The UK remained one of the top development partners of Nepal with the annual British aid on an increasing trend. Tourism, trade, education, and British Gurkha connection remained the key dimensions of the bilateral relations.

Since then, relations between the two countries have continued to grow, with a new Treay of Perpetual Peace and Friendship signed in 1950 which expanded areas of cooperation and an exchange of State Visits. Amicable relations continue today; Nepal continues to be the source of recruitment of Gurkha soldiers into the British army -- a tradition dating back to the nineteeth century but still an essential part of Britain's modern army -- and the UK remains one of the most significant providers of development assistance to Nepal.

As one of Great Britain's allies during the two world wars, Nepalese soldiers in hundreds of thousands fought and sacrificed their lives in many battlefields of the world, thus leaving a heritage of deep and sincere friendship in the history of two countries. The Gurkha soldiers are the most visible bridge between Nepal and the United Kingdom. It is to state that the Gurkha's service in the British army started on April 24, 1815, which is continuing till the date spanning over more than two hundred years. The solid foundation of the relation is built on the history of service, sacrifice and bravery of these Gurkha soldiers.

Nepal and United Kingdom are celebrating the bicentenary of bilateral relations between two countries. To mark the 200th year of the beginning of official contacts between the two countries in 1816, a number of programs and events from both sides have been organized in 2016-2017. There have been intermittent to and fro high level visits between Nepal and United Kindom since as early as 1960.

\textbf{Treaties/Memorandum of Understandings/Agreements between the UK and Nepal}

Bilateral consultation mechanism

With a view to enhancing the friendly cooperation and further promoting the bilateral relations Nepal and the United Kingdom have signed a MoU on 7th January, 2014. On behalf of Nepal and UK, Foreign secretary Mr. Arjun Bahadur Thapa and UK's Permanent Under Secretary Sir Simon Frazer signed the MoU for establishing the Consultation Mechanism at the Foreign and Commonwealth Office, London respectively. As per the MoU, both the parties will hold consultations at jointly decided intervals alternately in the capital of either country at an appropriate level as may be decided upon and hold discussion on all aspects of bilateral relations as well as on international or regional issues of mutual interest.

Finalized Agreements/MoU

- TIFA Agreement signed in 2011.
- An agreement on promotion and protection of investment between Nepal and UK signed on March 2, 1993
- Development Partnership Agreement between Nepal and UK, signed on December 2013. It supersedes the Technical cooperation agreement desiring to strengthen the traditional cooperation and cordial relation between Nepal and the UK signed on 31 May 1994.
- Trade Agreement signed in 1965.

Agreements/MoU in Progress

- Bilateral Agreement on Avoidance of Double Taxation

\textbf{Economic and Technical Cooperation (Development Cooperation)}

The UK aid to Nepal in various fields of activities started in 1961. The United Kingdom has been offering fellowships to the Government of Nepal since 1950s. British volunteers are engaged in Nepal since 1964. These programmes have contributed to Nepal's need for specialised and trained manpower and also have developed important links between the people of the two countries. In order to better understand and address the issues of poverty in Nepal, in April 1999 Department for International Development (DFID) established an in-country office staffed by a multi-disciplinary team of both UK nationals and staff appointed in country. DFID presence in Nepal has greatly assisted in working with Government of Nepal in the development of understanding and establishment of networks that will promote opportunities for change.

Recently Development Partnership Arrangement to provide a transparent and mutual accountability framework between Nepal and the United Kingdom on development assistance was signed on December 2013. The assistance of British Government generally comes through the Umbrella Agreement. The first such agreement of Pound Sterling 12 million was signed on 23 September 1979 for the implementation of various projects. The UK Government’s Department for International Development (DFID) prepared an Operational Plan for Nepal during 2011-2016. This plan commits up to Pound Sterling 413 million of UK official development assistance during the period 2011-2016. All bilateral aid with Nepal is on a grant basis.

As per Development Cooperation Report 2014-2015 by Ministry of Finance, UK remained the top ODA provider (based on disbursement) with a total assistance of US dollars 168.07 million in the Fiscal Year 2014/2015.

DFID Technical Support has been instrumental in promoting foreign direct investment especially in large hydropower project through Investment Board Nepal for the last few years. This very programme has provisioned of supporting IBN to promote private investment in the country. Likewise there are provisions of continuing support to the second and third phases of Financial Sector Reform Programme of the Government initiated few years back. The program will last for six years from 2015 to 2020.

The DFID Nepal's operational Plan is divided into four main areas: governance and security, inclusive wealth creation, human development (basic services including education and health), and climate change/disaster management. The UK has assisted Nepal in the areas of livelihoods, e.g. agriculture, forestry, transport and communications, local development; basic services, e.g. education, health, water supply and sanitation; good governance, human rights and peace building efforts.

\textbf{Earthquake, 2015}

Nepal received generous support and assistance from the Government of the United Kingdom in the aftermath of the earthquake of 25 April 2015. It responded to the Nepal earthquakes immediately with the deployment of Eight Disaster Response Specialists under Rapid Response Facility and with the release of GBP 5 million. The United Kingdom Government pledged USD 110 million for the reconstruction and rebuilding of Nepal. 

\textbf{Trade}

Nepal and UK signed trade agreement in 1965. Despite our extensive engagements and linkages, both at the level of government and, people in the backdrop of our long-standing friendly and cooperative relations, trade and investment relations between the two countries have largely remained way below its actual potential.

With high imbalance and elements of unpredictability, this is, of course, a typical feature of a trade relation between a developed and least developed country. Nepal faced a trade deficit of 142,182,856 Rupees in the first eight month of 2016. The total export to UK stands at 1,256,018,669 Rupees while the total import stands at 1,398,201,525. It is for this reason that Nepal places high importance to directing foreign development cooperation in areas which help enhancing our productive capacity. Aid for trade is, therefore, another preferred area for us in our development cooperation. 

The major items of import from UK are: Motor car, vehicle, parts of aeroplane and helicopter, Whiskies, Malt not roasted, Sweets biscuits, Chocolate in blocks, slab or bar etc. Major Nepalese exports to UK are Pashmina shawls, goatskin, leather goods, Nepalese paper and paper products, woollen carpets, handicrafts, ready-made garments, silverware and jewellery. Likewise, major imports from UK are copper scrap, hard drinks, cosmetics, medicine and medical equipment, textiles, copper wire rod, machinery and parts, aircraft and spare parts, scientific research equipment, office equipment and stationery.

Exchange of visits by trade delegations from Nepal and Britain has added a new dimension to the commercial relations between Nepal and Britain. Nepal-Britain Chamber of Commerce and Industry is active in promoting trade and investment between Nepal and UK.

\textbf{Tourism}

A sizeable number of British tourists come to Nepal every year for trekking and mountaineering and other leisurely activities. Given our historical linkages and increasing people-to-people contacts, the prospects for increasing the tourist arrival from UK in Nepal are quite promising.

UK's support and cooperation in line with our broader policy priorities will be much appreciated, especially in construction and upgradation of tourism infrastructure with focus on tourist safety. DIFID’s recent involvement and interest in tourist security, trekkers tracking, and maintenance of damaged trekking trails, including the Great Himalayan Trail, is most welcome.

Embassy of Nepal in London is planning to organize the many tourism promotional programme to increase the number of tourists from UK.

\textbf{Consular activities}

Embassy of Nepal in London involves in various consular service ranging from passports, visas, Travel Document, NRN Card, Verification, Attestation and Power of Attorney. During Fiscal Year 2075/76, Embassy of Nepal in London generated GBP 542048.72 (NRS 7,73,71,510.41) revenue.

\textbf{Investment}

Foreign investment commitment from UK stands at 1663 million rupees till 2013/14. The major investment is in the areas of banking, tourism, education, technology. An agreement for the promotion and protection of investment between Nepal and UK was signed on March 2, 1993.

There are some British joint ventures in the areas of hotel, travel and trekking, tea production, education, garment, bio-technology and consultancy. These are significant figures, but do not reflect the true potential. There are many British and Non Resident Nepalese (NRN) entrepreneurs who are still making trade and investment successes in Nepal. Standard Chartered Bank and Unilever are two major British companies who have made a huge return from their investment in Nepal in the last 20 years.

An agreement on promotion and protection of investment between Nepal and UK signed on 02 March 1993 provides a framework to further expand cooperation in this field. In this context, the proposal for signing another agreement on avoidance of double taxation will further help boost the flow of FDI between the two countries.

\textbf{Foreign employment and British Gurkha}

Thousands of Nepalese are working at service sectors like hospitals, university, hotels and restaurants in UK. About 200 Nepalese are recruited as Maritime Security Guards in UK. About 150,000 Nepalese are living in UK. Many Nepalese belong to Ex-Gurkha Servicemen and their families.

British Gurkhas is a fully integrated part of the British Armed Forces. It constitutes an important element in Nepal-Britain relations.  Over 160,000 Gurkhas were enlisted in the Gurkha Brigade and other units of the Indian Army during the First and Second World Wars. In recognition of their distinguished service, the British Gorkhas servicemen from Nepal have won 13 Victoria Crosses (VC), the highest British gallantry honour.

\textbf{Cultural Cooperation}

British Council in Nepal is working to extend cultural relations between UK and Nepal. Similarly, Nepalese diaspora and Gurkha soldiers are visible bond who have interlinked the Nepalese culture and traditions in UK.

\textbf{Education cooperation}

UK is supporting US dollars 260,012 for the education of marginalised girl in Kailali district of Nepal. DFID was the core education funding agency in Nepal, supporting the foundations of education values. Active as a School Sector Reform Plan (SSRP) pooling Development Partner. With Education for All overall goals, the SSRP objectives were to 
\begin{enumerate}
\item ensure access and equity in primary education; 
\item improve the efficiency and institutional capacity of primary education; and 
\item enhance the quality and relevance of basic primary education for children and illiterate adults.
\end{enumerate}

In 2013, DFID committed around GBP 2 million to establish 810 additional girls' toilet and water supplies/WASH facilities across 18 districts selected based on low indicators in hygiene and sanitation with an implementation date which was extended to July 2015.

The British Government has been regularly providing scholarships in different areas for the development of human resource in Nepal since 1950s. Britain is offering Chevening Award to Nepalese from many years. A lot of Nepalese students are pursuing higher level studies in the UK. 

\textbf{Security cooperation}

UK has provided support for upgrading of Nepal Army Birendra Peace Operations Training Centre (BPOTC) located at Panchkhal (45 km East of Kathmandu). It has also provided Explosive Ordinance Device (EOD) related Improvised Explosive Device Disposal (IEDD) related Special Training to Nepal Army. UK is continually providing the special training courses to Nepal Army officials.

High level visit from Nepal Army and British Army are regularly taking place in UK and Nepal respectively.

\textbf{Nepalese students in UK}

\end{solution}

2. Discuss about the bilateral US relations with Nepal

**Solution:**

The United States recognized the Kingdom of Nepal on April 27, 1947, when Joseph C. Satterthwaite presented a letter from President Harry S. Truman to then king Tribhubana of Nepal. Satterthwaite was on a special diplomatic mission to Nepal as personal representative of the President. The two countries established diplomatic relations in 1948. Bilateral relations are friendly, and U.S. policy objectives toward Nepal center on helping it build a peaceful, prosperous, resilient and democratic society. Primary U.S. objectives in Nepal include supporting a stable, democratic Nepal that respects the rule of law; promoting investor-friendly economic development; and improving disaster risk management systems.

The United States enjoys a strong and positive relationship with Nepal. Years of diplomatic, development, and military engagement have advanced U.S. interests in transitioning Nepal into a more peaceful, stable democracy with significant economic potential. Since the end of its 10-year civil war in 2006 and the devastating earthquakes of 2015, Nepal has successfully transitioned into a full-fledged constitutional federal republic grounded in a constitution promulgated in 2015. With the recent formation of a new government, Nepal may now be on the cusp of a period of much-needed political stability.

Nepal has sought a bilateral consultation mechanism with the United States and has signed similar agreements with other countries like Australia in summer 2017. Such a forum might include as topics: Millenium Challenge Cooperation (MCC) coordination and implementation; trade (TPS, technical assistance, etc.); security and defense cooperation; and humanitarian assistance and disaster response.

To improve Nepal's economic situation, the MCC signed, in 2017, a \$500 million Compact with Nepal to expand Nepal’s electricity transmission infrastructure and improve its road maintenance regime. The Nepali government has committed another \$130 million for a program total of \$630 million. The Compact will build 300 kilometers (km) of high-voltage electric transmission lines, three substations, perform enhanced road maintenance on 305 km of strategic highways, and provide technical assistance to the national electric utility, the new electricity regulator, and the Department of Roads. The Compact is entered into force after the end of FY 2019.

U.S. assistance has been critical to helping Nepal rebuild after the devastating 2015 earthquake and remains committed to building Nepal's resilience in the event of any future disaster. The United States has provided over \$190 million for earthquake relief, recovery, and reconstruction. While these commitments have exceeded our initial pledge of \$130 million made at the International Conference on Nepal's Reconstruction in June 2015, significant recovery needs remain. To date, the United States has built 36 schools and hospitals; has directly helped rebuild over 16,000 homes; trained 260,000 people in safer construction; and developed policies, systems, and controls to ensure that $8.6 billion in reconstruction results in safer structures for all. In addition to rebuilding a safer Nepal, we are working to help Nepal institutionalize all lessons learned from the 2015 earthquake. We are also helping Nepal to implement its new disaster management law and stand up a new National Disaster Management Authority. We are further supporting Nepal as it introduces federalism by working with newly elected local governments to implement their own disaster management plans—thus helping local authorities meet commitments made to their constituencies. These efforts will help Nepal on its journey to self-reliance.

The United States has also committed security assistance to Nepal, working with the Nepali Army to strengthen their peacekeeping and disaster response capabilities.

The United States and Nepal have signed a trade and investment framework agreement, providing a forum for bilateral talks to enhance trade and investment, discuss specific trade issues, and promote more comprehensive trade agreements between the two countries. In 2016, Nepal became one of few countries in the world with a single-country trade preference program with the United States. Principal U.S. exports to Nepal include agricultural products, aircraft parts, optic and medical instruments and machinery. U.S. imports from Nepal include carpets, apparel and jewelry.

## Miscellaneous

### Impacts of climate change

What are the impacts of climate change ? Discuss on detail about the initiative taken by Nepal Government in tackling the impacts of climate change.

Climate change is now affecting every country in every continent. It is disrupting national economies and affecting lives, of individuals, communities and countries today and is likely to do so in the ensuing future. Weather pattern are changing, snowline are melting due to global warming and sea level are rising. 

**Negative impacts**

1. Numerous mountains are seeing accelerated the melting of snow due to global warming.
2. Changes in monsoon patterns, such as erratic rainfall with alternate drought periods are the serious impacts of climate change in Nepal which hampers the agricultural production. The changes in monsoon pattern in Nepal greatly exacerbate the situation of poverty and food insecurity in the country.
3. Flood, landslides, forest fires, snow melt are the disasters that occur as a result of climate change without any prewarning and causes heavy loss.
4. High hills and mountains are more vulnerable to climate change compared to terai, thus climate change make large scale economic and ecolgical disruption in upper latitudes.
5. Loss of plant and animal biodiversity.
6. Increase in number of insects, pest and diseases and sporadic outbreaks becoming much frequent.

The positive impact of global warming, a characteristic feature of climate change, is that the crops that could in the past be only growth in certain low altitudes, requiring prolonged exposure to heat can now be growth in higher regions such as Manang and Mustang districts of Nepal. The is evidently due to increase average temperature of the region, thus allowing the crop to perform well. However, as noted earlier, there are more prominent negative impacts than the benifits we can derive from the climate change.

Nepal became signatory of UN Framework Convention on Climate Change (UNFCC) in 1994. Ever since then the issue of climate change has been routinely incorporated into policy and strategic frameworks by Government of Nepal. Following policies and strategies are currently in place to help mitigate the impacts of climate change in Nepal:

1. Constitutional approach

Article 30(1) of the constitution of Nepal guarantees a "clean environment" as fundamental right and elaborates that "every citizen shall have the right to live in a clean and healthy environment." Article 30(2) of the constitution also encourages the state to formulate necessary legal framework to balance environment and development.

2. Approaches in periodic plans
3. Climate change policy, 2067 (2011)
4. National climate change policy, 2076 (2019)
6. National adaptation plan of action (NAPA, 2010)
7. Framework for local adaptation plan for action (LAPA, 2011)

### Highlights of NCCP, 2076

Reasons for Nepal's vulnerability to climate change are Topographical fragility, fragile geological structure, sensitive ecosystems and diversity of climate and micro-climate zones prevalence in Nepal. Additionally, poverty, illiteracy, social disparity and high dependence of communities on natural resources for livelihood.

- Nepal National Red Strategy implemented to check greenhouse gases emmission from deforestation and forest destruction.
- Provision of climate change coordination committee in Chairmanship of Prime minister.
- Establishment of Climate change management section under Ministry of Forest and Environment as focal point for implementation of international treaty.
- Provision of Climate budget code by National planning commission.
- Alternative energy promotion center has established climate and carbon unit for executing Clean Development Mechanism (CDM) related activities.

**Goal**

To contribute to socio-economic prosperity of the nation by building a climate resilient society.

**Objectives**

a. To enhance climate change adaptation capacity of persons, families, groups and communities vulnerable to, and at risk of, climate change;
b. To build resilience of ecosystems that are at risk of adverse impacts of climate change;
c. To promote green economy by adopting the concept of low carbon emmission development;
d. To mobilize national and international financial resources for climate change mitigation and adaptation in just manner;
e. To conduct research, make effective technology development and information service delivery related to climate change;
f. To mainstream or integrate climate change issues into policies, strategies, plans and programs at all levels of State and sectoral areas;
g. To mainstream gender equality and social inclusion (GESI) into climate change mitigation and adaptation programs.

**Policies** (Sectoral policy, strategies and working policies)

1. Agriculture and food security
2. Forest, biodiversity and watershed conservation
3. Water resources and energy
4. Rural and urban habitats
5. Industry, transport and physical infrastructure
6. Tourism and natural and cultural heritage
7. Health and drinking water and sanitation
8. Disaster risk reduction and management
9. Gender equality and social inclusion, livelihoods and good governance
10. Awareness raising and capacity development
11. Research, technology development and expansion
12. Climate finance management

Policy number 9-12 are inter-thematic areas.

### Population census, 2068

```{r}
dplyr::tribble(~"Indicator", ~"Unit", ~"Value", 
               "Male", "count", 12849041,
               "Female", "count", 13645463,
               "Annual population growth rate", "percentage", 1.35,
               "Population density", "individuals per msq", 180,
               "Total engagement in agriculture", "percentage", 65.6,
               "Male engagement in agriculture", "percentage", 60.2,
               "Female engagement in agriculture", "percentage", 72.8, 
               "Crude birth rate", "per thousand", 21.8,
               "Crude death rate", "per thousand", 7.3,
               "Child death rate", "per thousand", 40.5,
               "Net fertility rate", "percentage", 2.5, 
               "Average life expectancy", "years", 66.6) %>% 
  knitr::kable(caption = "Population census, 2068 summary", booktabs = TRUE, longtable = T)

```

### Hulaki rajmarga (Hulaki highway)

http://www.karma99.com/
http://www.karma99.com/2015/09/hulaki-rajmarga.html

### Demographics of Nepal

- Refer to wikipedia link \url{https://en.wikipedia.org/wiki/Demographics_of_Nepal}, for demographic statistics of Nepal. This includes
  - Population growth
  - Vital statistics: Live births per year, Deaths per year, Natural change per year, CBR, CDR, NC, TFR, IMR
  - Population structure: Based on age group
  - Ethnicity statisics
  - Languages
  - CIA factbook summary
  - Religion, etc.

### Madan puraskar

Madan puraskar is a literary honor in Nepal which Madan Puraskar Guthi confers annually for an outstanding book in Nepali language published within the calendar year. It is considered the most prestigious literature award in Nepal.

It was established by daughter in law of Prime Minister Chandra Shumsher JBR Queen Jagadamba Kumari Devi, in memory of her husband late General Madan Shumsher JBR, on September 26, 1955. Since then it has been presented every year, except in 1974 and 1977. At the time of its establishment, the award prize was NRS 4000. Since the Golden Jubilee year 2005 the award prize has been increased to NRs. 400,000.

List of winners

```{r winners-list}
# madan_puraskar <- read_html("https://en.wikipedia.org/wiki/Madan_Puraskar") %>% 
#   html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[2]')
# 
# madan_puraskar %>% 
#   .[[1]] %>%
#   html_table() %>% 
#   as_tibble() %>% 
#   write_csv("./madan_puraskar_awardees.csv", na = "")

madan_puraskar <- read_csv("./data/madan_puraskar_awardees.csv")

madan_puraskar %>% 
  knitr::kable(booktabs = TRUE, caption = "List of Madan Puraskar winners till 2075", longtable = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped", font_size = 8) %>% 
  kableExtra::column_spec(column = 2:3, width = c("14em", "24em"))
```


### Terrorism

Terrorism in its broadest sense, describes the use of intentionally indiscriminate violence as a means to create terror, or fear, to achieve political, religious or ideological aim. It is used in this regard primarily to refer to violence against peacetime targets or in war against non-combatants.

Terrorism has been practiced by political organizations with both rightist and leftist objectives, by nationalistic and religious groups, by revolutionarie, and even by state institutions such as armies, intelligence services, and police.

Terrorism is not legally defined in all jurisdictions; the statutes that do exist, however, generally share some common elements. Terrorism involves the use of threat of violence and seeks to create fear, not just within the direct victims but among a wide audience. The degree to which it relies on fear distinguishes terrorism from both conventional and guerilla warfare. Although conventional militiary foreces invariably engage in psychological warfare against the enemy, their principal means of victory is strength of arms. Similarly, guerrilla forces, which often rely on acts of terror and other forms of propaganda, aim at military victory and occassionally succeed (e.g. the Viet Cong in Vietnam and the Khmer Rouge in Cambodia). Terrorism proper is thus the systematic use of violence to generate fear, and thereby to achieve political goals, when direct military victory is not possible. This has led to some social scientists to refer to guerrilla warfare as the "weapon of the weak" and terrorism as the "weapon of the weakest".

Terrorism has been described variously as both a tactic and strategy; a crime and a holy duty; a justified reaction to oppression and an inexcusable abomination. Obviously, a lot depends on whose point of view is being represented. Terrorism has often been an effective tactic for the weaker side in a conflict. As an asymmetric form of conflict, it confers coercive power with many of the advantages of military force at a fraction of the cost. Due to secretive nature and small size of terrorist organizations, they often offer opponents no clear organization to defend against or to deter.

This is why pre-emption is being considered to be so important. In some cases, terrorism has been a means to carry out a conflict without the adversary realizing the nature of the threat, mistaking terrorism for criminal activity. Because of these characteristics, terrorism has become increasingly common among those pursuing extreme goals throughout the world. But despite its popularity, terrorism can be a nebulous concept.

The US Department of Defense defines terrorism as "the calculated use of unlawful violence or threat of unlawful violence to inculcate fear; intended to coerce or to intimidate governments or societies in pursuit of goals that are generally political, religious, or ideological."

The UN produced the following definition of terrorism in 1992; "An anxiety inspiring method of repeated violent action, employed by semi-clandestine individual, group or state actors, for idiosyncratic, criminal or political reasons, whereby -- in contrast to assassination -- the direct targets of violence are not the main targets."

The strategy of terrorists is to commit acts of violence that draws the attention of the local populace, the government, and the world to their cause. The terrorists plan their attack to obtain the greatest publicity, choosing targets that symbolize what they oppose.

The effectiveness of the terrorist act lies not in the act itself, but in the public's or governments reaction to the act. For example, in 1972 at the Munich Olympics, the Black September organization killed 11 Israelis. The Israelis were the immediate victims but the true target was the estimated 1 billion people watching the televised event. Those billion people watching were to be introduced to fear -- which is terrorism's ultimate goal. The introduction of this fear can be from the threat of physical harm/a grizzly death, financial terrorism from the fear of losing money or negative effects on the economy, cyber terrorism harming the critical technological infrastructures of society and psychological terrorism designed to influence people's behavior.

There are three perspectives of terrorism: the terrorist's, the victim's, and the general public's. The phrase "one man's terrorist is another man's freedom fighter" is a view terrorists themselves would gladly accept. Terrorists do not see themseleves as evil. They believe they are legitimate combatants, fighting for what they believe in, by whatever means possible to attain their goals. A victim of a terrorist act sees the terrorist as a criminal with no regard for human life. The general public's view though can be the most unstable. the terrorists tke great pains to foster a Robin Hood image in hope of swaying the general public's point of view towards their cause. This sympathetic view of terrorism has become an integral part of their psychological warfare and has been countered vigourously by governments, the media and other organizations.

Technolgical advances, such as automatic weapons and compact electrically detonated explosives, gave terrorists a new mobility and lethality, and the growth of air travel provided new methods and opportunities. Terrorism was virtually an official policy in totalitarian states such as those of Nazi Germany under Adolf Hitler and the Soviet Union under Stalin. In these states arrest, imprisonment, tourture, and execution were carried out without legal guidance or restraints to create a climate of fear and to encourage adherece to the national ideology and the declared economic, social, and political goals of the state.

One of the deadliest terrorist strikes to date were the September 11 attacks of 2001, in which suicide terrorists associated with al-Qaeda hijacked four commercial airplanes, crashing two of them into the twin towers of the World Trade Center complex in the New York city and the third into the Pentagon building near Washington D.C.; the fourth plane crashed near Pittsburgh, Pennsylvania. The crashes destroyed much of the World Trade Center complex and a large portion of one side of the Pentagon and killed more than 3000 people.

Terrorism appears to be an enduring feature of political life. Even prior to the September 11 attacks, there was widespread concern that terrorist might escalate their destructive power to vastly greater proportions by using weapons of mass destruction -- including nuclear, biological, or chemical weapons -- as was done by the Japanese doomsday cult AUM Shirnrikyo, which released nerve gas into a Tokyo subway in 1995.

Factors conducive to terrorism

1. Poverty
2. Globalization
3. Legitimate grievances and failure of governments.
4. Humiliation
5. Lack of democracy, and widespread and systematic violation of human rights
6. Lopsided foreign policy
7. Failed state governance and authority

Goals, strategies, adn weapons of terrorism

1. Social and political justice
2. Self determination
3. Racial superiority
4. Foreign policy
5. Publicty
6. Demoralized governments

Costs associated with terrorism are so widespread, complex, and intangible that they are virtually impossible to measure. Individuals, families, governments, companies and non-state actors bear the time, money and resource costs of terrorism to varying degrees. Migration, trade, travel and interpersonal relations are affected. It is generally believed that the financial crisis and the global recesssion were caused in by policies adopted by the Bush administration to fight global terrorism, including wars in Afganistan and Iraq.

Kinds of terrorism

Although the types of terrorism tend to overlap, they vary in their implications and affect us in different ways. For example, the indiscriminate nature of global terrorism contrasts sharply with domestic terrorism aimed at specific groups or governments.

1. Domestic terrorism occurs within the borders of a particular country and is associated with extremist groups.
2. Nationalist terrorism is closely associated with struggles for political autonomy and independence.
3. Religious terrorism grows out of extreme fundamentalist religous groups that believe that God is on their side and that their violence is divinely justified and inspired.
4. State terrorism is a cold, calculated, efficient and extremely destructive form of terrorism, partly because of the overwhelming power at the disposal of governments.

Leading terrorist organizations

- ISIS (part of northern Iraq and parts of Western Syria)
- Al-Qaeda (Afganistan)
- Taliban
- Boko Haram (now sworn to ISIS; Nigeria)
- Lashkar-e-toa (Pakstan)
- Hezbollah (Lebanon)
- Al-Shabaab
- Hamas (Palestine)
- Naxal/Naxalites (India)
- Irish Republican Army

No cause justified terrorism. The world must repond and fight vigillantly with this evil that is intent on threatening and destroying our basic feedoms and way of life. The complexity of terrorists requires employing a wide variety of instruments to combat it. It addition to military, traditional law enforcement, intelligent responses, and global cooperation has become obvious to counter terrorist attacks.

The most prevalent response to terrorism is the use of force, both domestically and internationally. Besides, reduction in the production of weapons of mass destruction and control over media and technology can be effective.

We must retaliate forcefully and successfully in order to reassure the world of our confidence as a great power, and our ability to retain that status. Lastly, and practically, we must explore the possibilities of terrorism that exist and attempt to grasp the actual threat that these possibilities realistically pose for us.

### Globalization

Globalisation is the process of interaction and integration among people, companies, and governments worldwide. As a complex and multifaceted phenomenon, globalization is considered by some as a form of capitalist expansion which entails the integration of local and national economies into a global, unregulated market economy. Globalization has grown due to advances in transportation and communication technology. With the increased global interactions comes the growth of international trade, ideas, and culture. Globalization is primarily an economic process of interaction and integration that's associated with social and cultural aspects. However, conflicts and diplomacy are also large parts of the history of globalization, and modern globalization.

Economically, globalization involves goods, services, the economic resources of capital, technology, and data. Also, the expansions of global markets liberalize the economic activities of the exchange of goods and funds. Removal of cross-border trade barriers has made formation of global markets more feasible.[citation needed] The steam locomotive, steamship, jet engine, and container ships are some of the advances in the means of transport while the rise of the telegraph and its modern offspring, the Internet and mobile phones show development in telecommunications infrastructure. All of these improvements have been major factors in globalization and have generated further interdependence of economic and cultural activities around the globe.

Though many scholars place the origins of globalization in modern times, others trace its history long before the European Age of Discovery and voyages to the New World, some even to the third millennium BC. Large-scale globalization began in the 1820s. In the late 19th century and early 20th century, the connectivity of the world's economies and cultures grew very quickly. The term globalization is recent, only establishing its current meaning in the 1970s.

Three major dimensions of globalization can be differentiated. They are:

Economic globalization

It is the increasing economic interdependence of national economies across the world through a rapid increase in cross-border movement of goods, services, technology, and capital. Whereas the globalization of business is centered around the diminution of international trade regulations as well as tariffs, taxes, and other impediments that suppresses global trade, economic globalization is the process of increasing economic integration between countries, leading to the emergence of a global marketplace or a single world market. Depending on the paradigm, economic globalization can be viewed as either a positive or a negative phenomenon. Economic globalization broadly comprises: globalization of production and market but also includes competition, technology, and corporations and industries.

Cultural globalization

Cultural globalization refers to the transmission of ideas, meanings, and values around the world in such a way as to extend and intensify social relations. This process is marked by the common consumption of cultures that have been diffused by the Internet, popular culture media, and international travel. This has added to processes of commodity exchange and colonization which have a longer history of carrying cultural meaning around the globe. The circulation of cultures enables individuals to partake in extended social relations that cross national and regional borders. The creation and expansion of such social relations is not merely observed on a material level. Cultural globalization involves the formation of shared norms and knowledge with which people associate their individual and collective cultural identities. It brings increasing interconnectedness among different populations and cultures. Cross-cultural communication, cultural diffusion and religious thought-schools are the major aspects of cultural globalization. Sports too bear the spirit of globalization; for eg., the FIFA World Cup is the most widely viewed and followed sporting event in the world, exceeding even the Olympic Games; a ninth of the entire population of the planet watched the 2006 FIFA World Cup Final.

Political globalization

Political globalization refers to the growth of the worldwide political system, both in size and complexity. That system includes national governments, their governmental and intergovernmental organizations as well as government-independent elements of global civil society such as international non-governmental organizations and social movement organizations. One of the key aspects of the political globalization is the declining importance of the nation-state and the rise of other actors on the political scene. Adherence to or practice of intergovernmentalism, multi-level governance and multiple citizenship reflect today's globalized politics, although it should also be noted that some countries, like North Korea, have embraced isolationist policies.

Aside from the the above mentioned aspects, globalization touches upon a wide range of issues including population migration, information exchange, environmental protection.

\begin{questions}

\question How does the management information system help in decision making process ?

Management information system (MIS) is an information system used for decision-making, and for the coordination, control, analysis, and visualization of information in an organization.

In public governance, MIS are the tools to promote transparency, accountability and participation. More specifically, recent approaches to delivering bureaucratic services are expeceted to meet their goals of public orientation and result based performace evaluation aided through information system. Establishment of computerised Integrated Financial Management Information System (IFMIS) is being mulled over to implement single platform central accouting system as a part of a public finance management reform initiative by Government of Nepal with support of World Bank. For example, under this system information about expenditure made by the Department of Roads and the revenue collected by the Inland Revenue Department will be disseminated via the same platform. Also the concept of digital National Identity Card based is a part of broader vison of government, currently under delegation of the National ID Management Center (NIDMC) under aegis of Ministry of Home Affairs, to aggregate vital information database of the citizens.

In a corporate setting, the ultimate goal of the use of a management information system is to increase the value and profits of the business. This is done by providing managers with timely and appropriate information allowing them to make effective decisions within a shorter period of time.

While management information systems can be used by any and every level of management, the decision of which systems to implement generally falls upon the chief information officers (CIO) and chief technology officers (CTO). These officers are generally responsible for the overall technology strategy of an organization including evaluating how new technology can help their organization. Once decisions have been made, IT directors, including MIS directors, are in charge of the technical implementation of the system. Upon implementation, the assigned users will have the appropriate access to relevant information. It is important to note that not everyone inputting data into MIS need necessarily be management level.

The role of MIS in public and private governance is further invigorated by widespread networking facilities made possible through internet. There are also multiple cases of MIS platform sharing across multiple business centres to benefit from cross enterprise information exchange.

The MIS represents the electronic automation of several different kinds of counting, tallying, record-keeping, and accounting techniques of which the by far oldest, of course, was the ledger on which the business owner kept track of his or her business. Automation emerged in the 1880s in the form of tabulating cards which could be sorted and counted. These were the punch-cards still remembered by many: they captured elements of information keyed in on punch-card machines; the cards were then processed by other machines some of which could print out results of tallies. 

Waves of innovation spread the fundamental virtues of coherent information systems across all corporate functions and to all sizes of businesses in the 1970s, 80s, and 90s. Within companies major functional areas developed their own MIS capabilities; often these were not yet connected: engineering, manufacturing, and inventory systems developed side by side sometimes running on specialized hardware. Personal computers ("micros," PCs) appeared in the 70s and spread widely in the 80s. Some of these were used as free-standing "seeds" of MIS systems serving sales, marketing, and personnel systems, with summarized data from them transferred to the "mainframe." In the 1980s networked PCs appeared and developed into powerful systems in their own right in the 1990s in many companies displacing midsized and small computers. Equipped with powerful database engines, such networks were in turn organized for MIS purposes. Simultaneously, in the 90s, the World Wide Web came of age, morphed into the Internet with a visual interface, connecting all sorts of systems to one another.

Midway through the first decade of the 21st century the narrowly conceived idea of the MIS has become somewhat fuzzy. Management information systems, of course, are still doing their jobs, but their function is now one among many others that feed information to people in business to help them manage. Systems are available for computer assisted design and manufacturing (CAD-CAM); computers supervise industrial processes in power, chemicals, petrochemicals, pipelines, transport systems, etc. Systems manage and transfer money worldwide and communicate worldwide. Virtually all major administrative functions are supported by automated system. Many people now file their taxes over the Internet and have their refunds credited (or money owning deducted) from bank accounts automatically. MIS was thus the first major system of the Information Age. 

Advantages and disadvantages

The following are some of the benefits that can be attained using MIS:

\begin{itemize}
\item Improve an organization's operational efficiency, add value to existing products, engender innovation and new product development, and help managers make better decisions.
\item Companies are able to identify their strengths and weaknesses due to the presence of revenue reports, employee performance records etc. Identifying these aspects can help a company improve its business processes and operations.
\item Giving an overall picture of the company.
\item Acting as a communication and planning tool.
\item The availability of customer data and feedback can help the company to align its business processes according to the needs of its customers. The effective management of customer data can help the company to perform direct marketing and promotion activities.
\item MIS can help a company gain a competitive advantage.
\item MIS reports can help with decision-making as well as reduce downtime for actionable items.
\end{itemize}

Some of the disadvantages of MIS systems:
\begin{itemize}
\item Retrieval and dissemination are dependent on technology hardware and software.
\item Potential for inaccurate information.
\end{itemize}

source: \url{https://www.inc.com/encyclopedia/management-information-systems-mis.html}; \url{https://en.wikipedia.org/wiki/Management_information_system}

\question What do you mean by inventory management ("Jinsi byabasthapan") ? Highlight about its importance in Nepalese public administration.

\question Write short notes on:
\begin{enumerate}[label*=\alph*., leftmargin=1cm]
\item Tax reduction at source
\item Civil service 
\end{enumerate}

source: \url{https://kathmandupost.com/money/2017/06/07/integrated-financial-info-system-mulled}

\question What are the geographical features of Nepal ? Mention in brief.


\question (Long answer type) What is energy crisis ? What are the measures that can be taken to solve the crisis ? What are the different sectors affected by energy crisis ?

\question Explain briefly on the prospect of hydroelectricity development in Nepal.

\question What is human development index (HDI) ? Explain in brief.

\question Discuss in brief the role of private sector in economic development of Nepal. 

\question What is poverty ? What is the difference between absolute and relative poverty ?

\question How are indigeneous and local ethnic groups prioritized in Nepal ? Mention what strategies are taken by government to that end.

\question (Long answer type) "Foreign employment significantly contibutes to customs earning of Nepal". Briefly elaborate the statement.

\question What are the different types of climates occuring in Nepal based on the geographical features ?

\question Write in brief about the importance of water resources of Nepal.

\question What are the major causes of deforestation in Nepal. Discuss.

\question What is the status of environmental pollution in Nepal ? Discuss about the efforts made by Government of Nepal in reducing effects of environmental pollution and preparing for climate change ?

\question What is meant by standard time ? Mention how is standard time of Nepal determined ?

\question Justify: Unplanned urbanization negatively impacts development of nation.

\end{questions}

# World history

## French revolution

The social and political upheavals in France, spreading subsequently to most Europe over the generation after 1789, constitute the classic revolution of modern times. 

The trigger for the French revoultion was a financial crisis. The government of King Louis XVI, already burdened by substantial debts incurred to finance midcentury wars, intervened between 1778 and 1783 in the American War of Independence in the hope of destroying the power of Great Britain. Taxing too little and borrowing too much on disadvantageous terms the king found himself by 1786 facing bankrupcy. To avoid it, an ambitious plan of reform was formulated, but when the plan was laid before a handpicked Assembly of Notables in 1787, members claimed, after much detailed criticism, that the only body competent to approve such far-reaching changes was the elected Estates-General, which had not met since 1614. From the dissolution of the Notables (May 1787) until August 1788 the government struggled to avoid convening a body that seemed bound to diminish hitherto absolute royal power. Financial paralysis, however, forced Louis XVI to convoke the Estates for 1789.

The composition being three houses representing the orders of clergy, nobility, and third estate (comprising everyone else in the kingdom), numerically equal in membership but voting by order. Thus the first two "privileged orders" could outvote the third estate, even though it represented the vast majority of the nation. By the time of the elections in spring 1789, in reponse to massive publicity campaign, double representation had been conceded to the third estate, but it remained undecided whether votes would be counted by order, preserving the dominance of the clergy and nobility, or by head, which would give the third estate -- now doubled -- an equal vote. The issue, unresolved until six weeks after the Estates convened, dominated the elections, producing a groundswell of hostility toward the nobility.

Frustrated by noble ineptitude, on June 17 the third estate, reinforced by a few renegade clergy, declared itself the National Assembly, with full soverign powers. Attempts by the king and certain nobles to resist the claims of the third estate, ultimately with the threat of military force, produced rising popular alarm, which culminated on July 14 with a revolt in Paris as a terrified, hungry populace scoured the city for arms. With the help of military mutineers, the people stormed the main state prison, the Bastille. Unsure of his troops' reliability, the king abandoned his resistance to the seizure of soverignty by the French nation's elected representatives.

... continue

source: \url{The Encyclopedia of Political Revolutions}

## American revolution

The American Revolution was a colonial revolt which occurred between 1765 and 1783. The American Patriots in the Thirteen Colonies defeated the British in the American Revolutionary War (1775–1783) with the assistance of France, winning independence from Great Britain and establishing the United States of America.

The American colonials proclaimed "no taxation without representation" starting with the Stamp Act Congress in 1765. They had no representatives in the British Parliament and so rejected Parliament's authority to tax them. Protests steadily escalated to the Boston Massacre in 1770 and the burning of the Gaspee in Rhode Island in 1772, followed by the Boston Tea Party in December 1773. The British responded by closing Boston Harbor and enacting a series of punitive laws which effectively rescinded Massachusetts Bay Colony's rights of self-government. The other colonies rallied behind Massachusetts, and a group of American Patriot leaders set up their own government in late 1774 at the Continental Congress to coordinate their resistance of Britain; other colonists retained their allegiance to the Crown and were known as Loyalists or Tories.

Tensions erupted into battle between Patriot militia and British regulars when King George's forces attempted to destroy American military supplies at Lexington and Concord on April 19, 1775. The conflict quickly escalated into war, during which the Patriots (and later their French allies) fought the British and Loyalists in the Revolutionary War. Each of the thirteen colonies formed a Provincial Congress which assumed power from the former colonial governments, suppressed Loyalism, and recruited a Continental Army led by General George Washington. The Continental Congress declared King George a tyrant who trampled the colonists' rights as Englishmen, and they declared the colonies free and independent states on July 2, 1776. The Patriot leadership professed the political philosophies of liberalism and republicanism to reject monarchy and aristocracy, and they proclaimed that all men are created equal.

The Patriots attempted to invade Canada during the winter of 1775–76 without success, expecting like-minded colonists in British Canada to rally to the cause. The newly created Continental Army forced the British military out of Boston in March 1776, but the British captured New York City and its strategic harbor that summer, which they held for the duration of the war. The Royal Navy blockaded ports and captured other cities for brief periods, but they failed to destroy Washington's forces. The Continental Army captured a British army at the Battle of Saratoga in October 1777, and France then entered the war as an ally of the United States. Britain then refocused its war to make France the main enemy. Britain also attempted to hold the Southern states with the anticipated aid of Loyalists, and the war moved south. Charles Cornwallis captured an army at Charleston, South Carolina in early 1780, but he failed to enlist enough volunteers from Loyalist civilians to take effective control of the territory. Finally, a combined American and French force captured a second British army at Yorktown in the fall of 1781, effectively ending the war. The Treaty of Paris was signed September 3, 1783, formally ending the conflict and confirming the new nation's complete separation from the British Empire. The United States took possession of nearly all the territory east of the Mississippi River and south of the Great Lakes, with the British retaining control of Canada, and Spain taking Florida.

Among the significant results of the Revolution were American independence and friendly economic trade with Britain. The Americans adopted the United States Constitution, establishing a strong national government which included an elected executive, a national judiciary, and an elected bicameral Congress representing states in the Senate and the population in the House of Representatives. Around 60,000 Loyalists migrated to other British territories, particularly to British North America (Canada), but the great majority remained in the United States. 
