\documentclass[a4paper,12pt]{exam}

\usepackage[onehalfspacing]{setspace}
\usepackage{graphicx}
\usepackage{amsmath,amssymb,amsfonts}

\title{\vspace{-1.5cm}{\huge \textbf{Quiz contest, 2076}}\\[2mm]
\textbf{Gokuleshwor Agriculture and Animal Science College, Gokuleshwor, Baitadi} \\[2mm]
\hrule
}

\date{}

% print correct answers and fillins

\printanswers
\CorrectChoiceEmphasis{\bfseries\boldmath}

% print marks weight in the right
\pointsinrightmargin
\bracketedpoints

% adding gradetable
\addpoints

% the \llap command to correct alignment in a question with parts.
\renewcommand{\questionlabel}{\llap{Q}\,\thequestion.}
\unframedsolutions % put solution inside a box

\begin{document}

\maketitle

\section*{Rounds}

\subsection*{Choose your topic Round}

{\Large \textit{Instructions}}

\begin{itemize}
\item	Each team selects a topic they would like to attempt.
\item All topics are related to the field of Agriculture Sciences.
\item	Each topic will pose 10 questions for the team.
\item	10 points will be awarded on every single question, upon correctly answering. Hence a team can secure a maximum of 100 points in this round.
\item No deduction in score will be made for answering the question incorrectly.
\end{itemize}

Option topics are:

\begin{enumerate}
  \item Agricultural economics and extension
  \item Plant pathology
  \item Entomology
  \item Horticulture
  \item Agronomy, genetics and plant breeding
  \item Agricultural engineering and soil science
  \item Animal sciences and industrial agriculture
\end{enumerate}

\subsection*{General knowledge Round}

{\Large \textit{Instructions}}

\begin{itemize}
\item Each team will be asked question for the respective choice of number they make.
\item There will be a total of N x 5 questions asked.
\item Correctly answering a question on first hand will reward the team with 15 marks.
\item Incorrectly answering a question will reduce 5 marks in the team's score whereas passing the question will not lead to deduction of score.
\item Question will be passed to the next team upon answering incorrectly with a reduction of 5 marks in the previous value of question.
\end{itemize}


\subsection*{Double trouble Round}

{\Large \textit{Instructions}}

\begin{itemize}
\item Question topics will be general type. 
\item Each question carries 10 points.
\item Each team can choose to double the points on offer. 
\item If they get the answer correctly, points will be doubled. However, if the answer is wrong 5 points will be deducted from their score.
\item The question will not be passed to the other teams.
\end{itemize}

\subsection*{Get a clue Round}

{\Large \textit{Instructions}}

Each team will be asked a (typically difficult) question. The team tries to answer in following way:

\begin{itemize}
\item There are 2 x N questions available.
\item	No score deduction will be made in case a team gets answer wrong.
\item	The questions will be passed to the next team for the current value of answer.
\item	Team can choose not to obtain clue and get 15 points for correctly answering.
\item	Team can choose to obtain first clue and get 10 points for correctly answering.
\item	Team can choose to obtain second clue and get 5 points for correctly answering.
\end{itemize}

\subsection*{Just list it Round}

{\Large \textit{Instructions}}

\begin{itemize}
\item Each team will be asked a question that demands multiple answers in a list. 
\item Each question carries 15 marks.
\item Each correct item in the answer to the question will be awarded equal points.
\item When decimal score is obtained it it rounded to nearest value of 5 in 1st digit place.
\end{itemize}


\subsection*{Connect question Round}

{\Large \textit{Instructions}}

\begin{itemize}
\item Pictures or audible clues will be provided and question related to the clue will be asked. 
\item There will be a total of N x 2 questions.
\item Each question carries 15 marks.
\item Question will not be passed if answered incorrectly.
\end{itemize}

\subsection*{Rapid fire Round}

{\Large \textit{Instructions}}

\begin{itemize}
\item There will be a total of 10 questions for a team to be answered within a 2 minutes time period. 
\item Questions will be homogenous and of similar lengths.
\item There will be no deduction of score for getting answer incorrectly.
\item Each question carries 10 marks. Hence the maximum score a team can secure in this round is 100. 
\end{itemize}

\end{document}
