\documentclass[10pt,a4paper,answers]{exam}
\usepackage{amsmath,amsthm,amsfonts,amssymb}
\usepackage[left=1in,right=0.5in,top=0.75in,bottom=0.75in]{geometry}
\usepackage{ifthen}

\printanswers
\CorrectChoiceEmphasis{\bfseries\boldmath}

\begin{document}
\begin{center}
{\textbf {\LARGE Get a clue}}
\end{center}

\begin{questions}
  \question State the year of release of the first commercial national Hybrid cereal of Nepal ? \fillin[2069 BS or 2012 AD][3cm]
  \begin{solution}
  The commercial hybrid belongs to the Maize crop.
  The commercial hybrid name is RH-2.
  \end{solution}

\question Name the micro nutrient that has a significant role in pollination and seed development. \fillin[Boron][3cm].
  \begin{solution}
  The element is also required for effective nodulation in legume crops.
  The deficiency of this causes muliple cobs or barren cobs in maize.
  It is also reponsible for hollow heart of cauliflower.
  \end{solution}
  
\question When was the DDT first ever banned ? \fillin[1972][3cm].
  \begin{solution}
  The DDT was banned first in US
  It was banned after 10 years that Rachael Carson published her famous book.
  \end{solution}

\question For which crop is the Scripo-lure used ? \fillin[Rice][3cm].
  \begin{solution}
  The larval stage of the pest is the most harmful stage for crop.
  The lure is effective against yellow stem borer insect.
  \end{solution}

\question BIMSTEC's five of six core areas of cooperation includes agriculture, energy, tourism, trade and transportation. What is the missing area ? \fillin[Fishery][3cm].
  \begin{solution}
  The missing area is mostly benefitted when a country is not-landlocked
  The missing area is related to obtaining food.
  \end{solution}

\question Except honey bees and flies, yet another insect assists in pollination of Litchi. What is it ? \fillin[Ant][3cm].
  \begin{solution}
  It is primarily a ground dweller.
  It belongs to order Hymenoptera.
  \end{solution}

\question Anethum graveolens is the scientific name of a agriculural crop. What is the english or nepali name of the crop ? \fillin[Dill][3cm].
  \begin{solution}
  It is a spice crop.
  It belongs to family umbelliferae.
  \end{solution}

\question SALT-2 technology lays major emphasis to what component of agriculture ? \fillin[Livestock][3cm].
  \begin{solution}
  It is best suited for small holder farmers
  It is short for Simple Agro-Livestock Land Technology.
  \end{solution}

\question Silver revolution is related to ? \fillin[Poultry and egg production][4cm].
  \begin{solution}
  It advocated for more animal protein nutrition in human diet.
  It is connected to poultry.
  \end{solution}

\question What is Cyperus rotundus mostly propagated by ? \fillin[stolon][3cm].
  \begin{solution}
  It is a subsurface structure.
  It has very long survival period underground.  
  \end{solution}

\question The average productivity of vegetables produced in Nepal in 2015/16 AD was ? \fillin[13992 kg/ha][3cm]. (A margin of error of 500 kgs will be tolerated)
  \begin{solution}
  It is almost 4.5 times to that of rice.
  The yield is marginally (3000 kg) less than that of Banana. 
  \end{solution}

\question Which fish has drastically different source of major nutrition (i.e. is a carnivore at early stages and a herbivore later) on different stages of maturity ? \fillin[Sahar][3cm].
  \begin{solution}
  It is mostly regarded as cold water fish.
  It is abundantly found in Far-western midhill region.
  \end{solution}

\end{questions}

\end{document}
