\documentclass[10pt,a4paper,answers]{exam}
% \documentclass[10pt,a4paper]{exam}
\usepackage{amsmath,amsthm,amsfonts,amssymb}
\usepackage[left=1in,right=0.5in,top=0.75in,bottom=0.75in]{geometry}
\usepackage{ifthen}

% \printanswers
\CorrectChoiceEmphasis{\bfseries\boldmath}

\begin{document}
\begin{center}
{\textbf {\LARGE Rapid fire}}
\end{center}

{\Large
\textit{Instructions}
}

\begin{enumerate}
  \item There will be a total of 10 questions for a team to be answered within a 2 minutes time period. 
  \item Questions will be homogenous and of similar lengths.
  \item There will be no deduction of score for getting answer incorrectly.
  \item Each question carries 10 marks. Hence the maximum score a team can secure in this round is 100. 
\end{enumerate}

\vspace{1cm}

{\Large \textbf{Questions} \\[1.5mm] 
\hrule
}

\section*{Set 1}

\begin{questions}
\question What is the name of longest river of Nepal ? \fillin[Karnali][3cm].
\question In which country is the southernmost city of the world, except Antarctica, located ? \fillin[Chile][3cm].
\question "Baraha Kshetra" is in which district of Nepal ? \fillin[Sunsari][3cm]. 
\question What imparts the characteristic yellow coloration of Ginger ? \fillin[Curcumin][3cm].
\question The cooperative movement was first started in which country ? \fillin[England][3cm].
\question The country of origin of soybean is ? \fillin[China][3cm].
\question One acre is equivalent to how much hectares ? \fillin[0.404][3cm].
\question What is the seed rate for tobaccoo ? \fillin[2-3 kg per ha][3cm].
\question Dew, mist, rain and fog, all occur in which layer of atmosphere ? \fillin[Trophosphere][3cm].
\question Entry of water from air to soil is called \fillin[Infilteration][3cm].
\end{questions}

\section*{Set 2}

\begin{questions}
\question What is the largest freshwater lake of the world ? \fillin[Lake superior][3cm].
\question As of most recent survey, how many agriculture cooperatives are there currently in Nepal? \fillin[9965][3cm].
\question What is the process of delivering artificial heating condition for the hatching of egg called ? \fillin[Incubation][3cm].
\question What is the scientific name of Danphe ? \fillin[Lophophorus impejanus][3cm].
\question Removal of terminal bud in tobaccoo is called \fillin[Topping][3cm].
\question The distress occuring due to lack of dissolved oxygen in pond is called \fillin[Asphyxiation][3cm].
\question Hybrid vigor was suggested by \fillin[GH Shull][3cm].
\question Pyrilla is a major insect of which crop ? \fillin[Sugarcane][3cm].
\question Onion belongs to the family ? \fillin[Amaryllidaceae][3cm].
\question Hexagonal method of orchard layout accomodates how many percentage more seedlings than square method ? \fillin[15\%][3cm].
\end{questions}

\section*{Set 3}

\begin{questions}
\question "Seto Dharati" nobel was written by which nobelist ? \fillin[Amar Neupane][3cm].
\question What is the electricity production capacity of "Large" hydroelectricity projects in Nepal ? \fillin[100-1000 Mega Watt][3cm].
\question Who is also known as "Father of Mutation" ? \fillin[Hugo de Vries][3cm].
\question When was structure of DNA explained first ? \fillin[1953][3cm].
\question Optimum temperature for growth and development of rainbow trout is \fillin[16-18 degree Centigrade][3cm].
\question Canopy temperature is measured by \fillin[Infra-red thermometer][4cm].
\question The species of rhizobium contained in root nodules of the soybean is \fillin[Bradyrhizobium japonicum][4cm].
\question Carbohydrate content in potato is \fillin[70\%][3cm].
\question Give the name of any chemical that is used for hairy caterpillar. \fillin[Dimethoate, Chlorpyriphos][4cm]. 
\question Chiasmata formation occurs during which stage of prophase \fillin[Diplotene][3cm].
\end{questions}

\section*{Set 4}

\begin{questions}
\question What is the name of national park nearest to Kathmandu ? \fillin[Shivapuri National Park][5cm].
\question COP organized in 2019 at Santiago, Chile was of which rank ? \fillin[25th][3cm]. 
\question What is "Agenda-21" related to ? \fillin[Environment and Development][5cm].
\question There are more than 1500 species of Fungi in Nepal (True or False). \fillin[True][3cm].
\question Meiosis was first observed by \fillin[Farmer and Moore][3cm].
\question Where did cabbage originate ? \fillin[West europe][3cm].
\question What is LSD drug obtained from fungus used as ? \fillin[Stimulant][3cm].
\question Translation is the process of \fillin[Protein synthesis][5cm].
\question Black tip of mango is caused by \fillin[Smokes from brick klin][4cm].
\question Fungi is better suited to alkaline pH condition. True or False. \fillin[False][5cm].
\end{questions}

\section*{Set 5}
\begin{questions}
\question What is the process of digestion of cell by another cell called ? \fillin[Phagocytosis][3cm].
\question In which year was Nepal Red Cross society first established ? \fillin[2020 BS][3cm].
\question ECG is an instrument used in diagnosis of which human organ ? \fillin[Heart][3cm].
\question What is the full form of OLED in Television technology ? \fillin[Organic Light Emitting Diode][5cm].
\question Which chemical is used in the induction of polypoloidy ? \fillin[Colchicine][3cm].
\question What is the scientific name of Bitter gourd ? \fillin[Momordica charantia][3cm].
\question Yeast doesn't produce mycellium ? True or False \fillin[True][3cm].
\question mRNA codon for synthesis of methionine is \fillin[AUG][3cm].
\question Silkworm's best source of food is \fillin[Mulberry][3cm] crop.
\question Corona virus is assumed to have transmitted to humans from which organism ? \fillin[Bat][3cm].
\end{questions}


\end{document}