% \documentclass[10pt,a4paper,answers]{exam}
\documentclass[10pt,a4paper]{exam}
\usepackage{amsmath,amsthm,amsfonts,amssymb}
\usepackage[left=1in,right=0.5in,top=0.75in,bottom=0.75in]{geometry}
\usepackage{ifthen}

% \printanswers
\CorrectChoiceEmphasis{\bfseries\boldmath}

\begin{document}
\begin{center}
{\textbf {\LARGE Just list it}}
\end{center}

{\Large
\textit{Instructions}
}

\begin{enumerate}
\item Each team will be asked a question that demands multiple answers in a list. 
\item Each question carries 15 marks.
\item Each correct item in the answer to the question will be awarded equal points.
\item When decimal score is obtained it it rounded to nearest value of 5 in 1st digit place.
\end{enumerate}

\vspace{1cm}

{\Large
\textbf{Questions} \\[1mm]
\hrule
}


\begin{questions}

\question Human Development Index (HDI) is a composite of three indicators. What are those?

\begin{solution}
\begin{enumerate}
\item Life expectancy
\item Education
\item Per capita income
\end{enumerate}
\end{solution}

\question Which 3 countries occur among the top 5 producers of paddy in the world ?
  \begin{solution}
  \begin{enumerate}
  \item China
  \item India
  \item Indonesia
  \item Bangladesh
  \item Vietnam
  \end{enumerate}
  \end{solution}

\question Enlist 3 plants and animals and animal species with their number of chromosomes.
\question	Enlist 6 of top highest 10 mountain peaks of Nepal.
\question Enlist 3 of top 5 biggest deserts of the world.

\question Enlist 3 are the biological pesticides or biocontrol agents ?
  \begin{solution}
  \begin{enumerate}
  \item Beauveria bassiana
  \item Metarhizium anisopliae
  \item Bacillus thuringiensis kurstaki
  \item NPV (Heli and Spodo)
  \item Trichoderma viridae and T. herzianum
  \item Pseudomonas fluorescence
  \end{enumerate}
  \end{solution}

\question Enlist any 3 crops with specification of their optimum pH ranges ?
  \begin{solution}
  \begin{enumerate}
  \item Paddy: 4-6.5
  \item Wheat: 5.5-7.5
  \item Maize: 5.5-7.5
  \item Barley: 6.5-8.5
  \item Oat: 5.0-7.0
  \item Linseed: 6.0-7.0
  \item Groundnut: 5.0-6.0
  \item Soybean: 6.0-6.8
  \item Mustard: 6.0-7.5
  \item Tea: 4.5-5.5
  \end{enumerate}
  \end{solution}

\question Enlist any 6 pesticides that are banned in Nepal ?
  \begin{solution}
  \begin{enumerate}
  \item Chlorodane 2001 AD
  \item DDT 2001 AD
  \item Dieldrin 2001 AD
  \item Endrin 2001 AD
  \item Aldrin 2001 AD
  \item Mirex 2001 AD
  \item Toxaphen 2001 AD
  \item BHC 2001 AD
  \item Lindane 2001 AD
  \item Phosphamidan 2001 AD
  \item Organo mercuric compounds 2001 AD
  \item Methyl parathion 2007 AD
  \item Monocrotophos 2007 AD
  \item Endosulphan 2014 AD
  \item Phorate 2016
  \end{enumerate}
  \end{solution}

\question Enlist any 6 released rice varieties of Nepal.

\question Enlist any 6 cross pollinated crops.

\question Enlist any 3 varieties of wheat recommended for terai domain of Nepal.

\end{questions}

\end{document}